### 2013-12-18 paper.tex
initialised repo

\documentclass[final,leqno]{siamltex}
%\documentclass[draftcls, onecolumn, 11pt]{../bib/IEEEtran}
%\documentclass[journal]{../bib/IEEEtran}
%\documentclass[conference]{../bib/IEEEtran}

\usepackage{../bib/mathbf-abbrevs}

\input{defs}

\title{Finding a closest point in a lattice of Voronoi's first kind}

\author{Robby~G.~McKilliam, Alex Grant and I. Vaughan L. Clarkson}

\begin{document}

% make the title area 
\maketitle

 \begin{abstract} 
We show that for those lattices of Voronoi's first kind with known obtuse superbasis, a closest lattice point can be computed in $O(n^4)$ operations where $n$ is the dimension of the lattice.  To achieve this a series of relevant lattice vectors that converges to a closest lattice point is found.  We show that the series converges after at most $n$ terms.  Each vector in the series can be efficiently computed in $O(n^3)$ operations using an algorithm to compute a minimum cut in an undirected flow network.  %We discuss potential applications of this algorithm as suggest some future research questions.
\end{abstract}

%\begin{IEEEkeywords}
\begin{keywords}
Lattices, closest point algorithm, closest vector problem.
\end{keywords}
%\end{IEEEkeywords}

% The paper headers
\pagestyle{myheadings}
\thispagestyle{plain} 
\markboth{Finding a closest point in a lattice of Voronoi's first kind}{DRAFT \today}

\section{Introduction}\label{sec:introduction}

An $n$-dimensional \term{lattice} $\Lambda$ is a discrete set of vectors from $\reals^m$, $m \geq n$, formed by the integer linear combinations of a set of linearly independent basis vectors $b_1, \dots, b_n$ from $\reals^m$~\cite{SPLAG}.  That is, $\Lambda$ consists of all those vectors, or \emph{lattice points}, $x \in \reals^m$ satisfying
\[
  x = b_1 u_1 + b_2u_2 + \dots + b_n u_n \qquad u_1, \dots , u_n \in \ints. 
\] 
%In this paper vectors are assumed to be column vectors unless otherwise stated.
Given a lattice $\Lambda$ in $\reals^m$ and a vector $y \in \reals^m$, a problem of interest is to find a lattice point $x \in \Lambda$ such that the squared Euclidean norm
\[
\| y - x \|^2 = \sum_{i=1}^m (y_i - x_i)^2
\] 
is minimised.  This is called the \emph{closest lattice point problem} (or \emph{closest vector problem}) and a solution is called a \emph{closest lattice point} (or simply \emph{closest point}) to $y$. %Equivalently we wish to minimise
%\[
%\| y - \sum_{i=1}^n b_i w_i \|^2
%\]
%over integers $w_1,\dots,w_n$.  
A related problem is to find a lattice point of minimum nonzero Euclidean length, that is, a lattice point of length
\[
\min_{x\in \Lambda \backslash \{ \zerobf \} } \| x \|^2,
\]
where $\Lambda \backslash  \{\zerobf\}$ denotes the set of lattice point not equal to the origin $\zerobf$.  This is called the \emph{shortest vector problem}.

The closest lattice point problem and the shortest vector problem have interested mathematicians and computer scientists due to their relationship with integer programming~\cite{Lenstra_integerprogramming1983,Kannan1987_fast_general_np,Babai1986}, the factoring of polynomials~\cite{Lenstra1982}, and cryptanalysis~\cite{Joux_toolbox_cryptanal1998,NyguyenStern_two_faces_crypto,Micciancio_lattice_based_post_quantum_crypto}.  
Solutions of the closest lattice point problem also have engineering applications.  For example, if a lattice is used as a vector quantiser then the closest lattice point corresponds to the minimum distortion point~\cite{Conway1983VoronoiCodes,Conway1982VoronoiRegions,Conway1982FastQuantDec}.  If the lattice is used as a code, then the closest lattice point corresponds to what is called \emph{lattice decoding}\index{lattice decoding} and has been shown to yield arbitrarily good codes~\cite{Erex2004_lattice_decoding,Erez2005}.  The closest lattice point problem also occurs in communications systems involving multiple antennas~\cite{Ryan2008,Wubben_2011}.  The unwrapping of phase data for location estimation can also be posed as a closest lattice point problem and this has been applied to the global positioning system~\cite{Teunissen_GPS_1995,Hassibi_GPS_1998}.  The problem has also found applications to circular statistics~\cite{McKilliam_mean_dir_est_sq_arc_length2010}, single frequency estimation~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}, and related signal processing problems~\cite{McKilliam2007,Clarkson2007,McKilliam2009IndentifiabliltyAliasingPolyphase,Quinn_sparse_noisy_SSP_2012}.

The closest lattice point problem is known to be NP-hard under certain conditions when the lattice itself, or rather a basis thereof, is considered as an additional input parameter~\cite{micciancio_hardness_2001, Dinur2003_approximating_CVP_NP_hard, Jalden2005_sphere_decoding_complexity}. Nevertheless, algorithms exist that can compute a closest lattice point in reasonable time if the dimension is small~\cite{Pohst_sphere_decoder_1981,Kannan1987_fast_general_np,Agrell2002}.  These algorithms all require a number of operations that grows as $O(n^{O(n)})$ where $n$ is the dimension of the lattice.  Recently, Micciancio~\cite{Micciancio09adeterministic} described a solution for the closest lattice point problem that requires a number of operations that grows as $O(2^{2n})$.  This single exponential growth in complexity is the best known.

Although the problem is NP-hard in general, fast algorithms are known for specific highly regular lattices, such as the root lattices $A_n$ and $D_n$, their dual lattices $A_n^*$ and $D_n^*$, the integer lattice $\ints^{n}$, and the Leech lattice~\cite[Chap. 4]{SPLAG}\cite{Conway1982FastQuantDec,Conway1986SoftDecLeechGolay, Clarkson1999:Anstar, McKilliam2008, McKilliam2008b, McKilliam2009CoxeterLattices,Vardy1993_leech_lattice_MLD}.  In this paper we consider a particular class of lattices, those of \emph{Voronoi's first kind}~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases,Valentin2003_coverings_tilings_low_dimension,Voronoi1908_main_paper}.  Each lattice of Voronoi's first kind has what is called an \emph{obtuse superbasis}.  We show that if the obtuse superbasis is known, then a closest lattice point can be computed in $O(n^4)$ operations.  This is achieved by enumerating a series of \emph{relevant vectors} of the lattice.  Each relevant vector in the series can be computed in $O(n^3)$ operations using an algorithm for computing a minimum cut in an undirected flow network.  We show that the series converges to a closest lattice point after at most $n$ terms, resulting in $O(n^4)$ operations in total.  This result extends upon a recent result by some of the authors showing that a short vector in a lattice of Voronoi's first kind can be found by computing a minimum cut in a weighted graph~\cite{McKilliam_short_vectors_first_type_isit_2012}.

The paper is structured as follows.  Section~\ref{sec:voron-cells-relev} describes the relevant vectors and the \emph{Voronoi cell} of a lattice. Section~\ref{sec:iterative-slicer} describes a procedure to find a closest lattice point by enumerating a series of relevant vectors.  The series is guaranteed to converge to a closest lattice point after a finite number of terms.  In general the procedure might be computationally expensive because the number of terms required might be large and because computation of each relevant vector in the series might be expensive.  Section~\ref{sec:latt-voron-first} describes lattices of Voronoi's first kind and their obtuse superbasis.  It is shown that for these lattices, the series of relevant vectors results in a closest lattice point after at most $n$ terms.  Section~\ref{sec:comp-clos-relev} shows that each relevant vector in the series can be computed in $O(n^3)$ operations by computing a minimum cut in an undirected flow network.  Section~\ref{sec:discussion} discusses some potential applications of this algorithm and poses some interesting questions for future research.

\section{Voronoi cells and relevant vectors}\label{sec:voron-cells-relev}
\newcommand{\calR}{\mathcal{R}}
The (closed) \term{Voronoi cell}, denoted $\vor(\Lambda)$, of a lattice $\Lambda$ in $\reals^m$ is the subset of $\reals^m$ containing all points nearer or of equal distance (here with respect to the Euclidean norm) to the lattice point at the origin than to any other lattice point. The Voronoi cell is an $m$-dimensional convex polytope that is symmetric about the origin. %Here we will always assume the Euclidean norm (or 2-norm), so $\vor(\Lambda)$ contains those points nearer in Euclidean distance to the origin. %If $x \in \Lambda$ it follows that $\vor(\Lambda) + x$ is the subset of $\reals^n$ that is nearer to $x$ than any other lattice point in $\Lambda$. Figure \ref{lattices:fig:vorregion} is an example of the Voronoi cell.

Equivalently the Voronoi cell can be defined as the intersection of the half spaces 
\begin{align*}
H_{v} &= \{x \in \reals^n \mid \|x\| \leq \|x - v\| \} \\
&= \{x \in \reals^n \mid \dotprod{x}{v} \leq \tfrac{1}{2}\dotprod{v}{v} \}
\end{align*}
for all $v \in \Lambda \backslash  \{\zerobf\}$.  %We denote by $x^\prime$ the transpose of the vector $x$ and so $\dotprod{x}{v}$ is the inner product between column vectors $x$ and $v$.  
We denote by $\dotprod{x}{v}$ the inner product between vectors $x$ and $v$.
It is not necessary to consider all $v \in \Lambda \backslash  \{\zerobf\}$ to define the Voronoi cell.   %Figure \ref{lattices:fig:vorregion} depicts the relevant vectors of the lattice with generator given by~\eqref{lattices:eq:genB}.  A lattice point in $\calR$ is called \term{relevant}. The following remark follow trivially from the preceding discussion.
The \emph{relevant vectors} $v \in \Lambda$ are those for which  
\[
\dotprod{v}{x} \leq \dotprod{x}{x}
\]
for all $x \in \Lambda$ with $x \neq v$ and $x \neq \zerobf$.  We denote by $\relevant(\Lambda)$ the set of relevant vectors of the lattice $\Lambda$.  The Voronoi cell is the intersection of the halfspaces corresponding with the relevant vectors, that is, 
\[
%\begin{equation}\label{eq:defvortermsrelvecs}
\vor(\Lambda) = \cap_{v\in\relevant(\Lambda)}{H_{v}}.
%\end{equation}
\]
The closest lattice point problem and the Voronoi cell are related in that $x\in\Lambda$ is a closest lattice point to $y$ if and only $y - x \in \vor(\Lambda)$, that is, if and only if
\begin{equation}\label{eq:relvectnearpointieq}
\dotprod{(y - x)}{v} \leq \tfrac{1}{2} \dotprod{v}{v}
\end{equation} 
for all $v \in \relevant(\Lambda)$.

% Every short vector in a lattice is also a relevant vector, because, if a lattice point $s$ is not relevant there exists a lattice point $x$ not equal to $s$ or $\zerobf$ such that $\dotprod{x}{s} \geq \dotprod{x}{x}$ or equivalently $\dotprod{\tfrac{x}{\|x\|}}{s} \geq \|x\|$ after dividing by $\|x\|$.  Since both $\tfrac{x}{\|x\|}$ and $\tfrac{s}{\|s\|}$ are unit vectors and $s \neq x$ then,
% \[
% \|s\| = \dotprod{\tfrac{s}{\|s\|}}{s} >  \dotprod{\tfrac{x}{\|x\|}}{s} \geq \|x\|,
% \]
% i.e, $\|s\| > \|x\|$, so $s$ is not a short vector.

If $s$ is a short vector in a lattice $\Lambda$ then 
\[
\rho = \frac{\|s\|}{2} = \frac{1}{2} \min_{x \in \Lambda / \{\zerobf\} } \|x\|
\]
is called the \emph{packing radius} (or \emph{inradius}) of $\Lambda$~\cite{SPLAG}.  The packing radius is the minimum distance between the boundary of the Voronoi cell and the origin.  It is also the radius of the largest sphere that can be placed at every lattice point such that no two spheres intersect (see Figure~\ref{lattices:fig:vorregion}).  The following well known result will be useful.

%  \begin{proposition}\label{prop:latticepointinscaledtranslatedvorcell}
%  Let $\Lambda \subset \reals^{m}$ be an $n$-dimensional lattice with Voronoi cell $\vor(\Lambda)$.  Let $D$ be a positive real number and let $\ceil{D}$ denote the smallest integer larger than $D$.  Denote by 
%  \[
%   D\vor(\Lambda) + t = \left\{ x \in \reals^{m} \ \mid \frac{x-t}{D} \in \vor(\Lambda) \right\}
%  \]
% the Voronoi cell scaled by $D$ and translated by $t \in \reals^{m}$.  The number of lattice points from $\Lambda$ in $ D\vor(\Lambda) + t$ is less than or equal to $\ceil{D}^{n}$. 
%  \end{proposition}
%  \begin{proof}
% See, for example,~\cite[Lemma~3.7]{Micciancio09adeterministic}.  The result also follows from~\cite[vol.~134, p.~277]{Voronoi1908_main_paper} or \cite[Theorem~2]{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.
%  \end{proof}

\begin{proposition}\label{eq:latticepointsinsphere}
Let $\Lambda \subset \reals^{m}$ be an $n$-dimensional lattice with packing radius $\rho$.  Let $S$ be an $m$-dimensional hypersphere of radius $r$ centered at $t \in \reals^{m}$.  The number of lattice points from $\Lambda$ in the sphere $S$ is less than $\ceil{r/\rho}^{n}$ where $\ceil{\cdot}$ denotes the smallest integer strictly larger than its argument. 
 \end{proposition}
 \begin{proof}
Since the packing radius $\rho$ is the point on the boundary of the Voronoi that is closest to the origin, the hypersphere $S$ is a subset of the Voronoi cell scaled by $r/\rho$ and translated by $t$.  That is, $S \subset r/\rho \vor(\Lambda) + t$.  The proof follows because the number of lattice points in $r/\rho \vor(\Lambda) + t$ is less than $\ceil{r/\rho}^n$ (this follows from, for example, \cite[Theorem~2]{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}, see also~\cite[Lemma~3.7]{Micciancio09adeterministic}).
\end{proof}

% \footnote{These are the `strict' relevant vectors according to Conway and Sloane~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.  If the inequality $\dotprod{v}{x} < \dotprod{x}{x}$ is replaced by $\dotprod{v}{x} \leq \dotprod{x}{x}$ then this would also include the `lax' relevant vectors.  The short vectors are always strict so we only have use of the strict relevant vectors here.}

\begin{figure}[tp] 
	\centering      
		\includegraphics{figs/latticefigures-1.mps} 
		\caption{The $2$-dimensional lattice with basis vectors $(3,0.6)$ and $(0.6,3)$.  The lattice points are represented by dots and the relevant vectors are circled.  The Voronoi cell $\vor(\Lambda)$ is the shaded region and the packing radius $\rho$ and corresponding sphere packing (circles) are depicted.
}     
		\label{lattices:fig:vorregion}   
\end{figure}

\section{Finding a closest lattice point by a series of relevant vectors} \label{sec:iterative-slicer}

Let $\Lambda$ be a lattice in $\reals^m$ and let $y \in \reals^m$. A simple method to compute the lattice point $x \in \Lambda$ closest to $y$ is as follows.  Let $x_0$ be some lattice point from $\Lambda$, for example the origin.  Consider the following iteration,
\begin{align}\
x_{k+1} &= x_k + v_k \nonumber \\
v_k &= \arg\min_{ v \in \relevant(\Lambda) \cup \{\zerobf\} } \|y - x_k - v \|, \label{eq:relvectsminimslicer}
\end{align} 
where $\relevant(\Lambda) \cup \{\zerobf\}$ is the set of relevant vectors of $\Lambda$ including the origin.  Observe that the minimum over $\relevant(\Lambda) \cup \{\zerobf\}$ may not be unique.  That is, there may be multiple vectors from $\relevant(\Lambda) \cup \{\zerobf\}$ that are closest to $y - x_k$.  In this case, any one of the minimisers may be chosen.  The results that we will describe do not depend on this choice. We make the following straightforward propositions.

\begin{proposition}\label{obs:1}
%After a finite number, say $K$ iterations, the above proceedure The above proceedure reaches a stationary point after a finite number $K$ of iterations
At the $k$th iteration either $x_k$ is a closest lattice point to $y$ or $\|y - x_k\| > \| y - x_{k+1} \|$.
\end{proposition}
\begin{proof}
If $x_k$ is a closest lattice point to $y$ then $\|y - x_k\| \leq \| y - x_{k+1} \|$ by definition.  On the other hand if $x_k$ is not a closest lattice point to $y$ we have $y - x_k \notin \vor(\Lambda)$ and from~\eqref{eq:relvectnearpointieq} there exists a relevant vector $v$ such that
\[
0 > \dotprod{v}{v} - 2\dotprod{(y - x_k)}{v}.
\]
Adding $\|y - x_k\|^2$ to both sides of this inequality gives
\begin{align*}
\|y - x_k\|^2 &> \dotprod{v}{v} - 2\dotprod{(y - x_k)}{v} + \|y - x_k\|^2 \\
&= \|y - x_k - v\|^2 \\
&\geq \arg\min_{ v \in \relevant(\Lambda) \cup \{\zerobf\}}\|y - x_k - v \|^2 \\
&= \|y - x_k - v_k \|^2 \\
&= \|y - x_{k+1}\|^2. 
\end{align*}
\end{proof}

% \begin{observation}\label{obs:1}
% At the $k$th iteration either $x_k$ is a closest lattice point to $y$ or 
% \[
%  \max_{ v \in \relevant(\Lambda)}\frac{\dotprod{(y - x_k)}{v}}{\dotprod{v}{v}} > \frac{1}{2}.
% \]
% \end{observation}
% \begin{proof}
% If $x_k$ is a closest lattice point to $y$ then $\dotprod{(y - x)}{v} \leq \tfrac{1}{2} \dotprod{v}{v}$ for all $v \in \relevant(\Lambda)$ by definition of the Voronoi cell~\eqref{eq:relvecsvorcellnearpoint} and so
% \[
%  \max_{ v \in \relevant(\Lambda)}\frac{\dotprod{(y - x_k)}{v}}{\dotprod{v}{v}} \leq \frac{1}{2}.
% \]
% Conversely if $x_k$ is not the closest point to $y$ there exists a relevant vector $r$ such that $\dotprod{(y - x)}{r} > \tfrac{1}{2} \dotprod{r}{r}$ and so
% \[
%  \max_{ v \in \relevant(\Lambda)}\frac{\dotprod{(y - x_k)}{v}}{\dotprod{v}{v}} \geq \frac{\dotprod{(y - x_k)}{r}}{\dotprod{r}{r}} > \frac{1}{2}.
% \]
% \end{proof}

 \begin{proposition}\label{obs:2}
 There is a finite number $K$ such that $x_K, x_{K+1}, x_{K+2}, \dots$ are all closest points to $y$.
 \end{proposition}
 \begin{proof}
Suppose no such finite $K$ exists, then
\[
\|y - x_0\| >  \|y - x_1\| > \|y - x_2\| > \dots
\]
and so $x_0,x_1,\dots$ is an infinite sequence of distinct (due to the strict inequality) lattice points  all contained inside an $n$-dimensional hypersphere of radius $r = \|y - x_0\|$ centered at $y$.  This is a contradiction since, if $\rho$ is the packing radius of the lattice, then less than $\ceil{r/\rho}^n$ lattice points lie inside this sphere by Proposition~\ref{eq:latticepointsinsphere}. 
\end{proof}

Proposition~\ref{obs:2} above asserts that after some finite number $K$ of iterations the procedure arrives at $x_K$, a closest lattice point to $y$.  Using Proposition~\ref{obs:1} we can detect that $x_K$ is a closest lattice point by asserting that $\|y - x_K\| \leq \| y - x_{K+1} \|$.
This simple iterative approach to compute a closest lattice point is related to what is called the \emph{iterative slicer}~\cite{Shalvi_iterativeslicer_2009}.  Micciancio~\cite{Micciancio09adeterministic} describes a related, but more sophisticated, iterative algorithm that can compute a closest lattice point in a number of operations that grows exponentially as $O(2^{2 n})$.  This single exponential growth in complexity is the best known.  %Most popular algorithms for computed closest lattice points require a number of operations of order $O(n^n)$\cite{Agrell2002,Viterbo_sphere_decoder_1999,Pohst_sphere_decoder_1981}.

Two factors contribute to the computational complexity of this iterative approach to compute a closest lattice point.  The first factor is computing the minimum over the set $\relevant(\Lambda) \cup \{\zerobf\}$ in~\eqref{eq:relvectsminimslicer}.  In general a lattice can have as many as $2^{n+1}-2$ relevant vectors and so computing a minimiser directly can require a number of operations that grows exponentially with $n$.  To add to this it is typically the case that the set of relevant vectors $\relevant(\Lambda)$ must be stored in memory, and so the algorithm can require an amount of memory that grows exponentially with $n$.  We will show that, for a lattice of Voronoi's first kind, the set of relevant vectors has a compact representation in terms of what is called its \emph{obtuse superbasis}.  To store the obtuse superbasis requires an amount of memory of order $O(n^2)$ in the worst case.  We also show that for a lattice of Voronoi's first kind the minimisation over $\relevant(\Lambda) \cup \{\zerobf\}$ in~\eqref{eq:relvectsminimslicer} can be solved efficiently by computing a minimum cut in an undirected flow network.  Using known algorithms a minimiser can be computed in $O(n^3)$ operations~\cite{Goldberg:1986:NAM:12130.12144,EdmondsKarp_max_flow,Cormen2001}.

The other factor affecting the complexity is the number of iterations required before the algorithm arrives at a closest lattice point, that is, the size of $K$.  Propositon~\ref{eq:latticepointsinsphere} suggests that this number can be as large as $\ceil{r/\rho}^n$ where $r = \|y - x_0\|^2$ and $\rho$ is the packing radius of the lattice.  Thus, the number of iterations required can potentially grow exponentially with $n$.  The number of iterations required depends on the lattice point that starts the iteration $x_0$.  It is helpful for $x_0$ to be, in some sense, a close approximation of the closest lattice point $x_K$.  Unfortunately, computing close approximations of a closest lattice point is known to be computationally difficult~\cite{feige_inapproximability_2004}.  We will show that for a lattice of Voronoi's first kind a simple and easy to compute choice for $x_0$ ensures that a closest lattice point is reached in at most $n$ iterations, and so $K \leq n$.  Combining this with the fact that each iteration of the algorithm requires $O(n^3)$ operations results in an algorithm that requires $O(n^4)$ operations to compute a closest point in a lattice of Voronoi's first kind.

\begin{figure}[tp] 
	\centering      
		\includegraphics{figs/iterativeseriesexample-1.mps} 
		\caption{Example of the iterative procedure described in~\eqref{eq:relvectsminimslicer} to compute a closest lattice point to $y = (4,3.5)$ (marked with a cross) in the $2$-dimensional lattice generated by basis vectors $(2,0.4)$ and $(0.4,2)$.  The initial lattice point for the iteration is $x_0 = (-4.4,-2.8)$.  The shaded region is the Voronoi cell translated about the closest lattice point $x_6 = (4.4,2.8)$.}       
		\label{lattices:fig:iterativeexample} 
\end{figure}

\section{Lattices of Voronoi's first kind} \label{sec:latt-voron-first}

An $n$-dimensional lattice $\Lambda$ is said to be of \emph{Voronoi's first kind} if it has what is called an \emph{obtuse superbasis}~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.  That is, there exists a set of $n+1$ vectors $b_1,\dots,b_{n+1}$ such that $b_1,\dots,b_n$ are a basis for $\Lambda$,
\begin{equation}\label{eq:superbasecond}
b_1 + b_2 \dots + b_{n+1} = 0
\end{equation}
(the \emph{superbasis} condition), and the inner products satisfy
\begin{equation}\label{eq:obtusecond}
q_{ij} = b_i \cdot b_j \leq 0, \qquad \text{for} \qquad i,j = 1,\dots,n+1, i \neq j
\end{equation}
(the \emph{obtuse} condition).  The $q_{ij}$ are called the \emph{Selling parameters}~\cite{Selling1874}.  It is known that all lattices in dimensions less than $4$ are of Voronoi's first kind~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.  An interesting property of lattices of Voronoi's first kind is that their relevant vectors have a straightforward description.

\begin{theorem} \label{thm:revvecssuperbase} (Conway and Sloane~\cite[Theorem~3]{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases})
The relevant vectors of $\Lambda$ are of the form,
\[
\sum_{i \in I} b_i
\]
where $I$ is a strict subset of $\{1, 2, \dots, n+1\}$ that is not empty, i.e. $I \subset \{1, 2, \dots, n+1\}$ and $I \neq \emptyset$.
\end{theorem}

We are interested in solving the closest lattice point problem for lattices of Voronoi's first kind.  Let $\Lambda \subset \reals^m$ be an $n$ dimensional lattice of Voronoi's first kind with obtuse superbasis $b_1,\dots,b_{n+1}$ and let $y \in \reals^m$.  We want to find $n$ integers $w_1,\dots,w_n$ that minimise
\[
\| y - \sum_{i=1}^n b_i w_i \|^2.
\]
We can equivalently find $n+1$ integers $w_1,\dots,w_{n+1}$ that minimise
\[
\| y - \sum_{i=1}^{n+1} b_i w_i \|^2.
\]
We will use the iterative procedure described in~\eqref{eq:relvectsminimslicer} to do this.  In what follows we will assume that $y$ lies in the space spanned by the basis vectors $b_1,\dots,b_{n}$.  This assumption is without loss of generality because $x$ is closest lattice point to $y$ if and only if $x$ is a the closest lattice point to the projection of $y$ orthogonal into the space spanned by $b_1,\dots,b_{n+1}$.  Let
\begin{equation}\label{eq:matrxBobtusebasis}
B = (b_1\,\,b_2\,\,\dots\,\,b_{n+1})
\end{equation}
be the matrix with columns given by $b_1,\dots,b_{n+1}$ and let $z \in \reals^{n+1}$ be a column vector such that $y = Bz$.  We now want to find a column vector $w = (w_1,\dots,w_{n+1})^\prime$ of integers such that
\begin{equation}\label{eq:tominimise}
\| B(z  -  w) \|^2
\end{equation}
is minimised.  Define the column vector $u_0 = \floor{z}$ where $\floor{\cdot}$ denotes the largest integer less than or equal to its argument and operates on vectors elementwise. In view of Theorem~\ref{thm:revvecssuperbase} the iterative procedure~\eqref{eq:relvectsminimslicer} to compute a closest lattice point can be written in the form
\begin{align}
x_{k+1} &= B u_{k+1} \label{eq:xseqfirsttype}  \\
u_{k+1} &= u_k + p_k \nonumber \\
p_k &= \arg\min_{p \in \{0,1\}^{n+1}}\| B(z - u_k - p) \|^2, \label{eq:pvecmin}
\end{align}
where $\{0,1\}^{n+1}$ denotes the set of column vectors of length $n+1$ with elements equal to zero or one.  Observe that the procedure is initialised at the lattice point $x_0 = Bu_0 = B\floor{z}$.  This choice of initial lattice point is important.  In Section~\ref{sec:comp-clos-relev} we show how the minimisation over $\{0,1\}^{n+1}$ in~\eqref{eq:pvecmin} can be computed efficiently in $O(n^3)$ operations by computing a minimum cut in an undirected flow network.  In the remainder of this section we prove that this iterative procedure results in a closest lattice point after at most $n$ iterations.  That is, we show that there exists a positive integer $K \leq n$ such that $x_K$ is a closest lattice point to $y = Bz$.

\newcommand{\rng}{\operatorname{rng}}
\newcommand{\subrng}{\operatorname{subr}}
\newcommand{\decrng}{\operatorname{decrng}}

It is necessary to introduce some notation.  Given a subset $S$ of indices from $\{1,\dots,n+1\}$ we denote by $I_S$ the vector of length $n+1$ with elements equal to zero except for those with indices from $S$ that are equal to one.  That is, if $g = I_S$ then $g$ is the vector with elements
\[
g_i = \begin{cases}
1 & i \in S\\
0 & i \notin S.
\end{cases}
\]
If $p \in \reals^{n+1}$ then $p - I_S$ denotes the vector $p$ with those elements with index from $S$ decremented by one.  Similarly, $p + I_S$ denotes the vector $p$ with those elements with index from $S$ incremented by one.

\begin{lemma}\label{lem:decSellings}
Let $p \in \reals^{n+1}$, let $S$ be a subset of the indices of $p$, and let $B$ be the matrix from~\eqref{eq:matrxBobtusebasis} with column given by the vectors $b_1,\dots,b_{n+1}$ from the obtuse superbasis.  The following equalities hold:
\begin{enumerate}
\item  ${\displaystyle \|Bp\|^2 - \|B(p - I_S)\|^2 = \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_j - 2p_i)}$, \label{eq:lem:decSellingsdec}
\item  ${\displaystyle \|Bp\|^2 - \|B(p + I_S)\|^2 = \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_i - 2p_j)}$, \label{eq:lem:decSellingsinc}
\end{enumerate}
where $q_{ij} = \dotprod{b_i}{b_j}$ are the Selling parameters.
\end{lemma}
\begin{proof}
We give a proof for part~\ref{eq:lem:decSellingsdec}.  The proof for part~\ref{eq:lem:decSellingsinc} is similar.  From Selling's formula~\cite[Proposition 2.3.1]{Valentin2003_coverings_tilings_low_dimension}~\cite{Selling1874},
\begin{align*}
\|Bp\|^2 &= \| \sum_{i=1}^{n+1} b_i p_i \|^2 \\
&= - \sum_{i=1}^{n+1}\sum_{j=i+1}^{n+1}q_{ij}(p_i - p_j)^2 \\
&= - \frac{1}{2} \sum_{i=1}^{n+1}\sum_{j=1}^{n+1}q_{ij}(p_i - p_j)^2, 
\end{align*}
where $q_{ij} = q_{ji} = b_i \cdot b_j$ are the Selling parameters.  Put $g = p - I_S$.  It follows that
\[
\|Bp\|^2 - \|Bg\|^2 = \frac{1}{2} \sum_{i=1}^{n+1}\sum_{j=1}^{n+1}q_{ij}d_{ij},
\]
where $d_{ij} = (g_i - g_j)^2 - (p_i - p_j)^2$.  If both $i \notin S$ and $j \notin S$ then $p_i = g_i$, $p_j = g_j$ and so $d_{ij} = 0$.  Similarly if both $i \in S$ and $j \in S$ then 
\[
d_{ij} = (p_i-1 - p_j+1)^2 - (p_i - p_j)^2 = 0.
\]
If $i \in S$ and $j \notin S$ then
\[
d_{ij} = (p_i-1 - p_j)^2 - (p_i - p_j)^2= 1 + 2p_j - 2p_i,
\]
while if $i \notin S$ and $j \in S$ then
\[
d_{ij} = (p_i - p_j+1)^2 - (p_i - p_j)^2 = 1 + 2p_i - 2p_j.
\]
We thus have
\begin{align*}
\|Bp\|^2 - \|Bg\|^2 &= \frac{1}{2} \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_j - 2p_i) + \frac{1}{2} \sum_{i \notin S}\sum_{j \in S}q_{ij}(1 + 2p_i - 2p_j) \\
&= \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_j - 2p_i),
\end{align*}
where the last line holds because $q_{ij} = q_{ji}$.
\end{proof}

We denote by $\min(p)$ and $\max(p)$ the minimum and maximum values obtained by the elements of the vector $p$.  We define the function
\[
\rng(p) = \max(p) - \min(p)
\] 
to return the difference between the maximum and minimum of $p$.  For example, if $p = (2,-1,4) \in \ints^3$ then $\rng(p) = 4 - (-1)=5$.  We might also write simply $\rng(4,-1,5)=5$.  Observe that $\rng(p)$ cannot be negative and that if $\rng(p) = 0$ then all of the elements of $p$ are equal.  We define the function $\subrng(p)$ to return the largest subset, say $S$, of the indices of $p$ such that $\min\{p_i, i \in S\} - \max\{p_i, i \notin S\} \geq 2.$  If no such subset exists then $\subrng(p)$ is the empty set $\emptyset$.  For example, 
\[
\subrng(2,-1,4) = \{1,3\}, \qquad \subrng(2,1,3) = \emptyset, \quad  \subrng(1,3,1) = \{2\}.
\]  
To make the definition of $\subrng$ clear we give the following alternative and equivalent definition.  Let $p \in \ints^n$ and let $\sigma$ be the permutation of the indices $\{1,\dots,n\}$ that puts the elements of $p$ in ascending order, that is
\[
p_{\sigma(1)} \leq p_{\sigma(2)} \leq \dots \leq p_{\sigma(n)}.
\]  
Let $T$ be the smallest integer from $\{2,\dots,n\}$ such that $p_{\sigma(T)} - p_{\sigma(T-1)} \geq 2$.  If no such integer $T$ exists then $\subrng(p) = \emptyset$.  Otherwise 
\[
\subrng(p) =  \{ \sigma(T), \sigma(T+1), \dots, \sigma(n) \}.
\]
The following straightforward property of $\subrng$ will be useful.

\begin{proposition}\label{prop:subrrngsmall}
Let $p \in \ints^{n+1}$.  If $\subrng(p) = \emptyset$ then $\rng(p) \leq n$.
\end{proposition}
\begin{proof}
Let $\sigma$ be the permutation of the indices $\{1,\dots,n+1\}$ that puts the elements of $p$ in ascending order.  Because $\subrng(p) = \emptyset$ and because the elements of $p$ are integers we have $p_{\sigma(i+1)} \leq p_{\sigma(i)} + 1$ for all $i=1,\dots,n$.  It follows that
\[
p_{\sigma(n+1)} \leq p_{\sigma(n)} + 1 \leq p_{\sigma(n-1)} + 2 \leq \dots \leq p_{\sigma(1)} + n.
\]
and so $\rng(p) = p_{\sigma(n+1)} - p_{\sigma(1)} \leq n$.
\end{proof}

Finally we define the function
\[
\decrng(p) = p -  I_{\subrng(p)}
\]
that decrements those elements from $p$ with indices from $\subrng(p)$.  If $\subrng(p) = \emptyset$, then $\decrng(p) = p$, that is, $\decrng$ does not modify $p$.  On the other hand, if $\subrng(p) \neq \emptyset$ then
\[
\rng\big(\decrng(p)\big) = \rng(p) - 1
\]
because $\subrng(p)$ contains all those indices $i$ such that $p_i = \max(p)$.  Observe that by repeatedly applying $\decrng$ to a vector one eventually obtains a vector for which further application of $\decrng$ has no effect.  For example,
\begin{align*}
\decrng(2,-1,4) &= (2,-1,4) - I_{\subrng(2,-1,4)} = (2,-1,4) - I_{\{1,3\}} = (1,-1,3) \\ 
\decrng(1,-1,3) &= (1,-1,3) - I_{\{1,3\}} = (0,-1,2) \\ 
\decrng(0,-1,2) &= (0,-1,2) - I_{\{3\}} = (0,-1,1) \\ 
\decrng(0,-1,1) &= (0,-1,1) - I_{\emptyset} = (0,-1,1).
\end{align*}
This will be a useful property so we state it formally in the following proposition.

\begin{proposition} \label{lem:repeatappdecrange} Let $p \in \reals^{n+1}$ and define the infinite sequence $d_0,d_1,d_2,\dots$ of vectors according to $d_0=p$ and $d_{k+1} = \decrng(d_k)$.  There is a finite integer $T$ such that $d_T=d_{T+1}=d_{T+2}=\dots$.
\end{proposition}
\begin{proof}
Assume that no such $T$ exists.  Then $\decrng(d_k) \neq d_k$ for all positive integers $k$ and so 
\[
\rng(d_{k}) = \rng(d_{k-1}) - 1 = \rng(d_{k-2}) - 2 = \dots = \rng(p) - k.
\]  
Choosing $k > \rng(p)$ we have $\rng(d_{k}) < 0$ contradicting that $\rng(d_k)$ is nonegative.
\end{proof}

We are now ready to study properties of a closest lattice point in a lattice of Voronoi's first kind.

\begin{lemma}\label{lem:decrngpreservesclosestpoints}
If $v \in \ints^{n+1}$ such that $B(\floor{z} + v)$ is a closest lattice point to $y = Bz$, then $B\big(\floor{z} + \decrng(v)\big)$ is also a closest lattice point to $y$.
\end{lemma}
\begin{proof}
The lemma is trivial if $\subrng(v) = \emptyset$ so that $\decrng(v) = v$.  It remains to prove the lemma when $\subrng(v) \neq \emptyset$.  In this case put $S = \subrng(v)$ and put $u = \decrng(v) = v - I_S$.  Let $\zeta = z - \floor{z}$ be the column vector containing the fractional parts of the elements of $z$.  We have $\zeta - u = \zeta - v + I_S$.  Applying part~\ref{eq:lem:decSellingsinc} of Lemma~\ref{lem:decSellings} with $p = \zeta - v$ we obtain
\begin{align}
\|B(\zeta - v)\|^2 - \|B(\zeta - u)\|^2 &= \sum_{i \in S}\sum_{j \notin S}q_{ij}\big(1 + 2(\zeta_i - v_i)  - 2(\zeta_j - v_j) \big) \nonumber \\
&= \sum_{i \in S}\sum_{j \notin S}q_{ij}\big(1 + 2(\zeta_i-\zeta_j) - 2(v_i - v_j)\big). \label{eq:sumsumBzBu}
\end{align}
Observe that $\zeta_i =  z_i - \floor{z_i} \in [0,1)$ for all $i=1,\dots,n+1$ and so $-1 < \zeta_i-\zeta_j < 1$ for all $i,j=1,\dots,n+1$.  Also, for $i \in S$ and $j\notin S$ we have 
\[
v_i - v_j \geq \min\{ v_i, i \in S\} - \max\{v_j, j \notin S\} \geq 2
\]
by definition of $\subrng(v) = S$.  Thus,
\[
1 + 2(\zeta_i-\zeta_j) - 2(v_i - v_j) < 1 + 2 - 4 = -1 < 0.
\]
Substituting this inequality into~\eqref{eq:sumsumBzBu} and using that $q_{ij} \leq 0$ (the obtuse condition~\eqref{eq:obtusecond}) we find that
\[
\|B(z - \floor{z} - v)\|^2 - \|B(z - \floor{z} - u)\|^2 \geq 0,
\]
and so $B(\floor{z} + u) = B\big(\floor{z} + \decrng(v)\big)$ is a closest lattice point to $y = Bz$ whenever $B(\floor{z} + v)$ is.
\end{proof}

\begin{lemma}\label{lem:roundzclose}
There exists a closest lattice point to $y = Bz$ in the form $B(\floor{z} + v)$ where $v \in \ints^{n+1}$ with $\rng(v) \leq n$.
\end{lemma}
\begin{proof}
Let $d_0 \in \ints^{n+1}$ be such that $B(\floor{z} + d_0)$ is a closest lattice point to $y$. Define the sequence of vectors $d_0,d_1,\dots$ from $\ints^{n+1}$ according to the recursion $d_{k+1} = \decrng(d_k)$.  It follows from Lemma~\ref{lem:decrngpreservesclosestpoints} that $B(\floor{z} + d_{k})$ is a closest lattice point for all positive integers $k$.  By Proposition~\ref{lem:repeatappdecrange} there is a finite $T$ such that 
\[
d_{T+1}=d_T=\decrng(d_T).
\]  
It follows that $\subrng(d_T) = \emptyset$ and so $\rng(d_T) \leq n$ by Proposition~\ref{prop:subrrngsmall}.  The proof follows with $v = d_T$.   
\end{proof}

Let $\ell$ be a nonegative integer.  We will say that a lattice point $x$ is $\ell$-\emph{close} to $y$ if there exists a $v \in \ints^{n+1}$ with $\rng(v) = \ell$ such that $x + Bv$ is a closest lattice point to $y$.  Lemma~\ref{lem:roundzclose} asserts that the lattice point $x_0 = B\floor{z}$ that initialises the iterative procedure~\eqref{eq:xseqfirsttype} is $K$-close to $y$ where $K \leq n$.  From Lemma~\ref{lem:rngdecreases} stated below it will follow that if the lattice point $x_k$ obtained on the $k$th iteration of the procedure is $\ell$-close, then the lattice point $x_{k+1}$ obtained on the next iteration is $(\ell-1)$-close.  Since $x_0$ is $K$-close it will then follow that after $K \leq n$ iterations the lattice point $x_K$ is $0$-close.  At this stage it is guaranteed that $x_{K}$ itself is a closest lattice point to $y$.  This is shown in the following lemma.

\begin{lemma}\label{lem:rngzeroclosestpoint}
If $x$ is a lattice point that is $0$-close to $y$, then $x$ is a closest lattice point to $y$.
\end{lemma}
\begin{proof}
Because $x$ is $0$-close there exists a $v \in \ints^{n+1}$ with $\rng(v) = 0$ such that $x + Bv$ is a closest lattice point to $y$.  Because $\rng(v) = 0$ all elements from $v$ are identical.  That is $v_1=v_2=\dots=v_{n+1}$.  In this case $Bu = \sum_{i=1}^{n+1} v_n b_n = v_1\sum_{i=1}^{n+1}b_n = 0$
as a result of the superbasis condition~\eqref{eq:superbasecond}.  Thus $x = x + Bv$ is a closest point to $y$. 
\end{proof}

\begin{lemma}\label{lem:rngdecreases}
Let $Bu$ with $u \in \ints^{n+1}$ be a lattice point that is $\ell$-close to $y = Bz$ where $\ell > 0$.  Let $g \in \{0,1\}^{n+1}$ be such that
\begin{equation}\label{eq:qismin}
\|B(z - u - g)\|^2 = \min_{p \in \{0,1\}^{n+1}}\|B(z - u - p)\|^2.
\end{equation}
The lattice point $B(u+g)$ is $(\ell-1)$-close to $y$.
\end{lemma}
\begin{proof}
Because $Bu$ is $\ell$-close to $y$ there exists $v \in \ints^{n+1}$ with $\rng(v) = \ell$ such that $B(u+v)$ is a closest lattice point to $y$.  Let $A$ be the (possibly empty) subset of indices $\{1,\dots,n+1\}$ such that $g_i = 1$ and $v_i = \min(v)$ and let $D$ be the (possibly empty) subset of indices such that $g_i = 0$ and $v_i = \max(v)$. That is,
\[
A = \{i \mid g_i = 1, v_i = \min(v) \}, \qquad D = \{i \mid g_i = 0, v_i = \max(v) \}.
\]
Put $d = g - I_A + I_D$ and put $w = v - d$.  Observe that $d_i = 0$ whenever $v_i = \min(v)$ and so $\min(w) = \min(v-d) = \min(v)$.  Also, $d_i = 1$ whenever $v_i = \max(v)$ and so $\max(w) = \max(v-d) = \max(v) - 1$.  Thus,
\[
\rng(w) = \max(v) - 1 - \min(v) = \rng(v) - 1.
\]
The lemma will follow if we show that $B(u+g+w)$ is a closest lattice point to $y$ since then $B(u+g)$ with be $(\ell-1)$-close to $y$.  We show this in two stages.  First put $h = v - g + I_A$.  Suppose that $B(u+g+h)$ is not a closest lattice point to $y = Bz$.  That is, suppose that 
\[
\|B(z-u-g-h)\|^2 = \|B(z-u-v-I_A)\|^2 > \|B(z-u-v)\|^2.
\]
In this case, Lemma~\ref{lem:vqcontramin} (proved below) with $t = z-u$ and $S = A$ shows that 
\begin{equation}\label{eq:lem:rng:contag}
\|B(z-u - g)\|^2 > \min_{p \in \{0,1\}^{n+1}}\|B(z- u - p)\|^2,
\end{equation}
contradicting~\eqref{eq:qismin}.  Thus, $B(u+g+h)$ is a closest lattice point to $y$.  Now suppose that $B(u+g+w) = B(u+g+h-I_D)$ is not a closest lattice point to $y$.  That is, suppose that
\[
\|B(z-u-g-h+I_D)\|^2 > \|B(z-u-h-g)\|^2.
\]
Lemma~\ref{lem:vqcontramax} (proved below) with $t = z-u$ and $S = D$ shows that~\eqref{eq:lem:rng:contag} occurs again, contradicting~\eqref{eq:qismin}.  Thus, $B(u+g+w)$ is a closest lattice point to $y$ and, since $\rng(w) = \ell-1$, the lattice point $B(u+g)$ is $(\ell-1)$-close to $y$.
 \end{proof}

The proof of the previous Lemma~\ref{lem:rngdecreases} relies upon Lemmas~\ref{lem:vqcontramin} and~\ref{lem:vqcontramax}.  Before we prove these lemmas we state the following theorem asserting that the iterative procedure~\eqref{eq:xseqfirsttype} converges to a closest lattice point in $K \leq n$ iterations.  This is the primary result of this section.

\begin{theorem}
Let $x_0,x_1,\dots$ be the sequence of lattice points given by the iterative procedure~\eqref{eq:xseqfirsttype}.  There exists $K \leq n$ such that $x_K$ is a closest lattice point to $y = Bz$.
\end{theorem}
\begin{proof}
Let $x_k = B u_k$ with $u_k \in \ints^{n+1}$ be the lattice point obtained on the $k$th iteration of the procedure.  Suppose that $x_k$ is $\ell$-close to $y=Bz$ with $\ell > 0$.  The procedure computes $p_{k} \in \{0,1\}^{n+1}$ satisfying
\[
\|B(z - u_k - p_{k})\|^2 = \min_{p \in \{0,1\}^{n+1}}\|B(z - u_k - p)\|^2
\]
and puts $x_{k+1} = B(u_k + p_k)$.  It follows from Lemma~\ref{lem:rngdecreases} that $x_{k+1}$ is $(\ell-1)$-close to $y$.  By Lemma~\ref{lem:roundzclose} the lattice point that initialises the procedure $x_0 = B \floor{z}$ is $K$-close to $y$ where $K \leq n$.  Thus, $x_1$ is $(K-1)$-close, $x_2$ is $(K-2)$-close, and so on until $x_K$ is $0$-close.  That $x_K$ is a closest lattice point to $y$ follows from Lemma~\ref{lem:rngzeroclosestpoint}.
\end{proof}

We now provide proof of Lemmas~\ref{lem:vqcontramin} and~\ref{lem:vqcontramax} that were used in the proof of Lemma~\ref{lem:rngdecreases}.

\begin{lemma}\label{lem:vqcontramin}
Let $t \in \reals^{n+1}$ and $v \in \ints^{n+1}$ be such that $Bv$ is a closest lattice point to $Bt$.  Let $g \in \{0,1\}^{n+1}$ and let $S$ be the subset of indices $\{1,\dots,n+1\}$ such that $g_i=1$ and $v_i = \min(v)$ for $i \in S$.  If $\|B(t-v)\|^2 < \|B(t-v-I_S)\|^2$, then
\[
\|B(t - g)\|^2 > \min_{p \in \{0,1\}^{n+1}}\|B(t - p)\|^2.
\]
\end{lemma}
\begin{proof}
From our hypothesis and part~\ref{eq:lem:decSellingsdec} of Lemma~\ref{lem:decSellings} with $p = t-v$ we have
\[
\|B(t-v)\|^2 - \|B(t-v-I_S)\|^2 = \sum_{i \in S}\sum_{j \notin S}q_{ij}\big(1 + \gamma_{ij} + 2(v_i - v_j)\big) < 0,
\]
where $\gamma_{ij} = 2(t_j - t_i)$.  Put $h = g - I_S$ and observe that $h_i = 0$ when $v_i = \min(v)$ and so, since $S \subset \{i \mid v_i = \min(v) \}$, we have
\[
h_i - h_j \geq v_i - v_j \qquad \text{when $i \in S$ and $j=1,\dots,n+1$.}
\]
Thus, since $q_{ij} \leq 0$,
\[
q_{ij}\big(1 + \gamma_{ij} + 2(h_i - h_j)\big) \leq q_{ij}\big(1 + \gamma_{ij} + 2(v_i - v_j)\big)
\]
when $i \in S$ and $j =1,\dots,n+1$.  By part~\ref{eq:lem:decSellingsdec} of Lemma~\ref{lem:decSellings} we have
\begin{align*}
\|B(t-h)\|^2 - \|B(t-g)\|^2 &= \|B(t-h)\|^2 - \|B(t-h-I_S)\|^2 \\
&= \sum_{i \in S}\sum_{j \notin S}q_{ij}\big(1 + \gamma_{ij} + 2(h_i - h_j)\big) \\
&\leq \sum_{i \in S}\sum_{j \notin S}q_{ij}\big(1 + \gamma_{ij} + 2(v_i - v_j)\big) < 0
\end{align*}
from which it follows that
\[
\|B(t-g)\|^2 > \|B(t-h)\|^2 \geq \min_{p \in \{0,1\}^{n+1}}\|B(t  - p)\|^2
\]
as required.
\end{proof}

\begin{lemma}\label{lem:vqcontramax}
Let $t \in \reals^{n+1}$ and $v \in \ints^{n+1}$ be such that $Bv$ is a closest lattice point to $Bt$.  Let $g \in \{0,1\}^{n+1}$ and let $S$ be the subset of indices $\{1,\dots,n+1\}$ such that $g_i=0$ and $v_i = \max(v)$ for $i \in S$.  If $\|B(t-v)\|^2 < \|B(t-v+I_S)\|^2$, then
\[
\|B(t - g)\|^2 > \min_{p \in \{0,1\}^{n+1}}\|B(t - p)\|^2.
\]
\end{lemma}
\begin{proof}
The proof mimics that of Lemma~\ref{lem:vqcontramin} with only minor modification.  We give the proof in full for completeness, and since the small modifications that do exist might not be obvious. From our hypothesis and part~\ref{eq:lem:decSellingsinc} of Lemma~\ref{lem:decSellings} with $p = t-v$ we have
\[
\|B(t-v)\|^2 - \|B(t-v+I_S)\|^2 = \sum_{i \in S}\sum_{j \notin S}q_{ij}\big(1 + \gamma_{ji} + 2(v_j - v_i)\big) < 0,
\]
where $\gamma_{ij} = 2(t_j - t_i)$.  Put $h = g + I_S$ and observe that $h_i = 1$ when $v_i = \max(v)$ and so, since $S \subset \{i \mid v_i = \max(v) \}$, we have
\[
h_j - h_i \geq v_j - v_i \qquad \text{for $i \in S$ and $j=1,\dots,n+1$.}
\]
Thus, since $q_{ij} \leq 0$,
\[
q_{ij}\big(1 + \gamma_{ji} + 2(h_j - h_i)\big) \leq q_{ij}\big(1 + \gamma_{ji} + 2(v_j - v_i)\big)
\]
when $i \in S$ and $j =1,\dots,n+1$.  By part~\ref{eq:lem:decSellingsdec} of Lemma~\ref{lem:decSellings} we have
\begin{align*}
\|B(t-h)\|^2 - \|B(t-g)\|^2 &= \|B(t-h)\|^2 - \|B(t-h+I_S)\|^2 \\
&= \sum_{i \in S}\sum_{j \notin S}q_{ij}\big(1 + \gamma_{ji} + 2(h_j - h_i)\big) \\
&\leq \sum_{i \in S}\sum_{j \notin S}q_{ij}\big(1 + \gamma_{ji} + 2(v_j - v_i)\big) < 0
\end{align*}
from which it follows that
\[
\|B(t-g)\|^2 > \|B(t-h)\|^2 \geq \min_{p \in \{0,1\}^{n+1}}\|B(t  - p)\|^2
\]
as required.
\end{proof}

% \begin{lemma}
% Let $x \in \ints^{n+1}$.  If there exists a $v \in \ints^{n+1}$ with $\rng(v) = 0$ such that $B(x + v)$ is a closest lattice point to $y = Bz$, then $Bx$ is also a closest lattice point to $y$.  
% \end{lemma}
% \begin{proof}
% Because $\rng(v) = 0$ all of the elements $v_1=v_2=\dots=v_{n+1} = V$ for some integer $V$. Now
% \begin{align*}
% \|B(z - x - v)\|^2 &= \| \sum_{i=1}^{n+1} b_i (z_i-x_i-v_i) \|^2 \\
% &= \| \sum_{i=1}^{n+1} b_i (z_i-x_i) -V\sum_{i=1}^{n+1}b_i \|^2 \\
% &=  \| \sum_{i=1}^{n+1} b_i (z_i-x_i) \|^2 \\
% &=  \|B(z-x)\|^2
% \end{align*}
% because $\sum_{i=1}^{n+1}b_i = 0$ due to the superbase condition~\eqref{eq:superbasecond}.  So $Bx$ is a closest point to $y$ whenever $B(x+v)$ is.
% \end{proof}

% \begin{lemma}\label{lem:minrelvecsisonesonh}
% Let $x \in \ints^{n+1}$ be such that $w_i = x_i + h_i$ are minimisers of~\eqref{eq:tominimise} and such that $h_1,\dots,h_{n+1}$ starts at zero and increments in step of one.  Let $s$ be the largest integer such that $h_s = 0$ and let
% \[
% \alpha = \min_{p \in \{0,1\}^{n+1}} \| B(z - x - p) \|^2.
% \]
% The first $s$ elements of the vector $\alpha$ are equal to zero, that is, $\alpha_1,\alpha_2,\dots,\alpha_s = 0$.
% \end{lemma}
% \begin{proof}
% Let $S$ be some nonempty subset of the indices $\{1, 2,\dots,s\}$ and let $\beta$ be the column vector with elements satisfying
% \[
% \beta_i = \begin{cases}
% 1 & i \in S \\
% \alpha_i & \text{otherwise}.
% \end{cases}
% \]
% Since the subset $S$ is arbitrary, our proof holds if we show that
% \[
% A = \| B(z - x - \alpha) \|^2 - \| B(z - x - \beta) \|^2 \leq 0.
% \]

% Let $g$ be the column of length $n+1$ with elements
% \[
% g_i = \begin{cases}
% 1 & i \in S \\
% h_i & \text{otherwise}.
% \end{cases}
% \]
% Since $w_i = x_i + h_i$ are minimsers of~\eqref{eq:tominimise} we have
% \[
% \| B(z - x - h) \|^2  \leq \| B(z - x - g) \|^2
% \]
% and by Selling's formula
% \begin{equation}\label{eq:sellingswithhg}
%  - \frac{1}{2} \sum_{i=1}^{n+1}\sum_{j=1}^{n+1}q_{ij}d_{ij}^2 \leq 0,
% \end{equation}
% where 
% \[
% d_{ij} = d_{ji} = (\gamma_{ij} - h_j + h_i)^2 - (\gamma_{ij} - g_j + g_i)^2,
% \]
% and where $\gamma_{ij} = z_i - x_i - z_j + x_i$.  When $i, j \in S$ or $i,j \notin S$ we have $d_{ij} = 0$.  When $i \in S, j \notin S$ we have
% \[
% d_{ij} = (\gamma_{ij} - h_j)^2 - (\gamma_{ij} - h_j + 1)^2 = -2(\gamma_{ij} - h_j) - 1.
% \]
% and when $i \notin S, j \in S$ we have
% \[
% d_{ij} = (\gamma_{ij} + h_i)^2 - (\gamma_{ij} - 1 + h_i)^2 = 2(\gamma_{ij} + h_i) - 1.
% \]
% Substituting these values for $d_{ij}$ into~\eqref{eq:sellingswithhg} gives
% \begin{align*}
% -\frac{1}{2} \sum_{i \in S, j \notin S} q_{ij}(2h_j - 2\gamma_{ij} - 1) -  &\frac{1}{2} \sum_{i \notin S, j \in S} q_{ij}( 2h_i + 2\gamma_{ij} - 1 ) \\
% &= -\sum_{i \in S, j \notin S} q_{ij}(2h_j - 2\gamma_{ij} - 1) \leq 0.
% \end{align*}
% where equality on the second line hold because $q_{ij}=q_{ji}$ and $\gamma_{ij}=\gamma_{ji}$.  Use of Sellings formula and similar working to previously shows that
% \[
% \| B(z - x - \alpha) \|^2 - \| B(z - x - \beta) \|^2 = - \sum_{i \in S, j \notin S} q_{ij}(2\alpha_j - 2\gamma_{ij} - 1) 
% \]
% and, since $h_j \geq \alpha_j$ for all $j = 1,\dots,n+1$ and $q_{ij} \leq 0$ for all $i \neq j$ we have 
% \[
% q_{ij}(2\alpha_j - 2\gamma_{ij} - 1) \geq q_{ij}(2h_j - 2\gamma_{ij} - 1)
% \]
% for all $i \neq j$ and so,
% \begin{align*}
% \| B(z - x - \alpha) \|^2 - \| B(z - x - \beta) \|^2 &= - \sum_{i \in S, j \notin S} q_{ij}(2\alpha_j - 2\gamma_{ij} - 1)  \\
% &\leq - \sum_{i \in S, j \notin S} q_{ij}(2h_j - 2\gamma_{ij} - 1) \leq 0.
% \end{align*}
% \end{proof}

% \begin{theorem}
% Let $\Lambda \subset \reals^m$ be an $n$ dimensional lattice of Voronoi's first kind with obtuse superbasis $b_1,\dots,b_{n+1}$ and let $y \in \reals^m$.  Define the vectors $x_1,x_2,\dots$ according to~\eqref{eq:xseqfirsttype}.  There is a positive integer $K \leq (n+1)^2$ such that $x_K$ is a closest lattice point to $y$. 
% \end{theorem}

\section{Computing a closest relevant vector}\label{sec:comp-clos-relev}

In the previous section we showed that the iterative procedure~\eqref{eq:xseqfirsttype} results in a closest lattice point in at most $n$ iterations.  It remains to show that each iteration of the procedure can be computed efficiently.  Specifically, it remains to show that the minimisation over the set of binary vectors $\{0,1\}^{n+1}$ described in~\eqref{eq:pvecmin} can be computed efficiently.  Putting $t = z - u_k$ in~\eqref{eq:pvecmin} we require an efficient method to compute a $p \in \{0,1\}^{n+1}$ such that the binary quadratic form
\[
\| B(t - p) \|^2 = \| \sum_{i=1}^{n+1} b_i (t_i - p_i) \|^2
\]
is minimised.  Expanding this quadratic form gives
\[
\| \sum_{i=1}^{n+1} b_i (t_i - p_i) \|^2 =  \sum_{i=1}^{n+1}\sum_{j=1}^{n+1} q_{ij}t_i t_j -  2\sum_{i=1}^{n+1}\sum_{j=1}^{n+1} q_{ij}t_j p_i + \sum_{i=1}^{n+1}\sum_{j=1}^{n+1} q_{ij} p_i p_j.
\]
The first sum above is independent of $p$ and can be ignored for the purpose of minimisation.  Letting $s_i = \sum_{j=1}^{n+1} q_{ij}t_j$, we can equivalently minimise the binary quadratic form
\begin{equation}\label{eq:quadformnp}
Q(p) = \sum_{i=1}^{n+1} s_i p_i + \sum_{i=1}^{n+1}\sum_{j=1}^{n+1} q_{ij} p_i p_j.
\end{equation}
We will show that a minimiser of $Q(p)$ can be found efficiently be computing a minimum cut in an undirected flow network.
%\section{Quadratic $\{0,1\}$ programs and minimium cuts in graphs}
%This section describes how a binary quadratic program can be mapped into the problem of computing a minimum cut in a weighted graph.  
This technique has appeared previously~\cite{Picard_min_cuts_1974,Sankaran_solving_CDMA_mincut_1998,Ulukus_cdma_mincut_1998,Cormen2001} but we include the derivation here so that this paper is self contained.

Let $G$ be an undirected graph with $n+3$ vertices $v_0, \dots, v_{n+2}$ contained in the set $V$ and edges $e_{ij}$ connecting $v_i$ to $v_j$.  To each edge we assign a \emph{weight} $w_{ij} \in \reals$.  The graph is undirected so the weights are symmetric, that is, $w_{ij} = w_{ji}$.  By calling the vertex $v_0$ the \emph{source} and the vertex $v_{n+2}$ the \emph{sink} the graph $G$ is what is called a \emph{flow network}.  The flow network is \emph{undirected} since the weights assigned to each edge are undirected.  A \emph{cut} in the flow network $G$ is a subset $C \subset V$ of vertices with its complement $\bar{C} \subset V$ such that the source vertex $v_0 \in C$ and the sink vertex $v_{n+2} \in \bar{C}$.  %The vertex $v_0$ is typically called the \emph{source} and the vertex $v_{n+1}$ is typically called the \emph{sink}.

The weight of a cut is
\[
W(C,\bar{C}) = \sum_{i \in I} \sum_{j \in J} w_{ij}, 
\]
where $I = \{ i \mid v_i \in C\}$ and $J = \{j \mid v_j \in \bar{C}\}$.  That is, $W(C,\bar{C})$ is the sum of the weights on the edges crossing from the vertices in $C$ to the vertices in $\bar{C}$.  In what follows we will often drop the argument and write $W$ rather than $W(C,\bar{C})$.  The \emph{minimum cut} is the $C$ and $\bar{C}$ that minimise the weight $W$.  If all of the edge weights $w_{ij}$ for $i \neq j$ are nonnegative, the minimum cut can be computed in 
order $O(n^3)$ arithmetic operations~\cite{Cormen2001,Even_graph_algorithms_1979}.

We require some properties of the weights $w_{ij}$ in relation to $W$.  If the graph is allowed to contain loops, that is, edges from a vertex to itself, then the weight of these edges $w_{ii}$ have no effect on the weight of any cut.  We may choose any values for the $w_{ii}$ without affecting $W$.  We will find it convenient to set $w_{0,0} = w_{n+2,n+2} = 0$.  The remaining $w_{ii}$ we shall specify shortly.  The edge $e_{0,n+2}$ is in every cut.  If a constant is added to the weight of this edge, that is, $w_{0,n+2}$ is replaced by $w_{0,n+2} + c$ then $W$ is replaced by $W + c$ for every $C$ and $\bar{C}$.  In particular, the subsets $C$ and $\bar{C}$ corresponding to the minimum cut are not changed.  We will find it convenient to choose $w_{0,n+2} = w_{n+2,0} = 0$.

If vertex $v_i$ is in $C$ then edge $e_{i,n+2}$ contributes to the weight of the cut.  If $v_i \notin C$, i.e., $v_i \in \bar{C}$, then edge $e_{0,i}$ contributes to the weight of the cut.  So, either $e_{0,i}$ or $e_{i,n+2}$ \emph{but not both} contribute to every cut.  If a constant, say $c$, is added to the weights of these edges, that is, $w_{0,i}$ and $w_{i,n+2}$ are replaced by $w_{0,i} + c$ and $w_{i,n+2} + c$, then $W$ is replaced by $W + c$ for every $C$ and $\bar{C}$.  The $C$ and $\bar{C}$ corresponding the minimum cut are unchanged.  In this way, the minimum cut is only affected by the differences 
\[
d_i = w_{i,n+2} - w_{0,i}
\]
for each $i$ and not the specific values of the weights $w_{i,n+2}$ and $w_{0,i}$.  %For the purpose of computing the minimum cut it is necessary to choose the difference so that the weights $w_{0,i}$ and $w_{i,n+1}$ are nonnegative.

% It is convenient to transform $W(C)$ as follows.  Let $D = C \backslash 0$ and $\bar{D} = \bar{C} \backslash (n+1)$ where $C \backslash 0$ denotes the set $C$ with the element $0$ removed and $\bar{C} \backslash (n+1)$ denotes the set $\bar{C}$ with the element $n+1$ removed.  Now we can write the weight as,
% \[
% W(C) =  \sum_{i \in D}w_{i,n+1} + \sum_{j \in \bar{d}}w_{0,j} +  \sum_{i \in D} \sum_{j \in \bar{D}} w_{ij}
% \]
% becuase $w_{0,0},w_{0,n+2}$ and $w_{n+2,n+2}$ are zero.

We now show how $W(C,\bar{C})$ can be represented as a binary quadratic form.  Put $p_0 = 1$ and $p_{n+2} = 0$ and
\[
p_i = \begin{cases}
1, & i \in C \\
0, & i \in \bar{C}
\end{cases}
\]
for $i = 1,2,\dots,n+1$.  Observe that
\[
p_i(1 - p_j) = \begin{cases}
1, & i \in C, j \in \bar{C} \\
0, & \text{otherwise}.
\end{cases}
\]
The weight can now be written as
\begin{align*}
W(C,\bar{C}) = \sum_{i \in C} \sum_{j \in \bar{C}} w_{ij} = \sum_{i =0}^{n+2} \sum_{j =0}^{n+2} w_{ij} p_i (1 - p_j) = F(p)
\end{align*}
say, where $p_0 = 1$ and $p_{n+2} = 0$. Finding a minimum cut is equivalent to finding the binary vector $p = (p_1, \dots, p_{n+1})$ that minimises $F(p)$.  Write,
\[
F(p) =  \sum_{i=0}^{n+2} \sum_{j =0}^{n+2} w_{ij}p_i - \sum_{i=0}^{n+2} \sum_{j =0}^{n+2} w_{ij} p_ip_j.
\]
Letting $k_i = \sum_{j =0}^{n+2} w_{ij}$, and using that $p_0 = 1$ and $p_{n+2} = 0$,
\[
F(p) = \sum_{i=0}^{n+1}k_ip_i  - w_{00} - \sum_{i=1}^{n+1} w_{i0} p_i - \sum_{j=1}^{n+1} w_{0j} p_j - \sum_{i=1}^{n+1} \sum_{j =1}^{n+1} w_{ij} p_ip_j.
\]
Because $w_{00} = 0$ and $w_{ij} = w_{ji}$ we have
\[
F(p) = k_0 + \sum_{i=1}^{n+1}( k_i  - 2 w_{i0}) p_i  - \sum_{i=1}^{n+1} \sum_{j =1}^{n+1} w_{ij} p_ip_j.
\]
The constant term $k_0$ is unimportant for the purpose of minimisation, and so, finding the minimum cut is equivalent to minimising the binary quadratic form
\[
\sum_{i=1}^{n+1}g_i p_i  - \sum_{i=1}^{n+1} \sum_{j =1}^{n+1} w_{ij} p_ip_j,
\]
where $g_i = k_i  - 2 w_{i0} = d_i + \sum_{j=1}^{n+1} w_{ij}$.  It only remains to observe the equivalence of this quadratic form and $Q(p)$ from~(\ref{eq:quadformnp}) when the weights are assigned to satisfy,
\begin{align*}
&q_{ij} = - w_{ij} \qquad i,j = 1,\dots,n+1 \\
&s_i = g_i = d_i + \sum_{j=1}^{n+1} w_{ij}.
\end{align*}
Because the $q_{ij}$ are nonpositive for $i \neq j$ the weights $w_{ij}$ are nonnegative for all $i \neq j$ with $i,j = 1,\dots,n+1$.  As discussed the value of the weights $w_{ii}$ have no effect on the weight of any cut $W$ so setting $q_{ii} = - w_{ii}$ for  $i = 1,\dots,n+1$ is of no consequence.  Finally the weights $w_{i,n+2}$ and $w_{0,i}$ can be chosen so that both are nonnegative and 
\[
w_{i,n+2} - w_{0,i} = d_i = s_i + \sum_{j=1}^{n+1} q_{ij} = s_i
\]  
because $\sum_{j=1}^{n+1} q_{ij} = 0$ due to the superbase condition~\eqref{eq:superbasecond}.  That is, we choose $w_{i,n+2} = s_i$ and $w_{0,i} = 0$ when $s_i \geq 0$ and $w_{i,n+2}=0$ and $w_{0,i} = -s_i$ when $s_i < 0$.  With these choices, all the weights $w_{ij}$ for $i \neq j$ are nonnegative.  A minimiser of $Q(p)$, and correspondingly a solution of~\eqref{eq:pvecmin} can be computed in $O(n^3)$ operations by computing a minimum cut in the undirected flow network $G$ assigned with these nonnegative weights~\cite{Picard_min_cuts_1974,Sankaran_solving_CDMA_mincut_1998,Ulukus_cdma_mincut_1998,Cormen2001}.  %The closest lattice point is then computed using~\eqref{eq:compnp}.

\section{Discussion}\label{sec:discussion}

The closest lattice point problem has a number of applications, for example, channel coding and data quantisation~\cite{Conway1983VoronoiCodes,Conway1982VoronoiRegions,Conway1982FastQuantDec,Erex2004_lattice_decoding,Erez2005}.  A significant hurdle in the practical application of lattices as codes or as quantisers is that computing a closest lattice point is computationally difficult in general~\cite{micciancio_hardness_2001}.  The best known general purpose algorithms require a number of operations of order $O(2^{2n})$~\cite{Micciancio09adeterministic}.

In this paper we have focused on the class of lattices of Voronoi's first kind.  We have shown that computing a closest point in a lattice of Voronoi's first kind can be achieved in a comparatively modest number of operations of order $O(n^4)$.  Besides being of theoretical interest, the algorithm has potential for practical application.  A question of immediate interest is: do there exist lattices of Voronoi's first kind that produce good codes or good quantisers?  Since lattices that produce good codes and quantisers often also describe dense~\emph{sphere packings}~\cite{SPLAG}, a related question is: do there exist lattices of Voronoi's first kind that produce dense sphere packings?  These question do not appear to have trivial answers.  The questions have heightened importance due to the algorithm described in this paper.

A final remark is that our algorithm assumes that the~\emph{obtuse superbasis} is known in advance.  It is known that all lattices of dimension less than 4 are of Voronoi's first kind and an algorithm exists to recover the obtuse superbasis in this case~\cite{SPLAG}.  Lattices of dimension larger than 4 need not be of Voronoi's first kind.  An interesting question is: given a lattice, is it possible to efficiently decide whether it is of Voronoi's first kind?  A related question is: is it possible to efficiently find an obtuse superbasis if it exists?

%In some applications, for example, the detection of communications signals involving multiple anntanea~\cite{Ryan2008,Wubben_2011} the lattice basis is not known in advance.  It is then of interest whether the

\section{Conclusion}

The paper describes an algorithm to compute a closest lattice point in a lattice of Voronoi's first kind when the obtuse superbasis is known~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.  The algorithm requires $O(n^4)$ operations where $n$ is the dimension of the lattice.  The algorithm iteratively computes a series of relevant vectors that converge to a closest lattice point after at most $n$ terms.   Each relevant vector in the series can be efficiently computed in $O(n^3)$ operations by computing a minimum cut in an undirected flow network.  The algorithm has potential application in communications engineering problems such as coding and quantisation.  An interesting problem for future research is to find lattices of Voronoi's first kind that produce good codes, good quantisers, or dense sphere packings~\cite{SPLAG,Conway1982VoronoiRegions}.

%\small
\bibliography{../bib/bib}

\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2013-12-18 figs/iterativeseriesexample.rb

@@ -1,0 +1,116 @@
+require 'rubypost'
+
+require 'matrix'
+require 'mathn'
+
+include RubyPost
+
+file = RubyPost::File.new('iterativeseriesexample')
+
+pic = Picture.new('picL')
+
+M = Matrix[ [2,0.4], [0.4, 2] ]
+(-7..7).each do |x|
+	(-7..7).each do |y|
+		v = M*Vector[x,y]
+    pic.add_drawable(Fill.new(Circle.new()).scale(0.07.cm).translate(v[0].cm,v[1].cm).colour(0,0,0))
+	end
+end
+
+#matrix with columns given by vector from obtuse superbasis
+B = Matrix[ 
+           [2.0, -1.6, -0.4], 
+           [0.4, 1.6, -2.0] 
+]
+
+@fig1 = Figure.new
+
+#draw a picture of a bigshaded square that we will clip to the Voronoi region
+vorPic = Picture.new('vorregion')
+vorPic.add_drawable(Fill.new(Square.new).scale(15.cm).colour(0.7,0.7,0.7))
+
+#relevant vector given by the appropriate combinations of the obtuse superbasis
+(0..1).each do |x|
+  (0..1).each do |y|
+    (0..1).each do |z|
+      if( ((x+y+z)!=0) && ((x+y+z)!=3) )
+        v = (B*Vector[x,y,z])*0.5
+        pv = Vector[1/v[0], -1/v[1]]*30
+        clipp = Path.new
+        clipp.add_pair(Pair.new((pv+v)[0].cm, (pv+v)[1].cm))
+        clipp.add_pair(Pair.new((pv*(-1)+v)[0].cm, (pv*(-1)+v)[1].cm))
+        clipp.add_pair(Pair.new((pv*(-1)-v)[0].cm, (pv*(-1)-v)[1].cm))
+        clipp.add_pair(Pair.new((pv-v)[0].cm, (pv-v)[1].cm))
+        clipp.add_pair('cycle')
+        @fig1.add_drawable(Clip.new(clipp, vorPic))
+      end
+    end
+  end
+end
+
+clippath = "((-6.5cm,-4.1cm)--(6.5cm,-4.1cm)--(6.5cm,4.1cm)--(-6.5cm,4.1cm)--cycle)"
+v = M*Vector[2,1]
+@fig1.add_drawable(Draw.new(vorPic).translate(v[0].cm,v[1].cm))
[…]

### 2013-12-18 figs/latticefigures.rb

@@ -1,0 +1,95 @@
+require 'rubypost'
+
+require 'matrix'
+require 'mathn'
+
+include RubyPost
+
+file = RubyPost::File.new('latticefigures')
+
+pic = Picture.new('picL')
+
+M = Matrix[ [3,0.6], [0.6, 3] ]
+(-7..7).each do |x|
+	(-7..7).each do |y|
+		v = M*Vector[x,y]
+pic.add_drawable(Fill.new(Circle.new()).scale(0.07.cm).translate(v[0].cm,v[1].cm).colour(0,0,0))
+	end
+end
+
+#matrix with columns given by vector from obtuse superbasis
+B = Matrix[ 
+           [3.0, -2.4, -0.6], 
+           [0.6, 2.4, -3.0] 
+]
+
+fig1 = Figure.new
+
+#draw a picture of a bigshaded square that we will clip to the Voronoi region
+vorPic = Picture.new('vorregion')
+vorPic.add_drawable(Fill.new(Square.new).scale(15.cm).colour(0.7,0.7,0.7))
+
+#relevant vector given by the appropriate combinations of the obtuse superbasis
+(0..1).each do |x|
+  (0..1).each do |y|
+    (0..1).each do |z|
+      if( ((x+y+z)!=0) && ((x+y+z)!=3) )
+        v = (B*Vector[x,y,z])*0.5
+        pv = Vector[1/v[0], -1/v[1]]*30
+        clipp = Path.new
+        clipp.add_pair(Pair.new((pv+v)[0].cm, (pv+v)[1].cm))
+        clipp.add_pair(Pair.new((pv*(-1)+v)[0].cm, (pv*(-1)+v)[1].cm))
+        clipp.add_pair(Pair.new((pv*(-1)-v)[0].cm, (pv*(-1)-v)[1].cm))
+        clipp.add_pair(Pair.new((pv-v)[0].cm, (pv-v)[1].cm))
+        clipp.add_pair('cycle')
+        fig1.add_drawable(Clip.new(clipp, vorPic))
+      end
+    end
+  end
+end
+
+clippath = "((-6.5cm,-4cm)--(6.5cm,-4cm)--(6.5cm,4cm)--(-6.5cm,4cm)--cycle)"
+fig1.add_drawable(Clip.new(clippath, pic))
+fig1.add_drawable(Clip.new(clippath, vorPic))
[…]

### 2013-12-18 paper.tex
minor
added derivation of Sellings formula

@@
 
-\usepackage{../bib/mathbf-abbrevs}
+\usepackage{mathbf-abbrevs}
 
@@
 \begin{lemma}\label{lem:decSellings}
-Let $p \in \reals^{n+1}$, let $S$ be a subset of the indices of $p$, and let $B$ be the matrix from~\eqref{eq:matrxBobtusebasis} with column given by the vectors $b_1,\dots,b_{n+1}$ from the obtuse superbasis.
+Let $p \in \reals^{n+1}$, let $S$ be a subset of the indices of $p$, and let $B$ be the matrix from~\eqref{eq:matrxBobtusebasis} with column given by the superbasis vectors $b_1,\dots,b_{n+1}$.
 The following equalities hold:
@@
 \begin{proof}
-We give a proof for part~\ref{eq:lem:decSellingsdec}.
-The proof for part~\ref{eq:lem:decSellingsinc} is similar.
-From Selling's formula~\cite[Proposition 2.3.1]{Valentin2003_coverings_tilings_low_dimension}~\cite{Selling1874},
+We give a proof for part~\ref{eq:lem:decSellingsdec}.  The proof for part~\ref{eq:lem:decSellingsinc} is similar.  Observe that
+\[
+\sum_{i=1}^{n+1}\sum_{j=1}^{n+1} q_{ij} p_i^2 =  \sum_{i=1}^{n+1}p_i^2\sum_{j=1}^{n+1} q_{ij} = 0
+\]
+and
+\[
+\sum_{i=1}^{n+1}\sum_{j=1}^{n+1} q_{ij} p_j^2 =  \sum_{j=1}^{n+1}p_j^2\sum_{i=1}^{n+1} q_{ij} = 0
+\]
+since $\sum_{i=1}^{n+1} q_{ij} = 0$ as a result of the superbasis condition~\eqref{eq:superbasecond}.  Now
 \begin{align*}
-\|Bp\|^2 &= \| \sum_{i=1}^{n+1} b_i p_i \|^2 \\
-&= - \sum_{i=1}^{n+1}\sum_{j=i+1}^{n+1}q_{ij}(p_i - p_j)^2 \\
-&= - \frac{1}{2} \sum_{i=1}^{n+1}\sum_{j=1}^{n+1}q_{ij}(p_i - p_j)^2, 
+\|Bp\|^2 &= \sum_{i=1}^{n+1} \sum_{j=1}^{n+1} q_{ij} p_i p_j \\
+&= \sum_{i=1}^{n+1} \sum_{j=1}^{n+1} q_{ij} (\tfrac{1}{2}p_j^2 + p_i p_j -  \tfrac{1}{2}p_i^2) \\
+&= -\frac{1}{2} \sum_{i=1}^{n+1} \sum_{j=1}^{n+1} q_{ij} (p_i - p_j)^2.
 \end{align*}
-where $q_{ij} = q_{ji} = b_i \cdot b_j$ are the Selling parameters.  Put $g = p - I_S$.  It follows that
+This is closely related to what is known as Sellings formula~\cite[Proposition 2.3.1]{Valentin2003_coverings_tilings_low_dimension}~\cite{Selling1874}.  Put $g = p - I_S$.  It follows that
 \[
@@
 %\small
-\bibliography{../bib/bib}
+\bibliography{bib}
 

### 2014-01-02 paper.tex
pulled out arguments from Lemmas 4.10 and 4.11 (now 4.12 and 4.13)

@@
 \]
-where $\Lambda \backslash  \{\zerobf\}$ denotes the set of lattice point not equal to the origin $\zerobf$.  This is called the \emph{shortest vector problem}.
+where $\Lambda \backslash  \{\zerobf\}$ denotes the set of lattice points not equal to the origin $\zerobf$.  This is called the \emph{shortest vector problem}.
 
@@
  \begin{proof}
-Since the packing radius $\rho$ is the point on the boundary of the Voronoi that is closest to the origin, the hypersphere $S$ is a subset of the Voronoi cell scaled by $r/\rho$ and translated by $t$.
+The packing radius $\rho$ is the point on the boundary of the Voronoi cell $\vor(\Lambda)$ that is closest to the origin.
+Therefore, the sphere $S$ is a subset of $\vor(\Lambda)$ scaled by $r/\rho$ and translated by $t$.
 That is, $S \subset r/\rho \vor(\Lambda) + t$.
@@
 To add to this it is typically the case that the set of relevant vectors $\relevant(\Lambda)$ must be stored in memory, and so the algorithm can require an amount of memory that grows exponentially with $n$.
-We will show that, for a lattice of Voronoi's first kind, the set of relevant vectors has a compact representation in terms of what is called its \emph{obtuse superbasis}.
+We will show that for a lattice of Voronoi's first kind the set of relevant vectors has a compact representation in terms of what is called its \emph{obtuse superbasis}.
 To store the obtuse superbasis requires an amount of memory of order $O(n^2)$ in the worst case.
@@
 In what follows we will assume that $y$ lies in the space spanned by the basis vectors $b_1,\dots,b_{n}$.
-This assumption is without loss of generality because $x$ is closest lattice point to $y$ if and only if $x$ is a the closest lattice point to the projection of $y$ orthogonal into the space spanned by $b_1,\dots,b_{n+1}$.
+This assumption is without loss of generality because $x$ is closest lattice point to $y$ if and only if $x$ is a the closest lattice point to the projection of $y$ orthogonal into the space spanned by $b_1,\dots,b_{n}$.
 Let
@@
 It is necessary to introduce some notation.
-Given a subset $S$ of indices from $\{1,\dots,n+1\}$ we denote by $I_S$ the vector of length $n+1$ with elements equal to zero except for those with indices from $S$ that are equal to one.
-That is, if $g = I_S$ then $g$ is the vector with elements
-\[
-g_i = \begin{cases}
-1 & i \in S\\
-0 & i \notin S.
-\end{cases}
-\]
-If $p \in \reals^{n+1}$ then $p - I_S$ denotes the vector $p$ with those elements with index from $S$ decremented by one.
-Similarly, $p + I_S$ denotes the vector $p$ with those elements with index from $S$ incremented by one.
+For $S$ and subset of the indices $\{1,\dots,n+1\}$ let $I_S$ denote the column vector of length $n+1$ with $i$th element equal to one if $i \in S$ and $0$ otherwise.
 
 \begin{lemma}\label{lem:decSellings}
-Let $p \in \reals^{n+1}$, let $S$ be a subset of the indices of $p$, and let $B$ be the matrix from~\eqref{eq:matrxBobtusebasis} with column given by the superbasis vectors $b_1,\dots,b_{n+1}$.
+Let $p \in \reals^{n+1}$, let $S$ be a subset of the indices of $p$, and let $B$ be the matrix from~\eqref{eq:matrxBobtusebasis} with columns given by the superbasis vectors $b_1,\dots,b_{n+1}$.
 The following equalities hold:
@@
 \end{enumerate}
-where $q_{ij} = \dotprod{b_i}{b_j}$ are the Selling parameters.
+where $q_{ij} = \dotprod{b_i}{b_j} \leq 0$ are the Selling parameters.
 \end{lemma}
@@
 \]
-since $\sum_{i=1}^{n+1} q_{ij} = 0$ as a result of the superbasis condition~\eqref{eq:superbasecond}.  Now
+since $\sum_{i=1}^{n+1} q_{ij} = \sum_{j=1}^{n+1} q_{ij} = 0$ as a result of the superbasis condition~\eqref{eq:superbasecond}.  Now
 \begin{align*}
@@
 \[
-\rng(w) = \max(v) - 1 - \min(v) = \rng(v) - 1.
+\rng(w) = \max(v) - 1 - \min(v) = \rng(v) - 1 = \ell - 1.
 \]
-The lemma will follow if we show that $B(u+g+w)$ is a closest lattice point to $y$ since then $B(u+g)$ with be $(\ell-1)$-close to $y$.
+The lemma will follow if we show that $B(u+g+w)$ is a closest lattice point to $y$ since then the lattice point $B(u+g)$ with be $(\ell-1)$-close to $y$.
 We show this in two stages.
@@
 \[
-\|B(z-u-g-h+I_D)\|^2 > \|B(z-u-h-g)\|^2.
+\|B(z-u-g-h+I_D)\|^2 > \|B(z-u-g-h)\|^2.
 \]
@@
 \begin{proof}
-Let $x_k = B u_k$ with $u_k \in \ints^{n+1}$ be the lattice point obtained on the $k$th iteration of the procedure.
+Let $x_k = B u_k$ be the lattice point obtained on the $k$th iteration of the procedure.
 Suppose that $x_k$ is $\ell$-close to $y=Bz$ with $\ell > 0$.
@@
 
-We now provide proof of Lemmas~\ref{lem:vqcontramin} and~\ref{lem:vqcontramax} that were used in the proof of Lemma~\ref{lem:rngdecreases}.
+We now provide proof of Lemmas~\ref{lem:vqcontramin} and~\ref{lem:vqcontramax} that were used in the proof of Lemma~\ref{lem:rngdecreases}.  We first require the following two simple Lemmas.
+
+\begin{lemma}\label{lem:simplevminh}
+Let $v \in \ints^{n+1}$ and $h \in \{0,1\}^{n+1}$ be such that $h_i = 0$ whenever $v_i = \min(v)$.
+If $S \subseteq \{ i \mid v_i = \min(v) \}$ then $h_i - h_j \geq v_i - v_j$ for all $i \in S$ and $j = 1,\dots,n+1$.
+\end{lemma}
+\begin{proof}
+For $i \in S$ we have $v_i = \min(v)$ and $h_i = 0$ and so the lemma is true if we show that $-h_j \geq \min(v) - v_j$ for all $j=1,\dots,n+1$.
+If $v_j = \min(v)$ then $h_j = 0$ so that $-h_j = 0 = \min(v) - v_j$.
+Otherwise, if $v_j \neq \min(v)$ then
+\[
+-h_j \geq -1 \geq \min(v) - v_j
+\] 
+because $h_j \in \{0,1\}$ and $v_j$ and $\min(v)$ are integers with $v_j > \min(v)$.
+\end{proof}
+
+\begin{lemma}\label{lem:simplevmaxh}
+Let $v \in \ints^{n+1}$ and $h \in \{0,1\}^{n+1}$ be such that $h_i = 1$ whenever $v_i = \max(v)$.
+If $S \subseteq \{ i \mid v_i = \max(v) \}$ then $h_j - h_i \geq v_j - v_i$ for all $i \in S$ and $j = 1,\dots,n+1$.
+\end{lemma}
+\begin{proof}
+For $i \in S$ we have $v_i = \max(v)$ and $h_i = 1$ and so the lemma is true if we show that $h_j-1 \geq v_j - \max(v)$ for all $j=1,\dots,n+1$.
+If $v_j = \max(v)$ then $h_j = 1$ and $h_j-1 = 0 = v_j - \max(v)$.
+Otherwise, if $v_j \neq \max(v)$ then
+\[
+h_j-1 \geq -1 \geq v_j - \max(v)
+\] 
+because $h_j \in \{0,1\}$ and $v_j$ and $\max(v)$ are integers with $v_j < \max(v)$.
+\end{proof}
 
@@
 \]
-where $\gamma_{ij} = 2(t_j - t_i)$.  Put $h = g - I_S$ and observe that $h_i = 0$ when $v_i = \min(v)$ and so, since $S \subset \{i \mid v_i = \min(v) \}$, we have
+where $\gamma_{ij} = 2(t_j - t_i)$.  Put $h = g - I_S$ and observe that $h_i = 0$ whenever $v_i = \min(v)$ and so, since $S \subset \{i \mid v_i = \min(v) \}$, we have
 \[
-h_i - h_j \geq v_i - v_j \qquad \text{when $i \in S$ and $j=1,\dots,n+1$.}
+h_i - h_j \geq v_i - v_j \qquad \text{when $i \in S$ and $j=1,\dots,n+1$}
 \]
-Thus, since $q_{ij} \leq 0$,
+by Lemma~\ref{lem:simplevminh}.  Thus, since $q_{ij} \leq 0$,
 \[
@@
 The proof mimics that of Lemma~\ref{lem:vqcontramin} with only minor modification.
-We give the proof in full for completeness, and since the small modifications that do exist might not be obvious.
+We give the proof in full for completeness and because the small modifications that do exist might not be obvious.
 From our hypothesis and part~\ref{eq:lem:decSellingsinc} of Lemma~\ref{lem:decSellings} with $p = t-v$ we have
@@
 \[
-h_j - h_i \geq v_j - v_i \qquad \text{for $i \in S$ and $j=1,\dots,n+1$.}
+h_j - h_i \geq v_j - v_i \qquad \text{for $i \in S$ and $j=1,\dots,n+1$}
 \]
-Thus, since $q_{ij} \leq 0$,
+by Lemma~\ref{lem:simplevmaxh}.  Thus, since $q_{ij} \leq 0$,
 \[

### 2014-01-08 paper.tex
minor
minor mods

@@
 It is necessary to introduce some notation.
-For $S$ and subset of the indices $\{1,\dots,n+1\}$ let $I_S$ denote the column vector of length $n+1$ with $i$th element equal to one if $i \in S$ and $0$ otherwise.
+For $S$ a subset of the indices $\{1,\dots,n+1\}$ let $I_S$ denote the column vector of length $n+1$ with $i$th element equal to one if $i \in S$ and $0$ otherwise.
+Thus, for example, $p+I_S$ denotes the vector $p$ with those elements having index from $S$ incremented by one.
 
@@
 The algorithm requires $O(n^4)$ operations where $n$ is the dimension of the lattice.
-The algorithm iteratively computes a series of relevant vectors that converge to a closest lattice point after at most $n$ terms.
+The algorithm iteratively computes a series of relevant vectors that converges to a closest lattice point after at most $n$ terms.
 Each relevant vector in the series can be efficiently computed in $O(n^3)$ operations by computing a minimum cut in an undirected flow network.

### 2014-01-09 paper.tex
minor

@@
 This is achieved by enumerating a series of \emph{relevant vectors} of the lattice.
-Each relevant vector in the series can be computed in $O(n^3)$ operations using an algorithm for computing a minimum cut in an undirected flow network.
+Each relevant vector in the series can be computed in $O(n^3)$ operations using an algorithm for computing a minimum cut in an undirected flow network~\cite{Picard_min_cuts_1974,Sankaran_solving_CDMA_mincut_1998,Ulukus_cdma_mincut_1998,Cormen2001}.
 We show that the series converges to a closest lattice point after at most $n$ terms, resulting in $O(n^4)$ operations in total.
@@
 where $\relevant(\Lambda) \cup \{\zerobf\}$ is the set of relevant vectors of $\Lambda$ including the origin.
-Observe that the minimum over $\relevant(\Lambda) \cup \{\zerobf\}$ may not be unique.
-That is, there may be multiple vectors from $\relevant(\Lambda) \cup \{\zerobf\}$ that are closest to $y - x_k$.
+The minimum over $\relevant(\Lambda) \cup \{\zerobf\}$ may not be unique, that is, there may be multiple vectors from $\relevant(\Lambda) \cup \{\zerobf\}$ that are closest to $y - x_k$.
 In this case, any one of the minimisers may be chosen.
@@
 and so $x_0,x_1,\dots$ is an infinite sequence of distinct (due to the strict inequality) lattice points  all contained inside an $n$-dimensional hypersphere of radius $r = \|y - x_0\|$ centered at $y$.
-This is a contradiction since, if $\rho$ is the packing radius of the lattice, then less than $\ceil{r/\rho}^n$ lattice points lie inside this sphere by Proposition~\ref{eq:latticepointsinsphere}.
+This is a contradiction because, if $\rho$ is the packing radius of the lattice, then less than $\ceil{r/\rho}^n$ lattice points lie inside this sphere by Proposition~\ref{eq:latticepointsinsphere}.
 \end{proof}
@@
 In general a lattice can have as many as $2^{n+1}-2$ relevant vectors and so computing a minimiser directly can require a number of operations that grows exponentially with $n$.
-To add to this it is typically the case that the set of relevant vectors $\relevant(\Lambda)$ must be stored in memory, and so the algorithm can require an amount of memory that grows exponentially with $n$.
+To add to this it is typically the case that the set of relevant vectors $\relevant(\Lambda)$ must be stored in memory and so the algorithm can require an amount of memory that grows exponentially with $n$.
 We will show that for a lattice of Voronoi's first kind the set of relevant vectors has a compact representation in terms of what is called its \emph{obtuse superbasis}.
@@
 The other factor affecting the complexity is the number of iterations required before the algorithm arrives at a closest lattice point, that is, the size of $K$.
-Propositon~\ref{eq:latticepointsinsphere} suggests that this number can be as large as $\ceil{r/\rho}^n$ where $r = \|y - x_0\|^2$ and $\rho$ is the packing radius of the lattice.
-Thus, the number of iterations required can potentially grow exponentially with $n$.
+Propositon~\ref{eq:latticepointsinsphere} suggests that this number might be as large as $\ceil{r/\rho}^n$ where $r = \|y - x_0\|^2$ and $\rho$ is the packing radius of the lattice.
+Thus, the number of iterations required might grow exponentially with $n$.
 The number of iterations required depends on the lattice point that starts the iteration $x_0$.
@@
 Unfortunately, computing close approximations of a closest lattice point is known to be computationally difficult~\cite{feige_inapproximability_2004}.
-We will show that for a lattice of Voronoi's first kind a simple and easy to compute choice for $x_0$ ensures that a closest lattice point is reached in at most $n$ iterations, and so $K \leq n$.
+We will show that for a lattice of Voronoi's first kind a simple and easy to compute choice for $x_0$ ensures that a closest lattice point is reached in at most $n$ iterations and so $K \leq n$.
 Combining this with the fact that each iteration of the algorithm requires $O(n^3)$ operations results in an algorithm that requires $O(n^4)$ operations to compute a closest point in a lattice of Voronoi's first kind.
@@
 \]
-where $I$ is a strict subset of $\{1, 2, \dots, n+1\}$ that is not empty, i.e. $I \subset \{1, 2, \dots, n+1\}$ and $I \neq \emptyset$.
+where $I$ is a strict subset of $\{1, 2, \dots, n+1\}$ that is not empty, i.e., $I \subset \{1, 2, \dots, n+1\}$ and $I \neq \emptyset$.
 \end{theorem}
@@
 \]
-We will use the iterative procedure described in~\eqref{eq:relvectsminimslicer} to do this.
-In what follows we will assume that $y$ lies in the space spanned by the basis vectors $b_1,\dots,b_{n}$.
-This assumption is without loss of generality because $x$ is closest lattice point to $y$ if and only if $x$ is a the closest lattice point to the projection of $y$ orthogonal into the space spanned by $b_1,\dots,b_{n}$.
+The iterative procedure described in~\eqref{eq:relvectsminimslicer} will be used to do this.
+In what follows it is assumed that $y$ lies in the space spanned by the basis vectors $b_1,\dots,b_{n}$.
+This assumption is without loss of generality because $x$ is a closest lattice point to $y$ if and only if $x$ is a the closest lattice point to the orthogonalprojection of $y$ into the space spanned by $b_1,\dots,b_{n}$.
 Let
@@
 \item  ${\displaystyle \|Bp\|^2 - \|B(p + I_S)\|^2 = \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_i - 2p_j)}$, \label{eq:lem:decSellingsinc}
+%\item  ${\displaystyle \|Bp\|^2 - \|B(p + I_S + I_T)\|^2 = \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_i - 2p_j)}$, \label{eq:lem:decSellingsincanddec}
 \end{enumerate}
@@
 
+% \begin{lemma}\label{lem:decSellingsincanddec}
+% With the same symbols as in Lemma~\ref{lem:decSellings} and with $T$ a subset of the indices of $p$,
+% \[
+% \|Bp\|^2 - \|B(p + I_S + I_T)\|^2 = \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_i - 2p_j) + \sum_{i \in T}\sum_{j \notin T}q_{ij}(1 + 2p_i - 2p_j) + \sum_{i \in S}\sum_{j \in T} 2q_{ij}
+% \]
+% \end{lemma}
+
 We denote by $\min(p)$ and $\max(p)$ the minimum and maximum values obtained by the elements of the vector $p$.  We define the function
@@
 To make the definition of $\subrng$ clear we give the following alternative and equivalent definition.
-Let $p \in \ints^n$ and let $\sigma$ be the permutation of the indices $\{1,\dots,n\}$ that puts the elements of $p$ in ascending order, that is
+Let $p \in \reals^n$ and let $\sigma$ be the permutation of the indices $\{1,\dots,n\}$ that puts the elements of $p$ in ascending order, that is
 \[
@@
 \[
-1 + 2(\zeta_i-\zeta_j) - 2(v_i - v_j) < 1 + 2 - 4 = -1 < 0.
+1 + 2(\zeta_i-\zeta_j) - 2(v_i - v_j) < 1 + 2 - 4 = -1 < 0 \qquad \text{for $i \in S$ and $j \notin S$}.
 \]
@@
 \[
-\|B(z - \floor{z} - v)\|^2 - \|B(z - \floor{z} - u)\|^2 \geq 0,
+\|B(z - \floor{z} - v)\|^2 - \|B(z - \floor{z} - u)\|^2 \geq 0
 \]
@@
 That is $v_1=v_2=\dots=v_{n+1}$.
-In this case $Bu = \sum_{i=1}^{n+1} v_n b_n = v_1\sum_{i=1}^{n+1}b_n = 0$
+In this case $Bv = \sum_{i=1}^{n+1} v_n b_n = v_1\sum_{i=1}^{n+1}b_n = 0$
 as a result of the superbasis condition~\eqref{eq:superbasecond}.  Thus $x = x + Bv$ is a closest point to $y$. 
@@
 \]
+
 In this case, Lemma~\ref{lem:vqcontramin} (proved below) with $t = z-u$ and $S = A$ shows that 
 \begin{equation}\label{eq:lem:rng:contag}
-\|B(z-u - g)\|^2 > \min_{p \in \{0,1\}^{n+1}}\|B(z- u - p)\|^2,
+\|B(z- u - g)\|^2 > \min_{p \in \{0,1\}^{n+1}}\|B(z- u - p)\|^2,
 \end{equation}

### 2014-01-09 paper.tex
minor.  Need to cleanup this notation

@@
 \begin{lemma}\label{lem:decSellings}
-Let $p \in \reals^{n+1}$, let $S$ be a subset of the indices of $p$, and let $B$ be the matrix from~\eqref{eq:matrxBobtusebasis} with columns given by the superbasis vectors $b_1,\dots,b_{n+1}$.
+Let $p \in \reals^{n+1}$, let $S$ and $T$ be subsets of the indices of $p$, and let $B$ be the matrix from~\eqref{eq:matrxBobtusebasis} with columns given by the superbasis vectors $b_1,\dots,b_{n+1}$.
 The following equalities hold:
@@
 \item  ${\displaystyle \|Bp\|^2 - \|B(p + I_S)\|^2 = \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_i - 2p_j)}$, \label{eq:lem:decSellingsinc}
-%\item  ${\displaystyle \|Bp\|^2 - \|B(p + I_S + I_T)\|^2 = \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_i - 2p_j)}$, \label{eq:lem:decSellingsincanddec}
+% \item  \label{eq:lem:decSellingsincanddec} \begin{align*}
+% \|Bp\|^2 - \|B(p + I_S + I_T)\|^2 = \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_j - 2p_i) + \sum_{i \notin T}\sum_{j \in T}q_{ij}(1 + 2p_j - 2p_i) + \sum_{i\in S}\sum_{j\in S} q_{ij}, 
+% \end{align*}
 \end{enumerate}
@@
 \begin{proof}
-Because $Bu$ is $\ell$-close to $y$ there exists $v \in \ints^{n+1}$ with $\rng(v) = \ell$ such that $B(u+v)$ is a closest lattice point to $y$.
-Let $A$ be the (possibly empty) subset of indices $\{1,\dots,n+1\}$ such that $g_i = 1$ and $v_i = \min(v)$ and let $D$ be the (possibly empty) subset of indices such that $g_i = 0$ and $v_i = \max(v)$.
-That is,
+Because $Bu$ is $\ell$-close to $y$ there exists $v \in \ints^{n+1}$ with $\rng(v) = \ell$ such that $B(u+v)$ is a closest lattice point to $y$.  Define subsets of indices
 \[
-A = \{i \mid g_i = 1, v_i = \min(v) \}, \qquad D = \{i \mid g_i = 0, v_i = \max(v) \}.
+A = \{i \mid g_i = 1, v_i = \min(v) \}, \qquad D = \{i \mid g_i = 0, v_i = \max(v) \}
 \]
-Put $d = g - I_A + I_D$ and put $w = v - d$.
+and put $d = g - I_A + I_D \in \{0,1\}^{n+1}$ and $w = v - d$.
 Observe that $d_i = 0$ whenever $v_i = \min(v)$ and so $\min(w) = \min(v-d) = \min(v)$.
@@
 We show this in two stages.
-First put $h = v - g + I_A$.
+First put $f = g - I_A \in \{0,1\}^{n+1}$ and  $h = v - f = v - g + I_A$.
 Suppose that $B(u+g+h)$ is not a closest lattice point to $y = Bz$.
-That is, suppose that
+That is, with $t = z-u$, suppose that
 \[
-\|B(z-u-g-h)\|^2 = \|B(z-u-v-I_A)\|^2 > \|B(z-u-v)\|^2.
+\|B(t-g-h)\|^2 = \|B(t-v-I_A)\|^2 > \|B(t-v)\|^2.
+\]
+By part~\ref{eq:lem:decSellingsdec} of Lemma~\ref{lem:decSellings} we obtain the inequality
+\[
+\|B(t-v)\|^2 - \|B(t-v-I_A)\|^2 =  \sum_{i \in A}\sum_{j \notin A}q_{ij}(1 + 2(t_j-t_i) + 2(v_i-v_j)) < 0.
+\]
+Because $v_i = \min(v)$ only when $f_i= g_i - 1 = 0$ and because $v_i = \min(v)$ for $i \in A$ we have $f_i - f_j \geq v_i - v_j$ for all $i \in S$ and $j =1,\dots,n+1$.
+Now, since $q_{ij} \leq 0$ for $i\neq j$,
+\begin{align*}
+\|B(z-u-f)\|^2 - \|B(z-u-g)\|^2 &= \|B(t-f)\|^2 - \|B(t-f-I_A)\|^2 \\
+&= \sum_{i \in A}\sum_{j \notin A}q_{ij}(1 + 2(t_j-t_i) + 2(f_i-f_j)) \\
+&\leq  \sum_{i \in A}\sum_{j \notin A}q_{ij}(1 + 2(t_j-t_i) + 2(v_i-v_j)) < 0,
+\end{align*}
+contradicting~\eqref{eq:qismin} because $f \in \{0,1\}^{n+1}$.  Thus, $B(u+g+h)$ is a closest lattice point to $y = Bz$.
+
+Now suppose that $B(u+g+w) = B(u+g+h-I_D)$ is not a closest lattice point to $y$.  That is, suppose that
+\[
+\|B(t-g-w)\|^2 = \|B(t-g-h+I_D)\|^2 > \|B(t-g-h)\|^2.
+\]
+By part~\ref{eq:lem:decSellingsinc} of Lemma~\ref{lem:decSellings} we obtain the inequality
+\[
+\|B(t-v)\|^2 - \|B(t-v-I_A)\|^2 =  \sum_{i \in A}\sum_{j \notin A}q_{ij}(1 + 2(t_j-t_i) + 2(v_i-v_j)) < 0.
 \]
 
-In this case, Lemma~\ref{lem:vqcontramin} (proved below) with $t = z-u$ and $S = A$ shows that 
-\begin{equation}\label{eq:lem:rng:contag}
-\|B(z- u - g)\|^2 > \min_{p \in \{0,1\}^{n+1}}\|B(z- u - p)\|^2,
-\end{equation}
-contradicting~\eqref{eq:qismin}.  Thus, $B(u+g+h)$ is a closest lattice point to $y$.  Now suppose that $B(u+g+w) = B(u+g+h-I_D)$ is not a closest lattice point to $y$.  That is, suppose that
-\[
-\|B(z-u-g-h+I_D)\|^2 > \|B(z-u-g-h)\|^2.
-\]
-Lemma~\ref{lem:vqcontramax} (proved below) with $t = z-u$ and $S = D$ shows that~\eqref{eq:lem:rng:contag} occurs again, contradicting~\eqref{eq:qismin}.
+% In this case, Lemma~\ref{lem:vqcontramin} (proved below) with $t = z-u$ and $S = A$ shows that 
+% \begin{equation}\label{eq:lem:rng:contag}
+% \|B(z- u - g)\|^2 > \min_{p \in \{0,1\}^{n+1}}\|B(z- u - p)\|^2,
+% \end{equation}
+% contradicting~\eqref{eq:qismin}.  Thus, $B(u+g+h)$ is a closest lattice point to $y$.  Now suppose that $B(u+g+w) = B(u+g+h-I_D)$ is not a closest lattice point to $y$.  That is, suppose that
+% \[
+% \|B(z-u-g-h+I_D)\|^2 > \|B(z-u-g-h)\|^2.
+% \]
+% Lemma~\ref{lem:vqcontramax} (proved below) with $t = z-u$ and $S = D$ shows that~\eqref{eq:lem:rng:contag} occurs again, contradicting~\eqref{eq:qismin}.
+
 Thus, $B(u+g+w)$ is a closest lattice point to $y$ and, since $\rng(w) = \ell-1$, the lattice point $B(u+g)$ is $(\ell-1)$-close to $y$.

### 2014-01-10 paper.tex
added to paper.  This is nicer, but I need to fix a few bugs

--- 2014-01-09 paper.tex
@@
 x_{k+1} &= B u_{k+1} \label{eq:xseqfirsttype}  \\
-u_{k+1} &= u_k + p_k \nonumber \\
-p_k &= \arg\min_{p \in \{0,1\}^{n+1}}\| B(z - u_k - p) \|^2, \label{eq:pvecmin}
+u_{k+1} &= u_k + t_k \nonumber \\
+t_k &= \arg\min_{t \in \{0,1\}^{n+1}}\| B(z - u_k - t) \|^2, \label{eq:pvecmin}
 \end{align}
@@
 It is necessary to introduce some notation.
-For $S$ a subset of the indices $\{1,\dots,n+1\}$ let $I_S$ denote the column vector of length $n+1$ with $i$th element equal to one if $i \in S$ and $0$ otherwise.
-Thus, for example, $p+I_S$ denotes the vector $p$ with those elements having index from $S$ incremented by one.
+For $S$ a subset of the indices $\{1,\dots,n+1\}$ let $\onebf_S$ denote the column vector of length $n+1$ with $i$th element equal to one if $i \in S$ and $0$ otherwise.
+%Thus, for example, if $p \in \reals^{n+1}$, then $p+\onebf_S$ denotes the vector $p$ with those elements having index from $S$ incremented by one.
+We define the function
+\[
+\Phi(S, p) = \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_i - 2p_j)
+\]
+where $q_{ij} = \dotprod{b_i}{b_j}$ are the Selling parameters.  We denote by $\bar{S}$ the complement of the set of indices $S$, that is $\bar{S} = \{ i =1,\dots,n+1 \mid i \notin S\}$.
 
 \begin{lemma}\label{lem:decSellings}
-Let $p \in \reals^{n+1}$, let $S$ be a subset of the indices of $p$, and let $B$ be the matrix from~\eqref{eq:matrxBobtusebasis} with columns given by the superbasis vectors $b_1,\dots,b_{n+1}$.
-The following equalities hold:
+Let $p \in \reals^{n+1}$ and let $S$ and $T$ be subsets of the indices of $p$.  The following equalities hold:
 \begin{enumerate}
-\item  ${\displaystyle \|Bp\|^2 - \|B(p - I_S)\|^2 = \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_j - 2p_i)}$, \label{eq:lem:decSellingsdec}
-\item  ${\displaystyle \|Bp\|^2 - \|B(p + I_S)\|^2 = \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_i - 2p_j)}$, \label{eq:lem:decSellingsinc}
-%\item  ${\displaystyle \|Bp\|^2 - \|B(p + I_S + I_T)\|^2 = \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_i - 2p_j)}$, \label{eq:lem:decSellingsincanddec}
+\item  $\|Bp\|^2 - \|B(p + \onebf_S)\|^2 = \Phi(S, p)$, \label{eq:lem:decSellingsinc}
+\item  $\|Bp\|^2 - \|B(p - \onebf_S)\|^2 = \Phi(\bar{S}, p)$, \label{eq:lem:decSellingsdec}
+\item  ${\displaystyle \|Bp\|^2 - \|B(p + \onebf_S - \onebf_T)\|^2 = \Phi(S, p) + \Phi(\bar{T},p) + \sum_{i\in S}\sum_{j\in T} q_{ij} }$.  \label{eq:lem:decSellingsincdec}
 \end{enumerate}
-where $q_{ij} = \dotprod{b_i}{b_j} \leq 0$ are the Selling parameters.
 \end{lemma}
 \begin{proof}
-We give a proof for part~\ref{eq:lem:decSellingsdec}.  The proof for part~\ref{eq:lem:decSellingsinc} is similar.  Observe that
-\[
-\sum_{i=1}^{n+1}\sum_{j=1}^{n+1} q_{ij} p_i^2 =  \sum_{i=1}^{n+1}p_i^2\sum_{j=1}^{n+1} q_{ij} = 0
-\]
-and
-\[
-\sum_{i=1}^{n+1}\sum_{j=1}^{n+1} q_{ij} p_j^2 =  \sum_{j=1}^{n+1}p_j^2\sum_{i=1}^{n+1} q_{ij} = 0
-\]
-since $\sum_{i=1}^{n+1} q_{ij} = \sum_{j=1}^{n+1} q_{ij} = 0$ as a result of the superbasis condition~\eqref{eq:superbasecond}.  Now
+Part~\ref{eq:lem:decSellingsincdec} follows immediately from parts~\ref{eq:lem:decSellingsinc} and~\ref{eq:lem:decSellingsdec} because
 \begin{align*}
-\|Bp\|^2 &= \sum_{i=1}^{n+1} \sum_{j=1}^{n+1} q_{ij} p_i p_j \\
-&= \sum_{i=1}^{n+1} \sum_{j=1}^{n+1} q_{ij} (\tfrac{1}{2}p_j^2 + p_i p_j -  \tfrac{1}{2}p_i^2) \\
-&= -\frac{1}{2} \sum_{i=1}^{n+1} \sum_{j=1}^{n+1} q_{ij} (p_i - p_j)^2.
+\|Bp\|^2 - &\|B(p + \onebf_S - \onebf_T)\|^2 \\
+&= \|Bp\|^2-\|B(p + \onebf_S)\|^2 +  \|Bp\|^2-\|B(p - \onebf_T)\|^2 + \sum_{i\in S}\sum_{j\in T} q_{ij}.
 \end{align*}
-This is closely related to what is known as Sellings formula~\cite[Proposition 2.3.1]{Valentin2003_coverings_tilings_low_dimension}~\cite{Selling1874}.  Put $g = p - I_S$.  It follows that
-\[
-\|Bp\|^2 - \|Bg\|^2 = \frac{1}{2} \sum_{i=1}^{n+1}\sum_{j=1}^{n+1}q_{ij}d_{ij},
-\]
-where $d_{ij} = (g_i - g_j)^2 - (p_i - p_j)^2$.  If both $i \notin S$ and $j \notin S$ then $p_i = g_i$, $p_j = g_j$ and so $d_{ij} = 0$.  Similarly if both $i \in S$ and $j \in S$ then 
-\[
-d_{ij} = (p_i-1 - p_j+1)^2 - (p_i - p_j)^2 = 0.
-\]
-If $i \in S$ and $j \notin S$ then
-\[
-d_{ij} = (p_i-1 - p_j)^2 - (p_i - p_j)^2= 1 + 2p_j - 2p_i,
-\]
-while if $i \notin S$ and $j \in S$ then
-\[
-d_{ij} = (p_i - p_j+1)^2 - (p_i - p_j)^2 = 1 + 2p_i - 2p_j.
-\]
-We thus have
+We give a proof for part~\ref{eq:lem:decSellingsinc}.  The proof for part~\ref{eq:lem:decSellingsdec} is similar.  
+Put $Q = B^\prime B$ where supercript $^\prime$ indicates the vector or matrix transpose.
+Denote by $\onebf$ the column vector of length $n+1$ containing all ones.
+We have $B\onebf = \sum_{i=1}^{n+1}b_i = \zerobf$ as a result of the superbasis condition~\eqref{eq:superbasecond} and so $Q\onebf = 0$.
+Observe that $\onebf_S = \onebf - \onebf_{\bar{S}}$ and so $Q\onebf_S = -Q\onebf_{\bar{S}}$.
+With $\circ$ the elementwise vector product, i.e., the Schur or Hadamard product, we have
  \begin{align*}
-\|Bp\|^2 - \|Bg\|^2 &= \frac{1}{2} \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_j - 2p_i) + \frac{1}{2} \sum_{i \notin S}\sum_{j \in S}q_{ij}(1 + 2p_i - 2p_j) \\
-&= \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_j - 2p_i),
+ \|Bp\|^2 - \|B(p + \onebf_S)\|^2 &= -\onebf^\prime_S Q \onebf_S - 2 p^\prime Q \onebf_S \\
+ &= \onebf^\prime_S Q \onebf_{\bar{S}} - 2 p^\prime Q \onebf_S \\
+&= \onebf^\prime_S Q \onebf_{\bar{S}} - 2 (p\circ\onebf_{\bar{S}})^\prime Q \onebf_S - 2 (p\circ\onebf_S)^\prime Q \onebf_S \\
+&= \onebf^\prime_S Q \onebf_{\bar{S}} - 2 (p\circ\onebf_{\bar{S}})^\prime Q \onebf_{S} + 2 (p\circ\onebf_S)^\prime Q \onebf_{\bar{S}} \\
  \end{align*}
-where the last line holds because $q_{ij} = q_{ji}$.
+which is precisely $\Phi(S,p)$.
+% Observe that
+% \[
+% \sum_{i=1}^{n+1}\sum_{j=1}^{n+1} q_{ij} p_i^2 =  \sum_{i=1}^{n+1}p_i^2\sum_{j=1}^{n+1} q_{ij} = 0
+% \]
+% and
+% \[
+% \sum_{i=1}^{n+1}\sum_{j=1}^{n+1} q_{ij} p_j^2 =  \sum_{j=1}^{n+1}p_j^2\sum_{i=1}^{n+1} q_{ij} = 0
+% \]
+% since $\sum_{i=1}^{n+1} q_{ij} = \sum_{j=1}^{n+1} q_{ij} = 0$ as a result of the superbasis condition~\eqref{eq:superbasecond}.  Now
+% \begin{align*}
+% \|Bp\|^2 &= \sum_{i=1}^{n+1} \sum_{j=1}^{n+1} q_{ij} p_i p_j \\
+% &= \sum_{i=1}^{n+1} \sum_{j=1}^{n+1} q_{ij} (\tfrac{1}{2}p_j^2 + p_i p_j -  \tfrac{1}{2}p_i^2) \\
+% &= -\frac{1}{2} \sum_{i=1}^{n+1} \sum_{j=1}^{n+1} q_{ij} (p_i - p_j)^2.
+% \end{align*}
+% This is closely related to what is known as Sellings formula~\cite[Proposition 2.3.1]{Valentin2003_coverings_tilings_low_dimension}~\cite{Selling1874}.  Put $g = p - \onebf_S$.  It follows that
+% \[
+% \|Bp\|^2 - \|Bg\|^2 = \frac{1}{2} \sum_{i=1}^{n+1}\sum_{j=1}^{n+1}q_{ij}d_{ij},
+% \]
+% where $d_{ij} = (g_i - g_j)^2 - (p_i - p_j)^2$.  If both $i \notin S$ and $j \notin S$ then $p_i = g_i$, $p_j = g_j$ and so $d_{ij} = 0$.  Similarly if both $i \in S$ and $j \in S$ then 
+% \[
+% d_{ij} = (p_i-1 - p_j+1)^2 - (p_i - p_j)^2 = 0.
+% \]
+% If $i \in S$ and $j \notin S$ then
+% \[
+% d_{ij} = (p_i-1 - p_j)^2 - (p_i - p_j)^2= 1 + 2p_j - 2p_i,
+% \]
+% while if $i \notin S$ and $j \in S$ then
+% \[
+% d_{ij} = (p_i - p_j+1)^2 - (p_i - p_j)^2 = 1 + 2p_i - 2p_j.
+% \]
+% We thus have
+% \begin{align*}
+% \|Bp\|^2 - \|Bg\|^2 &= \frac{1}{2} \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_j - 2p_i) + \frac{1}{2} \sum_{i \notin S}\sum_{j \in S}q_{ij}(1 + 2p_i - 2p_j) \\
+% &= \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_j - 2p_i),
+% \end{align*}
+% where the last line holds because $q_{ij} = q_{ji}$.
 \end{proof}
@@
 % \[
-% \|Bp\|^2 - \|B(p + I_S + I_T)\|^2 = \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_i - 2p_j) + \sum_{i \in T}\sum_{j \notin T}q_{ij}(1 + 2p_i - 2p_j) + \sum_{i \in S}\sum_{j \in T} 2q_{ij}
+% \|Bp\|^2 - \|B(p + \onebf_S + \onebf_T)\|^2 = \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_i - 2p_j) + \sum_{i \in T}\sum_{j \notin T}q_{ij}(1 + 2p_i - 2p_j) + \sum_{i \in S}\sum_{j \in T} 2q_{ij}
 % \]
@@
 
-We denote by $\min(p)$ and $\max(p)$ the minimum and maximum values obtained by the elements of the vector $p$.  We define the function
+Denote by $\min(p)$ and $\max(p)$ the minimum and maximum values obtained by the elements of the vector $p$ and define the function
 \[
@@
 \] 
-to return the difference between the maximum and minimum of $p$.
-For example, if $p = (2,-1,4) \in \ints^3$ then $\rng(p) = 4 - (-1)=5$.
-We might also write simply $\rng(4,-1,5)=5$.
+to return the difference between the maximum and minimum of $p$.  %For example, if $p = (2,-1,4) \in \ints^3$ then $\rng(p) = 4 - (-1)=5$.  We might also 
+%For example $\rng(2,-1,4)= 4 - (-1) = 5$.  
 Observe that $\rng(p)$ cannot be negative and that if $\rng(p) = 0$ then all of the elements of $p$ are equal.
@@
 \[
-\decrng(p) = p -  I_{\subrng(p)}
+\decrng(p) = p -  \onebf_{\subrng(p)}
 \]
@@
 \begin{align*}
-\decrng(2,-1,4) &= (2,-1,4) - I_{\subrng(2,-1,4)} = (2,-1,4) - I_{\{1,3\}} = (1,-1,3) \\ 
-\decrng(1,-1,3) &= (1,-1,3) - I_{\{1,3\}} = (0,-1,2) \\ 
-\decrng(0,-1,2) &= (0,-1,2) - I_{\{3\}} = (0,-1,1) \\ 
-\decrng(0,-1,1) &= (0,-1,1) - I_{\emptyset} = (0,-1,1).
+\decrng(2,-1,4) &= (2,-1,4) - \onebf_{\subrng(2,-1,4)} = (2,-1,4) - \onebf_{\{1,3\}} = (1,-1,3) \\ 
+\decrng(1,-1,3) &= (1,-1,3) - \onebf_{\{1,3\}} = (0,-1,2) \\ 
+\decrng(0,-1,2) &= (0,-1,2) - \onebf_{\{3\}} = (0,-1,1) \\ 
+\decrng(0,-1,1) &= (0,-1,1) - \onebf_{\emptyset} = (0,-1,1).
 \end{align*}
@@
 It remains to prove the lemma when $\subrng(v) \neq \emptyset$.
-In this case put $S = \subrng(v)$ and put $u = \decrng(v) = v - I_S$.
+In this case put $S = \subrng(v)$ and put $u = \decrng(v) = v - \onebf_S$.
 Let $\zeta = z - \floor{z}$ be the column vector containing the fractional parts of the elements of $z$.
-We have $\zeta - u = \zeta - v + I_S$.
+We have $\zeta - u = \zeta - v + \onebf_S$.
 Applying part~\ref{eq:lem:decSellingsinc} of Lemma~\ref{lem:decSellings} with $p = \zeta - v$ we obtain
 \begin{align}
-\|B(\zeta - v)\|^2 - \|B(\zeta - u)\|^2 &= \sum_{i \in S}\sum_{j \notin S}q_{ij}\big(1 + 2(\zeta_i - v_i)  - 2(\zeta_j - v_j) \big) \nonumber \\
+\|B(\zeta - v)\|^2 - \|B(\zeta - u)\|^2 &= \Phi(S, \zeta-v) \nonumber \\
 &= \sum_{i \in S}\sum_{j \notin S}q_{ij}\big(1 + 2(\zeta_i-\zeta_j) - 2(v_i - v_j)\big). \label{eq:sumsumBzBu}
@@
 
+Before giving the proof of Lemma~\ref{lem:rngdecreases} we require the following simple result.
+
+\begin{lemma}\label{eq:integergreaterless}
+Let $h \in \{0,1\}^{n+1}$ and $v \in \ints^{n+1}$.  Suppose that
+$h_i = 0$ whenever $v_i = \min(v)$ and that $h_i = 1$ whenever $v_i = \max(v)$.  Then
+\[
+h_i - h_j \geq v_i - v_j
+\]
+when either $v_i = \min(v)$ or $v_j = \max(v)$.
+\end{lemma}
+\begin{proof}
+If $v_i = \min(v)$ then $h_i = 0$ and we need only show that $-h_j \geq \min(v) - v_j$ for all $j$.
+If $v_j = \min(v)$ then $h_j = 0 = \min(v) - v_j$ and the results holds.
+Otherwise if $v_j > \min(v)$ then $-h_j \geq -1 \geq \min(v) - v_j$ because $\min(v)$ and $v_j$ are integers.
+
+Now, if $v_j = \max(v)$ then $h_j = 1$ and we need only show that $h_i - 1 \geq v_i - \max(v)$ for all $i$.
+If $v_i = \max(v)$ then $h_i = 1$ and $h_i - 1 = 0 = v_i - \max(v)$.
+Otherwise if $v_i < \max(v)$ then $h_i - 1 \geq -1 \geq v_i - \max(v)$.
+\end{proof}
+
 \begin{lemma}\label{lem:rngdecreases}
@@
 \begin{equation}\label{eq:qismin}
-\|B(z - u - g)\|^2 = \min_{p \in \{0,1\}^{n+1}}\|B(z - u - p)\|^2.
+\|B(z - u - g)\|^2 = \min_{t \in \{0,1\}^{n+1}}\|B(z - u - t)\|^2.
 \end{equation}
@@
 \begin{proof}
-Because $Bu$ is $\ell$-close to $y$ there exists $v \in \ints^{n+1}$ with $\rng(v) = \ell$ such that $B(u+v)$ is a closest lattice point to $y$.
-Let $A$ be the (possibly empty) subset of indices $\{1,\dots,n+1\}$ such that $g_i = 1$ and $v_i = \min(v)$ and let $D$ be the (possibly empty) subset of indices such that $g_i = 0$ and $v_i = \max(v)$.
-That is,
+Because $Bu$ is $\ell$-close to $y$ there exists $v \in \ints^{n+1}$ with $\rng(v) = \ell$ such that $B(u+v)$ is a closest lattice point to $y$.  Define subsets of indices
 \[
-A = \{i \mid g_i = 1, v_i = \min(v) \}, \qquad D = \{i \mid g_i = 0, v_i = \max(v) \}.
+S = \{i \mid g_i = 0, v_i = \max(v) \}, \qquad T = \{i \mid g_i = 1, v_i = \min(v) \}, 
 \]
-Put $d = g - I_A + I_D$ and put $w = v - d$.
-Observe that $d_i = 0$ whenever $v_i = \min(v)$ and so $\min(w) = \min(v-d) = \min(v)$.
-Also, $d_i = 1$ whenever $v_i = \max(v)$ and so $\max(w) = \max(v-d) = \max(v) - 1$.
+and put $h = g + \onebf_S - \onebf_T \in \{0,1\}^{n+1}$ and $w = v - h$.
+Observe that $h_i = 0$ whenever $v_i = \min(v)$ and so $\min(w) = \min(v-h) = \min(v)$.
+Also, $h_i = 1$ whenever $v_i = \max(v)$ and so $\max(w) = \max(v-h) = \max(v) - 1$.
 Thus,
@@
 The lemma will follow if we show that $B(u+g+w)$ is a closest lattice point to $y$ since then the lattice point $B(u+g)$ with be $(\ell-1)$-close to $y$.
-We show this in two stages.
-First put $h = v - g + I_A$.
-Suppose that $B(u+g+h)$ is not a closest lattice point to $y = Bz$.
-That is, suppose that
+The proof is by contradiction.
+Suppose that $B(u+g+w)$ is not a closest point to $y$, that is, suppose that
 \[
-\|B(z-u-g-h)\|^2 = \|B(z-u-v-I_A)\|^2 > \|B(z-u-v)\|^2.
+\|B(z-u-g-w)\|^2 > |B(z-u-v)\|^2.
 \]
+Putting $p = z-u-v$ we have
+\[
+\|B(p+\onebf_S-\onebf_T)\|^2 > \|Bp\|^2.
+\]
+By part~\ref{eq:lem:decSellingsincdec} of Lemma~\ref{lem:decSellings} we obtain
+\begin{equation}\label{eq:BpBpdecincineq}
+\|Bp\|^2 - \|B(p+\onebf_S-\onebf_T)\|^2 = \Phi(S,p) + \Phi(\bar{T}, p) + \sum_{i\in S}\sum_{j \in S} q_{ij} < 0.
+\end{equation}
 
-In this case, Lemma~\ref{lem:vqcontramin} (proved below) with $t = z-u$ and $S = A$ shows that 
-\begin{equation}\label{eq:lem:rng:contag}
-\|B(z- u - g)\|^2 > \min_{p \in \{0,1\}^{n+1}}\|B(z- u - p)\|^2,
+As stated already $h_i = 0$ whenever $v_i = \min(v)$ and $h_i = 1$ whenever $v_i = \max(v)$.  It follows by Lemma~\ref{eq:integergreaterless} that
+\begin{equation}\label{eq:hijvijieq}
+h_{i} - h_j \geq v_{i} - v_j
 \end{equation}
-contradicting~\eqref{eq:qismin}.  Thus, $B(u+g+h)$ is a closest lattice point to $y$.  Now suppose that $B(u+g+w) = B(u+g+h-I_D)$ is not a closest lattice point to $y$.  That is, suppose that
+when either $v_i = \min(v)$ or $v_j = \max(v)$.  Since $v_i = \min(v)$ for $i \in T$ and $v_j = \max(v)$ for $j \in S$ the inequality~\eqref{eq:hijvijieq} holds whenever $i \in T$ or $j \in S$.
+
+Put $r = z - u - h$.  By~\eqref{eq:hijvijieq} we have $r_i - r_j \leq p_{i} - p_{j}$ when either $i \in T$ or $j \in S$.  Now, since $q_{ij} \leq 0$ for $i \neq j$,
 \[
-\|B(z-u-g-h+I_D)\|^2 > \|B(z-u-g-h)\|^2.
+\Phi(S,r) = \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2r_i - 2r_j) \geq \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_i - 2p_j) = \Phi(S,p).
 \]
-Lemma~\ref{lem:vqcontramax} (proved below) with $t = z-u$ and $S = D$ shows that~\eqref{eq:lem:rng:contag} occurs again, contradicting~\eqref{eq:qismin}.
-Thus, $B(u+g+w)$ is a closest lattice point to $y$ and, since $\rng(w) = \ell-1$, the lattice point $B(u+g)$ is $(\ell-1)$-close to $y$.
+Similarly
+\[
+\Phi(\bar{T},r) = \sum_{i \notin T}\sum_{j \in T}q_{ij}(1 + 2r_i - 2r_j) \geq \sum_{i \notin T}\sum_{j \in T}q_{ij}(1 + 2p_i - 2p_j) = \Phi(\bar{T},p).
+\]\
+Now
+\begin{align*}
+\|B(z-u-h)\|^2 - \|B(z-u-g)\|^2 &= \|Br\|^2 - \|B(r+\onebf_S - \onebf_T)\|^2 \\
+&= \Phi(S,r) + \Phi(\bar{T},r) + \sum_{i\in S}\sum_{j \in S} q_{ij} \\
+&\leq \Phi(S,p) + \Phi(\bar{T}, p) + \sum_{i\in S}\sum_{j \in S} q_{ij} < 0. 
+\end{align*}
+as a result of~\eqref{eq:BpBpdecincineq}.  However, $d \in \{0,1\}^{n+1}$ and so
+\[
+\|B(z-u-g)\|^2 > \|B(z-u-h)\|^2 \geq \min_{t \in \{0,1\}^{n+1}}\|B(z - u - t)\|^2
+\]
+contradicting~\eqref{eq:qismin}.  Thus, $B(u+g+w)$ is a closest lattice point to $y$ and, since $\rng(w) = \ell-1$, the lattice point $B(u+g)$ is $(\ell-1)$-close to $y$.
  \end{proof}
 
-The proof of the previous Lemma~\ref{lem:rngdecreases} relies upon Lemmas~\ref{lem:vqcontramin} and~\ref{lem:vqcontramax}.
-Before we prove these lemmas we state the following theorem asserting that the iterative procedure~\eqref{eq:xseqfirsttype} converges to a closest lattice point in $K \leq n$ iterations.
-This is the primary result of this section.
+The next theorem asserts that the iterative procedure~\eqref{eq:xseqfirsttype} converges to a closest lattice point in $K \leq n$ iterations.  This is the primary result of this section.
 
@@
 Suppose that $x_k$ is $\ell$-close to $y=Bz$ with $\ell > 0$.
-The procedure computes $p_{k} \in \{0,1\}^{n+1}$ satisfying
+The procedure computes $t_{k} \in \{0,1\}^{n+1}$ satisfying
 \[
-\|B(z - u_k - p_{k})\|^2 = \min_{p \in \{0,1\}^{n+1}}\|B(z - u_k - p)\|^2
+\|B(z - u_k - t_{k})\|^2 = \min_{t \in \{0,1\}^{n+1}}\|B(z - u_k - t)\|^2
 \]
-and puts $x_{k+1} = B(u_k + p_k)$.
+and puts $x_{k+1} = B(u_k + t_k)$.
 It follows from Lemma~\ref{lem:rngdecreases} that $x_{k+1}$ is $(\ell-1)$-close to $y$.
@@
 \end{proof}
-
-We now provide proof of Lemmas~\ref{lem:vqcontramin} and~\ref{lem:vqcontramax} that were used in the proof of Lemma~\ref{lem:rngdecreases}.  We first require the following two simple Lemmas.
-
-\begin{lemma}\label{lem:simplevminh}
-Let $v \in \ints^{n+1}$ and $h \in \{0,1\}^{n+1}$ be such that $h_i = 0$ whenever $v_i = \min(v)$.
-If $S \subseteq \{ i \mid v_i = \min(v) \}$ then $h_i - h_j \geq v_i - v_j$ for all $i \in S$ and $j = 1,\dots,n+1$.
-\end{lemma}
-\begin{proof}
-For $i \in S$ we have $v_i = \min(v)$ and $h_i = 0$ and so the lemma is true if we show that $-h_j \geq \min(v) - v_j$ for all $j=1,\dots,n+1$.
-If $v_j = \min(v)$ then $h_j = 0$ so that $-h_j = 0 = \min(v) - v_j$.
-Otherwise, if $v_j \neq \min(v)$ then
-\[
--h_j \geq -1 \geq \min(v) - v_j
-\] 
-because $h_j \in \{0,1\}$ and $v_j$ and $\min(v)$ are integers with $v_j > \min(v)$.
-\end{proof}
-
-\begin{lemma}\label{lem:simplevmaxh}
-Let $v \in \ints^{n+1}$ and $h \in \{0,1\}^{n+1}$ be such that $h_i = 1$ whenever $v_i = \max(v)$.
-If $S \subseteq \{ i \mid v_i = \max(v) \}$ then $h_j - h_i \geq v_j - v_i$ for all $i \in S$ and $j = 1,\dots,n+1$.
-\end{lemma}
-\begin{proof}
-For $i \in S$ we have $v_i = \max(v)$ and $h_i = 1$ and so the lemma is true if we show that $h_j-1 \geq v_j - \max(v)$ for all $j=1,\dots,n+1$.
-If $v_j = \max(v)$ then $h_j = 1$ and $h_j-1 = 0 = v_j - \max(v)$.
-Otherwise, if $v_j \neq \max(v)$ then
-\[
-h_j-1 \geq -1 \geq v_j - \max(v)
-\] 
-because $h_j \in \{0,1\}$ and $v_j$ and $\max(v)$ are integers with $v_j < \max(v)$.
-\end{proof}
-
-\begin{lemma}\label{lem:vqcontramin}
-Let $t \in \reals^{n+1}$ and $v \in \ints^{n+1}$ be such that $Bv$ is a closest lattice point to $Bt$.
-Let $g \in \{0,1\}^{n+1}$ and let $S$ be the subset of indices $\{1,\dots,n+1\}$ such that $g_i=1$ and $v_i = \min(v)$ for $i \in S$.
-If $\|B(t-v)\|^2 < \|B(t-v-I_S)\|^2$, then
-\[
-\|B(t - g)\|^2 > \min_{p \in \{0,1\}^{n+1}}\|B(t - p)\|^2.
-\]
-\end{lemma}
-\begin{proof}
-From our hypothesis and part~\ref{eq:lem:decSellingsdec} of Lemma~\ref{lem:decSellings} with $p = t-v$ we have
-\[
-\|B(t-v)\|^2 - \|B(t-v-I_S)\|^2 = \sum_{i \in S}\sum_{j \notin S}q_{ij}\big(1 + \gamma_{ij} + 2(v_i - v_j)\big) < 0, 
-\]
-where $\gamma_{ij} = 2(t_j - t_i)$.  Put $h = g - I_S$ and observe that $h_i = 0$ whenever $v_i = \min(v)$ and so, since $S \subset \{i \mid v_i = \min(v) \}$, we have
-\[
-h_i - h_j \geq v_i - v_j \qquad \text{when $i \in S$ and $j=1,\dots,n+1$}
-\]
-by Lemma~\ref{lem:simplevminh}.  Thus, since $q_{ij} \leq 0$,
-\[
-q_{ij}\big(1 + \gamma_{ij} + 2(h_i - h_j)\big) \leq q_{ij}\big(1 + \gamma_{ij} + 2(v_i - v_j)\big)
-\]
-when $i \in S$ and $j =1,\dots,n+1$.  By part~\ref{eq:lem:decSellingsdec} of Lemma~\ref{lem:decSellings} we have
-\begin{align*}
-\|B(t-h)\|^2 - \|B(t-g)\|^2 &= \|B(t-h)\|^2 - \|B(t-h-I_S)\|^2 \\
-&= \sum_{i \in S}\sum_{j \notin S}q_{ij}\big(1 + \gamma_{ij} + 2(h_i - h_j)\big) \\
-&\leq \sum_{i \in S}\sum_{j \notin S}q_{ij}\big(1 + \gamma_{ij} + 2(v_i - v_j)\big) < 0
-\end{align*}
-from which it follows that
-\[
-\|B(t-g)\|^2 > \|B(t-h)\|^2 \geq \min_{p \in \{0,1\}^{n+1}}\|B(t  - p)\|^2
-\]
-as required.
-\end{proof}
-
-\begin{lemma}\label{lem:vqcontramax}
-Let $t \in \reals^{n+1}$ and $v \in \ints^{n+1}$ be such that $Bv$ is a closest lattice point to $Bt$.
-Let $g \in \{0,1\}^{n+1}$ and let $S$ be the subset of indices $\{1,\dots,n+1\}$ such that $g_i=0$ and $v_i = \max(v)$ for $i \in S$.
-If $\|B(t-v)\|^2 < \|B(t-v+I_S)\|^2$, then
-\[
-\|B(t - g)\|^2 > \min_{p \in \{0,1\}^{n+1}}\|B(t - p)\|^2.
-\]
-\end{lemma}
-\begin{proof}
-The proof mimics that of Lemma~\ref{lem:vqcontramin} with only minor modification.
-We give the proof in full for completeness and because the small modifications that do exist might not be obvious.
-From our hypothesis and part~\ref{eq:lem:decSellingsinc} of Lemma~\ref{lem:decSellings} with $p = t-v$ we have
-\[
-\|B(t-v)\|^2 - \|B(t-v+I_S)\|^2 = \sum_{i \in S}\sum_{j \notin S}q_{ij}\big(1 + \gamma_{ji} + 2(v_j - v_i)\big) < 0,
-\]
-where $\gamma_{ij} = 2(t_j - t_i)$.  Put $h = g + I_S$ and observe that $h_i = 1$ when $v_i = \max(v)$ and so, since $S \subset \{i \mid v_i = \max(v) \}$, we have
-\[
-h_j - h_i \geq v_j - v_i \qquad \text{for $i \in S$ and $j=1,\dots,n+1$}
-\]
-by Lemma~\ref{lem:simplevmaxh}.  Thus, since $q_{ij} \leq 0$,
-\[
-q_{ij}\big(1 + \gamma_{ji} + 2(h_j - h_i)\big) \leq q_{ij}\big(1 + \gamma_{ji} + 2(v_j - v_i)\big)
-\]
-when $i \in S$ and $j =1,\dots,n+1$.  By part~\ref{eq:lem:decSellingsdec} of Lemma~\ref{lem:decSellings} we have
-\begin{align*}
-\|B(t-h)\|^2 - \|B(t-g)\|^2 &= \|B(t-h)\|^2 - \|B(t-h+I_S)\|^2 \\
-&= \sum_{i \in S}\sum_{j \notin S}q_{ij}\big(1 + \gamma_{ji} + 2(h_j - h_i)\big) \\
-&\leq \sum_{i \in S}\sum_{j \notin S}q_{ij}\big(1 + \gamma_{ji} + 2(v_j - v_i)\big) < 0
-\end{align*}
-from which it follows that
-\[
-\|B(t-g)\|^2 > \|B(t-h)\|^2 \geq \min_{p \in \{0,1\}^{n+1}}\|B(t  - p)\|^2
-\]
-as required.
-\end{proof}
-
-% \begin{lemma}
-% Let $x \in \ints^{n+1}$.  If there exists a $v \in \ints^{n+1}$ with $\rng(v) = 0$ such that $B(x + v)$ is a closest lattice point to $y = Bz$, then $Bx$ is also a closest lattice point to $y$.  
-% \end{lemma}
-% \begin{proof}
-% Because $\rng(v) = 0$ all of the elements $v_1=v_2=\dots=v_{n+1} = V$ for some integer $V$. Now
-% \begin{align*}
-% \|B(z - x - v)\|^2 &= \| \sum_{i=1}^{n+1} b_i (z_i-x_i-v_i) \|^2 \\
-% &= \| \sum_{i=1}^{n+1} b_i (z_i-x_i) -V\sum_{i=1}^{n+1}b_i \|^2 \\
-% &=  \| \sum_{i=1}^{n+1} b_i (z_i-x_i) \|^2 \\
-% &=  \|B(z-x)\|^2
-% \end{align*}
-% because $\sum_{i=1}^{n+1}b_i = 0$ due to the superbase condition~\eqref{eq:superbasecond}.  So $Bx$ is a closest point to $y$ whenever $B(x+v)$ is.
-% \end{proof}
-
-% \begin{lemma}\label{lem:minrelvecsisonesonh}
-% Let $x \in \ints^{n+1}$ be such that $w_i = x_i + h_i$ are minimisers of~\eqref{eq:tominimise} and such that $h_1,\dots,h_{n+1}$ starts at zero and increments in step of one.
-Let $s$ be the largest integer such that $h_s = 0$ and let
-% \[
-% \alpha = \min_{p \in \{0,1\}^{n+1}} \| B(z - x - p) \|^2.
-% \]
-% The first $s$ elements of the vector $\alpha$ are equal to zero, that is, $\alpha_1,\alpha_2,\dots,\alpha_s = 0$.
-% \end{lemma}
-% \begin{proof}
-% Let $S$ be some nonempty subset of the indices $\{1, 2,\dots,s\}$ and let $\beta$ be the column vector with elements satisfying
-% \[
-% \beta_i = \begin{cases}
-% 1 & i \in S \\
-% \alpha_i & \text{otherwise}.
-% \end{cases}
-% \]
-% Since the subset $S$ is arbitrary, our proof holds if we show that
-% \[
-% A = \| B(z - x - \alpha) \|^2 - \| B(z - x - \beta) \|^2 \leq 0.
-% \]
-
-% Let $g$ be the column of length $n+1$ with elements
-% \[
-% g_i = \begin{cases}
-% 1 & i \in S \\
-% h_i & \text{otherwise}.
-% \end{cases}
-% \]
-% Since $w_i = x_i + h_i$ are minimsers of~\eqref{eq:tominimise} we have
-% \[
-% \| B(z - x - h) \|^2  \leq \| B(z - x - g) \|^2
-% \]
-% and by Selling's formula
-% \begin{equation}\label{eq:sellingswithhg}
-%  - \frac{1}{2} \sum_{i=1}^{n+1}\sum_{j=1}^{n+1}q_{ij}d_{ij}^2 \leq 0,
-% \end{equation}
-% where 
-% \[
-% d_{ij} = d_{ji} = (\gamma_{ij} - h_j + h_i)^2 - (\gamma_{ij} - g_j + g_i)^2,
-% \]
-% and where $\gamma_{ij} = z_i - x_i - z_j + x_i$.  When $i, j \in S$ or $i,j \notin S$ we have $d_{ij} = 0$.  When $i \in S, j \notin S$ we have
-% \[
-% d_{ij} = (\gamma_{ij} - h_j)^2 - (\gamma_{ij} - h_j + 1)^2 = -2(\gamma_{ij} - h_j) - 1.
-% \]
-% and when $i \notin S, j \in S$ we have
-% \[
-% d_{ij} = (\gamma_{ij} + h_i)^2 - (\gamma_{ij} - 1 + h_i)^2 = 2(\gamma_{ij} + h_i) - 1.
-% \]
-% Substituting these values for $d_{ij}$ into~\eqref{eq:sellingswithhg} gives
-% \begin{align*}
-% -\frac{1}{2} \sum_{i \in S, j \notin S} q_{ij}(2h_j - 2\gamma_{ij} - 1) -  &\frac{1}{2} \sum_{i \notin S, j \in S} q_{ij}( 2h_i + 2\gamma_{ij} - 1 ) \\
-% &= -\sum_{i \in S, j \notin S} q_{ij}(2h_j - 2\gamma_{ij} - 1) \leq 0.
-% \end{align*}
-% where equality on the second line hold because $q_{ij}=q_{ji}$ and $\gamma_{ij}=\gamma_{ji}$.  Use of Sellings formula and similar working to previously shows that
-% \[
-% \| B(z - x - \alpha) \|^2 - \| B(z - x - \beta) \|^2 = - \sum_{i \in S, j \notin S} q_{ij}(2\alpha_j - 2\gamma_{ij} - 1) 
-% \]
-% and, since $h_j \geq \alpha_j$ for all $j = 1,\dots,n+1$ and $q_{ij} \leq 0$ for all $i \neq j$ we have 
-% \[
-% q_{ij}(2\alpha_j - 2\gamma_{ij} - 1) \geq q_{ij}(2h_j - 2\gamma_{ij} - 1)
-% \]
-% for all $i \neq j$ and so,
-% \begin{align*}
-% \| B(z - x - \alpha) \|^2 - \| B(z - x - \beta) \|^2 &= - \sum_{i \in S, j \notin S} q_{ij}(2\alpha_j - 2\gamma_{ij} - 1)  \\
-% &\leq - \sum_{i \in S, j \notin S} q_{ij}(2h_j - 2\gamma_{ij} - 1) \leq 0.
-% \end{align*}
-% \end{proof}
-
-% \begin{theorem}
-% Let $\Lambda \subset \reals^m$ be an $n$ dimensional lattice of Voronoi's first kind with obtuse superbasis $b_1,\dots,b_{n+1}$ and let $y \in \reals^m$.
-Define the vectors $x_1,x_2,\dots$ according to~\eqref{eq:xseqfirsttype}.
-There is a positive integer $K \leq (n+1)^2$ such that $x_K$ is a closest lattice point to $y$.
-% \end{theorem}
 

### 2014-01-10 paper.tex
added minor

@@
 In what follows it is assumed that $y$ lies in the space spanned by the basis vectors $b_1,\dots,b_{n}$.
-This assumption is without loss of generality because $x$ is a closest lattice point to $y$ if and only if $x$ is a the closest lattice point to the orthogonalprojection of $y$ into the space spanned by $b_1,\dots,b_{n}$.
+This assumption is without loss of generality because $x$ is a closest lattice point to $y$ if and only if $x$ is a the closest lattice point to the orthogonal projection of $y$ into the space spanned by $b_1,\dots,b_{n}$.
 Let
@@
 where $\{0,1\}^{n+1}$ denotes the set of column vectors of length $n+1$ with elements equal to zero or one.
-Observe that the procedure is initialised at the lattice point $x_0 = Bu_0 = B\floor{z}$.
+The procedure is initialised at the lattice point $x_0 = Bu_0 = B\floor{z}$.
 This choice of initial lattice point is important.
@@
 It is necessary to introduce some notation.
-For $S$ a subset of the indices $\{1,\dots,n+1\}$ let $\onebf_S$ denote the column vector of length $n+1$ with $i$th element equal to one if $i \in S$ and $0$ otherwise.
+For $S$ a subset of indices $\{1,\dots,n+1\}$ let $\onebf_S$ denote the column vector of length $n+1$ with $i$th element equal to one if $i \in S$ and $0$ otherwise.
 %Thus, for example, if $p \in \reals^{n+1}$, then $p+\onebf_S$ denotes the vector $p$ with those elements having index from $S$ incremented by one.
-We define the function
+For $S \subseteq \{1,\dots,n+1\}$ and $p \in \reals^{n+1}$ we define the function
 \[
@@
 \]
-where $q_{ij} = \dotprod{b_i}{b_j}$ are the Selling parameters.  We denote by $\bar{S}$ the complement of the set of indices $S$, that is $\bar{S} = \{ i =1,\dots,n+1 \mid i \notin S\}$.
+where $q_{ij} = \dotprod{b_i}{b_j}$ are the Selling parameters from~\eqref{eq:obtusecond}.
+We denote by $\bar{S}$ the complement of the set of indices $S$, that is $\bar{S} = \{ i \in \{1,\dots,n+1\} \mid i \notin S\}$.
 
@@
 \[
-\rng(p) = \max(p) - \min(p)
+\rng(p) = \max(p) - \min(p).
 \] 
-to return the difference between the maximum and minimum of $p$.  %For example, if $p = (2,-1,4) \in \ints^3$ then $\rng(p) = 4 - (-1)=5$.  We might also 
+%to return the difference between the maximum and minimum of $p$.  %For example, if $p = (2,-1,4) \in \ints^3$ then $\rng(p) = 4 - (-1)=5$.  We might also 
 %For example $\rng(2,-1,4)= 4 - (-1) = 5$.  
@@
 \begin{proof}
-The lemma is trivial if $\subrng(v) = \emptyset$ so that $\decrng(v) = v$.
-It remains to prove the lemma when $\subrng(v) \neq \emptyset$.
-In this case put $S = \subrng(v)$ and put $u = \decrng(v) = v - \onebf_S$.
+The lemma is trivial if $\subrng(v) = \emptyset$ so that $\decrng(v) = v$.  It remains to prove the lemma when $\subrng(v) \neq \emptyset$.  In this case put $S = \subrng(v)$ and put 
+\[
+u = \decrng(v) = v - \onebf_S.
+\] 
 Let $\zeta = z - \floor{z}$ be the column vector containing the fractional parts of the elements of $z$.
@@
 \]
-Substituting this inequality into~\eqref{eq:sumsumBzBu} and using that $q_{ij} \leq 0$ (the obtuse condition~\eqref{eq:obtusecond}) we find that
+Substituting this inequality into~\eqref{eq:sumsumBzBu} and using that $q_{ij} \leq 0$ for $i \neq j$ (the obtuse condition~\eqref{eq:obtusecond}) we find that
 \[
-\|B(z - \floor{z} - v)\|^2 - \|B(z - \floor{z} - u)\|^2 \geq 0
+\|B(z - \floor{z} - v)\|^2 - \|B(z - \floor{z} - u)\|^2 \geq 0.
 \]
-and so $B(\floor{z} + u) = B\big(\floor{z} + \decrng(v)\big)$ is a closest lattice point to $y = Bz$ whenever $B(\floor{z} + v)$ is.
+It follows that $B(\floor{z} + u) = B\big(\floor{z} + \decrng(v)\big)$ is a closest lattice point to $y = Bz$ whenever $B(\floor{z} + v)$ is.
 \end{proof}
@@
 \[
-h_i - h_j \geq v_i - v_j
+h_i - h_j \leq v_i - v_j
 \]
-when either $v_i = \min(v)$ or $v_j = \max(v)$.
+when either $v_i = \max(v)$ or $v_j = \min(v)$.
 \end{lemma}
 \begin{proof}
-If $v_i = \min(v)$ then $h_i = 0$ and we need only show that $-h_j \geq \min(v) - v_j$ for all $j$.
-If $v_j = \min(v)$ then $h_j = 0 = \min(v) - v_j$ and the results holds.
-Otherwise if $v_j > \min(v)$ then $-h_j \geq -1 \geq \min(v) - v_j$ because $\min(v)$ and $v_j$ are integers.
+If $v_i = \max(v)$ then $h_i = 1$ and we need only show that $1-h_j \leq \max(v) - v_j$ for all $j$.
+If $v_j = \max(v)$ then $h_j = 0 = \min(v) - v_j$ and the results holds.
+Otherwise if $v_j < \max(v)$ then $1-h_j \leq 1 \leq \max(v) - v_j$ because $\max(v)$ and $v_j$ are integers.
 
-Now, if $v_j = \max(v)$ then $h_j = 1$ and we need only show that $h_i - 1 \geq v_i - \max(v)$ for all $i$.
-If $v_i = \max(v)$ then $h_i = 1$ and $h_i - 1 = 0 = v_i - \max(v)$.
-Otherwise if $v_i < \max(v)$ then $h_i - 1 \geq -1 \geq v_i - \max(v)$.
+Now, if $v_j = \min(v)$ then $h_j = 0$ and we need only show that $h_i \leq v_i - \min(v)$ for all $i$.
+If $v_i = \min(v)$ then $h_i = 0 = v_i - \min(v)$ and the results holds.
+Otherwise if $v_i > \min(v)$ then $h_i \leq 1 \leq v_i - \min(v)$ because $\min(v)$ and $v_i$ are integers.
 \end{proof}
@@
 
-As stated already $h_i = 0$ whenever $v_i = \min(v)$ and $h_i = 1$ whenever $v_i = \max(v)$.  It follows by Lemma~\ref{eq:integergreaterless} that
+As stated already $h_i = 0$ whenever $v_i = \min(v)$ and $h_i = 1$ whenever $v_i = \max(v)$.  It follows from Lemma~\ref{eq:integergreaterless} that
 \begin{equation}\label{eq:hijvijieq}
-h_{i} - h_j \geq v_{i} - v_j
+h_{i} - h_j \leq v_{i} - v_j
 \end{equation}
-when either $v_i = \min(v)$ or $v_j = \max(v)$.  Since $v_i = \min(v)$ for $i \in T$ and $v_j = \max(v)$ for $j \in S$ the inequality~\eqref{eq:hijvijieq} holds whenever $i \in T$ or $j \in S$.
+when either $v_i = \max(v)$ or $v_j = \min(v)$.  Since $v_i = \max(v)$ for $i \in S$ and $v_j = \min(v)$ for $j \in T$ the inequality~\eqref{eq:hijvijieq} holds whenever $i \in S$ or $j \in T$.
 
-Put $r = z - u - h$.  By~\eqref{eq:hijvijieq} we have $r_i - r_j \leq p_{i} - p_{j}$ when either $i \in T$ or $j \in S$.  Now, since $q_{ij} \leq 0$ for $i \neq j$,
+Put $r = z - u - h$.  By~\eqref{eq:hijvijieq} we have $r_i - r_j \geq p_{i} - p_{j}$ when either $i \in S$ or $j \in T$.  Now, since $q_{ij} \leq 0$ for $i \neq j$,
 \[
-\Phi(S,r) = \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2r_i - 2r_j) \geq \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_i - 2p_j) = \Phi(S,p).
+\Phi(S,r) = \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2r_i - 2r_j) \leq \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_i - 2p_j) = \Phi(S,p).
 \]
-Similarly
+Similarly,
 \[
-\Phi(\bar{T},r) = \sum_{i \notin T}\sum_{j \in T}q_{ij}(1 + 2r_i - 2r_j) \geq \sum_{i \notin T}\sum_{j \in T}q_{ij}(1 + 2p_i - 2p_j) = \Phi(\bar{T},p).
+\Phi(\bar{T},r) = \sum_{i \notin T}\sum_{j \in T}q_{ij}(1 + 2r_i - 2r_j) \leq \sum_{i \notin T}\sum_{j \in T}q_{ij}(1 + 2p_i - 2p_j) = \Phi(\bar{T},p).
 \]\
@@
 &= \Phi(S,r) + \Phi(\bar{T},r) + \sum_{i\in S}\sum_{j \in S} q_{ij} \\
-&\leq \Phi(S,p) + \Phi(\bar{T}, p) + \sum_{i\in S}\sum_{j \in S} q_{ij} < 0. 
+&\leq \Phi(S,p) + \Phi(\bar{T}, p) + \sum_{i\in S}\sum_{j \in S} q_{ij} < 0 
 \end{align*}

### 2014-01-11 paper.tex
added Vaughan's edits.  It's looking pretty nice now!

@@
 The first factor is computing the minimum over the set $\relevant(\Lambda) \cup \{\zerobf\}$ in~\eqref{eq:relvectsminimslicer}.
-In general a lattice can have as many as $2^{n+1}-2$ relevant vectors and so computing a minimiser directly can require a number of operations that grows exponentially with $n$.
-To add to this it is typically the case that the set of relevant vectors $\relevant(\Lambda)$ must be stored in memory and so the algorithm can require an amount of memory that grows exponentially with $n$.
+In general a lattice can have as many as $2^{n+1}-2$ relevant vectors so computing a minimiser directly can require a number of operations that grows exponentially with $n$.
+To add to this it is typically the case that the set of relevant vectors $\relevant(\Lambda)$ must be stored in memory so the algorithm can require an amount of memory that grows exponentially with $n$.
 We will show that for a lattice of Voronoi's first kind the set of relevant vectors has a compact representation in terms of what is called its \emph{obtuse superbasis}.
@@
 The other factor affecting the complexity is the number of iterations required before the algorithm arrives at a closest lattice point, that is, the size of $K$.
-Propositon~\ref{eq:latticepointsinsphere} suggests that this number might be as large as $\ceil{r/\rho}^n$ where $r = \|y - x_0\|^2$ and $\rho$ is the packing radius of the lattice.
+Proposition~\ref{eq:latticepointsinsphere} suggests that this number might be as large as $\ceil{r/\rho}^n$ where $r = \|y - x_0\|^2$ and $\rho$ is the packing radius of the lattice.
 Thus, the number of iterations required might grow exponentially with $n$.
@@
 \end{equation}
-be the matrix with columns given by $b_1,\dots,b_{n+1}$ and let $z \in \reals^{n+1}$ be a column vector such that $y = Bz$.
+be the $n+1$ by $n+1$ matrix with columns given by $b_1,\dots,b_{n+1}$ and let $z \in \reals^{n+1}$ be a column vector such that $y = Bz$.
 We now want to find a column vector $w = (w_1,\dots,w_{n+1})^\prime$ of integers such that
@@
 This choice of initial lattice point is important.
-In Section~\ref{sec:comp-clos-relev} we show how the minimisation over $\{0,1\}^{n+1}$ in~\eqref{eq:pvecmin} can be computed efficiently in $O(n^3)$ operations by computing a minimum cut in an undirected flow network.
+In Section~\ref{sec:comp-clos-relev} we show how minimisation over $\{0,1\}^{n+1}$ in~\eqref{eq:pvecmin} can be computed efficiently in $O(n^3)$ operations by computing a minimum cut in an undirected flow network.
+The minimiser may not be unique corresponding with the existence of multiple minimum cuts.
+In this case any one of the minimisers may be chosen.
+Our results do not depend on this choice.
 In the remainder of this section we prove that this iterative procedure results in a closest lattice point after at most $n$ iterations.
@@
 Put $Q = B^\prime B$ where supercript $^\prime$ indicates the vector or matrix transpose.
+The $n+1$ by $n+1$ matrix $Q$ has elements given by the Selling parameters, that is, $Q_{ij} = q_{ij}=\dotprod{b_i}{b_j}$.
 Denote by $\onebf$ the column vector of length $n+1$ containing all ones.
-We have $B\onebf = \sum_{i=1}^{n+1}b_i = \zerobf$ as a result of the superbasis condition~\eqref{eq:superbasecond} and so $Q\onebf = 0$.
-Observe that $\onebf_S = \onebf - \onebf_{\bar{S}}$ and so $Q\onebf_S = -Q\onebf_{\bar{S}}$.
+Now $B\onebf = \sum_{i=1}^{n+1}b_i = \zerobf$ as a result of the superbasis condition~\eqref{eq:superbasecond} and so $Q\onebf = \zerobf$.
+Since $\onebf_S = \onebf - \onebf_{\bar{S}}$ it follows that $Q\onebf_S = -Q\onebf_{\bar{S}}$.
 With $\circ$ the elementwise vector product, i.e., the Schur or Hadamard product, we have
@@
 &= \onebf^\prime_S Q \onebf_{\bar{S}} - 2 (p\circ\onebf_{\bar{S}})^\prime Q \onebf_S - 2 (p\circ\onebf_S)^\prime Q \onebf_S \\
-&= \onebf^\prime_S Q \onebf_{\bar{S}} - 2 (p\circ\onebf_{\bar{S}})^\prime Q \onebf_{S} + 2 (p\circ\onebf_S)^\prime Q \onebf_{\bar{S}} \\
+&= \onebf^\prime_S Q \onebf_{\bar{S}} - 2 (p\circ\onebf_{\bar{S}})^\prime Q \onebf_{S} + 2 (p\circ\onebf_S)^\prime Q \onebf_{\bar{S}}
  \end{align*}
@@
 because $\subrng(p)$ contains all those indices $i$ such that $p_i = \max(p)$.
-Observe that by repeatedly applying $\decrng$ to a vector one eventually obtains a vector for which further application of $\decrng$ has no effect.
+By repeatedly applying $\decrng$ to a vector one eventually obtains a vector for which further application of $\decrng$ has no effect.
 For example,
@@
 \]  
-It follows that $\subrng(d_T) = \emptyset$ and so $\rng(d_T) \leq n$ by Proposition~\ref{prop:subrrngsmall}.  The proof follows with $v = d_T$.   
+Thus $\subrng(d_T) = \emptyset$ and $\rng(d_T) \leq n$ by Proposition~\ref{prop:subrrngsmall}.  The proof follows with $v = d_T$.   
 \end{proof}
@@
 Let $\ell$ be a nonegative integer.
-We will say that a lattice point $x$ is $\ell$-\emph{close} to $y$ if there exists a $v \in \ints^{n+1}$ with $\rng(v) = \ell$ such that $x + Bv$ is a closest lattice point to $y$.
+We say that a lattice point $x$ is $\ell$-\emph{close} to $y$ if there exists a $v \in \ints^{n+1}$ with $\rng(v) = \ell$ such that $x + Bv$ is a closest lattice point to $y$.
 Lemma~\ref{lem:roundzclose} asserts that the lattice point $x_0 = B\floor{z}$ that initialises the iterative procedure~\eqref{eq:xseqfirsttype} is $K$-close to $y$ where $K \leq n$.
@@
 Because $x$ is $0$-close there exists a $v \in \ints^{n+1}$ with $\rng(v) = 0$ such that $x + Bv$ is a closest lattice point to $y$.
-Because $\rng(v) = 0$ all elements from $v$ are identical.
-That is $v_1=v_2=\dots=v_{n+1}$.
+Because $\rng(v) = 0$ all elements from $v$ are identical, that is, $v_1=v_2=\dots=v_{n+1}$.
 In this case $Bv = \sum_{i=1}^{n+1} v_n b_n = v_1\sum_{i=1}^{n+1}b_n = 0$
@@
 If $v_i = \max(v)$ then $h_i = 1$ and we need only show that $1-h_j \leq \max(v) - v_j$ for all $j$.
-If $v_j = \max(v)$ then $h_j = 0 = \min(v) - v_j$ and the results holds.
+If $v_j = \max(v)$ then $h_j = 1$ and the results holds since $1 - h_j = 0 = \min(v) - v_j$.
 Otherwise if $v_j < \max(v)$ then $1-h_j \leq 1 \leq \max(v) - v_j$ because $\max(v)$ and $v_j$ are integers.
@@
 The proof is by contradiction.
-Suppose that $B(u+g+w)$ is not a closest point to $y$, that is, suppose that
+Suppose $B(u+g+w)$ is not a closest point to $y$, that is, suppose
 \[
@@
 \end{equation}
-when either $v_i = \max(v)$ or $v_j = \min(v)$.  Since $v_i = \max(v)$ for $i \in S$ and $v_j = \min(v)$ for $j \in T$ the inequality~\eqref{eq:hijvijieq} holds whenever $i \in S$ or $j \in T$.
+when either $v_i = \max(v)$ or $v_j = \min(v)$.  Since $v_i = \max(v)$ for $i \in S$ and $v_j = \min(v)$ for $j \in T$ the inequality~\eqref{eq:hijvijieq} holds when either $i \in S$ or $j \in T$.
 
@@
 \[
-\Phi(S,r) = \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2r_i - 2r_j) \leq \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_i - 2p_j) = \Phi(S,p).
+\Phi(S,r) = \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2r_i - 2r_j) \leq \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_i - 2p_j) = \Phi(S,p),
 \]
-Similarly,
+and
 \[
 \Phi(\bar{T},r) = \sum_{i \notin T}\sum_{j \in T}q_{ij}(1 + 2r_i - 2r_j) \leq \sum_{i \notin T}\sum_{j \in T}q_{ij}(1 + 2p_i - 2p_j) = \Phi(\bar{T},p).
-\]\
-Now
+\]
+Using part~\ref{eq:lem:decSellingsincdec} of Lemma~\ref{lem:decSellings} again,
 \begin{align*}
@@
 \end{align*}
-as a result of~\eqref{eq:BpBpdecincineq}.  However, $d \in \{0,1\}^{n+1}$ and so
+as a result of~\eqref{eq:BpBpdecincineq}.  However, $h \in \{0,1\}^{n+1}$ and so this implies
 \[
@@
 \]
-contradicting~\eqref{eq:qismin}.  Thus, $B(u+g+w)$ is a closest lattice point to $y$ and, since $\rng(w) = \ell-1$, the lattice point $B(u+g)$ is $(\ell-1)$-close to $y$.
+contradicting~\eqref{eq:qismin}.
+Thus, or original supposition is false and $B(u+g+w)$ is a closest lattice point to $y$.
+Because $\rng(w) = \ell-1$ the lattice point $B(u+g)$ is $(\ell-1)$-close to $y$.
  \end{proof}
@@
 Specifically, it remains to show that the minimisation over the set of binary vectors $\{0,1\}^{n+1}$ described in~\eqref{eq:pvecmin} can be computed efficiently.
-Putting $t = z - u_k$ in~\eqref{eq:pvecmin} we require an efficient method to compute a $p \in \{0,1\}^{n+1}$ such that the binary quadratic form
+Putting $p = z - u_k$ in~\eqref{eq:pvecmin} we require an efficient method to compute a $t \in \{0,1\}^{n+1}$ such that the binary quadratic form
 \[
-\| B(t - p) \|^2 = \| \sum_{i=1}^{n+1} b_i (t_i - p_i) \|^2
+\| B(p - t) \|^2 = \| \sum_{i=1}^{n+1} b_i (p_i - t_i) \|^2
 \]
@@
 \[
-\| \sum_{i=1}^{n+1} b_i (t_i - p_i) \|^2 =  \sum_{i=1}^{n+1}\sum_{j=1}^{n+1} q_{ij}t_i t_j -  2\sum_{i=1}^{n+1}\sum_{j=1}^{n+1} q_{ij}t_j p_i + \sum_{i=1}^{n+1}\sum_{j=1}^{n+1} q_{ij} p_i p_j.
+\| \sum_{i=1}^{n+1} b_i (p_i - t_i) \|^2 =  \sum_{i=1}^{n+1}\sum_{j=1}^{n+1} q_{ij}p_i p_j -  2\sum_{i=1}^{n+1}\sum_{j=1}^{n+1} q_{ij}p_j t_i + \sum_{i=1}^{n+1}\sum_{j=1}^{n+1} q_{ij} t_i t_j.
 \]
-The first sum above is independent of $p$ and can be ignored for the purpose of minimisation.  Letting $s_i = \sum_{j=1}^{n+1} q_{ij}t_j$, we can equivalently minimise the binary quadratic form
+The first sum above is independent of $t$ and can be ignored for the purpose of minimisation.  Letting $s_i = \sum_{j=1}^{n+1} q_{ij}p_j$, we can equivalently minimise the binary quadratic form
 \begin{equation}\label{eq:quadformnp}
-Q(p) = \sum_{i=1}^{n+1} s_i p_i + \sum_{i=1}^{n+1}\sum_{j=1}^{n+1} q_{ij} p_i p_j.
+Q(t) = \sum_{i=1}^{n+1} s_i t_i + \sum_{i=1}^{n+1}\sum_{j=1}^{n+1} q_{ij} t_i t_j.
 \end{equation}
-We will show that a minimiser of $Q(p)$ can be found efficiently be computing a minimum cut in an undirected flow network.
+We will show that a minimiser of $Q(t)$ can be found efficiently be computing a minimum cut in an undirected flow network.
 %\section{Quadratic $\{0,1\}$ programs and minimium cuts in graphs}
@@
 In what follows we will often drop the argument and write $W$ rather than $W(C,\bar{C})$.
-The \emph{minimum cut} is the $C$ and $\bar{C}$ that minimise the weight $W$.
-If all of the edge weights $w_{ij}$ for $i \neq j$ are nonnegative, the minimum cut can be computed in
+A \emph{minimum cut} is a $C$ and $\bar{C}$ that minimise the weight $W$.
+If all of the edge weights $w_{ij}$ for $i \neq j$ are nonnegative, a minimum cut can be computed in
 order $O(n^3)$ arithmetic operations~\cite{Cormen2001,Even_graph_algorithms_1979}.
@@
 If a constant is added to the weight of this edge, that is, $w_{0,n+2}$ is replaced by $w_{0,n+2} + c$ then $W$ is replaced by $W + c$ for every $C$ and $\bar{C}$.
-In particular, the subsets $C$ and $\bar{C}$ corresponding to the minimum cut are not changed.
+In particular, the subsets $C$ and $\bar{C}$ corresponding to a minimum cut are not changed.
 We will find it convenient to choose $w_{0,n+2} = w_{n+2,0} = 0$.
@@
 If a constant, say $c$, is added to the weights of these edges, that is, $w_{0,i}$ and $w_{i,n+2}$ are replaced by $w_{0,i} + c$ and $w_{i,n+2} + c$, then $W$ is replaced by $W + c$ for every $C$ and $\bar{C}$.
-The $C$ and $\bar{C}$ corresponding the minimum cut are unchanged.
+The $C$ and $\bar{C}$ corresponding to a minimum cut are unchanged.
 In this way, the minimum cut is only affected by the differences
@@
 
-We now show how $W(C,\bar{C})$ can be represented as a binary quadratic form.  Put $p_0 = 1$ and $p_{n+2} = 0$ and
+We now show how $W(C,\bar{C})$ can be represented as a binary quadratic form.  Put $t_0 = 1$ and $t_{n+2} = 0$ and
 \[
-p_i = \begin{cases}
+t_i = \begin{cases}
 1, & i \in C \\
@@
 \[
-p_i(1 - p_j) = \begin{cases}
+t_i(1 - t_j) = \begin{cases}
 1, & i \in C, j \in \bar{C} \\
@@
 \begin{align*}
-W(C,\bar{C}) = \sum_{i \in C} \sum_{j \in \bar{C}} w_{ij} = \sum_{i =0}^{n+2} \sum_{j =0}^{n+2} w_{ij} p_i (1 - p_j) = F(p)
+W(C,\bar{C}) = \sum_{i \in C} \sum_{j \in \bar{C}} w_{ij} = \sum_{i =0}^{n+2} \sum_{j =0}^{n+2} w_{ij} t_i (1 - t_j) = F(t),
 \end{align*}
-say, where $p_0 = 1$ and $p_{n+2} = 0$. Finding a minimum cut is equivalent to finding the binary vector $p = (p_1, \dots, p_{n+1})$ that minimises $F(p)$.  Write,
+say.  Finding a minimum cut is equivalent to finding the binary vector $t = (t_1, \dots, t_{n+1})$ that minimises $F(t)$.  Write,
 \[
-F(p) =  \sum_{i=0}^{n+2} \sum_{j =0}^{n+2} w_{ij}p_i - \sum_{i=0}^{n+2} \sum_{j =0}^{n+2} w_{ij} p_ip_j.
+F(t) =  \sum_{i=0}^{n+2} \sum_{j =0}^{n+2} w_{ij}t_i - \sum_{i=0}^{n+2} \sum_{j =0}^{n+2} w_{ij} t_it_j.
 \]
-Letting $k_i = \sum_{j =0}^{n+2} w_{ij}$, and using that $p_0 = 1$ and $p_{n+2} = 0$,
+Letting $k_i = \sum_{j =0}^{n+2} w_{ij}$, and using that $t_0 = 1$ and $t_{n+2} = 0$,
 \[
-F(p) = \sum_{i=0}^{n+1}k_ip_i  - w_{00} - \sum_{i=1}^{n+1} w_{i0} p_i - \sum_{j=1}^{n+1} w_{0j} p_j - \sum_{i=1}^{n+1} \sum_{j =1}^{n+1} w_{ij} p_ip_j.
+F(t) = \sum_{i=0}^{n+1}k_it_i  - w_{00} - \sum_{i=1}^{n+1} w_{i0} t_i - \sum_{j=1}^{n+1} w_{0j} t_j - \sum_{i=1}^{n+1} \sum_{j =1}^{n+1} w_{ij} t_it_j.
 \]
@@
 \[
-F(p) = k_0 + \sum_{i=1}^{n+1}( k_i  - 2 w_{i0}) p_i  - \sum_{i=1}^{n+1} \sum_{j =1}^{n+1} w_{ij} p_ip_j.
+F(t) = k_0 + \sum_{i=1}^{n+1}( k_i  - 2 w_{i0}) t_i  - \sum_{i=1}^{n+1} \sum_{j =1}^{n+1} w_{ij} t_it_j.
 \]
-The constant term $k_0$ is unimportant for the purpose of minimisation, and so, finding the minimum cut is equivalent to minimising the binary quadratic form
+The constant term $k_0$ is unimportant for the purpose of minimisation so finding a minimum cut is equivalent to minimising the binary quadratic form
 \[
-\sum_{i=1}^{n+1}g_i p_i  - \sum_{i=1}^{n+1} \sum_{j =1}^{n+1} w_{ij} p_ip_j,
+\sum_{i=1}^{n+1}g_i t_i  - \sum_{i=1}^{n+1} \sum_{j =1}^{n+1} w_{ij} t_it_j,
 \]
 where $g_i = k_i  - 2 w_{i0} = d_i + \sum_{j=1}^{n+1} w_{ij}$.
-It only remains to observe the equivalence of this quadratic form and $Q(p)$ from~(\ref{eq:quadformnp}) when the weights are assigned to satisfy,
+It only remains to observe the equivalence of this quadratic form and $Q(t)$ from~\eqref{eq:quadformnp} when the weights are assigned to satisfy,
 \begin{align*}
@@
 With these choices, all the weights $w_{ij}$ for $i \neq j$ are nonnegative.
-A minimiser of $Q(p)$, and correspondingly a solution of~\eqref{eq:pvecmin} can be computed in $O(n^3)$ operations by computing a minimum cut in the undirected flow network $G$ assigned with these nonnegative weights~\cite{Picard_min_cuts_1974,Sankaran_solving_CDMA_mincut_1998,Ulukus_cdma_mincut_1998,Cormen2001}.
+A minimiser of $Q(t)$, and correspondingly a solution of~\eqref{eq:pvecmin} can be computed in $O(n^3)$ operations by computing a minimum cut in the undirected flow network $G$ assigned with these nonnegative weights~\cite{Picard_min_cuts_1974,Sankaran_solving_CDMA_mincut_1998,Ulukus_cdma_mincut_1998,Cormen2001}.
 %The closest lattice point is then computed using~\eqref{eq:compnp}.
@@
 
-A final remark is that our algorithm assumes that the~\emph{obtuse superbasis} is known in advance.
+The root lattices $A_n$ and their dual lattices $A_n^*$ are classical examples of lattices of Voronoi's first kind~\cite{SPLAG}.
+The construction of an `arbitrary' lattice of Voronoi's first kind appears nontrivial.
+One approach would be to construct the matrix $Q = B^\prime B$ containing the Selling parameters $q_{ij}$.
+The matrix $Q$ requires to have nonegative off diagonal entries and to be positive semidefinite with null space containing the vector $\onebf$ of all ones.
+The explicit construction of such matrices is an interesting question for future research.
+A related question is to determine, in some sense, how `large' the set of matrices satisfying these criteria is.
+This would shed some light on how `large' the set of lattices of Voronoi's first kind is.
+
+A final remark is that our algorithm assumes that the obtuse superbasis is known in advance.
 It is known that all lattices of dimension less than 4 are of Voronoi's first kind and an algorithm exists to recover the obtuse superbasis in this case~\cite{SPLAG}.

### 2014-01-11 paper.tex
minor

@@
 One approach would be to construct the matrix $Q = B^\prime B$ containing the Selling parameters $q_{ij}$.
-The matrix $Q$ requires to have nonegative off diagonal entries and to be positive semidefinite with null space containing the vector $\onebf$ of all ones.
+The matrix $Q$ requires to have nonpositive off diagonal entries and to be positive semidefinite with null space containing the vector $\onebf$ of all ones.
 The explicit construction of such matrices is an interesting question for future research.

### 2014-01-13 paper.tex
added stuff

--- 2014-01-11 paper.tex
@@
  \begin{proof}
-The packing radius $\rho$ is the point on the boundary of the Voronoi cell $\vor(\Lambda)$ that is closest to the origin.
+The packing radius $\rho$ is the Euclidean length of a point on the boundary of the Voronoi cell $\vor(\Lambda)$ that is closest to the origin.
 Therefore, the sphere $S$ is a subset of $\vor(\Lambda)$ scaled by $r/\rho$ and translated by $t$.
@@
 Proposition~\ref{obs:2} above asserts that after some finite number $K$ of iterations the procedure arrives at $x_K$, a closest lattice point to $y$.
-Using Proposition~\ref{obs:1} we can detect that $x_K$ is a closest lattice point by asserting that $\|y - x_K\| \leq \| y - x_{K+1} \|$.
+Using Proposition~\ref{obs:1} we can detect that $x_K$ is a closest lattice point by checking whether $\|y - x_K\| \leq \| y - x_{K+1} \|$.
 This simple iterative approach to compute a closest lattice point is related to what is called the \emph{iterative slicer}~\cite{Shalvi_iterativeslicer_2009}.
@@
 \end{theorem}
+
+% A number of important lattices are of Voronoi's first kind.
+For example, the \emph{root lattices} $\ints^n, A_n, E_6, E_7$ and $E_8$ are of Voronoi's first kind~\cite{SPLAG}.
+The root lattice $E_8$ has obtuse superbasis
+% \[
+
+% \]
+
+% The root lattice $D_n$ has obtuse superbasis given by the vectors
+% \begin{align*}
+% b_1 &= -e_1 - e_2 \\
+% b_i &= e_{i} - e_{i+1}, \qquad i = 2,\dots,n \\
+% b_{n+1} &= e_{2} + e_{n}.
+% \end{align*}
+
+Classical examples of lattices of Voronoi's first kind are the $n$ dimensional root lattice $A_n$ and it's dual lattice $A_n^*$.
+These lattices have convenient bases represented in $n+1$ dimensional space.
+An obtuse superbasis for $A_n$ are $n+1$ vectors of length $n+1$ of the form
+\begin{align*}
+b_i &= e_i - e_{i+1}, \qquad i = 1, \dots, n \\
+b_{n+1} &= e_{n+1} - e_{1},
+\end{align*}
+where $e_i$ is a vector of length $n+1$ with elements equal to zero except for the $i$th element that is equal to one.
+The dual lattice $A_{n}^*$ is also of Voronoi's first kind.
+The $i$th vector $b_i$ of the obtuse superbasis is of length $n+1$ with $j$th element equal to
+\[
+b_{ij} = \begin{cases}
+\frac{n}{n+1} & j = i \\
+-\frac{1}{n+1} & j \neq i.
+\end{cases}
+\]
+For $A_n$ and $A_n^*$ there exists efficient algorithms that can compute a closest lattice point in $O(n)$ operations~\cite{McKilliam2009CoxeterLattices,Conway1982FastQuantDec}, and so, we do not recommend using the algorithm we describe here for these specific lattices.
+%However, there exist a large number of lattice of first
+%A lattice of Voronoi's first kind appears nontrivial.
+One approach would be to construct the matrix $Q = B^\prime B$ containing the Selling parameters $q_{ij}$.
+The matrix $Q$ requires to have nonpositive off diagonal entries and to be positive semidefinite with null space containing the vector $\onebf$ of all ones.
+The explicit construction of such matrices is an interesting question for future research.
+A related question is to determine, in some sense, how `large' the set of matrices satisfying these criteria is.
+This would shed some light on how `large' the set of lattices of Voronoi's first kind is.
+
+\section{A series of relevant vectors from a lattice of Voronoi's first kind}
 
@@
 
-The root lattices $A_n$ and their dual lattices $A_n^*$ are classical examples of lattices of Voronoi's first kind~\cite{SPLAG}.
-The construction of an `arbitrary' lattice of Voronoi's first kind appears nontrivial.
-One approach would be to construct the matrix $Q = B^\prime B$ containing the Selling parameters $q_{ij}$.
-The matrix $Q$ requires to have nonegative off diagonal entries and to be positive semidefinite with null space containing the vector $\onebf$ of all ones.
-The explicit construction of such matrices is an interesting question for future research.
-A related question is to determine, in some sense, how `large' the set of matrices satisfying these criteria is.
-This would shed some light on how `large' the set of lattices of Voronoi's first kind is.
-
 A final remark is that our algorithm assumes that the obtuse superbasis is known in advance.

### 2014-01-13 figs/latticefigures.rb

@@ -68,16 +68,16 @@ end
 picsphp = Picture.new('picsphp')
 v = M*Vector[1,0]/2.0
 inradp = Path.new
-alpha = 0.1
+alpha = 0.0
 inradp.add_pair(Pair.new((v[0]*alpha).cm, (v[1]*alpha).cm))
 inradp.add_pair(Pair.new((v[0]*(1-alpha)).cm, (v[1]*(1-alpha)).cm))
-picsphp.add_drawable(Draw.new(inradp))
-pv = Vector[1/v[0], -1/v[1]]*0.025
-lengthtickpair = Path.new
-lengthtickpair.add_pair(Pair.new(pv[0].cm,pv[1].cm))
-lengthtickpair.add_pair(Pair.new((-pv[0]).cm,(-pv[1]).cm))
-picsphp.add_drawable(Draw.new(lengthtickpair).translate((v[0]*alpha).cm, (v[1]*alpha).cm))
-picsphp.add_drawable(Draw.new(lengthtickpair).translate((v[0]*(1-alpha)).cm, (v[1]*(1-alpha)).cm))
+picsphp.add_drawable(Arrow.new(inradp))
+#pv = Vector[1/v[0], -1/v[1]]*0.025
+#lengthtickpair = Path.new
+#lengthtickpair.add_pair(Pair.new(pv[0].cm,pv[1].cm))
+#lengthtickpair.add_pair(Pair.new((-pv[0]).cm,(-pv[1]).cm))
+#picsphp.add_drawable(Draw.new(lengthtickpair).translate((v[0]*alpha).cm, (v[1]*alpha).cm))
+#picsphp.add_drawable(Draw.new(lengthtickpair).translate((v[0]*(1-alpha)).cm, (v[1]*(1-alpha)).cm))
 picsphp.add_drawable(Label.new(latex('$\displaystyle \rho$'), Pair.new((v[0]/2).cm, (v[1]/2).cm)) )
 (-7..7).each do |x|
   (-7..7).each do |y|

### 2014-01-13 paper.tex
added minor.  Needs a read through.

@@
 It is also the radius of the largest sphere that can be placed at every lattice point such that no two spheres intersect (see Figure~\ref{lattices:fig:vorregion}).
-The following well known result will be useful.
+The following well known results will be useful.
 
@@
 
+\begin{proposition}\label{eq:latticepointsinvorcvell}
+Let $\Lambda \subset \reals^{m}$ be an $n$-dimensional lattice.
+For $r\in \reals$ let $\ceil{r}$ denote the smallest integer strictly larger than $r$.
+Let $t \in \reals^m$.
+The number of lattice points inside the scaled and translated Voronoi cell $r \vor(\Lambda) + t$ is atmost $\ceil{r}^n$.
+\end{proposition}
+\begin{proof}
+It is convenient to modify the boundary of the Voronoi cell so that it tesselates $\reals^m$ under translations by $\Lambda$.
+With this aim we let $V \subset \vor(\Lambda)$ contain all those points from the interior of $\vor(\Lambda)$, but have opposing faces open and closed.
+That is, if $x \in V$ is on the boundary of $V$ then $-x \notin V$.
+With this definition it can be asserted that $V$ tesslates $\reals^m$ under translations by $\Lambda$.
+That is, $\reals^{m} = \cup_{x \in \Lambda}(V + x)$ and the intersection $(V + x)\cap(V+y)$ is empty for distinct lattice points $x$ and $y$~\cite[Sec.~2.2]{McKilliam2010thesis}.
+Now, for positive integer $k$, the scaled and translated cell $kV + t$ contain precisely one of each of the coset representatives of the group quotient $\Lambda/k\Lambda$~\cite[Sec.~2.4]{McKilliam2010thesis}.
+There are $k^n$ coset representatives.
+Thus, the number of lattice point inside $r \vor(\Lambda) + t \subset \ceil{r}V + t$ is at most $\ceil{r}^n$.
+\end{proof}
+
 \begin{proposition}\label{eq:latticepointsinsphere}
@@
 That is, $S \subset r/\rho \vor(\Lambda) + t$.
-The proof follows because the number of lattice points in $r/\rho \vor(\Lambda) + t$ is less than $\ceil{r/\rho}^n$ (this follows from, for example, \cite[Theorem~2]{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}, see also~\cite[Lemma~3.7]{Micciancio09adeterministic}).
+The proof follows because the number of lattice points in $r/\rho \vor(\Lambda) + t$ is less than $\ceil{r/\rho}^n$ by Proposition~\ref{eq:latticepointsinvorcvell}.
 \end{proof}
@@
 In general a lattice can have as many as $2^{n+1}-2$ relevant vectors so computing a minimiser directly can require a number of operations that grows exponentially with $n$.
-To add to this it is typically the case that the set of relevant vectors $\relevant(\Lambda)$ must be stored in memory so the algorithm can require an amount of memory that grows exponentially with $n$.
+To add to this it is often the case that the set of relevant vectors $\relevant(\Lambda)$ must be stored in memory so the algorithm can require an amount of memory that grows exponentially with $n$~\cite[Sec.~6]{Micciancio09adeterministic}\cite{Shalvi_iterativeslicer_2009}.
 We will show that for a lattice of Voronoi's first kind the set of relevant vectors has a compact representation in terms of what is called its \emph{obtuse superbasis}.
@@
 The initial lattice point for the iteration is $x_0 = (-4.4,-2.8)$.
-The shaded region is the Voronoi cell translated about the closest lattice point $x_6 = (4.4,2.8)$.}
+The shaded region is the Voronoi cell surrounding the closest lattice point $x_6 = (4.4,2.8)$.}
 		\label{lattices:fig:iterativeexample} 
@@
 
-Classical examples of lattices of Voronoi's first kind are the $n$ dimensional root lattice $A_n$ and it's dual lattice $A_n^*$.
+Classical examples of lattices of Voronoi's first kind are the $n$ dimensional root lattice $A_n$ and its dual lattice $A_n^*$~\cite{SPLAG}.
 These lattices have convenient bases represented in $n+1$ dimensional space.
-An obtuse superbasis for $A_n$ are $n+1$ vectors of length $n+1$ of the form
+An obtuse superbasis for $A_n$ is of the form
 \begin{align*}
@@
 The dual lattice $A_{n}^*$ is also of Voronoi's first kind.
-The $i$th vector $b_i$ of the obtuse superbasis is of length $n+1$ with $j$th element equal to
+The $i$th vector $b_i$ of the obtuse superbasis has $j$th element equal to
 \[
@@
 \]
-For $A_n$ and $A_n^*$ there exists efficient algorithms that can compute a closest lattice point in $O(n)$ operations~\cite{McKilliam2009CoxeterLattices,Conway1982FastQuantDec}, and so, we do not recommend using the algorithm we describe here for these specific lattices.
+For $A_n$ and $A_n^*$ there exists efficient algorithms that can compute a closest lattice point in $O(n)$ operations~\cite{McKilliam2009CoxeterLattices,Conway1982FastQuantDec} so we do not recommend using the algorithm described in this paper for $A_n$ and $A_n^*$.
+The fast algorithms for $A_n$ and $A_n^*$ exploit the special structure of these lattices and are not applicable to other lattices.
+In contrast, the algorithm we describe here works for all lattices of Voronoi's first kind.
+
 %However, there exist a large number of lattice of first
@@
 The best known general purpose algorithms require a number of operations of order $O(2^{2n})$~\cite{Micciancio09adeterministic}.
-
 In this paper we have focused on the class of lattices of Voronoi's first kind.
@@
 Besides being of theoretical interest, the algorithm has potential for practical application.
+
+ It is straightforward to construct an `arbitrary' lattice of Voronoi's first kind.
+One approach is to construct the $n+1$ by $n+1$ symmetric matrix $Q = B^\prime B$ with elements $Q_{ij} = q_{ij} = \dotprod{b_{i}}{b_j}$ given by the Selling parameters.
+Choose the off diagonal entries of $Q$ to be nonpositive with $q_{ij}=q_{ji}$ and set the diagonal elements $q_{ii} = -\sum_{j=1}^{n+1} q_{ij}$.
+%We then have $Q\onebf = \zerobf$ and so $\onebf$ is in the null space of $Q$.
+The matrix $Q$ is diagonally dominant, that is, $\abs{q_{ii}} \geq \sum_{j=1}^{n+1} \abs{q_{ij}}$, and so $Q$ is positive semidefinite.
+A rank deficient Cholesky decomposition~\cite{Higham90analysisof} can now be used to recover a matrix $B$ such that $B^\prime B  = Q$.
+The columns of $B$ are vectors of the obtuse superbasis.
+
 A question of immediate interest is: do there exist lattices of Voronoi's first kind that produce good codes or good quantisers?

### 2014-01-14 paper.tex
minor
added more edits.  Looking good

@@
 Section~\ref{sec:latt-voron-first} describes lattices of Voronoi's first kind and their obtuse superbasis.
-It is shown that for these lattices, the series of relevant vectors results in a closest lattice point after at most $n$ terms.
+In Section~\ref{sec:seri-relev-vect} it is shown that for these lattices, the series of relevant vectors results in a closest lattice point after at most $n$ terms.
 Section~\ref{sec:comp-clos-relev} shows that each relevant vector in the series can be computed in $O(n^3)$ operations by computing a minimum cut in an undirected flow network.
@@
 Let $t \in \reals^m$.
-The number of lattice points inside the scaled and translated Voronoi cell $r \vor(\Lambda) + t$ is atmost $\ceil{r}^n$.
+The number of lattice points inside the scaled and translated Voronoi cell $r \vor(\Lambda) + t$ is at most $\ceil{r}^n$.
 \end{proposition}
 \begin{proof}
-It is convenient to modify the boundary of the Voronoi cell so that it tesselates $\reals^m$ under translations by $\Lambda$.
+It is convenient to modify the boundary of the Voronoi cell so that it tessellates $\reals^m$ under translations by $\Lambda$.
 With this aim we let $V \subset \vor(\Lambda)$ contain all those points from the interior of $\vor(\Lambda)$, but have opposing faces open and closed.
 That is, if $x \in V$ is on the boundary of $V$ then $-x \notin V$.
-With this definition it can be asserted that $V$ tesslates $\reals^m$ under translations by $\Lambda$.
+With this definition it can be asserted that $V$ tessellates $\reals^m$ under translations by $\Lambda$.
 That is, $\reals^{m} = \cup_{x \in \Lambda}(V + x)$ and the intersection $(V + x)\cap(V+y)$ is empty for distinct lattice points $x$ and $y$~\cite[Sec.~2.2]{McKilliam2010thesis}.
-Now, for positive integer $k$, the scaled and translated cell $kV + t$ contain precisely one of each of the coset representatives of the group quotient $\Lambda/k\Lambda$~\cite[Sec.~2.4]{McKilliam2010thesis}.
+Now, for positive integer $k$, the scaled and translated cell $kV + t$ contains precisely one coset representative for each element of the quotient group $\Lambda/k\Lambda$~\cite[Sec.~2.4]{McKilliam2010thesis}.
 There are $k^n$ coset representatives.
-Thus, the number of lattice point inside $r \vor(\Lambda) + t \subset \ceil{r}V + t$ is at most $\ceil{r}^n$.
+Thus, the number of lattice points inside $r \vor(\Lambda) + t \subset \ceil{r}V + t$ is at most $\ceil{r}^n$.
 \end{proof}
@@
 Let $S$ be an $m$-dimensional hypersphere of radius $r$ centered at $t \in \reals^{m}$.
-The number of lattice points from $\Lambda$ in the sphere $S$ is less than $\ceil{r/\rho}^{n}$ where $\ceil{\cdot}$ denotes the smallest integer strictly larger than its argument.
+The number of lattice points from $\Lambda$ in the sphere $S$ is at most $\ceil{r/\rho}^{n}$ where $\ceil{\cdot}$ denotes the smallest integer strictly larger than its argument.
  \end{proposition}
@@
 That is, $S \subset r/\rho \vor(\Lambda) + t$.
-The proof follows because the number of lattice points in $r/\rho \vor(\Lambda) + t$ is less than $\ceil{r/\rho}^n$ by Proposition~\ref{eq:latticepointsinvorcvell}.
+The proof follows because the number of lattice points in $r/\rho \vor(\Lambda) + t$ is at most $\ceil{r/\rho}^n$ by Proposition~\ref{eq:latticepointsinvorcvell}.
 \end{proof}
@@
 Classical examples of lattices of Voronoi's first kind are the $n$ dimensional root lattice $A_n$ and its dual lattice $A_n^*$~\cite{SPLAG}.
-These lattices have convenient bases represented in $n+1$ dimensional space.
+% These lattices have convenient bases represented in $n+1$ dimensional space.
 An obtuse superbasis for $A_n$ is of the form
-\begin{align*}
-b_i &= e_i - e_{i+1}, \qquad i = 1, \dots, n \\
-b_{n+1} &= e_{n+1} - e_{1},
-\end{align*}
-where $e_i$ is a vector of length $n+1$ with elements equal to zero except for the $i$th element that is equal to one.
+% \begin{align*}
+% b_i &= e_i - e_{i+1}, \qquad i = 1, \dots, n \\
+% b_{n+1} &= e_{n+1} - e_{1},
+% \end{align*}
+% where $e_i$ is a vector of length $n+1$ with elements equal to zero except for the $i$th element that is equal to one.
 The dual lattice $A_{n}^*$ is also of Voronoi's first kind.
 The $i$th vector $b_i$ of the obtuse superbasis has $j$th element equal to
-\[
-b_{ij} = \begin{cases}
-\frac{n}{n+1} & j = i \\
--\frac{1}{n+1} & j \neq i.
-\end{cases}
-\]
-For $A_n$ and $A_n^*$ there exists efficient algorithms that can compute a closest lattice point in $O(n)$ operations~\cite{McKilliam2009CoxeterLattices,Conway1982FastQuantDec} so we do not recommend using the algorithm described in this paper for $A_n$ and $A_n^*$.
-The fast algorithms for $A_n$ and $A_n^*$ exploit the special structure of these lattices and are not applicable to other lattices.
+% \[
+% b_{ij} = \begin{cases}
+% \frac{n}{n+1} & j = i \\
+% -\frac{1}{n+1} & j \neq i.
+% \end{cases}
+% \]
+For $A_n$ and $A_n^*$ there exist efficient algorithms that can compute a closest lattice point in $O(n)$ operations~\cite{McKilliam2009CoxeterLattices,Conway1982FastQuantDec}.
+For this reason we do not recommend using the algorithm described in this paper for $A_n$ and $A_n^*$.
+The fast algorithms for $A_n$ and $A_n^*$ rely of the special structure of these lattices and are not applicable to other lattices.
 In contrast, the algorithm we describe here works for all lattices of Voronoi's first kind.
+Questions that arise are: how ``large'' (in some sense) is the set of lattices of Voronoi's first kind?
+Are there lattices of Voronoi's first kind that are useful in applications such as coding, quantisation, or signal processing?
+We discuss these questions in Section~\ref{sec:discussion}.
+We now focus on the problem of computing a closest lattice point in a lattice of Voronoi's first kind.
 
@@
 
-\section{A series of relevant vectors from a lattice of Voronoi's first kind}
+\section{A series of relevant vectors from a lattice of Voronoi's first kind}\label{sec:seri-relev-vect}
 
@@
 We give a proof for part~\ref{eq:lem:decSellingsinc}.  The proof for part~\ref{eq:lem:decSellingsdec} is similar.  
-Put $Q = B^\prime B$ where supercript $^\prime$ indicates the vector or matrix transpose.
+Put $Q = B^\prime B$ where superscript $^\prime$ indicates the vector or matrix transpose.
 The $n+1$ by $n+1$ matrix $Q$ has elements given by the Selling parameters, that is, $Q_{ij} = q_{ij}=\dotprod{b_i}{b_j}$.
@@
 contradicting~\eqref{eq:qismin}.
-Thus, or original supposition is false and $B(u+g+w)$ is a closest lattice point to $y$.
+Thus, our original supposition is false and $B(u+g+w)$ is a closest lattice point to $y$.
 Because $\rng(w) = \ell-1$ the lattice point $B(u+g)$ is $(\ell-1)$-close to $y$.
@@
 
+A question of immediate interest to communications engineers is: do there exist lattices of Voronoi's first kind that produce good codes or good quantisers?
+Since lattices that produce good codes and quantisers often also describe dense~\emph{sphere packings}~\cite{SPLAG}, a related question is: do there exist lattices of Voronoi's first kind that produce dense sphere packings?
+These question do not appear to have trivial answers.
+The questions have heightened importance due to the algorithm described in this paper.
+
 It is straightforward to construct an `arbitrary' lattice of Voronoi's first kind.
@@
 
-A question of immediate interest is: do there exist lattices of Voronoi's first kind that produce good codes or good quantisers?
-Since lattices that produce good codes and quantisers often also describe dense~\emph{sphere packings}~\cite{SPLAG}, a related question is: do there exist lattices of Voronoi's first kind that produce dense sphere packings?
-These question do not appear to have trivial answers.
-The questions have heightened importance due to the algorithm described in this paper.
+A number applications such as phase unwrapping~\cite{Teunissen_GPS_1995,Hassibi_GPS_1998}, single frequency estimation~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}, and related signal processing problems~\cite{McKilliam2007,Clarkson2007,McKilliam2009IndentifiabliltyAliasingPolyphase,Quinn_sparse_noisy_SSP_2012} also require computing a closest lattice point.
+In these applications the particular lattice arises from the signal processing problem under consideration.
+If that lattice happens to be of Voronoi's first kind then our algorithm can be used.
+An example where this occurs is the problem of computing the \emph{sample intrinsic mean} in circular statistics~\cite{McKilliam_mean_dir_est_sq_arc_length2010}.
+In this particular problem the lattice $A_n^*$ is involved.
+A fast closest point algorithm requiring only $O(n)$ operations exists for $A_n^*$~\cite{McKilliam2009CoxeterLattices,McKilliam2008b} and so the algorithm described in this paper is not needed in this particular case.
+However, there many exist other signal processing problems where lattices of Voronoi's first kind arise.
 

### 2014-01-15 paper.tex
minor

@@
 It is necessary to introduce some notation.
-For $S$ a subset of indices $\{1,\dots,n+1\}$ let $\onebf_S$ denote the column vector of length $n+1$ with $i$th element equal to one if $i \in S$ and $0$ otherwise.
+For $S$ a subset of indices $\{1,\dots,n+1\}$ let $\onebf_S$ denote the column vector of length $n+1$ with $i$th element equal to one if $i \in S$ and zero otherwise.
 %Thus, for example, if $p \in \reals^{n+1}$, then $p+\onebf_S$ denotes the vector $p$ with those elements having index from $S$ incremented by one.

### 2014-03-18 paper.tex
minor

--- 2014-01-13 paper.tex
@@
 Let $S$ be an $m$-dimensional hypersphere of radius $r$ centered at $t \in \reals^{m}$.
-The number of lattice points from $\Lambda$ in the sphere $S$ is less than $\ceil{r/\rho}^{n}$ where $\ceil{\cdot}$ denotes the smallest integer strictly larger than its argument.
+The number of lattice points from $\Lambda$ in the sphere $S$ is less than $\ceil{r/\rho}^{n}$.% where $\ceil{\cdot}$ denotes the smallest integer strictly larger than its argument.
  \end{proposition}
@@
 Classical examples of lattices of Voronoi's first kind are the $n$ dimensional root lattice $A_n$ and its dual lattice $A_n^*$~\cite{SPLAG}.  
-These lattices have convenient bases represented in $n+1$ dimensional space.
-An obtuse superbasis for $A_n$ is of the form
-\begin{align*}
-b_i &= e_i - e_{i+1}, \qquad i = 1, \dots, n \\
-b_{n+1} &= e_{n+1} - e_{1},
-\end{align*}
-where $e_i$ is a vector of length $n+1$ with elements equal to zero except for the $i$th element that is equal to one.
+% These lattices have convenient bases represented in $n+1$ dimensional space.  An obtuse superbasis for $A_n$ is of the form
+% \begin{align*}
+% b_i &= e_i - e_{i+1}, \qquad i = 1, \dots, n \\
+% b_{n+1} &= e_{n+1} - e_{1},
+% \end{align*}
+% where $e_i$ is a vector of length $n+1$ with elements equal to zero except for the $i$th element that is equal to one.
 The dual lattice $A_{n}^*$ is also of Voronoi's first kind.
 The $i$th vector $b_i$ of the obtuse superbasis has $j$th element equal to
-\[
-b_{ij} = \begin{cases}
-\frac{n}{n+1} & j = i \\
--\frac{1}{n+1} & j \neq i.
-\end{cases}
-\]
+% \[
+% b_{ij} = \begin{cases}
+% \frac{n}{n+1} & j = i \\
+% -\frac{1}{n+1} & j \neq i.
+% \end{cases}
+% \]
 For $A_n$ and $A_n^*$ there exists efficient algorithms that can compute a closest lattice point in $O(n)$ operations~\cite{McKilliam2009CoxeterLattices,Conway1982FastQuantDec} so we do not recommend using the algorithm described in this paper for $A_n$ and $A_n^*$.
@@
 Since lattices that produce good codes and quantisers often also describe dense~\emph{sphere packings}~\cite{SPLAG}, a related question is: do there exist lattices of Voronoi's first kind that produce dense sphere packings?
-These question do not appear to have trivial answers.
+These questions do not appear to have trivial answers.
 The questions have heightened importance due to the algorithm described in this paper.

### 2014-03-18 paper.tex

--- 2014-01-15 paper.tex
@@
 Let $S$ be an $m$-dimensional hypersphere of radius $r$ centered at $t \in \reals^{m}$.
-The number of lattice points from $\Lambda$ in the sphere $S$ is at most $\ceil{r/\rho}^{n}$ where $\ceil{\cdot}$ denotes the smallest integer strictly larger than its argument.
+The number of lattice points from $\Lambda$ in the sphere $S$ is at most $\ceil{r/\rho}^{n}$.
+%where $\ceil{\cdot}$ denotes the smallest integer strictly larger than its argument.
  \end{proposition}

### 2014-04-15 talk/presentation.tex
added start of talk

\documentclass{beamer}

\input{presentationdefs}

\input{../../../bib/mathbf-abbrevs.sty}

\usepackage{ulem}

\usetheme{default}

\setbeamertemplate{footline}[frame number]
\setbeamertemplate{navigation symbols}{}

%\usepackage{beamerthemesplit}

%\usetheme[width=2.8cm, hideothersubsections]{Goettingen}
%\usetheme[width=2.8cm]{Goettingen}

\setbeamercovered{transparent=20}

\title{Finding short vectors in a lattice of Voronoi's first kind}
\author{Robby McKilliam and Alex Grant \\ 
\vspace{0.5cm}
Institute for Telecommunications Research \\ University of South Australia
}

\date{\today}

\begin{document}

\frame{\titlepage}

\frame{\tableofcontents}

\section{Lattices and short vectors}

\frame{
\frametitle{Lattices and short vectors}
An $n$-dimensional \term{lattice} $\Lambda$ is a discrete set of vectors from $\reals^m$, $m \geq n$, given by
\[
  \Lambda = \{ x = b_1 u_1 + b_2u_2 + \dots + b_n u_n \mid u_1, \dots u_n \in \ints \},
\] 
where $b_1, \dots, b_n \in \reals^m$ are \term{basis vectors} of $\Lambda$.
}

\frame{
\begin{figure}[tp]
	\centering 
		\includegraphics[width=\linewidth]{preslatticefigures-1.mps}
		\caption{A 2-dimensional lattice.}
		\label{fig:latfig1}
\end{figure}
}

\frame{
\frametitle{Lattices and short vectors}
Those points in the lattice with smallest non-zero length are called \term{short vectors}.  That is, the short vectors have squared length
\[
\min_{x\in \Lambda \backslash \{ 0 \} } \| x \|^2.
\]
}

\frame{
\begin{figure}[tp]
	\centering 
		\includegraphics[width=\linewidth]{preslatticefigures-1.mps}
		\caption{A 2-dimensional lattice.}
		\label{fig:latfig2}
\end{figure}
}

\frame{
\begin{figure}[tp]
	\centering 
		\includegraphics[width=\linewidth]{preslatticefigures-3.mps}
		\caption{A 2-dimensional lattice.  There are 4 short vectors.}
		\label{fig:latfigandpacking1}
\end{figure}
}

\frame{
\frametitle{Lattices and short vectors}
\begin{itemize}
\item Computing a short vector is called the \term{shortest vector problem}.
\item Applications in cryptography and number theory.
\item NP-hard for arbitrary lattices.
\item Easier for specific lattices.
\item For example, short vectors are easy to find in the \term{root lattices} $\ints^n$, $A_n$ and $D_n$.
\item We will show that the problem is relatively easy to solve for lattices of \term{Voronoi's first kind}.
\end{itemize}
}

\section{Lattices of Voronoi's first kind}
\frame{
\frametitle{Lattices of Voronoi's first kind}
An $n$-dimensional lattice $\Lambda$ is of \term{Voronoi's first kind} if it has an \term{obtuse superbase}, that is, a set of $n+1$ vectors 
\[
b_1,\dots,b_{n+1}
\] 
such that
\begin{itemize}
\item $b_1,\dots,b_n$ are a basis for $\Lambda$,
\item $b_1 + b_2 \dots + b_{n+1} = 0$ \hspace{0.2cm} (the superbase condition),
\item $q_{ij} = b_i \cdot b_j \leq 0$ whenever $i \neq j$ \hspace{0.2cm} (the obtuse condition).
\end{itemize}
\vspace{0.2cm}
The $q_{ij}$ are called \term{Selling parameters}.
%\begin{itemize}
%\item Conway and Sloane have shown that every lattice of dimensional less than 3 is of Vornoi's first kind.
%\end{itemize}
}

\frame{
\frametitle{An example}
Consider the 3-dimensional lattice with basis
\[
\begin{split}
b_1 &= \left[ \begin{array}{rrr} 2 & -1 & 0 \end{array}\right] \\
b_2 &= \left[ \begin{array}{rrr}  -1 & 2 & 0  \end{array}\right] \\
b_3 &= \left[ \begin{array}{rrr}  0 & 0 & 2   \end{array}\right].
\end{split}
\]
Define a $4$th vector as
\[
b_4 = -b_1 - b_2 - b_3 =  \left[ \begin{array}{rrr} -1 &  -1  & -2 \end{array}\right],
\]
so that $b_1, b_2, b_3, b_4$ satisfy the superbase condition.
}

\frame{
\frametitle{An example}
The Selling parameters can be written in a matrix
\[
\left[ \begin{array}{cccc} 
q_{11} & q_{12} & q_{13} & q_{14}\\
q_{21} & q_{22} & q_{23} & q_{24} \\
q_{31} & q_{32} & q_{33} & q_{34} \\
q_{41} & q_{42} & q_{43} & q_{44}
 \end{array}\right]
=
\left[ \begin{array}{rrrr} 
5 & -4 & 0 & -1 \\
-4 & 5 & 0 & -1 \\
0 & 0 & 4 & -4 \\
-1 & -1 & -4 & 6
 \end{array}\right].
\]
The off diagonal elements are not positive so the obtuse condition is satisfied.
%\begin{itemize}
%\item All lattices of dimension less than 4 are of Vornoi's first kind.
%\end{itemize}
}

% \frame{
% \frametitle{Example: The root lattice $A_n$}
% \begin{itemize}
% \item The root lattice $A_n$ is of Voronoi's first kind.
% \item An obtuse superbase is all cyclic shifts of the vector 
% \[
% \left[ \begin{array}{rrrrrrr} 1 & -1 & 0 & 0 & \cdots & 0 \end{array}\right]
% \] 
% from $\reals^{n+1}$, that is,
% \begin{align*}
% b_1 &= \left[ \begin{array}{rrrrrrr} 1 & -1 & 0 & 0 & \cdots & 0 \end{array}\right] \\
% b_2 &= \left[ \begin{array}{rrrrrrr} 0 & 1 & -1 & 0 & \cdots & 0 \end{array}\right] \\
% &\vdots \\
% b_{n+1} &= \left[ \begin{array}{rrrrrrr} -1 & 0 & 0 & \cdots & 0 & 1 \end{array}\right].
% \end{align*}
% \item The Selling parameters are 
% \[
% q_{ij} = \dotprod{b_i}{b_j} = \begin{cases} 2, & i = j \\
% -1, & i - j \equiv 1 \bmod{n+1} \\
% 0, & \text{otherwise}.
% \end{cases}
% \]
% \end{itemize}
% }

\frame{
\frametitle{Lattices of Voronoi's first kind}
\begin{theorem}
Let $\Lambda$ be a $n$-dimensional lattice of Voronoi's first kind with obtuse superbase 
\[
b_1, \dots, b_{n+1}.
\]  
The short vectors in $\Lambda$ are of the form 
\[
\sum_{i \in I} b_i
\] 
where $I$ is a strict subset of $\{1, 2, \dots, n+1\}$ and $I$ is not empty.
\end{theorem}
}

\frame{
\frametitle{Lattices of Voronoi's first kind}
A na\"{i}ve way to compute a short vector is to compute
\[
\| \sum_{i \in I} b_i \|^2
\]
for all of the $2^{n+1} - 2$ possible subsets $I$.  
\begin{itemize}
\item Requires a number of operations that grows exponentially with the dimension $n$.
\item We can improve this using a \term{minimum cut algorithm}.
\end{itemize}
}

\section{Graphs, cuts, and minimum cuts}
\frame{
\frametitle{Graphs, cuts, and minimum cuts}
Let $G$ be a weighted graph with:
\begin{itemize}
\item $n+1$ vertices $v_1, \dots, v_{n+1}$,  
\item edges $e_{ij}$ connecting vertex $v_i$ to vertex $v_j$,
\item edge weights $w_{ij} \in \reals$.
\end{itemize} 
}

\frame{
\begin{figure}[tp]
	\centering 
        \includegraphics{graphs-3.mps}
		\caption{A graph with 4 vertices and 4 weighted edges.}
		\label{fig:graphexample}
\end{figure}
}

\frame{
\frametitle{Graphs, cuts, and minimum cuts}
A \term{cut} in $G$ is a partition of the vertices into two nonempty sets $C$ and its complement $\bar{C}$.
\begin{itemize}
\item The \term{weight} of a cut is the sum of the weights on the edges crossing from the vertices in $C$ to the vertices in $\bar{C}$.%, that is,
%\[
%W(C,\bar{C}) = \sum_{i \in I} \sum_{j \in J} w_{ij}, 
%\]
%where $I = \{ i \mid v_i \in C\}$ and $J = \{j \mid v_j \in \bar{C}\}$.
%\item \term{Loops} and zero weight edges don't affect the weight of any cut, so we ignore them.
\item A \term{minimum cut} is a pair ($C$,$\bar{C}$) with smallest weight. %that minimise $W(C,\bar{C})$.  
\end{itemize} 
}

\frame{
\begin{figure}[tp]
	\centering 
        \includegraphics{graphs-4.mps}
		\caption{A graph with 4 vertices and 4 weighted edges.}
		\label{fig:graphexample2}
\end{figure}
}

\frame{
\begin{figure}[tp]
	\centering 
        \includegraphics{graphs-5.mps} 
		\caption{The cut $C = \{v_2\}$ and $\bar{C} = \{v_1,v_3,v_4 \}$}
		\label{fig:graphcut1}
\end{figure}
}

\frame{
\begin{figure}[tp]
	\centering 
        \includegraphics{graphs-6.mps}
		\caption{The cut $C = \{v_2\}$ and $\bar{C} = \{v_1,v_3,v_4 \}$ has weight 5.}
		\label{fig:graphcut2}
\end{figure}
}

\frame{
\begin{figure}[tp]
	\centering 
        \includegraphics{graphs-7.mps}
		\caption{The minimum cut $C = \{v_3,v_4\}$ and $\bar{C} = \{v_1,v_2 \}$}
		\label{fig:graphcut3}
\end{figure}
}

\frame{
\begin{figure}[tp]
	\centering 
        \includegraphics{graphs-8.mps}
		\caption{The minimum cut $C = \{v_3,v_4\}$ and $\bar{C} = \{v_1,v_2 \}$ has weight 2.}
		\label{fig:graphcut4}
\end{figure}
}

\frame{
\frametitle{Graphs, cuts, and minimum cuts}
If the edge weights $w_{ij}$ are all nonnegative, a minimum cut can be computed:
\begin{itemize}
\item deterministically in $O(n^3)$ operations using the algorithm of Stoer and Wagner,
\item with high probability in $O(n^2 \log(n)^3)$ operations using the randomised algorithm of Karger and Stien.
\end{itemize}	
}

\frame{
\begin{theorem}
Let $\Lambda$ be a $n$-dimensional lattice of Voronoi's first kind with obtuse superbase 
\[
b_1, \dots, b_{n+1}.
\]  
Let $G$ be a graph with $n+1$ vertices $v_{1}, \dots, v_{n+1}$ and edge weights 
\[
w_{ij} = -q_{ij} = -b_i \cdot b_j \geq 0 \qquad i \neq j.
\]  
Let $(C, \bar{C})$ be a minimum cut in $G$.  A short vector in $\Lambda$ is 
\[
\sum_{i \in I} b_i \;\;\; \text{where} \;\;\; I = \{ i \mid v_i \in C\}.
\]  
The squared length of the short vector is given by the weight of the minimum cut. %$W(C,\bar{C})$.
\end{theorem}
}

\section{Examples}

% \frame{
% \frametitle{Example: The root lattice $A_n$}
% \begin{itemize} 
% \item Recall the root lattice $A_n$ of Voronoi's first kind.
% \item The Selling parameters are
% \[ 
% q_{ij} = \dotprod{b_i}{b_j} = \begin{cases} 2, & i = j \\
% -1, & i - j \equiv 1 \bmod{n+1} \\
% 0, & \text{otherwise}.
% \end{cases}
% \]  
% \item The corresponding graph has positive edge weights
% \[
% w_{ij} = 1 \qquad i - j \equiv 1 \bmod{n+1}.
% \]
% \item This is the cycle graph with $n+1$ vertices and weight 1 on every edge.
% \end{itemize}
% }

% \frame{
% \begin{figure}[tp]
% 	\centering 
%         \includegraphics{cyclegraph-1.mps}
% 		\caption{The cycle graph corresponding to $A_n$.}
% 		\label{fig:graphcycleAn}  
% \end{figure}
% }

% \frame{
% \begin{figure}[tp]
% 	\centering 
%         \includegraphics{cyclegraph-2.mps}
% 		\caption{Cut corresponding with short vector $b_1 = [1,-1,0,\dots,0]$}
% 		\label{fig:graphcycleAn}  
% \end{figure}
% }

% \frame{
% \begin{figure}[tp]
% 	\centering 
%         \includegraphics{cyclegraph-3.mps} 
% 		\caption{Cut corresponding with $b_1+b_{2}+b_3 = [1,0,0,-1\dots,0]$}
% 		\label{fig:graphcycleAn}  
% \end{figure}
% }

\frame{
\frametitle{An example}
Consider again the 3-dimensional lattice with obtuse superbase
\[
\begin{split}
b_1 &= \left[ \begin{array}{rrr} 2 & -1 & 0 \end{array}\right] \\
b_2 &= \left[ \begin{array}{rrr}  -1 & 2 & 0  \end{array}\right] \\
b_3 &= \left[ \begin{array}{rrr}  0 & 0 & 2   \end{array}\right] \\
b_4 &= \left[ \begin{array}{rrr} -1 &  -1  & -2 \end{array}\right].
\end{split} 
\]
The Selling parameters are given in matrix form as
\[
\left[ \begin{array}{cccc} 
q_{11} & q_{12} & q_{13} & q_{14}\\
q_{21} & q_{22} & q_{23} & q_{24} \\
q_{31} & q_{32} & q_{33} & q_{34} \\
q_{41} & q_{42} & q_{43} & q_{44}
 \end{array}\right]
=
\left[ \begin{array}{rrrr} 
5 & -4 & 0 & -1 \\
-4 & 5 & 0 & -1 \\
0 & 0 & 4 & -4 \\
-1 & -1 & -4 & 6
 \end{array}\right].
\]
}

\frame{
\begin{figure}[tp]
	\centering 
        \includegraphics{graphs-4.mps}
		\caption{We have seen this graph before!}
		%\label{fig:graphcut4}
\end{figure}
}

\frame{
\begin{figure}[tp]
	\centering 
        \includegraphics{graphs-8.mps}
		\caption{The minimum cut $C = \{v_3,v_4\}$ and $\bar{C} = \{v_1,v_2 \}$ has weight 2.}
		%\label{fig:graphcut4}
\end{figure}
}

\frame{
\frametitle{Another example}
The minimum cut corresponds with the short vectors
\[
b_1 + b_2 = [1,1,0]
\]
and 
\[
b_3 + b_4 = -b_1 - b_2 = [-1,-1,0]
\]
of squared length 2.
}

\section{What's next}
\frame{
\frametitle{What's next}

\begin{itemize}
\item Is it possible to efficiently decide whether a lattice is of Voronoi's first kind?
\item Is it possible to efficiently find an obtuse superbase if it exists?
\item Can a similar approach be taken to solve the \term{nearest lattice point problem}?
\end{itemize}

}

% \frame{
% \frametitle{The nearest point problem}
% \begin{itemize}
% \item Given some vector $y \in \reals^m$ we wish to find a lattice point $x \in \Lambda$ such that the squared Euclidean distance
% \[
% \| y - x \|^2
% \] 
% is minimised.  
% \item Equivalently we can minimise
% \[
% \| y - \sum_{i=1}^{n+1} b_i u_i \|^2
% \]
% over integers $u_1,\dots,u_{n+1}$.
% \end{itemize}
% }

% \frame{
% \begin{theorem}
% Let $z \in \reals^{n+1}$ be a column vector such that $y = \sum_{i=1}^{n+1}b_i z_i$.  A nearest point $x \in \Lambda$ to $y$ is of the form
% \[
% x = \sum_{i=1}^{n+1} b_i (\floor{z_i} + u_i)
% \]
% where the $u_i \in \{-1, 0, 1\}$ and $\floor{z_i}$ denotes the largest integer smaller that $z_i$.
% \end{theorem}

% %So for a lattice of Voronoi's first kind, the nearest point problem can be turned into a quadratic minimisation problem over vectors in $u_i \in \{-1,0,1\}$ rather that $u_i \in \ints$.

% }

\end{document}

### 2014-04-15 talk/figs/preslatticefigures.rb

@@ -1,0 +1,273 @@
+require 'rubygems'
+require 'rubypost'
+
+require 'matrix'
+require 'mathn'
+
+include RubyPost
+
+file = RubyPost::File.new('preslatticefigures')
+
+#create an axis peicture
+axes = Picture.new('axes')
+
+xaxis = Path.new
+xaxis.add_pair(Pair.new(-3.95.cm, 0))
+xaxis.add_pair(Pair.new(3.95.cm, 0))
+yaxis = Path.new
+yaxis.add_pair(Pair.new(0, -3.95.cm))
+yaxis.add_pair(Pair.new(0, 3.95.cm))
+
+#Axes.add_drawable(DoubleArrow.new(xaxis))
+#axes.add_drawable(DoubleArrow.new(yaxis))
+#axes.add_drawable(Label.new(latex("$y$"), Pair.new(0,3.95.cm)).bottom_right)
+#axes.add_drawable(Label.new(latex("$x$"), Pair.new(3.95.cm, 0)).top_left)
+
+
+
+pic = Picture.new('picL')
+pic.add_drawable(Draw.new(axes))
+
+M = Matrix[ [1,0.2], [0.2, 1] ]
+(-7..7).each do |x|
+	(-7..7).each do |y|
+		v = M*Vector[x,y]
+		pic.add_drawable(Fill.new(Circle.new()).scale(0.08.cm).translate(v[0].cm,v[1].cm).colour(0.7,0.0,0.0))
+	end
+end
+
+fL = Picture.new('fL')
+v = M*Vector[1,0]
+cyc = Path.new
+cyc.add_pair(Pair.new(0,0))
+cyc.add_pair(Pair.new(v[0].cm,v[1].cm))
+v2 = M*Vector[0,1]
+cyc.add_pair(Pair.new((v[0]+v2[0]).cm,(v[1]+v2[1]).cm))
+cyc.add_pair(Pair.new(v2[0].cm,v2[1].cm))
+cyc.add_pair('cycle')
+fL.add_drawable(Fill.new(cyc).colour(0.8,0.8,0.8))
+v = M*Vector[1,0]
+vec = Path.new
+vec.add_pair(Pair.new(0,0))
+vec.add_pair(Pair.new(v[0].cm,v[1].cm))
+fL.add_drawable(Arrow.new(vec))
+v = M*Vector[0,1]
+vec = Path.new
+vec.add_pair(Pair.new(0,0))
[…]

### 2014-04-15 talk/presentation.tex
added to presentation

@@
 
-\input{../../../bib/mathbf-abbrevs.sty}
+\input{../mathbf-abbrevs.sty}
 
@@
 
-\title{Finding short vectors in a lattice of Voronoi's first kind}
-\author{Robby McKilliam and Alex Grant \\ 
+\title{Finding shortest and closest vectors in a lattice of Voronoi's first kind}
+\author{Robby McKilliam, Alex Grant, and I. Vaughan L. Clarkson \\ 
 \vspace{0.5cm}
+{\small
 Institute for Telecommunications Research \\ University of South Australia 
+\\ \vspace{0.2cm}
+School of Information Technology and Electrical Engineering \\ The University of Queensland
+}
 }
@@
 
-\section{Lattices and short vectors}
+\section{Lattices}
 
 \frame{
-\frametitle{Lattices and short vectors}
+\frametitle{Lattices}
 An $n$-dimensional \term{lattice} $\Lambda$ is a discrete set of vectors from $\reals^m$, $m \geq n$, given by
 \[
-  \Lambda = \{ x = b_1 u_1 + b_2u_2 + \dots + b_n u_n \mid u_1, \dots u_n \in \ints \},
+  \Lambda = \{ b_1 u_1 + b_2u_2 + \dots + b_n u_n \mid u_1, \dots u_n \in \ints \},
 \] 
@@
 	\centering 
-		\includegraphics[width=\linewidth]{preslatticefigures-1.mps}
+		\includegraphics[width=\linewidth]{figs/preslatticefigures-1.mps}
 		\caption{A 2-dimensional lattice.}
@@
 \frame{
-\frametitle{Lattices and short vectors}
-Those points in the lattice with smallest non-zero length are called \term{short vectors}.  That is, the short vectors have squared length
+\frametitle{Short vectors}
+Those lattice points with smallest non-zero length are called \term{short vectors}.  That is, the short vectors have squared length
 \[
@@
 	\centering 
-		\includegraphics[width=\linewidth]{preslatticefigures-3.mps}
+		\includegraphics[width=\linewidth]{figs/preslatticefigures-3.mps}
 		\caption{A 2-dimensional lattice.  There are 4 short vectors.}
@@
 \frame{
-\frametitle{Lattices and short vectors}
+\frametitle{The Voronoi cell}
+}
+
+\frame{
+\frametitle{The shortest vector problem}
 \begin{itemize}
@@
 	\centering 
-        \includegraphics{graphs-3.mps}
+        \includegraphics{figs/graphs-3.mps}
 		\caption{A graph with 4 vertices and 4 weighted edges.}
@@
 	\centering 
-        \includegraphics{graphs-4.mps}
+        \includegraphics{figs/graphs-4.mps}
 		\caption{A graph with 4 vertices and 4 weighted edges.}
@@
 	\centering 
-        \includegraphics{graphs-5.mps} 
+        \includegraphics{figs/graphs-5.mps} 
 		\caption{The cut $C = \{v_2\}$ and $\bar{C} = \{v_1,v_3,v_4 \}$}
@@
 	\centering 
-        \includegraphics{graphs-6.mps}
+        \includegraphics{figs/graphs-6.mps}
 		\caption{The cut $C = \{v_2\}$ and $\bar{C} = \{v_1,v_3,v_4 \}$ has weight 5.}
@@
 	\centering 
-        \includegraphics{graphs-7.mps}
+        \includegraphics{figs/graphs-7.mps}
 		\caption{The minimum cut $C = \{v_3,v_4\}$ and $\bar{C} = \{v_1,v_2 \}$}
@@
 	\centering 
-        \includegraphics{graphs-8.mps}
+        \includegraphics{figs/graphs-8.mps}
 		\caption{The minimum cut $C = \{v_3,v_4\}$ and $\bar{C} = \{v_1,v_2 \}$ has weight 2.}
@@
 	\centering 
-        \includegraphics{graphs-4.mps}
+        \includegraphics{figs/graphs-4.mps}
 		\caption{We have seen this graph before!}
@@
 \frame{
-\frametitle{Another example}
+\frametitle{An example}
 The minimum cut corresponds with the short vectors
@@
 
-\section{What's next}
 \frame{
-\frametitle{What's next}
+\frametitle{Some questions we asked in 2012}
 
 \begin{itemize}
-\item Is it possible to efficiently decide whether a lattice is of Voronoi's first kind?
-\item Is it possible to efficiently find an obtuse superbase if it exists?
+\item Can we efficiently decide whether a lattice is of Voronoi's first kind?
+\item Can we efficiently find an obtuse superbase if it exists?
 \item Can a similar approach be taken to solve the \term{nearest lattice point problem}?
@@
 
-% \frame{
-% \frametitle{The nearest point problem}
-% \begin{itemize}
-% \item Given some vector $y \in \reals^m$ we wish to find a lattice point $x \in \Lambda$ such that the squared Euclidean distance
-% \[
-% \| y - x \|^2
-% \] 
-% is minimised.  
-% \item Equivalently we can minimise
-% \[
-% \| y - \sum_{i=1}^{n+1} b_i u_i \|^2
-% \]
-% over integers $u_1,\dots,u_{n+1}$.
-% \end{itemize}
-% }
+\frame{
+\frametitle{Some questions we asked in 2012}
 
-% \frame{
-% \begin{theorem}
-% Let $z \in \reals^{n+1}$ be a column vector such that $y = \sum_{i=1}^{n+1}b_i z_i$.  A nearest point $x \in \Lambda$ to $y$ is of the form
-% \[
-% x = \sum_{i=1}^{n+1} b_i (\floor{z_i} + u_i)
-% \]
-% where the $u_i \in \{-1, 0, 1\}$ and $\floor{z_i}$ denotes the largest integer smaller that $z_i$.
-% \end{theorem}
+\begin{itemize}
+\item Can we \sout{efficiently} decide whether a lattice is of Voronoi's first kind? {\color{red}\textbf{Yes}}
+\item Can we \sout{efficiently} find an obtuse superbase if it exists? {\color{red}\textbf{Yes}}
+\item Can a similar approach be taken to solve the \term{closest lattice point problem}? {\color{red}\textbf{$O(n^4)$}}
+\end{itemize}
 
-% %So for a lattice of Voronoi's first kind, the nearest point problem can be turned into a quadratic minimisation problem over vectors in $u_i \in \{-1,0,1\}$ rather that $u_i \in \ints$.
+}
 
-% }
+\frame{
+START TALKING ABOUT CLOSEST POINT ALGORITHM
+}
+
+\section{What now?}
+\frame{
+\frametitle{What now?}
+
+\begin{itemize}
+\item Can good codes or quantisers be constructed from lattices of Voronoi's first kind?
+\item Signal processing applications, phase unwrapping, circular statistics, etc.
+\end{itemize}
+
+}
 

### 2014-04-15 talk/figs/preslatticefigures.rb

@@ -270,4 +270,4 @@ fig.add_drawable(Draw.new(pic))
 file.add_figure(fig)
 
 file.compile
-file.view
+#file.view

### 2014-04-16 talk/presentation.tex
added minor

@@
 \frame{
+PICTURE OF VORONOI CELL
+}
+
+\frame{
+\frametitle{Relevant vectors}
+}
+
+\frame{
+PICTURE OF RELEVANT VECTORS
+}
+
+\frame{
+\frametitle{}
+}
+
+\frame{
 \frametitle{The shortest vector problem}
@@
 \end{itemize}
+}
+
+\frame{
+\frametitle{The closest lattice point problem}
+ADD THIS
 }
@@
 \begin{theorem}
-Let $\Lambda$ be a $n$-dimensional lattice of Voronoi's first kind with obtuse superbase 
-\[
-b_1, \dots, b_{n+1}.
-\]  
-The short vectors in $\Lambda$ are of the form 
+Let $\Lambda$ be a $n$-dimensional lattice of Voronoi's first kind with obtuse superbase $b_1, \dots, b_{n+1}$.  The relevant vectors in $\Lambda$ are of the form 
 \[
@@
 \end{theorem}
+
+\begin{corollary}
+Short vectors in $\Lambda$ are of the form $\sum_{i \in I} b_i$.
+\end{corollary}
 }
@@
 \item Can we \sout{efficiently} find an obtuse superbase if it exists? {\color{red}\textbf{Yes}}
+\item Can a similar approach be taken to solve the \term{closest lattice point problem}? 
+\end{itemize}
+
+}
+
+\frame{
+\frametitle{Some questions we asked in 2012}
+
+\begin{itemize}
+\item Can we \sout{efficiently} decide whether a lattice is of Voronoi's first kind? {\color{red}\textbf{Yes}}
+\item Can we \sout{efficiently} find an obtuse superbase if it exists? {\color{red}\textbf{Yes}}
 \item Can a similar approach be taken to solve the \term{closest lattice point problem}? {\color{red}\textbf{$O(n^4)$}}

### 2014-04-16 talk/build.sh

@@ -1,0 +1,7 @@
+cd figs
+ruby preslatticefigures.rb
+mpost graphs.mp
+cyclegraph.mp
+cd ..
+
+pdflatex presentation.tex

### 2014-04-21 talk/build.sh
added lattice figures

@@ -1,4 +1,5 @@
 cd figs
+ruby latticefigures.rb
 ruby preslatticefigures.rb
 mpost graphs.mp
 cyclegraph.mp

### 2014-04-21 talk/figs/latticefigures.rb

@@ -1,0 +1,49 @@
+require 'rubypost'
+
+require 'matrix'
+require 'mathn'
+
+include RubyPost
+
+#a picture containing lattice points zoomed out
+scale = 0.42
+pic = Picture.new('picL')
+M = Matrix[ [3.0,0.6], [0.6, 3.0] ]
+(-7..7).each do |x|
+  (-7..7).each do |y|
+    v = M*Vector[x,y]*scale
+    pic.add_drawable(Fill.new(Circle.new()).scale(0.1.cm).translate(v[0].cm,v[1].cm).colour(0.7,0,0))
+  end
+end
+
+#clippath for all figures.  Fits nicely on beamer screen
+clippath = "((-5.6cm,-3.83cm)--(5.6cm,-3.83cm)--(5.6cm,3.83cm)--(-5.6cm,3.83cm)--cycle)"
+
+#figure with just lattice points
+file = RubyPost::File.new('lattice')
+fig = Figure.new
+fig.add_drawable(Clip.new(clippath, pic))
+fig.add_drawable(Draw.new(pic))
+file.add_figure(fig)
+file.compile
+
+#a picture showing the sphere packing
+sphere_diameter = (M*Vector[1,0]).r * scale
+picsphp = Picture.new('picsphp')
+(-7..7).each do |x|
+  (-7..7).each do |y|
+    v = M*Vector[x,y]*scale
+    picsphp.add_drawable(Draw.new(Circle.new()).scale(sphere_diameter.cm).translate(v[0].cm,v[1].cm))
+  end
+end
+
+#figure with lattice points and sphere packing
+file = RubyPost::File.new('latticewithspherepacking')
+fig = Figure.new
+fig.add_drawable(Clip.new(clippath, pic))
+fig.add_drawable(Draw.new(pic))
+fig.add_drawable(Clip.new(clippath, picsphp))
+fig.add_drawable(Draw.new(picsphp))
+file.add_figure(fig)
+file.compile
+

### 2014-04-21 talk/figs/preslatticefigures.rb

@@ -23,8 +23,6 @@ yaxis.add_pair(Pair.new(0, 3.95.cm))
 #axes.add_drawable(Label.new(latex("$y$"), Pair.new(0,3.95.cm)).bottom_right)
 #axes.add_drawable(Label.new(latex("$x$"), Pair.new(3.95.cm, 0)).top_left)
 
-
-
 pic = Picture.new('picL')
 pic.add_drawable(Draw.new(axes))
 
@@ -157,7 +155,7 @@ end
 #this is a figure that shows a sphere packing and the inradius
 fig3 = Figure.new
 #fig3.add_drawable(Clip.new(Square.new, vorPic).scale(2.cm))
-fig.add_drawable(Clip.new("(-4.05cm,-3.1cm)--(4.05cm,-3.1cm)--(4.05cm,3.1cm)--(-4.05cm,3.1cm)--cycle", picsphp))
+fig3.add_drawable(Clip.new("(-4.05cm,-3.1cm)--(4.05cm,-3.1cm)--(4.05cm,3.1cm)--(-4.05cm,3.1cm)--cycle", picsphp))
 #fig3.add_drawable(Clip.new(Square.new, piccovering).scale(7.93.cm))
 #fig3.add_drawable(Draw.new(vorPic).scale(3))
 fig3.add_drawable(Draw.new(picsphp))
@@ -172,13 +170,14 @@ picrel = Picture.new('picrel')
 picrel.add_drawable(Draw.new(axes))
 (-4..4).each do |x|
 	(-4..4).each do |y|
-		v = M*Vector[x,y]*2
+		v = M*Vector[x,y]*1.6
 		picrel.add_drawable(Fill.new(Circle.new()).scale(0.05.cm).translate(v[0].cm,v[1].cm).colour(0.8,0.1,0.1))
 	end
 end
 
 #this is a figure that shows a sphere packing and the inradius
 fig4 = Figure.new
+fig4.add_drawable(Clip.new("(-4.5cm,-3.1cm)--(4.5cm,-3.1cm)--(4.5cm,3.1cm)--(-4.5cm,3.1cm)--cycle", picrel))
 fig4.add_drawable(Clip.new(Square.new, vorPic).scale(2.cm))
 fig4.add_drawable(Clip.new(Square.new, picrel).scale(7.93.cm))
 fig4.add_drawable(Draw.new(vorPic).scale(2.0))
[…]

### 2014-04-21 paper.tex
minor

@@
 \newcommand{\calR}{\mathcal{R}}
-The (closed) \term{Voronoi cell}, denoted $\vor(\Lambda)$, of a lattice $\Lambda$ in $\reals^m$ is the subset of $\reals^m$ containing all points nearer or of equal distance (here with respect to the Euclidean norm) to the lattice point at the origin than to any other lattice point.
+The (closed) \term{Voronoi cell}, denoted $\vor(\Lambda)$, of a lattice $\Lambda$ in $\reals^m$ is the subset of $\reals^m$ containing all points closer or of equal distance (here with respect to the Euclidean norm) to the lattice point at the origin than to any other lattice point.
 The Voronoi cell is an $m$-dimensional convex polytope that is symmetric about the origin.

### 2014-04-21 talk/presentation.tex
added lattice figures
minor

@@
 	\centering 
-		\includegraphics[width=\linewidth]{figs/preslatticefigures-1.mps}
+		\includegraphics{figs/lattice-1.mps}
 		\caption{A 2-dimensional lattice.}
@@
 	\centering 
-		\includegraphics[width=\linewidth]{figs/preslatticefigures-3.mps}
+		\includegraphics{figs/latticewithspherepacking-1.mps}
 		\caption{A 2-dimensional lattice.  There are 4 short vectors.}
@@
 \end{figure}
-}
-
-\frame{
-\frametitle{The Voronoi cell}
-}
-
-\frame{
-PICTURE OF VORONOI CELL
-}
-
-\frame{
-\frametitle{Relevant vectors}
-}
-
-\frame{
-PICTURE OF RELEVANT VECTORS
-}
-
-\frame{
-\frametitle{}
 }
@@
 \frame{ 
-\frametitle{The closest lattice point problem}
-ADD THIS
+\frametitle{The Voronoi cell}
+
+The \term{Voronoi cell} of a lattice $\Lambda \subset \reals^m$ is the subset of $\reals^m$ with points at least as close to the origin than to any lattice point,
+\[
+\vor(\Lambda) = \{ x \in \reals^m \mid \|x\| \leq \|x - y\|, y \in \Lambda \}.
+\]
+
+}
+
+\frame{ 
+\begin{figure}[tp]
+	\centering 
+		\includegraphics{figs/voronoicell-1.mps}
+		\caption{A 2-dimensional lattice and its Voronoi cell.}
+		\label{fig:latfigandpacking1} 
+\end{figure}
+}
+
+\frame{ 
+\frametitle{Relevant vectors}
+
+The \term{relevant vectors} of a lattice $\Lambda$ are those which contribute a face to the Voronoi cell.
+
+}
+
+\frame{
+\begin{figure}[tp] 
+	\centering 
+		\includegraphics{figs/relevantvectors-1.mps}  
+		\caption{A 2-dimensional lattice with 6 relevant vectors.}
+		\label{fig:latfigandpacking1}
+\end{figure}
+}
+
+\frame{ 
+\frametitle{Relevant vectors}
+\begin{itemize}
+\item The set of relevant vectors $\relevant(\Lambda)$ contains those $v \in \Lambda$ such that  
+\[
+\dotprod{v}{x} \leq \dotprod{x}{x}
+\]
+for all $x \in \Lambda$ with $x \neq v$ and $x \neq \zerobf$.  
+\item The Voronoi cell can be defined using the relevant vectors, 
+\[
+\vor(\Lambda) = \{ x \in\reals^m \mid \|x\| \leq \|x - v\|, v \in\relevant(\Lambda) \}
+\]
+\item Short vectors are relevant vectors.
+\end{itemize}
+}
+
+\frame{
+\begin{figure}[tp] 
+	\centering 
+		\includegraphics{figs/relevantvectorsandspherepacking-1.mps}   
+		\caption{A lattice with 6 relevant vectors and 4 short vectors.}
+		\label{fig:latfigandpacking1}
+\end{figure}
+}
+
+\frame{
+\frametitle{Closest lattice points}
+
+Given a lattice $\Lambda \subset \reals^m$ and a vector $y \in \reals^m$, a problem of interest is to find $x \in \Lambda$ such that
+\[
+\| y - x \|^2
+\] 
+is minimised.
+
+\begin{itemize}
+\item This is called the \term{closest lattice point problem} and a solution is called a \term{closest lattice point} to $y$
+\end{itemize}
+
 }

### 2014-04-21 talk/figs/latticefigures.rb

@@ -8,31 +8,31 @@ include RubyPost
 #a picture containing lattice points zoomed out
 scale = 0.42
 pic = Picture.new('picL')
-M = Matrix[ [3.0,0.6], [0.6, 3.0] ]
+M = Matrix[ [3.0,0.6], [0.6, 3.0] ]*scale
 (-7..7).each do |x|
   (-7..7).each do |y|
-    v = M*Vector[x,y]*scale
+    v = M*Vector[x,y]
     pic.add_drawable(Fill.new(Circle.new()).scale(0.1.cm).translate(v[0].cm,v[1].cm).colour(0.7,0,0))
   end
 end
 
 #clippath for all figures.  Fits nicely on beamer screen
-clippath = "((-5.6cm,-3.83cm)--(5.6cm,-3.83cm)--(5.6cm,3.83cm)--(-5.6cm,3.83cm)--cycle)"
+$clippath = "((-5.6cm,-3.83cm)--(5.6cm,-3.83cm)--(5.6cm,3.83cm)--(-5.6cm,3.83cm)--cycle)"
 
 #figure with just lattice points
 file = RubyPost::File.new('lattice')
 fig = Figure.new
-fig.add_drawable(Clip.new(clippath, pic))
+fig.add_drawable(Clip.new($clippath, pic))
 fig.add_drawable(Draw.new(pic))
 file.add_figure(fig)
 file.compile
 
 #a picture showing the sphere packing
-sphere_diameter = (M*Vector[1,0]).r * scale
+sphere_diameter = (M*Vector[1,0]).r
 picsphp = Picture.new('picsphp')
 (-7..7).each do |x|
   (-7..7).each do |y|
-    v = M*Vector[x,y]*scale
+    v = M*Vector[x,y]
     picsphp.add_drawable(Draw.new(Circle.new()).scale(sphere_diameter.cm).translate(v[0].cm,v[1].cm))
   end
 end
@@ -40,10 +40,89 @@ end
 #figure with lattice points and sphere packing
 file = RubyPost::File.new('latticewithspherepacking')
 fig = Figure.new
-fig.add_drawable(Clip.new(clippath, pic))
+fig.add_drawable(Clip.new($clippath, pic))
 fig.add_drawable(Draw.new(pic))
-fig.add_drawable(Clip.new(clippath, picsphp))
+fig.add_drawable(Clip.new($clippath, picsphp))
 fig.add_drawable(Draw.new(picsphp))
 file.add_figure(fig)
 file.compile
 
[…]

### 2014-04-21 talk/figs/latticefigureszoomed.rb

@@ -1,0 +1,69 @@
+require 'rubypost'
+
+require 'matrix'
+require 'mathn'
+
+include RubyPost
+
+#a picture containing lattice points
+scale = 1.0
+pic = Picture.new('picL')
+M = Matrix[ [3.0,0.6], [0.6, 3.0] ]
+(-7..7).each do |x|
+  (-7..7).each do |y|
+    v = M*Vector[x,y]*scale
+    pic.add_drawable(Fill.new(Circle.new()).scale(0.1.cm).translate(v[0].cm,v[1].cm).colour(0.7,0,0))
+  end
+end
+
+#clippath for all figures.  Fits nicely on beamer screen
+clippath = "((-5.6cm,-3.83cm)--(5.6cm,-3.83cm)--(5.6cm,3.83cm)--(-5.6cm,3.83cm)--cycle)"
+
+#a picture showing the sphere packing
+sphere_diameter = (M*Vector[1,0]).r * scale
+picsphp = Picture.new('picsphp')
+(-7..7).each do |x|
+  (-7..7).each do |y|
+    v = M*Vector[x,y]*scale
+    picsphp.add_drawable(Draw.new(Circle.new()).scale(sphere_diameter.cm).translate(v[0].cm,v[1].cm))
+  end
+end
+
+#matrix with columns given by vector from obtuse superbasis
+B = Matrix[ 
+           [3.0, -2.4, -0.6], 
+           [0.6, 2.4, -3.0] 
+]
+
+#draw a picture of a bigshaded square that we will clip to the Voronoi region
+fig = Figure.new
+vorPic = Picture.new('vorregion')
+vorPic.add_drawable(Fill.new(Square.new).scale(15.cm).colour(0.7,0.7,0.7))
+#relevant vector given by the appropriate combinations of the obtuse superbasis
+(0..1).each do |x|
+  (0..1).each do |y|
+    (0..1).each do |z|
+      if( ((x+y+z)!=0) && ((x+y+z)!=3) )
+        v = (B*Vector[x,y,z])*0.5
+        pv = Vector[1/v[0], -1/v[1]]*30
+        clipp = Path.new
+        clipp.add_pair(Pair.new((pv+v)[0].cm, (pv+v)[1].cm))
+        clipp.add_pair(Pair.new((pv*(-1)+v)[0].cm, (pv*(-1)+v)[1].cm))
[…]

### 2014-04-21 paper.tex
minor

@@
 The closest lattice point problem and the shortest vector problem have interested mathematicians and computer scientists due to their relationship with integer programming~\cite{Lenstra_integerprogramming1983,Kannan1987_fast_general_np,Babai1986}, the factoring of polynomials~\cite{Lenstra1982}, and cryptanalysis~\cite{Joux_toolbox_cryptanal1998,NyguyenStern_two_faces_crypto,Micciancio_lattice_based_post_quantum_crypto}.
-Solutions of the closest lattice point problem also have engineering applications.
+Solutions of the closest lattice point problem have engineering applications.
 For example, if a lattice is used as a vector quantiser then the closest lattice point corresponds to the minimum distortion point~\cite{Conway1983VoronoiCodes,Conway1982VoronoiRegions,Conway1982FastQuantDec}.
@@
 With this definition it can be asserted that $V$ tessellates $\reals^m$ under translations by $\Lambda$.
-That is, $\reals^{m} = \cup_{x \in \Lambda}(V + x)$ and the intersection $(V + x)\cap(V+y)$ is empty for distinct lattice points $x$ and $y$~\cite[Sec.~2.2]{McKilliam2010thesis}.
+That is, $\reals^{m} = \cup_{x \in \Lambda}(V + x)$ and the intersection $(V + x)\cap(V+y)$ is empty for distinct lattice points $x$ and $y$.
 Now, for positive integer $k$, the scaled and translated cell $kV + t$ contains precisely one coset representative for each element of the quotient group $\Lambda/k\Lambda$~\cite[Sec.~2.4]{McKilliam2010thesis}.
@@
 Let $\Lambda$ be a lattice in $\reals^m$ and let $y \in \reals^m$.
-A simple method to compute the lattice point $x \in \Lambda$ closest to $y$ is as follows.
+A simple method to compute a lattice point $x \in \Lambda$ closest to $y$ is as follows.
 Let $x_0$ be some lattice point from $\Lambda$, for example the origin.
 Consider the following iteration,
-\begin{align}\
+\begin{align}
 x_{k+1} &= x_k + v_k \nonumber \\

### 2014-04-21 talk/presentation.tex
minor

@@
 \title{Finding shortest and closest vectors in a lattice of Voronoi's first kind}
-\author{Robby McKilliam, Alex Grant, and I. Vaughan L. Clarkson \\ 
-\vspace{0.5cm}
+\author{Robby McKilliam \\
+\vspace{0.2cm}
+Alex Grant \\ Vaughan Clarkson \\
+\vspace{0.7cm}
 {\small
@@
 		\caption{A 2-dimensional lattice.}
-		\label{fig:latfig1}
+		%\label{fig:latfig1}
 \end{figure}
@@
 		\caption{A 2-dimensional lattice.}
-		\label{fig:latfig2}
+		%\label{fig:latfig2}
 \end{figure}
@@
 		\caption{A 2-dimensional lattice.  There are 4 short vectors.}
-		\label{fig:latfigandpacking1}
+		%\label{fig:latfigandpacking1}
 \end{figure}
@@
 \item Easier for specific lattices.
-\item For example, short vectors are easy to find in the \term{root lattices} $\ints^n$, $A_n$ and $D_n$.
+\item For example, short vectors are easy to find in the \term{root lattices} $\ints^n$, $A_n$, and $D_n$.
 \item We will show that the problem is relatively easy to solve for lattices of \term{Voronoi's first kind}.
@@
 		\caption{A 2-dimensional lattice and its Voronoi cell.}
-		\label{fig:latfigandpacking1} 
+		%\label{fig:latfigandpacking1} 
 \end{figure}
@@
 		\caption{A 2-dimensional lattice with 6 relevant vectors.}
-		\label{fig:latfigandpacking1}
+		%\label{fig:latfigandpacking1}
 \end{figure}
@@
 		\caption{A lattice with 6 relevant vectors and 4 short vectors.}
-		\label{fig:latfigandpacking1}
+		%\label{fig:latfigandpacking1}
 \end{figure}
@@
 \frame{
-\frametitle{Closest lattice points}
+\frametitle{The closest lattice point problem}
 
@@
 \begin{itemize}
-\item This is called the \term{closest lattice point problem} and a solution is called a \term{closest lattice point} to $y$
+\item This is called the \term{closest lattice point problem} and a solution is called a \term{closest lattice point} to $y$.
 \end{itemize}
 
+}
+
+\frame{
+\begin{figure}[tp]  
+	\centering 
+		\includegraphics{figs/closestpoint-1.mps}   
+		\caption{The closest lattice point.}
+		%\label{fig:latfigandpacking1} 
+\end{figure}
+}
+
+\frame{
+\frametitle{The closest lattice point problem}
+\begin{itemize}
+\item The lattice point $x \in \Lambda$ is closest to $y \in \reals^m$ if and only if
+\[
+y \in \vor(\Lambda) + x.
+\]
+\item Equivalently $x$ is closest to $y$ if and only if
+\[
+\dotprod{(y - x)}{v} \leq \tfrac{1}{2} \dotprod{v}{v}
+\]
+for all $v \in \relevant(\Lambda)$.
+\end{itemize}
+}
+
+\frame{
+\frametitle{The closest lattice point problem}
+Applications to:
+\begin{itemize}
+\item coding and quantisation,
+\item multi-antenna communications (MIMO),
+\item unwrapping of phase data for electronic distance measurement in GPS and surveying,
+\item single frequency estimation,
+\item polynomial phase estimation,
+\item circular statistics.
+\end{itemize}
+}
+
+\frame{
+\frametitle{The closest lattice point problem}
+\begin{itemize}
+\item NP-hard for arbitrary lattices.
+\item Easier for specific lattices.
+\item For example, fast algorithms exist for the \term{root lattices} $\ints^n$, $A_n$, and $D_n$.
+\item We will describe a fast algorithm to compute a closest point in lattices of \term{Voronoi's first kind}.
+\end{itemize}
 }
@@
 
-\section{Examples}
+%\section{Examples}
 
@@
 \item Can we efficiently find an obtuse superbase if it exists?
-\item Can a similar approach be taken to solve the \term{nearest lattice point problem}?
+\item Can a similar approach be taken to solve the \term{closest lattice point problem}?
 \end{itemize}
@@
 
+\section{A series of relevant vectors}
+
 \frame{
-START TALKING ABOUT CLOSEST POINT ALGORITHM
+\frametitle{A series of relevant vectors}
+
+Let $x_0$ be some lattice point from $\Lambda$ and consider the following iteration,
+\begin{align*}
+x_{k+1} &= x_k + v_k \\
+v_k &= \arg\min_{ v \in \relevant(\Lambda) \cup \{\zerobf\} } \|y - x_k - v \|
+\end{align*} 
+}
+
+\frame{
+\begin{figure}[tp]
+	\centering 
+        \includegraphics{figs/seriesofrelevantvectors-1.mps}
+		\caption{Computing a closest point by a series of relevant vectors.}
+		%\label{fig:graphcut4}
+\end{figure}
+}
+
+\newtheorem{proposition}{Proposition}
+
+\frame{
+\frametitle{A series of relevant vectors}
+ \begin{proposition}
+ There is a finite number $K$ such that 
+\[
+x_K, x_{K+1}, x_{K+2}, \dots
+\]
+ are all closest points to $y$.
+ \end{proposition}
+}
+
+\frame{
+\frametitle{A series of relevant vectors}
+\begin{itemize}
+\item The number $K$ of iterations might be large.
+\item The initial point $x_0$ affects $K$.
+\item Minimising over the set of relevant vectors, that is computing
+\[
+\arg\min_{ v \in \relevant(\Lambda) \cup \{\zerobf\} } \|y - x_k - v \|
+\]
+might be expensive.
+\item There are as many as $2^{n+1}-2$ relevant vectors.
+\end{itemize}
+}
+
+\frame{
+\frametitle{A series of relevant vectors}
+For a lattice of Voronoi's first kind:
+\begin{itemize}
+\item The intial point $x_0$ can be chosen to ensure that at most $K = n$ iterations is required before a closest point in found.
+\item Minimisation over the set of relevant vector can be performed by computing a minimimum cut in a flow network.
+\item Using known algorithms a minimum cut can be found in $O(n^3)$ operations.
+\item In total $O(n^4)$ operations are required to compute a closest lattice point.
+\end{itemize}
+}
+
+\frame{
+\frametitle{A series of relevant vectors}
+{\color{gray}For a lattice of Voronoi's first kind:}
+\begin{itemize}
+\item The intial point $x_0$ can be chosen to ensure that at most $K = n$ iterations is required before a closest point in found.
+\item {\color{gray}Minimisation over the set of relevant vector can be performed by computing a minimimum cut in a flow network.}
+\item {\color{gray}Using known algorithms a minimum cut can be found in $O(n^3)$ operations.}
+\item {\color{gray}In total $O(n^4)$ operations are required to compute a closest lattice point.}
+\end{itemize}
 }
@@
 \item Can good codes or quantisers be constructed from lattices of Voronoi's first kind?
-\item Signal processing applications, phase unwrapping, circular statistics, etc.
+\item Do applications, such as global position, phase unwrapping, circular statistics, involve lattices of Voronoi's first kind?
 \end{itemize}

### 2014-04-21 talk/figs/latticefigures.rb

@@ -6,9 +6,9 @@ require 'mathn'
 include RubyPost
 
 #a picture containing lattice points zoomed out
-scale = 0.42
+SCALE = 0.42
 pic = Picture.new('picL')
-M = Matrix[ [3.0,0.6], [0.6, 3.0] ]*scale
+M = Matrix[ [3.0,0.6], [0.6, 3.0] ]*SCALE
 (-7..7).each do |x|
   (-7..7).each do |y|
     v = M*Vector[x,y]
@@ -17,12 +17,12 @@ M = Matrix[ [3.0,0.6], [0.6, 3.0] ]*scale
 end
 
 #clippath for all figures.  Fits nicely on beamer screen
-$clippath = "((-5.6cm,-3.83cm)--(5.6cm,-3.83cm)--(5.6cm,3.83cm)--(-5.6cm,3.83cm)--cycle)"
+CLIPPATH = "((-5.6cm,-3.83cm)--(5.6cm,-3.83cm)--(5.6cm,3.83cm)--(-5.6cm,3.83cm)--cycle)"
 
 #figure with just lattice points
 file = RubyPost::File.new('lattice')
 fig = Figure.new
-fig.add_drawable(Clip.new($clippath, pic))
+fig.add_drawable(Clip.new(CLIPPATH, pic))
 fig.add_drawable(Draw.new(pic))
 file.add_figure(fig)
 file.compile
@@ -40,9 +40,9 @@ end
 #figure with lattice points and sphere packing
 file = RubyPost::File.new('latticewithspherepacking')
 fig = Figure.new
-fig.add_drawable(Clip.new($clippath, pic))
+fig.add_drawable(Clip.new(CLIPPATH, pic))
 fig.add_drawable(Draw.new(pic))
-fig.add_drawable(Clip.new($clippath, picsphp))
+fig.add_drawable(Clip.new(CLIPPATH, picsphp))
 fig.add_drawable(Draw.new(picsphp))
 file.add_figure(fig)
 file.compile
@@ -51,7 +51,7 @@ file.compile
 B = Matrix[ 
            [3.0, -2.4, -0.6], 
            [0.6, 2.4, -3.0] 
-]*scale
+]*SCALE
 
 #draws a picture of a bigshaded square that we will clip to the Voronoi region
 def voronoiCell(fig)
@@ -75,7 +75,8 @@ def voronoiCell(fig)
       end
     end
   end
-  fig.add_drawable(Clip.new($clippath, vorPic))
[…]

### 2014-04-21 talk/presentation.tex
minor

@@
 \item This is called the \term{closest lattice point problem} and a solution is called a \term{closest lattice point} to $y$.
+\item The lattice point $x \in \Lambda$ is closest to $y \in \reals^m$ if and only if
+\[
+y \in \vor(\Lambda) + x.
+\]
 \end{itemize}
@@
 
-\frame{
-\frametitle{The closest lattice point problem}
-\begin{itemize}
-\item The lattice point $x \in \Lambda$ is closest to $y \in \reals^m$ if and only if
-\[
-y \in \vor(\Lambda) + x.
-\]
-\item Equivalently $x$ is closest to $y$ if and only if
-\[
-\dotprod{(y - x)}{v} \leq \tfrac{1}{2} \dotprod{v}{v}
-\]
-for all $v \in \relevant(\Lambda)$.
-\end{itemize}
-}
+% \frame{
+% \frametitle{The closest lattice point problem}
+% \begin{itemize}
+% \item Equivalently $x$ is closest to $y$ if and only if
+% \[
+% \dotprod{(y - x)}{v} \leq \tfrac{1}{2} \dotprod{v}{v}
+% \]
+% for all $v \in \relevant(\Lambda)$.
+% \end{itemize}
+% }
 
@@
 \begin{itemize}
-\item The intial point $x_0$ can be chosen to ensure that at most $K = n$ iterations is required before a closest point in found.
+\item $x_0$ can be chosen to ensure that a closest lattice point in found after at most $K \leq n$ iterations.
 \item Minimisation over the set of relevant vector can be performed by computing a minimimum cut in a flow network.
@@
 \frametitle{A series of relevant vectors}
-{\color{gray}For a lattice of Voronoi's first kind:}
-\begin{itemize}
-\item The intial point $x_0$ can be chosen to ensure that at most $K = n$ iterations is required before a closest point in found.
-\item {\color{gray}Minimisation over the set of relevant vector can be performed by computing a minimimum cut in a flow network.}
-\item {\color{gray}Using known algorithms a minimum cut can be found in $O(n^3)$ operations.}
-\item {\color{gray}In total $O(n^4)$ operations are required to compute a closest lattice point.}
-\end{itemize}
+
+\begin{theorem}
+Let $\Lambda$ be a $n$-dimensional lattice of Voronoi's first kind with obtuse superbase $b_1, \dots, b_{n+1}$.  Let $z_1,\dots,z_{n+1} \in \reals$ minimise
+\[
+\| y - \sum_{n=1}^{n+1} b_i z_i \| 
+\]
+and put
+\[
+x_0 = \sum_{n=1}^{n+1} b_i \floor{z_i}.
+\]
+The iterative proceedure, initialized at $x_0$, converges to a closest lattice point in at most $K \leq n$ iterations.
+\end{theorem}
 }
@@
 \item Can good codes or quantisers be constructed from lattices of Voronoi's first kind?
-\item Do applications, such as global position, phase unwrapping, circular statistics, involve lattices of Voronoi's first kind?
+\item Do applications such as global position, phase unwrapping, circular statistics, etc., involve lattices of Voronoi's first kind?
 \end{itemize}

### 2014-04-21 paper.tex
minor

@@
 The following remark follow trivially from the preceding discussion.
-The \emph{relevant vectors} $v \in \Lambda$ are those for which  
+The \emph{relevant vectors} are those lattice points $v \in \Lambda \backslash  \{\zerobf\}$ for which  
 \[
-\dotprod{v}{x} \leq \dotprod{x}{x}
+\dotprod{v}{x} \leq \dotprod{x}{x} \qquad \text{for all $x \in \Lambda$}.
 \]
-for all $x \in \Lambda$ with $x \neq v$ and $x \neq \zerobf$.
-We denote by $\relevant(\Lambda)$ the set of relevant vectors of the lattice $\Lambda$.
-The Voronoi cell is the intersection of the halfspaces corresponding with the relevant vectors, that is,
+We denote by $\relevant(\Lambda)$ the set of relevant vectors of the lattice $\Lambda$.  The Voronoi cell is the intersection of the halfspaces corresponding with the relevant vectors, that is, 
 \[
@@
 \begin{equation}\label{eq:relvectnearpointieq}
-\dotprod{(y - x)}{v} \leq \tfrac{1}{2} \dotprod{v}{v}
+\dotprod{(y - x)}{v} \leq \tfrac{1}{2} \dotprod{v}{v} \qquad \text{for all $v \in \relevant(\Lambda)$}.  
 \end{equation}
-for all $v \in \relevant(\Lambda)$.
 

### 2014-04-21 talk/presentation.tex
minor

@@
 
-The \term{Voronoi cell} of a lattice $\Lambda \subset \reals^m$ is the subset of $\reals^m$ with points at least as close to the origin than to any lattice point,
+The \term{Voronoi cell} of a lattice $\Lambda \subset \reals^m$ is the subset of $\reals^m$ at least as close to the origin than to any lattice point,
 \[
@@
 \begin{itemize}
-\item The set of relevant vectors $\relevant(\Lambda)$ contains those $v \in \Lambda$ such that  
+\item The relevant vectors are those $v \in \Lambda \backslash \{0\}$ such that  
 \[
-\dotprod{v}{x} \leq \dotprod{x}{x}
+\dotprod{v}{x} \leq \dotprod{x}{x} \qquad \text{for all $x \in \Lambda$.}
 \]
-for all $x \in \Lambda$ with $x \neq v$ and $x \neq \zerobf$.  
 \item The Voronoi cell can be defined using the relevant vectors, 
 \[
-\vor(\Lambda) = \{ x \in\reals^m \mid \|x\| \leq \|x - v\|, v \in\relevant(\Lambda) \}
+\vor(\Lambda) = \{ x \in\reals^m \mid \|x\| \leq \|x - v\|, v \in\relevant(\Lambda) \}.
 \]
@@
 
-Given a lattice $\Lambda \subset \reals^m$ and a vector $y \in \reals^m$, a problem of interest is to find $x \in \Lambda$ such that
+Given a lattice $\Lambda \subset \reals^m$ and a vector $y \in \reals^m$ find $x \in \Lambda$ such that
 \[
@@
 x_{k+1} &= x_k + v_k \\
-v_k &= \arg\min_{ v \in \relevant(\Lambda) \cup \{\zerobf\} } \|y - x_k - v \|
+v_k &= \arg\min_{ v \in \relevant(\Lambda) \cup \{0\} } \|y - x_k - v \|
 \end{align*} 
@@
 \[
-\arg\min_{ v \in \relevant(\Lambda) \cup \{\zerobf\} } \|y - x_k - v \|
+\arg\min_{ v \in \relevant(\Lambda) \cup \{0\} } \|y - x_k - v \|
 \]

### 2014-04-22 talk/presentation.tex
added to talk

@@
 \author{Robby McKilliam \\
-\vspace{0.2cm}
+%\vspace{0.2cm}
 Alex Grant \\ Vaughan Clarkson \\
@@
 \item $x_0$ can be chosen to ensure that a closest lattice point in found after at most $K \leq n$ iterations.
-\item Minimisation over the set of relevant vector can be performed by computing a minimimum cut in a flow network.
+\item Minimisation over the set of relevant vectors can be performed by computing a minimum cut in a flow network.
 \item Using known algorithms a minimum cut can be found in $O(n^3)$ operations.
@@
 \frame{
-\frametitle{A series of relevant vectors}
+%\frametitle{A series of relevant vectors}
 
@@
 \]
-The iterative proceedure, initialized at $x_0$, converges to a closest lattice point in at most $K \leq n$ iterations.
+The iterative procedure, initialized at $x_0$, converges to a closest lattice point in at most $K \leq n$ iterations.
 \end{theorem}
@@
 \item Do applications such as global position, phase unwrapping, circular statistics, etc., involve lattices of Voronoi's first kind?
+\item Are there subfamilies of Voronoi's first kind that admit even faster algorithms?
+\item Are there other families of lattices for which similar techniques lead to polynomial time algorithms?
 \end{itemize}

### 2014-04-24 talk/presentation.tex
minor
removed chaff

@@
 
-\newtheorem{proposition}{Proposition}
+% \newtheorem{proposition}{Proposition}
 
-\frame{
-\frametitle{A series of relevant vectors}
- \begin{proposition}
- There is a finite number $K$ such that 
-\[
-x_K, x_{K+1}, x_{K+2}, \dots
-\]
- are all closest points to $y$.
- \end{proposition}
-}
+% \frame{
+% \frametitle{A series of relevant vectors}
+%  \begin{proposition}
+%  There is a finite number $K$ such that 
+% \[
+% x_K, x_{K+1}, x_{K+2}, \dots
+% \]
+%  are all closest points to $y$.
+%  \end{proposition}
+% }
 
@@
 \begin{itemize}
-\item The number $K$ of iterations might be large.
-\item The initial point $x_0$ affects $K$.
+\item The number of iterations depends on $x_0$ and might be large.
 \item Minimising over the set of relevant vectors, that is computing
@@
 \begin{itemize}
-\item $x_0$ can be chosen to ensure that a closest lattice point in found after at most $K \leq n$ iterations.
+\item $x_0$ can be chosen to ensure that a closest lattice point in found after at most $n$ iterations.
 \item Minimisation over the set of relevant vectors can be performed by computing a minimum cut in a flow network.
@@
 \[
-\| y - \sum_{n=1}^{n+1} b_i z_i \| 
+\| y - \sum_{i=1}^{n+1} b_i z_i \| 
 \]
@@
 \[
-x_0 = \sum_{n=1}^{n+1} b_i \floor{z_i}.
+x_0 = \sum_{i=1}^{n+1} b_i \floor{z_i}.
 \]
-The iterative procedure, initialized at $x_0$, converges to a closest lattice point in at most $K \leq n$ iterations.
+The iterative procedure, initialized at $x_0$, converges to a closest lattice point in at most $n$ iterations.
 \end{theorem}

### 2014-05-26 talk/presentation.tex
minor

@@
 \item Easier for specific lattices.
-\item For example, short vectors are easy to find in the \term{root lattices} $\ints^n$, $A_n$, and $D_n$.
+\item Short vectors are easy to find in the \term{root lattices} $\ints^n$, $A_n$, and $D_n$.
 \item We will show that the problem is relatively easy to solve for lattices of \term{Voronoi's first kind}.
@@
 \begin{itemize}
-\item The relevant vectors are those $v \in \Lambda \backslash \{0\}$ such that  
-\[
-\dotprod{v}{x} \leq \dotprod{x}{x} \qquad \text{for all $x \in \Lambda$.}
-\]
+\item Denote the set of relevant vectors by $\relevant(\Lambda)$.
+%\item The relevant vectors are those $v \in \Lambda \backslash \{0\}$ such that  
+%\[
+%\dotprod{v}{x} \leq \dotprod{x}{x} \qquad \text{for all $x \in \Lambda$.}
+%\]
 \item The Voronoi cell can be defined using the relevant vectors, 
@@
 \item coding and quantisation,
-\item multi-antenna communications (MIMO),
+\item multi-antenna communication (MIMO),
 \item unwrapping of phase data for electronic distance measurement in GPS and surveying,
@@
 \item Easier for specific lattices.
-\item For example, fast algorithms exist for the \term{root lattices} $\ints^n$, $A_n$, and $D_n$.
+\item Fast algorithms exist for the \term{root lattices} $\ints^n$, $A_n$, and $D_n$.
 \item We will describe a fast algorithm to compute a closest point in lattices of \term{Voronoi's first kind}.
@@
 \frametitle{Lattices of Voronoi's first kind}
-\begin{theorem}
+\begin{theorem}[Conway and Sloane (1992)]
 Let $\Lambda$ be a $n$-dimensional lattice of Voronoi's first kind with obtuse superbase $b_1, \dots, b_{n+1}$.  The relevant vectors in $\Lambda$ are of the form 
@@
 \begin{itemize}
-\item deterministically in $O(n^3)$ operations using the algorithm of Stoer and Wagner,
-\item with high probability in $O(n^2 \log(n)^3)$ operations using the randomised algorithm of Karger and Stien.
+\item deterministically in $O(n^3)$ operations using the algorithm of Stoer and Wagner (1997),
+\item with high probability in $O(n^2 \log(n)^3)$ operations using the randomised algorithm of Karger and Stien (1996).
 \end{itemize}	
@@
 \frame{
-\begin{theorem}
+\begin{theorem}[McKilliam and Grant (2012)]
 Let $\Lambda$ be a $n$-dimensional lattice of Voronoi's first kind with obtuse superbase 
@@
 \begin{itemize}
-\item $x_0$ can be chosen to ensure that a closest lattice point in found after at most $n$ iterations.
+\item $x_0$ can be chosen to ensure that a closest lattice point is found after at most $n$ iterations.
 \item Minimisation over the set of relevant vectors can be performed by computing a minimum cut in a flow network.
@@
 
-\begin{theorem}
+\begin{theorem}[McKilliam, Grant, Clarkson (2014)]
 Let $\Lambda$ be a $n$-dimensional lattice of Voronoi's first kind with obtuse superbase $b_1, \dots, b_{n+1}$.  Let $z_1,\dots,z_{n+1} \in \reals$ minimise
@@
 \item Can good codes or quantisers be constructed from lattices of Voronoi's first kind?
-\item Do applications such as global position, phase unwrapping, circular statistics, etc., involve lattices of Voronoi's first kind?
+\item Do applications such as global positioning, phase unwrapping, etc., involve lattices of Voronoi's first kind?
 \item Are there subfamilies of Voronoi's first kind that admit even faster algorithms?
@@
 
+% \frame{
+
+% \frametitle{An extra problem}
+
+% Choose real numbers $\lambda_1,\dots,\lambda_n$ to minimise
+% \[
+% \sum_{i=1}^n \frac{P^2}{\lambda_i^2}
+% \]
+% where $P  = \operatorname{lcm}(\lambda_1,\dots,\lambda_n)$ subject to the constraints
+% \[
+% P > P_{\text{min}}
+% \]
+% and 
+% \[
+% \lambda_{\text{min}} < \lambda_i < \lambda_{\text{max}} \qquad i = 1,\dots,n.
+% \]
+
+% }
+
 \end{document}

### 2014-06-02 findsuperbase/findsuperbase.tex
added start of document describing how to find a superbase

--- 2014-04-21 paper.tex
@@
-\documentclass[final,leqno]{siamltex}
-%\documentclass[draftcls, onecolumn, 11pt]{../bib/IEEEtran}
-%\documentclass[journal]{../bib/IEEEtran}
-%\documentclass[conference]{../bib/IEEEtran}
+\documentclass[a4paper,10pt]{article}
 
-\usepackage{mathbf-abbrevs}
+\usepackage{../mathbf-abbrevs}
 
@@
 
-\title{Finding a closest point in a lattice of Voronoi's first kind}
-
-\author{Robby~G.~McKilliam, Alex Grant and I. Vaughan L. Clarkson}
+%opening
+\title{Finding a obtuse superbase}
+\author{Robby McKilliam}
+\date{\today}
 
@@
 
-% make the title area 
 \maketitle
 
- \begin{abstract} 
-We show that for those lattices of Voronoi's first kind with known obtuse superbasis, a closest lattice point can be computed in $O(n^4)$ operations where $n$ is the dimension of the lattice.
-To achieve this a series of relevant lattice vectors that converges to a closest lattice point is found.
-We show that the series converges after at most $n$ terms.
-Each vector in the series can be efficiently computed in $O(n^3)$ operations using an algorithm to compute a minimum cut in an undirected flow network.
-%We discuss potential applications of this algorithm as suggest some future research questions.
-\end{abstract}
-
-%\begin{IEEEkeywords}
-\begin{keywords}
-Lattices, closest point algorithm, closest vector problem.
-\end{keywords}
-%\end{IEEEkeywords}
-
-% The paper headers
-\pagestyle{myheadings}
-\thispagestyle{plain} 
-\markboth{Finding a closest point in a lattice of Voronoi's first kind}{DRAFT \today}
-
-\section{Introduction}\label{sec:introduction}
-
-An $n$-dimensional \term{lattice} $\Lambda$ is a discrete set of vectors from $\reals^m$, $m \geq n$, formed by the integer linear combinations of a set of linearly independent basis vectors $b_1, \dots, b_n$ from $\reals^m$~\cite{SPLAG}.
-That is, $\Lambda$ consists of all those vectors, or \emph{lattice points}, $x \in \reals^m$ satisfying
-\[
-  x = b_1 u_1 + b_2u_2 + \dots + b_n u_n \qquad u_1, \dots , u_n \in \ints. 
-\] 
-%In this paper vectors are assumed to be column vectors unless otherwise stated.
-Given a lattice $\Lambda$ in $\reals^m$ and a vector $y \in \reals^m$, a problem of interest is to find a lattice point $x \in \Lambda$ such that the squared Euclidean norm
-\[
-\| y - x \|^2 = \sum_{i=1}^m (y_i - x_i)^2
-\] 
-is minimised.
-This is called the \emph{closest lattice point problem} (or \emph{closest vector problem}) and a solution is called a \emph{closest lattice point} (or simply \emph{closest point}) to $y$.
-%Equivalently we wish to minimise
-%\[
-%\| y - \sum_{i=1}^n b_i w_i \|^2
-%\]
-%over integers $w_1,\dots,w_n$.  
-A related problem is to find a lattice point of minimum nonzero Euclidean length, that is, a lattice point of length
-\[
-\min_{x\in \Lambda \backslash \{ \zerobf \} } \| x \|^2,
-\]
-where $\Lambda \backslash  \{\zerobf\}$ denotes the set of lattice points not equal to the origin $\zerobf$.  This is called the \emph{shortest vector problem}.
-
-The closest lattice point problem and the shortest vector problem have interested mathematicians and computer scientists due to their relationship with integer programming~\cite{Lenstra_integerprogramming1983,Kannan1987_fast_general_np,Babai1986}, the factoring of polynomials~\cite{Lenstra1982}, and cryptanalysis~\cite{Joux_toolbox_cryptanal1998,NyguyenStern_two_faces_crypto,Micciancio_lattice_based_post_quantum_crypto}.
-Solutions of the closest lattice point problem have engineering applications.
-For example, if a lattice is used as a vector quantiser then the closest lattice point corresponds to the minimum distortion point~\cite{Conway1983VoronoiCodes,Conway1982VoronoiRegions,Conway1982FastQuantDec}.
-If the lattice is used as a code, then the closest lattice point corresponds to what is called \emph{lattice decoding}\index{lattice decoding} and has been shown to yield arbitrarily good codes~\cite{Erex2004_lattice_decoding,Erez2005}.
-The closest lattice point problem also occurs in communications systems involving multiple antennas~\cite{Ryan2008,Wubben_2011}.
-The unwrapping of phase data for location estimation can also be posed as a closest lattice point problem and this has been applied to the global positioning system~\cite{Teunissen_GPS_1995,Hassibi_GPS_1998}.
-The problem has also found applications to circular statistics~\cite{McKilliam_mean_dir_est_sq_arc_length2010}, single frequency estimation~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}, and related signal processing problems~\cite{McKilliam2007,Clarkson2007,McKilliam2009IndentifiabliltyAliasingPolyphase,Quinn_sparse_noisy_SSP_2012}.
-
-The closest lattice point problem is known to be NP-hard under certain conditions when the lattice itself, or rather a basis thereof, is considered as an additional input parameter~\cite{micciancio_hardness_2001, Dinur2003_approximating_CVP_NP_hard, Jalden2005_sphere_decoding_complexity}.
-Nevertheless, algorithms exist that can compute a closest lattice point in reasonable time if the dimension is small~\cite{Pohst_sphere_decoder_1981,Kannan1987_fast_general_np,Agrell2002}.
-These algorithms all require a number of operations that grows as $O(n^{O(n)})$ where $n$ is the dimension of the lattice.
-Recently, Micciancio~\cite{Micciancio09adeterministic} described a solution for the closest lattice point problem that requires a number of operations that grows as $O(2^{2n})$.
-This single exponential growth in complexity is the best known.
-
-Although the problem is NP-hard in general, fast algorithms are known for specific highly regular lattices, such as the root lattices $A_n$ and $D_n$, their dual lattices $A_n^*$ and $D_n^*$, the integer lattice $\ints^{n}$, and the Leech lattice~\cite[Chap.
-4]{SPLAG}\cite{Conway1982FastQuantDec,Conway1986SoftDecLeechGolay, Clarkson1999:Anstar, McKilliam2008, McKilliam2008b, McKilliam2009CoxeterLattices,Vardy1993_leech_lattice_MLD}.
-In this paper we consider a particular class of lattices, those of \emph{Voronoi's first kind}~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases,Valentin2003_coverings_tilings_low_dimension,Voronoi1908_main_paper}.
-Each lattice of Voronoi's first kind has what is called an \emph{obtuse superbasis}.
-We show that if the obtuse superbasis is known, then a closest lattice point can be computed in $O(n^4)$ operations.
-This is achieved by enumerating a series of \emph{relevant vectors} of the lattice.
-Each relevant vector in the series can be computed in $O(n^3)$ operations using an algorithm for computing a minimum cut in an undirected flow network~\cite{Picard_min_cuts_1974,Sankaran_solving_CDMA_mincut_1998,Ulukus_cdma_mincut_1998,Cormen2001}.
-We show that the series converges to a closest lattice point after at most $n$ terms, resulting in $O(n^4)$ operations in total.
-This result extends upon a recent result by some of the authors showing that a short vector in a lattice of Voronoi's first kind can be found by computing a minimum cut in a weighted graph~\cite{McKilliam_short_vectors_first_type_isit_2012}.
-
-The paper is structured as follows.
-Section~\ref{sec:voron-cells-relev} describes the relevant vectors and the \emph{Voronoi cell} of a lattice.
-Section~\ref{sec:iterative-slicer} describes a procedure to find a closest lattice point by enumerating a series of relevant vectors.
-The series is guaranteed to converge to a closest lattice point after a finite number of terms.
-In general the procedure might be computationally expensive because the number of terms required might be large and because computation of each relevant vector in the series might be expensive.
-Section~\ref{sec:latt-voron-first} describes lattices of Voronoi's first kind and their obtuse superbasis.
-In Section~\ref{sec:seri-relev-vect} it is shown that for these lattices, the series of relevant vectors results in a closest lattice point after at most $n$ terms.
-Section~\ref{sec:comp-clos-relev} shows that each relevant vector in the series can be computed in $O(n^3)$ operations by computing a minimum cut in an undirected flow network.
-Section~\ref{sec:discussion} discusses some potential applications of this algorithm and poses some interesting questions for future research.
-
-\section{Voronoi cells and relevant vectors}\label{sec:voron-cells-relev}
-\newcommand{\calR}{\mathcal{R}}
-The (closed) \term{Voronoi cell}, denoted $\vor(\Lambda)$, of a lattice $\Lambda$ in $\reals^m$ is the subset of $\reals^m$ containing all points closer or of equal distance (here with respect to the Euclidean norm) to the lattice point at the origin than to any other lattice point.
-The Voronoi cell is an $m$-dimensional convex polytope that is symmetric about the origin.
-%Here we will always assume the Euclidean norm (or 2-norm), so $\vor(\Lambda)$ contains those points nearer in Euclidean distance to the origin.
-%If $x \in \Lambda$ it follows that $\vor(\Lambda) + x$ is the subset of $\reals^n$ that is nearer to $x$ than any other lattice point in $\Lambda$.
-Figure \ref{lattices:fig:vorregion} is an example of the Voronoi cell.
-
-Equivalently the Voronoi cell can be defined as the intersection of the half spaces 
-\begin{align*}
-H_{v} &= \{x \in \reals^n \mid \|x\| \leq \|x - v\| \} \\
-&= \{x \in \reals^n \mid \dotprod{x}{v} \leq \tfrac{1}{2}\dotprod{v}{v} \}
-\end{align*}
-for all $v \in \Lambda \backslash  \{\zerobf\}$.  %We denote by $x^\prime$ the transpose of the vector $x$ and so $\dotprod{x}{v}$ is the inner product between column vectors $x$ and $v$.  
-We denote by $\dotprod{x}{v}$ the inner product between vectors $x$ and $v$.
-It is not necessary to consider all $v \in \Lambda \backslash  \{\zerobf\}$ to define the Voronoi cell.
-%Figure \ref{lattices:fig:vorregion} depicts the relevant vectors of the lattice with generator given by~\eqref{lattices:eq:genB}.
-A lattice point in $\calR$ is called \term{relevant}.
-The following remark follow trivially from the preceding discussion.
-The \emph{relevant vectors} $v \in \Lambda$ are those for which  
-\[
-\dotprod{v}{x} \leq \dotprod{x}{x}
-\]
-for all $x \in \Lambda$ with $x \neq v$ and $x \neq \zerobf$.
-We denote by $\relevant(\Lambda)$ the set of relevant vectors of the lattice $\Lambda$.
-The Voronoi cell is the intersection of the halfspaces corresponding with the relevant vectors, that is,
-\[
-%\begin{equation}\label{eq:defvortermsrelvecs}
-\vor(\Lambda) = \cap_{v\in\relevant(\Lambda)}{H_{v}}.
-%\end{equation}
-\]
-The closest lattice point problem and the Voronoi cell are related in that $x\in\Lambda$ is a closest lattice point to $y$ if and only $y - x \in \vor(\Lambda)$, that is, if and only if
-\begin{equation}\label{eq:relvectnearpointieq}
-\dotprod{(y - x)}{v} \leq \tfrac{1}{2} \dotprod{v}{v}
-\end{equation} 
-for all $v \in \relevant(\Lambda)$.
-
-% Every short vector in a lattice is also a relevant vector, because, if a lattice point $s$ is not relevant there exists a lattice point $x$ not equal to $s$ or $\zerobf$ such that $\dotprod{x}{s} \geq \dotprod{x}{x}$ or equivalently $\dotprod{\tfrac{x}{\|x\|}}{s} \geq \|x\|$ after dividing by $\|x\|$.
-Since both $\tfrac{x}{\|x\|}$ and $\tfrac{s}{\|s\|}$ are unit vectors and $s \neq x$ then,
-% \[
-% \|s\| = \dotprod{\tfrac{s}{\|s\|}}{s} >  \dotprod{\tfrac{x}{\|x\|}}{s} \geq \|x\|,
-% \]
-% i.e, $\|s\| > \|x\|$, so $s$ is not a short vector.
-
-If $s$ is a short vector in a lattice $\Lambda$ then 
-\[
-\rho = \frac{\|s\|}{2} = \frac{1}{2} \min_{x \in \Lambda / \{\zerobf\} } \|x\|
-\]
-is called the \emph{packing radius} (or \emph{inradius}) of $\Lambda$~\cite{SPLAG}.
-The packing radius is the minimum distance between the boundary of the Voronoi cell and the origin.
-It is also the radius of the largest sphere that can be placed at every lattice point such that no two spheres intersect (see Figure~\ref{lattices:fig:vorregion}).
-The following well known results will be useful.
-
-%  \begin{proposition}\label{prop:latticepointinscaledtranslatedvorcell}
-%  Let $\Lambda \subset \reals^{m}$ be an $n$-dimensional lattice with Voronoi cell $\vor(\Lambda)$.
-Let $D$ be a positive real number and let $\ceil{D}$ denote the smallest integer larger than $D$.
-Denote by
-%  \[
-%   D\vor(\Lambda) + t = \left\{ x \in \reals^{m} \ \mid \frac{x-t}{D} \in \vor(\Lambda) \right\}
-%  \]
-% the Voronoi cell scaled by $D$ and translated by $t \in \reals^{m}$.  The number of lattice points from $\Lambda$ in $ D\vor(\Lambda) + t$ is less than or equal to $\ceil{D}^{n}$. 
-%  \end{proposition}
-%  \begin{proof}
-% See, for example,~\cite[Lemma~3.7]{Micciancio09adeterministic}.
-The result also follows from~\cite[vol.~134, p.~277]{Voronoi1908_main_paper} or \cite[Theorem~2]{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.
-%  \end{proof}
-
-\begin{proposition}\label{eq:latticepointsinvorcvell}
-Let $\Lambda \subset \reals^{m}$ be an $n$-dimensional lattice.
-For $r\in \reals$ let $\ceil{r}$ denote the smallest integer strictly larger than $r$.
-Let $t \in \reals^m$.
-The number of lattice points inside the scaled and translated Voronoi cell $r \vor(\Lambda) + t$ is at most $\ceil{r}^n$.
-\end{proposition}
-\begin{proof}
-It is convenient to modify the boundary of the Voronoi cell so that it tessellates $\reals^m$ under translations by $\Lambda$.
-With this aim we let $V \subset \vor(\Lambda)$ contain all those points from the interior of $\vor(\Lambda)$, but have opposing faces open and closed.
-That is, if $x \in V$ is on the boundary of $V$ then $-x \notin V$.
-With this definition it can be asserted that $V$ tessellates $\reals^m$ under translations by $\Lambda$.
-That is, $\reals^{m} = \cup_{x \in \Lambda}(V + x)$ and the intersection $(V + x)\cap(V+y)$ is empty for distinct lattice points $x$ and $y$.
-Now, for positive integer $k$, the scaled and translated cell $kV + t$ contains precisely one coset representative for each element of the quotient group $\Lambda/k\Lambda$~\cite[Sec.~2.4]{McKilliam2010thesis}.
-There are $k^n$ coset representatives.
-Thus, the number of lattice points inside $r \vor(\Lambda) + t \subset \ceil{r}V + t$ is at most $\ceil{r}^n$.
-\end{proof}
-
-\begin{proposition}\label{eq:latticepointsinsphere}
-Let $\Lambda \subset \reals^{m}$ be an $n$-dimensional lattice with packing radius $\rho$.
-Let $S$ be an $m$-dimensional hypersphere of radius $r$ centered at $t \in \reals^{m}$.
-The number of lattice points from $\Lambda$ in the sphere $S$ is at most $\ceil{r/\rho}^{n}$.
-%where $\ceil{\cdot}$ denotes the smallest integer strictly larger than its argument.
- \end{proposition}
- \begin{proof}
-The packing radius $\rho$ is the Euclidean length of a point on the boundary of the Voronoi cell $\vor(\Lambda)$ that is closest to the origin.
-Therefore, the sphere $S$ is a subset of $\vor(\Lambda)$ scaled by $r/\rho$ and translated by $t$.
-That is, $S \subset r/\rho \vor(\Lambda) + t$.
-The proof follows because the number of lattice points in $r/\rho \vor(\Lambda) + t$ is at most $\ceil{r/\rho}^n$ by Proposition~\ref{eq:latticepointsinvorcvell}.
-\end{proof}
-
-% \footnote{These are the `strict' relevant vectors according to Conway and Sloane~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.
-If the inequality $\dotprod{v}{x} < \dotprod{x}{x}$ is replaced by $\dotprod{v}{x} \leq \dotprod{x}{x}$ then this would also include the `lax' relevant vectors.
-The short vectors are always strict so we only have use of the strict relevant vectors here.}
-
-\begin{figure}[tp] 
-	\centering      
-		\includegraphics{figs/latticefigures-1.mps} 
-		\caption{The $2$-dimensional lattice with basis vectors $(3,0.6)$ and $(0.6,3)$.
-The lattice points are represented by dots and the relevant vectors are circled.
-The Voronoi cell $\vor(\Lambda)$ is the shaded region and the packing radius $\rho$ and corresponding sphere packing (circles) are depicted.
-}     
-		\label{lattices:fig:vorregion}   
-\end{figure}
-
-\section{Finding a closest lattice point by a series of relevant vectors} \label{sec:iterative-slicer}
-
-Let $\Lambda$ be a lattice in $\reals^m$ and let $y \in \reals^m$.
-A simple method to compute a lattice point $x \in \Lambda$ closest to $y$ is as follows.
-Let $x_0$ be some lattice point from $\Lambda$, for example the origin.
-Consider the following iteration,
-\begin{align}
-x_{k+1} &= x_k + v_k \nonumber \\
-v_k &= \arg\min_{ v \in \relevant(\Lambda) \cup \{\zerobf\} } \|y - x_k - v \|, \label{eq:relvectsminimslicer}
-\end{align} 
-where $\relevant(\Lambda) \cup \{\zerobf\}$ is the set of relevant vectors of $\Lambda$ including the origin.
-The minimum over $\relevant(\Lambda) \cup \{\zerobf\}$ may not be unique, that is, there may be multiple vectors from $\relevant(\Lambda) \cup \{\zerobf\}$ that are closest to $y - x_k$.
-In this case, any one of the minimisers may be chosen.
-The results that we will describe do not depend on this choice.
-We make the following straightforward propositions.
-
-\begin{proposition}\label{obs:1}
-%After a finite number, say $K$ iterations, the above proceedure The above proceedure reaches a stationary point after a finite number $K$ of iterations
-At the $k$th iteration either $x_k$ is a closest lattice point to $y$ or $\|y - x_k\| > \| y - x_{k+1} \|$.
-\end{proposition}
-\begin{proof}
-If $x_k$ is a closest lattice point to $y$ then $\|y - x_k\| \leq \| y - x_{k+1} \|$ by definition.
-On the other hand if $x_k$ is not a closest lattice point to $y$ we have $y - x_k \notin \vor(\Lambda)$ and from~\eqref{eq:relvectnearpointieq} there exists a relevant vector $v$ such that
-\[
-0 > \dotprod{v}{v} - 2\dotprod{(y - x_k)}{v}.
-\]
-Adding $\|y - x_k\|^2$ to both sides of this inequality gives
-\begin{align*}
-\|y - x_k\|^2 &> \dotprod{v}{v} - 2\dotprod{(y - x_k)}{v} + \|y - x_k\|^2 \\
-&= \|y - x_k - v\|^2 \\
-&\geq \arg\min_{ v \in \relevant(\Lambda) \cup \{\zerobf\}}\|y - x_k - v \|^2 \\
-&= \|y - x_k - v_k \|^2 \\
-&= \|y - x_{k+1}\|^2. 
-\end{align*}
-\end{proof}
-
-% \begin{observation}\label{obs:1}
-% At the $k$th iteration either $x_k$ is a closest lattice point to $y$ or 
-% \[
-%  \max_{ v \in \relevant(\Lambda)}\frac{\dotprod{(y - x_k)}{v}}{\dotprod{v}{v}} > \frac{1}{2}.
-% \]
-% \end{observation}
-% \begin{proof}
-% If $x_k$ is a closest lattice point to $y$ then $\dotprod{(y - x)}{v} \leq \tfrac{1}{2} \dotprod{v}{v}$ for all $v \in \relevant(\Lambda)$ by definition of the Voronoi cell~\eqref{eq:relvecsvorcellnearpoint} and so
-% \[
-%  \max_{ v \in \relevant(\Lambda)}\frac{\dotprod{(y - x_k)}{v}}{\dotprod{v}{v}} \leq \frac{1}{2}.
-% \]
-% Conversely if $x_k$ is not the closest point to $y$ there exists a relevant vector $r$ such that $\dotprod{(y - x)}{r} > \tfrac{1}{2} \dotprod{r}{r}$ and so
-% \[
-%  \max_{ v \in \relevant(\Lambda)}\frac{\dotprod{(y - x_k)}{v}}{\dotprod{v}{v}} \geq \frac{\dotprod{(y - x_k)}{r}}{\dotprod{r}{r}} > \frac{1}{2}.
-% \]
-% \end{proof}
-
- \begin{proposition}\label{obs:2}
- There is a finite number $K$ such that $x_K, x_{K+1}, x_{K+2}, \dots$ are all closest points to $y$.
- \end{proposition}
- \begin{proof}
-Suppose no such finite $K$ exists, then
-\[
-\|y - x_0\| >  \|y - x_1\| > \|y - x_2\| > \dots
-\]
-and so $x_0,x_1,\dots$ is an infinite sequence of distinct (due to the strict inequality) lattice points  all contained inside an $n$-dimensional hypersphere of radius $r = \|y - x_0\|$ centered at $y$.
-This is a contradiction because, if $\rho$ is the packing radius of the lattice, then less than $\ceil{r/\rho}^n$ lattice points lie inside this sphere by Proposition~\ref{eq:latticepointsinsphere}.
-\end{proof}
-
-Proposition~\ref{obs:2} above asserts that after some finite number $K$ of iterations the procedure arrives at $x_K$, a closest lattice point to $y$.
-Using Proposition~\ref{obs:1} we can detect that $x_K$ is a closest lattice point by checking whether $\|y - x_K\| \leq \| y - x_{K+1} \|$.
-This simple iterative approach to compute a closest lattice point is related to what is called the \emph{iterative slicer}~\cite{Shalvi_iterativeslicer_2009}.
-Micciancio~\cite{Micciancio09adeterministic} describes a related, but more sophisticated, iterative algorithm that can compute a closest lattice point in a number of operations that grows exponentially as $O(2^{2 n})$.
-This single exponential growth in complexity is the best known.
-%Most popular algorithms for computed closest lattice points require a number of operations of order $O(n^n)$\cite{Agrell2002,Viterbo_sphere_decoder_1999,Pohst_sphere_decoder_1981}.
-
-Two factors contribute to the computational complexity of this iterative approach to compute a closest lattice point.
-The first factor is computing the minimum over the set $\relevant(\Lambda) \cup \{\zerobf\}$ in~\eqref{eq:relvectsminimslicer}.
-In general a lattice can have as many as $2^{n+1}-2$ relevant vectors so computing a minimiser directly can require a number of operations that grows exponentially with $n$.
-To add to this it is often the case that the set of relevant vectors $\relevant(\Lambda)$ must be stored in memory so the algorithm can require an amount of memory that grows exponentially with $n$~\cite[Sec.~6]{Micciancio09adeterministic}\cite{Shalvi_iterativeslicer_2009}.
-We will show that for a lattice of Voronoi's first kind the set of relevant vectors has a compact representation in terms of what is called its \emph{obtuse superbasis}.
-To store the obtuse superbasis requires an amount of memory of order $O(n^2)$ in the worst case.
-We also show that for a lattice of Voronoi's first kind the minimisation over $\relevant(\Lambda) \cup \{\zerobf\}$ in~\eqref{eq:relvectsminimslicer} can be solved efficiently by computing a minimum cut in an undirected flow network.
-Using known algorithms a minimiser can be computed in $O(n^3)$ operations~\cite{Goldberg:1986:NAM:12130.12144,EdmondsKarp_max_flow,Cormen2001}.
-
-The other factor affecting the complexity is the number of iterations required before the algorithm arrives at a closest lattice point, that is, the size of $K$.
-Proposition~\ref{eq:latticepointsinsphere} suggests that this number might be as large as $\ceil{r/\rho}^n$ where $r = \|y - x_0\|^2$ and $\rho$ is the packing radius of the lattice.
-Thus, the number of iterations required might grow exponentially with $n$.
-The number of iterations required depends on the lattice point that starts the iteration $x_0$.
-It is helpful for $x_0$ to be, in some sense, a close approximation of the closest lattice point $x_K$.
-Unfortunately, computing close approximations of a closest lattice point is known to be computationally difficult~\cite{feige_inapproximability_2004}.
-We will show that for a lattice of Voronoi's first kind a simple and easy to compute choice for $x_0$ ensures that a closest lattice point is reached in at most $n$ iterations and so $K \leq n$.
-Combining this with the fact that each iteration of the algorithm requires $O(n^3)$ operations results in an algorithm that requires $O(n^4)$ operations to compute a closest point in a lattice of Voronoi's first kind.
-
-\begin{figure}[tp] 
-	\centering      
-		\includegraphics{figs/iterativeseriesexample-1.mps} 
-		\caption{Example of the iterative procedure described in~\eqref{eq:relvectsminimslicer} to compute a closest lattice point to $y = (4,3.5)$ (marked with a cross) in the $2$-dimensional lattice generated by basis vectors $(2,0.4)$ and $(0.4,2)$.
-The initial lattice point for the iteration is $x_0 = (-4.4,-2.8)$.
-The shaded region is the Voronoi cell surrounding the closest lattice point $x_6 = (4.4,2.8)$.}
-		\label{lattices:fig:iterativeexample} 
-\end{figure}
-
-\section{Lattices of Voronoi's first kind} \label{sec:latt-voron-first}
+This document describes an algorithm to determine whether a lattice, specified by a basis matrix, is of Voronoi's first kind.
+The algorithm returns an obtuse superbase for the lattice if the lattice is of Vornoi's first kind.
 
@@
 
-% A number of important lattices are of Voronoi's first kind.
-For example, the \emph{root lattices} $\ints^n, A_n, E_6, E_7$ and $E_8$ are of Voronoi's first kind~\cite{SPLAG}.
-The root lattice $E_8$ has obtuse superbasis
-% \[
-
-% \]
-
-% The root lattice $D_n$ has obtuse superbasis given by the vectors
-% \begin{align*}
-% b_1 &= -e_1 - e_2 \\
-% b_i &= e_{i} - e_{i+1}, \qquad i = 2,\dots,n \\
-% b_{n+1} &= e_{2} + e_{n}.
-% \end{align*}
-
-Classical examples of lattices of Voronoi's first kind are the $n$ dimensional root lattice $A_n$ and its dual lattice $A_n^*$~\cite{SPLAG}.
-% These lattices have convenient bases represented in $n+1$ dimensional space.
-An obtuse superbasis for $A_n$ is of the form
-% \begin{align*}
-% b_i &= e_i - e_{i+1}, \qquad i = 1, \dots, n \\
-% b_{n+1} &= e_{n+1} - e_{1},
-% \end{align*}
-% where $e_i$ is a vector of length $n+1$ with elements equal to zero except for the $i$th element that is equal to one.
-The dual lattice $A_{n}^*$ is also of Voronoi's first kind.
-The $i$th vector $b_i$ of the obtuse superbasis has $j$th element equal to
-% \[
-% b_{ij} = \begin{cases}
-% \frac{n}{n+1} & j = i \\
-% -\frac{1}{n+1} & j \neq i.
-% \end{cases}
-% \]
-For $A_n$ and $A_n^*$ there exist efficient algorithms that can compute a closest lattice point in $O(n)$ operations~\cite{McKilliam2009CoxeterLattices,Conway1982FastQuantDec}.
-For this reason we do not recommend using the algorithm described in this paper for $A_n$ and $A_n^*$.
-The fast algorithms for $A_n$ and $A_n^*$ rely of the special structure of these lattices and are not applicable to other lattices.
-In contrast, the algorithm we describe here works for all lattices of Voronoi's first kind.
-Questions that arise are: how ``large'' (in some sense) is the set of lattices of Voronoi's first kind?
-Are there lattices of Voronoi's first kind that are useful in applications such as coding, quantisation, or signal processing?
-We discuss these questions in Section~\ref{sec:discussion}.
-We now focus on the problem of computing a closest lattice point in a lattice of Voronoi's first kind.
-
-%However, there exist a large number of lattice of first
-%A lattice of Voronoi's first kind appears nontrivial.
-One approach would be to construct the matrix $Q = B^\prime B$ containing the Selling parameters $q_{ij}$.
-The matrix $Q$ requires to have nonpositive off diagonal entries and to be positive semidefinite with null space containing the vector $\onebf$ of all ones.
-The explicit construction of such matrices is an interesting question for future research.
-A related question is to determine, in some sense, how `large' the set of matrices satisfying these criteria is.
-This would shed some light on how `large' the set of lattices of Voronoi's first kind is.
-
-\section{A series of relevant vectors from a lattice of Voronoi's first kind}\label{sec:seri-relev-vect}
-
-We are interested in solving the closest lattice point problem for lattices of Voronoi's first kind.
-Let $\Lambda \subset \reals^m$ be an $n$ dimensional lattice of Voronoi's first kind with obtuse superbasis $b_1,\dots,b_{n+1}$ and let $y \in \reals^m$.
-We want to find $n$ integers $w_1,\dots,w_n$ that minimise
-\[
-\| y - \sum_{i=1}^n b_i w_i \|^2.
-\]
-We can equivalently find $n+1$ integers $w_1,\dots,w_{n+1}$ that minimise
-\[
-\| y - \sum_{i=1}^{n+1} b_i w_i \|^2.
-\]
-The iterative procedure described in~\eqref{eq:relvectsminimslicer} will be used to do this.
-In what follows it is assumed that $y$ lies in the space spanned by the basis vectors $b_1,\dots,b_{n}$.
-This assumption is without loss of generality because $x$ is a closest lattice point to $y$ if and only if $x$ is a the closest lattice point to the orthogonal projection of $y$ into the space spanned by $b_1,\dots,b_{n}$.
-Let
-\begin{equation}\label{eq:matrxBobtusebasis}
-B = (b_1\,\,b_2\,\,\dots\,\,b_{n+1})
-\end{equation}
-be the $n+1$ by $n+1$ matrix with columns given by $b_1,\dots,b_{n+1}$ and let $z \in \reals^{n+1}$ be a column vector such that $y = Bz$.
-We now want to find a column vector $w = (w_1,\dots,w_{n+1})^\prime$ of integers such that
-\begin{equation}\label{eq:tominimise}
-\| B(z  -  w) \|^2
-\end{equation}
-is minimised.
-Define the column vector $u_0 = \floor{z}$ where $\floor{\cdot}$ denotes the largest integer less than or equal to its argument and operates on vectors elementwise.
-In view of Theorem~\ref{thm:revvecssuperbase} the iterative procedure~\eqref{eq:relvectsminimslicer} to compute a closest lattice point can be written in the form
-\begin{align}
-x_{k+1} &= B u_{k+1} \label{eq:xseqfirsttype}  \\
-u_{k+1} &= u_k + t_k \nonumber \\
-t_k &= \arg\min_{t \in \{0,1\}^{n+1}}\| B(z - u_k - t) \|^2, \label{eq:pvecmin}
-\end{align}
-where $\{0,1\}^{n+1}$ denotes the set of column vectors of length $n+1$ with elements equal to zero or one.
-The procedure is initialised at the lattice point $x_0 = Bu_0 = B\floor{z}$.
-This choice of initial lattice point is important.
-In Section~\ref{sec:comp-clos-relev} we show how minimisation over $\{0,1\}^{n+1}$ in~\eqref{eq:pvecmin} can be computed efficiently in $O(n^3)$ operations by computing a minimum cut in an undirected flow network.
-The minimiser may not be unique corresponding with the existence of multiple minimum cuts.
-In this case any one of the minimisers may be chosen.
-Our results do not depend on this choice.
-In the remainder of this section we prove that this iterative procedure results in a closest lattice point after at most $n$ iterations.
-That is, we show that there exists a positive integer $K \leq n$ such that $x_K$ is a closest lattice point to $y = Bz$.
-
-\newcommand{\rng}{\operatorname{rng}}
-\newcommand{\subrng}{\operatorname{subr}}
-\newcommand{\decrng}{\operatorname{decrng}}
-
-It is necessary to introduce some notation.
-For $S$ a subset of indices $\{1,\dots,n+1\}$ let $\onebf_S$ denote the column vector of length $n+1$ with $i$th element equal to one if $i \in S$ and zero otherwise.
-%Thus, for example, if $p \in \reals^{n+1}$, then $p+\onebf_S$ denotes the vector $p$ with those elements having index from $S$ incremented by one.
-For $S \subseteq \{1,\dots,n+1\}$ and $p \in \reals^{n+1}$ we define the function
-\[
-\Phi(S, p) = \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_i - 2p_j)
-\]
-where $q_{ij} = \dotprod{b_i}{b_j}$ are the Selling parameters from~\eqref{eq:obtusecond}.
-We denote by $\bar{S}$ the complement of the set of indices $S$, that is $\bar{S} = \{ i \in \{1,\dots,n+1\} \mid i \notin S\}$.
-
-\begin{lemma}\label{lem:decSellings}
-Let $p \in \reals^{n+1}$ and let $S$ and $T$ be subsets of the indices of $p$.  The following equalities hold:
-\begin{enumerate}
-\item  $\|Bp\|^2 - \|B(p + \onebf_S)\|^2 = \Phi(S, p)$, \label{eq:lem:decSellingsinc}
-\item  $\|Bp\|^2 - \|B(p - \onebf_S)\|^2 = \Phi(\bar{S}, p)$, \label{eq:lem:decSellingsdec}
-\item  ${\displaystyle \|Bp\|^2 - \|B(p + \onebf_S - \onebf_T)\|^2 = \Phi(S, p) + \Phi(\bar{T},p) + \sum_{i\in S}\sum_{j\in T} q_{ij} }$.  \label{eq:lem:decSellingsincdec}
-\end{enumerate}
-\end{lemma}
-\begin{proof}
-Part~\ref{eq:lem:decSellingsincdec} follows immediately from parts~\ref{eq:lem:decSellingsinc} and~\ref{eq:lem:decSellingsdec} because
-\begin{align*}
-\|Bp\|^2 - &\|B(p + \onebf_S - \onebf_T)\|^2 \\
-&= \|Bp\|^2-\|B(p + \onebf_S)\|^2 +  \|Bp\|^2-\|B(p - \onebf_T)\|^2 + \sum_{i\in S}\sum_{j\in T} q_{ij}.
-\end{align*}
-We give a proof for part~\ref{eq:lem:decSellingsinc}.  The proof for part~\ref{eq:lem:decSellingsdec} is similar.  
-Put $Q = B^\prime B$ where superscript $^\prime$ indicates the vector or matrix transpose.
-The $n+1$ by $n+1$ matrix $Q$ has elements given by the Selling parameters, that is, $Q_{ij} = q_{ij}=\dotprod{b_i}{b_j}$.
-Denote by $\onebf$ the column vector of length $n+1$ containing all ones.
-Now $B\onebf = \sum_{i=1}^{n+1}b_i = \zerobf$ as a result of the superbasis condition~\eqref{eq:superbasecond} and so $Q\onebf = \zerobf$.
-Since $\onebf_S = \onebf - \onebf_{\bar{S}}$ it follows that $Q\onebf_S = -Q\onebf_{\bar{S}}$.
-With $\circ$ the elementwise vector product, i.e., the Schur or Hadamard product, we have
- \begin{align*}
- \|Bp\|^2 - \|B(p + \onebf_S)\|^2 &= -\onebf^\prime_S Q \onebf_S - 2 p^\prime Q \onebf_S \\
- &= \onebf^\prime_S Q \onebf_{\bar{S}} - 2 p^\prime Q \onebf_S \\
-&= \onebf^\prime_S Q \onebf_{\bar{S}} - 2 (p\circ\onebf_{\bar{S}})^\prime Q \onebf_S - 2 (p\circ\onebf_S)^\prime Q \onebf_S \\
-&= \onebf^\prime_S Q \onebf_{\bar{S}} - 2 (p\circ\onebf_{\bar{S}})^\prime Q \onebf_{S} + 2 (p\circ\onebf_S)^\prime Q \onebf_{\bar{S}}
- \end{align*}
-which is precisely $\Phi(S,p)$.
-% Observe that
-% \[
-% \sum_{i=1}^{n+1}\sum_{j=1}^{n+1} q_{ij} p_i^2 =  \sum_{i=1}^{n+1}p_i^2\sum_{j=1}^{n+1} q_{ij} = 0
-% \]
-% and
-% \[
-% \sum_{i=1}^{n+1}\sum_{j=1}^{n+1} q_{ij} p_j^2 =  \sum_{j=1}^{n+1}p_j^2\sum_{i=1}^{n+1} q_{ij} = 0
-% \]
-% since $\sum_{i=1}^{n+1} q_{ij} = \sum_{j=1}^{n+1} q_{ij} = 0$ as a result of the superbasis condition~\eqref{eq:superbasecond}.  Now
-% \begin{align*}
-% \|Bp\|^2 &= \sum_{i=1}^{n+1} \sum_{j=1}^{n+1} q_{ij} p_i p_j \\
-% &= \sum_{i=1}^{n+1} \sum_{j=1}^{n+1} q_{ij} (\tfrac{1}{2}p_j^2 + p_i p_j -  \tfrac{1}{2}p_i^2) \\
-% &= -\frac{1}{2} \sum_{i=1}^{n+1} \sum_{j=1}^{n+1} q_{ij} (p_i - p_j)^2.
-% \end{align*}
-% This is closely related to what is known as Sellings formula~\cite[Proposition 2.3.1]{Valentin2003_coverings_tilings_low_dimension}~\cite{Selling1874}.  Put $g = p - \onebf_S$.  It follows that
-% \[
-% \|Bp\|^2 - \|Bg\|^2 = \frac{1}{2} \sum_{i=1}^{n+1}\sum_{j=1}^{n+1}q_{ij}d_{ij},
-% \]
-% where $d_{ij} = (g_i - g_j)^2 - (p_i - p_j)^2$.  If both $i \notin S$ and $j \notin S$ then $p_i = g_i$, $p_j = g_j$ and so $d_{ij} = 0$.  Similarly if both $i \in S$ and $j \in S$ then 
-% \[
-% d_{ij} = (p_i-1 - p_j+1)^2 - (p_i - p_j)^2 = 0.
-% \]
-% If $i \in S$ and $j \notin S$ then
-% \[
-% d_{ij} = (p_i-1 - p_j)^2 - (p_i - p_j)^2= 1 + 2p_j - 2p_i,
-% \]
-% while if $i \notin S$ and $j \in S$ then
-% \[
-% d_{ij} = (p_i - p_j+1)^2 - (p_i - p_j)^2 = 1 + 2p_i - 2p_j.
-% \]
-% We thus have
-% \begin{align*}
-% \|Bp\|^2 - \|Bg\|^2 &= \frac{1}{2} \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_j - 2p_i) + \frac{1}{2} \sum_{i \notin S}\sum_{j \in S}q_{ij}(1 + 2p_i - 2p_j) \\
-% &= \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_j - 2p_i),
-% \end{align*}
-% where the last line holds because $q_{ij} = q_{ji}$.
-\end{proof}
-
-% \begin{lemma}\label{lem:decSellingsincanddec}
-% With the same symbols as in Lemma~\ref{lem:decSellings} and with $T$ a subset of the indices of $p$,
-% \[
-% \|Bp\|^2 - \|B(p + \onebf_S + \onebf_T)\|^2 = \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_i - 2p_j) + \sum_{i \in T}\sum_{j \notin T}q_{ij}(1 + 2p_i - 2p_j) + \sum_{i \in S}\sum_{j \in T} 2q_{ij}
-% \]
-% \end{lemma}
-
-Denote by $\min(p)$ and $\max(p)$ the minimum and maximum values obtained by the elements of the vector $p$ and define the function
-\[
-\rng(p) = \max(p) - \min(p).
-\] 
-%to return the difference between the maximum and minimum of $p$.  %For example, if $p = (2,-1,4) \in \ints^3$ then $\rng(p) = 4 - (-1)=5$.  We might also 
-%For example $\rng(2,-1,4)= 4 - (-1) = 5$.  
-Observe that $\rng(p)$ cannot be negative and that if $\rng(p) = 0$ then all of the elements of $p$ are equal.
-We define the function $\subrng(p)$ to return the largest subset, say $S$, of the indices of $p$ such that $\min\{p_i, i \in S\} - \max\{p_i, i \notin S\} \geq 2.$  If no such subset exists then $\subrng(p)$ is the empty set $\emptyset$.
-For example,
-\[
-\subrng(2,-1,4) = \{1,3\}, \qquad \subrng(2,1,3) = \emptyset, \quad  \subrng(1,3,1) = \{2\}.
-\]  
-To make the definition of $\subrng$ clear we give the following alternative and equivalent definition.
-Let $p \in \reals^n$ and let $\sigma$ be the permutation of the indices $\{1,\dots,n\}$ that puts the elements of $p$ in ascending order, that is
-\[
-p_{\sigma(1)} \leq p_{\sigma(2)} \leq \dots \leq p_{\sigma(n)}.
-\]  
-Let $T$ be the smallest integer from $\{2,\dots,n\}$ such that $p_{\sigma(T)} - p_{\sigma(T-1)} \geq 2$.  If no such integer $T$ exists then $\subrng(p) = \emptyset$.  Otherwise 
-\[
-\subrng(p) =  \{ \sigma(T), \sigma(T+1), \dots, \sigma(n) \}.
-\]
-The following straightforward property of $\subrng$ will be useful.
-
-\begin{proposition}\label{prop:subrrngsmall}
-Let $p \in \ints^{n+1}$.  If $\subrng(p) = \emptyset$ then $\rng(p) \leq n$.
-\end{proposition}
-\begin{proof}
-Let $\sigma$ be the permutation of the indices $\{1,\dots,n+1\}$ that puts the elements of $p$ in ascending order.
-Because $\subrng(p) = \emptyset$ and because the elements of $p$ are integers we have $p_{\sigma(i+1)} \leq p_{\sigma(i)} + 1$ for all $i=1,\dots,n$.
-It follows that
-\[
-p_{\sigma(n+1)} \leq p_{\sigma(n)} + 1 \leq p_{\sigma(n-1)} + 2 \leq \dots \leq p_{\sigma(1)} + n.
-\]
-and so $\rng(p) = p_{\sigma(n+1)} - p_{\sigma(1)} \leq n$.
-\end{proof}
-
-Finally we define the function
-\[
-\decrng(p) = p -  \onebf_{\subrng(p)}
-\]
-that decrements those elements from $p$ with indices from $\subrng(p)$.
-If $\subrng(p) = \emptyset$, then $\decrng(p) = p$, that is, $\decrng$ does not modify $p$.
-On the other hand, if $\subrng(p) \neq \emptyset$ then
-\[
-\rng\big(\decrng(p)\big) = \rng(p) - 1
-\]
-because $\subrng(p)$ contains all those indices $i$ such that $p_i = \max(p)$.
-By repeatedly applying $\decrng$ to a vector one eventually obtains a vector for which further application of $\decrng$ has no effect.
-For example,
-\begin{align*}
-\decrng(2,-1,4) &= (2,-1,4) - \onebf_{\subrng(2,-1,4)} = (2,-1,4) - \onebf_{\{1,3\}} = (1,-1,3) \\ 
-\decrng(1,-1,3) &= (1,-1,3) - \onebf_{\{1,3\}} = (0,-1,2) \\ 
-\decrng(0,-1,2) &= (0,-1,2) - \onebf_{\{3\}} = (0,-1,1) \\ 
-\decrng(0,-1,1) &= (0,-1,1) - \onebf_{\emptyset} = (0,-1,1).
-\end{align*}
-This will be a useful property so we state it formally in the following proposition.
-
-\begin{proposition} \label{lem:repeatappdecrange} Let $p \in \reals^{n+1}$ and define the infinite sequence $d_0,d_1,d_2,\dots$ of vectors according to $d_0=p$ and $d_{k+1} = \decrng(d_k)$.
-There is a finite integer $T$ such that $d_T=d_{T+1}=d_{T+2}=\dots$.
-\end{proposition}
-\begin{proof}
-Assume that no such $T$ exists.  Then $\decrng(d_k) \neq d_k$ for all positive integers $k$ and so 
-\[
-\rng(d_{k}) = \rng(d_{k-1}) - 1 = \rng(d_{k-2}) - 2 = \dots = \rng(p) - k.
-\]  
-Choosing $k > \rng(p)$ we have $\rng(d_{k}) < 0$ contradicting that $\rng(d_k)$ is nonegative.
-\end{proof}
-
-We are now ready to study properties of a closest lattice point in a lattice of Voronoi's first kind.
-
-\begin{lemma}\label{lem:decrngpreservesclosestpoints}
-If $v \in \ints^{n+1}$ such that $B(\floor{z} + v)$ is a closest lattice point to $y = Bz$, then $B\big(\floor{z} + \decrng(v)\big)$ is also a closest lattice point to $y$.
-\end{lemma}
-\begin{proof}
-The lemma is trivial if $\subrng(v) = \emptyset$ so that $\decrng(v) = v$.  It remains to prove the lemma when $\subrng(v) \neq \emptyset$.  In this case put $S = \subrng(v)$ and put 
-\[
-u = \decrng(v) = v - \onebf_S.
-\] 
-Let $\zeta = z - \floor{z}$ be the column vector containing the fractional parts of the elements of $z$.
-We have $\zeta - u = \zeta - v + \onebf_S$.
-Applying part~\ref{eq:lem:decSellingsinc} of Lemma~\ref{lem:decSellings} with $p = \zeta - v$ we obtain
-\begin{align}
-\|B(\zeta - v)\|^2 - \|B(\zeta - u)\|^2 &= \Phi(S, \zeta-v) \nonumber \\
-&= \sum_{i \in S}\sum_{j \notin S}q_{ij}\big(1 + 2(\zeta_i-\zeta_j) - 2(v_i - v_j)\big). \label{eq:sumsumBzBu}
-\end{align}
-Observe that $\zeta_i =  z_i - \floor{z_i} \in [0,1)$ for all $i=1,\dots,n+1$ and so $-1 < \zeta_i-\zeta_j < 1$ for all $i,j=1,\dots,n+1$.  Also, for $i \in S$ and $j\notin S$ we have 
-\[
-v_i - v_j \geq \min\{ v_i, i \in S\} - \max\{v_j, j \notin S\} \geq 2
-\]
-by definition of $\subrng(v) = S$.  Thus,
-\[
-1 + 2(\zeta_i-\zeta_j) - 2(v_i - v_j) < 1 + 2 - 4 = -1 < 0 \qquad \text{for $i \in S$ and $j \notin S$}.
-\]
-Substituting this inequality into~\eqref{eq:sumsumBzBu} and using that $q_{ij} \leq 0$ for $i \neq j$ (the obtuse condition~\eqref{eq:obtusecond}) we find that
-\[
-\|B(z - \floor{z} - v)\|^2 - \|B(z - \floor{z} - u)\|^2 \geq 0.
-\]
-It follows that $B(\floor{z} + u) = B\big(\floor{z} + \decrng(v)\big)$ is a closest lattice point to $y = Bz$ whenever $B(\floor{z} + v)$ is.
-\end{proof}
-
-\begin{lemma}\label{lem:roundzclose}
-There exists a closest lattice point to $y = Bz$ in the form $B(\floor{z} + v)$ where $v \in \ints^{n+1}$ with $\rng(v) \leq n$.
-\end{lemma}
-\begin{proof}
-Let $d_0 \in \ints^{n+1}$ be such that $B(\floor{z} + d_0)$ is a closest lattice point to $y$.
-Define the sequence of vectors $d_0,d_1,\dots$ from $\ints^{n+1}$ according to the recursion $d_{k+1} = \decrng(d_k)$.
-It follows from Lemma~\ref{lem:decrngpreservesclosestpoints} that $B(\floor{z} + d_{k})$ is a closest lattice point for all positive integers $k$.
-By Proposition~\ref{lem:repeatappdecrange} there is a finite $T$ such that
-\[
-d_{T+1}=d_T=\decrng(d_T).
-\]  
-Thus $\subrng(d_T) = \emptyset$ and $\rng(d_T) \leq n$ by Proposition~\ref{prop:subrrngsmall}.  The proof follows with $v = d_T$.   
-\end{proof}
-
-Let $\ell$ be a nonegative integer.
-We say that a lattice point $x$ is $\ell$-\emph{close} to $y$ if there exists a $v \in \ints^{n+1}$ with $\rng(v) = \ell$ such that $x + Bv$ is a closest lattice point to $y$.
-Lemma~\ref{lem:roundzclose} asserts that the lattice point $x_0 = B\floor{z}$ that initialises the iterative procedure~\eqref{eq:xseqfirsttype} is $K$-close to $y$ where $K \leq n$.
-From Lemma~\ref{lem:rngdecreases} stated below it will follow that if the lattice point $x_k$ obtained on the $k$th iteration of the procedure is $\ell$-close, then the lattice point $x_{k+1}$ obtained on the next iteration is $(\ell-1)$-close.
-Since $x_0$ is $K$-close it will then follow that after $K \leq n$ iterations the lattice point $x_K$ is $0$-close.
-At this stage it is guaranteed that $x_{K}$ itself is a closest lattice point to $y$.
-This is shown in the following lemma.
-
-\begin{lemma}\label{lem:rngzeroclosestpoint}
-If $x$ is a lattice point that is $0$-close to $y$, then $x$ is a closest lattice point to $y$.
-\end{lemma}
-\begin{proof}
-Because $x$ is $0$-close there exists a $v \in \ints^{n+1}$ with $\rng(v) = 0$ such that $x + Bv$ is a closest lattice point to $y$.
-Because $\rng(v) = 0$ all elements from $v$ are identical, that is, $v_1=v_2=\dots=v_{n+1}$.
-In this case $Bv = \sum_{i=1}^{n+1} v_n b_n = v_1\sum_{i=1}^{n+1}b_n = 0$
-as a result of the superbasis condition~\eqref{eq:superbasecond}.  Thus $x = x + Bv$ is a closest point to $y$. 
-\end{proof}
-
-Before giving the proof of Lemma~\ref{lem:rngdecreases} we require the following simple result.
-
-\begin{lemma}\label{eq:integergreaterless}
-Let $h \in \{0,1\}^{n+1}$ and $v \in \ints^{n+1}$.  Suppose that
-$h_i = 0$ whenever $v_i = \min(v)$ and that $h_i = 1$ whenever $v_i = \max(v)$.  Then
-\[
-h_i - h_j \leq v_i - v_j
-\]
-when either $v_i = \max(v)$ or $v_j = \min(v)$.
-\end{lemma}
-\begin{proof}
-If $v_i = \max(v)$ then $h_i = 1$ and we need only show that $1-h_j \leq \max(v) - v_j$ for all $j$.
-If $v_j = \max(v)$ then $h_j = 1$ and the results holds since $1 - h_j = 0 = \min(v) - v_j$.
-Otherwise if $v_j < \max(v)$ then $1-h_j \leq 1 \leq \max(v) - v_j$ because $\max(v)$ and $v_j$ are integers.
-
-Now, if $v_j = \min(v)$ then $h_j = 0$ and we need only show that $h_i \leq v_i - \min(v)$ for all $i$.
-If $v_i = \min(v)$ then $h_i = 0 = v_i - \min(v)$ and the results holds.
-Otherwise if $v_i > \min(v)$ then $h_i \leq 1 \leq v_i - \min(v)$ because $\min(v)$ and $v_i$ are integers.
-\end{proof}
-
-\begin{lemma}\label{lem:rngdecreases}
-Let $Bu$ with $u \in \ints^{n+1}$ be a lattice point that is $\ell$-close to $y = Bz$ where $\ell > 0$.  Let $g \in \{0,1\}^{n+1}$ be such that
-\begin{equation}\label{eq:qismin}
-\|B(z - u - g)\|^2 = \min_{t \in \{0,1\}^{n+1}}\|B(z - u - t)\|^2.
-\end{equation}
-The lattice point $B(u+g)$ is $(\ell-1)$-close to $y$.
-\end{lemma}
-\begin{proof}
-Because $Bu$ is $\ell$-close to $y$ there exists $v \in \ints^{n+1}$ with $\rng(v) = \ell$ such that $B(u+v)$ is a closest lattice point to $y$.  Define subsets of indices
-\[
-S = \{i \mid g_i = 0, v_i = \max(v) \}, \qquad T = \{i \mid g_i = 1, v_i = \min(v) \}, 
-\]
-and put $h = g + \onebf_S - \onebf_T \in \{0,1\}^{n+1}$ and $w = v - h$.
-Observe that $h_i = 0$ whenever $v_i = \min(v)$ and so $\min(w) = \min(v-h) = \min(v)$.
-Also, $h_i = 1$ whenever $v_i = \max(v)$ and so $\max(w) = \max(v-h) = \max(v) - 1$.
-Thus,
-\[
-\rng(w) = \max(v) - 1 - \min(v) = \rng(v) - 1 = \ell - 1.
-\]
-The lemma will follow if we show that $B(u+g+w)$ is a closest lattice point to $y$ since then the lattice point $B(u+g)$ with be $(\ell-1)$-close to $y$.
-The proof is by contradiction.
-Suppose $B(u+g+w)$ is not a closest point to $y$, that is, suppose
-\[
-\|B(z-u-g-w)\|^2 > |B(z-u-v)\|^2.
-\]
-Putting $p = z-u-v$ we have
-\[
-\|B(p+\onebf_S-\onebf_T)\|^2 > \|Bp\|^2.
-\]
-By part~\ref{eq:lem:decSellingsincdec} of Lemma~\ref{lem:decSellings} we obtain
-\begin{equation}\label{eq:BpBpdecincineq}
-\|Bp\|^2 - \|B(p+\onebf_S-\onebf_T)\|^2 = \Phi(S,p) + \Phi(\bar{T}, p) + \sum_{i\in S}\sum_{j \in S} q_{ij} < 0.
-\end{equation}
-
-As stated already $h_i = 0$ whenever $v_i = \min(v)$ and $h_i = 1$ whenever $v_i = \max(v)$.  It follows from Lemma~\ref{eq:integergreaterless} that
-\begin{equation}\label{eq:hijvijieq}
-h_{i} - h_j \leq v_{i} - v_j
-\end{equation}
-when either $v_i = \max(v)$ or $v_j = \min(v)$.  Since $v_i = \max(v)$ for $i \in S$ and $v_j = \min(v)$ for $j \in T$ the inequality~\eqref{eq:hijvijieq} holds when either $i \in S$ or $j \in T$.
-
-Put $r = z - u - h$.  By~\eqref{eq:hijvijieq} we have $r_i - r_j \geq p_{i} - p_{j}$ when either $i \in S$ or $j \in T$.  Now, since $q_{ij} \leq 0$ for $i \neq j$,
-\[
-\Phi(S,r) = \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2r_i - 2r_j) \leq \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_i - 2p_j) = \Phi(S,p),
-\]
-and
-\[
-\Phi(\bar{T},r) = \sum_{i \notin T}\sum_{j \in T}q_{ij}(1 + 2r_i - 2r_j) \leq \sum_{i \notin T}\sum_{j \in T}q_{ij}(1 + 2p_i - 2p_j) = \Phi(\bar{T},p).
-\]
-Using part~\ref{eq:lem:decSellingsincdec} of Lemma~\ref{lem:decSellings} again,
-\begin{align*}
-\|B(z-u-h)\|^2 - \|B(z-u-g)\|^2 &= \|Br\|^2 - \|B(r+\onebf_S - \onebf_T)\|^2 \\
-&= \Phi(S,r) + \Phi(\bar{T},r) + \sum_{i\in S}\sum_{j \in S} q_{ij} \\
-&\leq \Phi(S,p) + \Phi(\bar{T}, p) + \sum_{i\in S}\sum_{j \in S} q_{ij} < 0 
-\end{align*}
-as a result of~\eqref{eq:BpBpdecincineq}.  However, $h \in \{0,1\}^{n+1}$ and so this implies
-\[
-\|B(z-u-g)\|^2 > \|B(z-u-h)\|^2 \geq \min_{t \in \{0,1\}^{n+1}}\|B(z - u - t)\|^2
-\]
-contradicting~\eqref{eq:qismin}.
-Thus, our original supposition is false and $B(u+g+w)$ is a closest lattice point to $y$.
-Because $\rng(w) = \ell-1$ the lattice point $B(u+g)$ is $(\ell-1)$-close to $y$.
- \end{proof}
-
-The next theorem asserts that the iterative procedure~\eqref{eq:xseqfirsttype} converges to a closest lattice point in $K \leq n$ iterations.  This is the primary result of this section.
-
-\begin{theorem}
-Let $x_0,x_1,\dots$ be the sequence of lattice points given by the iterative procedure~\eqref{eq:xseqfirsttype}.  There exists $K \leq n$ such that $x_K$ is a closest lattice point to $y = Bz$.
-\end{theorem}
-\begin{proof}
-Let $x_k = B u_k$ be the lattice point obtained on the $k$th iteration of the procedure.
-Suppose that $x_k$ is $\ell$-close to $y=Bz$ with $\ell > 0$.
-The procedure computes $t_{k} \in \{0,1\}^{n+1}$ satisfying
-\[
-\|B(z - u_k - t_{k})\|^2 = \min_{t \in \{0,1\}^{n+1}}\|B(z - u_k - t)\|^2
-\]
-and puts $x_{k+1} = B(u_k + t_k)$.
-It follows from Lemma~\ref{lem:rngdecreases} that $x_{k+1}$ is $(\ell-1)$-close to $y$.
-By Lemma~\ref{lem:roundzclose} the lattice point that initialises the procedure $x_0 = B \floor{z}$ is $K$-close to $y$ where $K \leq n$.
-Thus, $x_1$ is $(K-1)$-close, $x_2$ is $(K-2)$-close, and so on until $x_K$ is $0$-close.
-That $x_K$ is a closest lattice point to $y$ follows from Lemma~\ref{lem:rngzeroclosestpoint}.
-\end{proof}
-
-\section{Computing a closest relevant vector}\label{sec:comp-clos-relev}
-
-In the previous section we showed that the iterative procedure~\eqref{eq:xseqfirsttype} results in a closest lattice point in at most $n$ iterations.
-It remains to show that each iteration of the procedure can be computed efficiently.
-Specifically, it remains to show that the minimisation over the set of binary vectors $\{0,1\}^{n+1}$ described in~\eqref{eq:pvecmin} can be computed efficiently.
-Putting $p = z - u_k$ in~\eqref{eq:pvecmin} we require an efficient method to compute a $t \in \{0,1\}^{n+1}$ such that the binary quadratic form
-\[
-\| B(p - t) \|^2 = \| \sum_{i=1}^{n+1} b_i (p_i - t_i) \|^2
-\]
-is minimised.  Expanding this quadratic form gives
-\[
-\| \sum_{i=1}^{n+1} b_i (p_i - t_i) \|^2 =  \sum_{i=1}^{n+1}\sum_{j=1}^{n+1} q_{ij}p_i p_j -  2\sum_{i=1}^{n+1}\sum_{j=1}^{n+1} q_{ij}p_j t_i + \sum_{i=1}^{n+1}\sum_{j=1}^{n+1} q_{ij} t_i t_j.
-\]
-The first sum above is independent of $t$ and can be ignored for the purpose of minimisation.  Letting $s_i = \sum_{j=1}^{n+1} q_{ij}p_j$, we can equivalently minimise the binary quadratic form
-\begin{equation}\label{eq:quadformnp}
-Q(t) = \sum_{i=1}^{n+1} s_i t_i + \sum_{i=1}^{n+1}\sum_{j=1}^{n+1} q_{ij} t_i t_j.
-\end{equation}
-We will show that a minimiser of $Q(t)$ can be found efficiently be computing a minimum cut in an undirected flow network.
-%\section{Quadratic $\{0,1\}$ programs and minimium cuts in graphs}
-%This section describes how a binary quadratic program can be mapped into the problem of computing a minimum cut in a weighted graph.  
-This technique has appeared previously~\cite{Picard_min_cuts_1974,Sankaran_solving_CDMA_mincut_1998,Ulukus_cdma_mincut_1998,Cormen2001} but we include the derivation here so that this paper is self contained.
-
-Let $G$ be an undirected graph with $n+3$ vertices $v_0, \dots, v_{n+2}$ contained in the set $V$ and edges $e_{ij}$ connecting $v_i$ to $v_j$.
-To each edge we assign a \emph{weight} $w_{ij} \in \reals$.
-The graph is undirected so the weights are symmetric, that is, $w_{ij} = w_{ji}$.
-By calling the vertex $v_0$ the \emph{source} and the vertex $v_{n+2}$ the \emph{sink} the graph $G$ is what is called a \emph{flow network}.
-The flow network is \emph{undirected} since the weights assigned to each edge are undirected.
-A \emph{cut} in the flow network $G$ is a subset $C \subset V$ of vertices with its complement $\bar{C} \subset V$ such that the source vertex $v_0 \in C$ and the sink vertex $v_{n+2} \in \bar{C}$.
-%The vertex $v_0$ is typically called the \emph{source} and the vertex $v_{n+1}$ is typically called the \emph{sink}.
-
-The weight of a cut is
-\[
-W(C,\bar{C}) = \sum_{i \in I} \sum_{j \in J} w_{ij}, 
-\]
-where $I = \{ i \mid v_i \in C\}$ and $J = \{j \mid v_j \in \bar{C}\}$.
-That is, $W(C,\bar{C})$ is the sum of the weights on the edges crossing from the vertices in $C$ to the vertices in $\bar{C}$.
-In what follows we will often drop the argument and write $W$ rather than $W(C,\bar{C})$.
-A \emph{minimum cut} is a $C$ and $\bar{C}$ that minimise the weight $W$.
-If all of the edge weights $w_{ij}$ for $i \neq j$ are nonnegative, a minimum cut can be computed in
-order $O(n^3)$ arithmetic operations~\cite{Cormen2001,Even_graph_algorithms_1979}.
-
-We require some properties of the weights $w_{ij}$ in relation to $W$.
-If the graph is allowed to contain loops, that is, edges from a vertex to itself, then the weight of these edges $w_{ii}$ have no effect on the weight of any cut.
-We may choose any values for the $w_{ii}$ without affecting $W$.
-We will find it convenient to set $w_{0,0} = w_{n+2,n+2} = 0$.
-The remaining $w_{ii}$ we shall specify shortly.
-The edge $e_{0,n+2}$ is in every cut.
-If a constant is added to the weight of this edge, that is, $w_{0,n+2}$ is replaced by $w_{0,n+2} + c$ then $W$ is replaced by $W + c$ for every $C$ and $\bar{C}$.
-In particular, the subsets $C$ and $\bar{C}$ corresponding to a minimum cut are not changed.
-We will find it convenient to choose $w_{0,n+2} = w_{n+2,0} = 0$.
-
-If vertex $v_i$ is in $C$ then edge $e_{i,n+2}$ contributes to the weight of the cut.
-If $v_i \notin C$, i.e., $v_i \in \bar{C}$, then edge $e_{0,i}$ contributes to the weight of the cut.
-So, either $e_{0,i}$ or $e_{i,n+2}$ \emph{but not both} contribute to every cut.
-If a constant, say $c$, is added to the weights of these edges, that is, $w_{0,i}$ and $w_{i,n+2}$ are replaced by $w_{0,i} + c$ and $w_{i,n+2} + c$, then $W$ is replaced by $W + c$ for every $C$ and $\bar{C}$.
-The $C$ and $\bar{C}$ corresponding to a minimum cut are unchanged.
-In this way, the minimum cut is only affected by the differences
-\[
-d_i = w_{i,n+2} - w_{0,i}
-\]
-for each $i$ and not the specific values of the weights $w_{i,n+2}$ and $w_{0,i}$.
-%For the purpose of computing the minimum cut it is necessary to choose the difference so that the weights $w_{0,i}$ and $w_{i,n+1}$ are nonnegative.
-
-% It is convenient to transform $W(C)$ as follows.
-Let $D = C \backslash 0$ and $\bar{D} = \bar{C} \backslash (n+1)$ where $C \backslash 0$ denotes the set $C$ with the element $0$ removed and $\bar{C} \backslash (n+1)$ denotes the set $\bar{C}$ with the element $n+1$ removed.
-Now we can write the weight as,
-% \[
-% W(C) =  \sum_{i \in D}w_{i,n+1} + \sum_{j \in \bar{d}}w_{0,j} +  \sum_{i \in D} \sum_{j \in \bar{D}} w_{ij}
-% \]
-% becuase $w_{0,0},w_{0,n+2}$ and $w_{n+2,n+2}$ are zero.
-
-We now show how $W(C,\bar{C})$ can be represented as a binary quadratic form.  Put $t_0 = 1$ and $t_{n+2} = 0$ and
-\[
-t_i = \begin{cases}
-1, & i \in C \\
-0, & i \in \bar{C}
-\end{cases}
-\]
-for $i = 1,2,\dots,n+1$.  Observe that
-\[
-t_i(1 - t_j) = \begin{cases}
-1, & i \in C, j \in \bar{C} \\
-0, & \text{otherwise}.
-\end{cases}
-\]
-The weight can now be written as
-\begin{align*}
-W(C,\bar{C}) = \sum_{i \in C} \sum_{j \in \bar{C}} w_{ij} = \sum_{i =0}^{n+2} \sum_{j =0}^{n+2} w_{ij} t_i (1 - t_j) = F(t),
-\end{align*}
-say.  Finding a minimum cut is equivalent to finding the binary vector $t = (t_1, \dots, t_{n+1})$ that minimises $F(t)$.  Write,
-\[
-F(t) =  \sum_{i=0}^{n+2} \sum_{j =0}^{n+2} w_{ij}t_i - \sum_{i=0}^{n+2} \sum_{j =0}^{n+2} w_{ij} t_it_j.
-\]
-Letting $k_i = \sum_{j =0}^{n+2} w_{ij}$, and using that $t_0 = 1$ and $t_{n+2} = 0$,
-\[
-F(t) = \sum_{i=0}^{n+1}k_it_i  - w_{00} - \sum_{i=1}^{n+1} w_{i0} t_i - \sum_{j=1}^{n+1} w_{0j} t_j - \sum_{i=1}^{n+1} \sum_{j =1}^{n+1} w_{ij} t_it_j.
-\]
-Because $w_{00} = 0$ and $w_{ij} = w_{ji}$ we have
-\[
-F(t) = k_0 + \sum_{i=1}^{n+1}( k_i  - 2 w_{i0}) t_i  - \sum_{i=1}^{n+1} \sum_{j =1}^{n+1} w_{ij} t_it_j.
-\]
-The constant term $k_0$ is unimportant for the purpose of minimisation so finding a minimum cut is equivalent to minimising the binary quadratic form
-\[
-\sum_{i=1}^{n+1}g_i t_i  - \sum_{i=1}^{n+1} \sum_{j =1}^{n+1} w_{ij} t_it_j,
-\]
-where $g_i = k_i  - 2 w_{i0} = d_i + \sum_{j=1}^{n+1} w_{ij}$.
-It only remains to observe the equivalence of this quadratic form and $Q(t)$ from~\eqref{eq:quadformnp} when the weights are assigned to satisfy,
-\begin{align*}
-&q_{ij} = - w_{ij} \qquad i,j = 1,\dots,n+1 \\
-&s_i = g_i = d_i + \sum_{j=1}^{n+1} w_{ij}.
-\end{align*}
-Because the $q_{ij}$ are nonpositive for $i \neq j$ the weights $w_{ij}$ are nonnegative for all $i \neq j$ with $i,j = 1,\dots,n+1$.
-As discussed the value of the weights $w_{ii}$ have no effect on the weight of any cut $W$ so setting $q_{ii} = - w_{ii}$ for  $i = 1,\dots,n+1$ is of no consequence.
-Finally the weights $w_{i,n+2}$ and $w_{0,i}$ can be chosen so that both are nonnegative and
-\[
-w_{i,n+2} - w_{0,i} = d_i = s_i + \sum_{j=1}^{n+1} q_{ij} = s_i
-\]  
-because $\sum_{j=1}^{n+1} q_{ij} = 0$ due to the superbase condition~\eqref{eq:superbasecond}.
-That is, we choose $w_{i,n+2} = s_i$ and $w_{0,i} = 0$ when $s_i \geq 0$ and $w_{i,n+2}=0$ and $w_{0,i} = -s_i$ when $s_i < 0$.
-With these choices, all the weights $w_{ij}$ for $i \neq j$ are nonnegative.
-A minimiser of $Q(t)$, and correspondingly a solution of~\eqref{eq:pvecmin} can be computed in $O(n^3)$ operations by computing a minimum cut in the undirected flow network $G$ assigned with these nonnegative weights~\cite{Picard_min_cuts_1974,Sankaran_solving_CDMA_mincut_1998,Ulukus_cdma_mincut_1998,Cormen2001}.
-%The closest lattice point is then computed using~\eqref{eq:compnp}.
-
-\section{Discussion}\label{sec:discussion}
-
-The closest lattice point problem has a number of applications, for example, channel coding and data quantisation~\cite{Conway1983VoronoiCodes,Conway1982VoronoiRegions,Conway1982FastQuantDec,Erex2004_lattice_decoding,Erez2005}.
-A significant hurdle in the practical application of lattices as codes or as quantisers is that computing a closest lattice point is computationally difficult in general~\cite{micciancio_hardness_2001}.
-The best known general purpose algorithms require a number of operations of order $O(2^{2n})$~\cite{Micciancio09adeterministic}.
-In this paper we have focused on the class of lattices of Voronoi's first kind.
-We have shown that computing a closest point in a lattice of Voronoi's first kind can be achieved in a comparatively modest number of operations of order $O(n^4)$.
-Besides being of theoretical interest, the algorithm has potential for practical application.
-
-A question of immediate interest to communications engineers is: do there exist lattices of Voronoi's first kind that produce good codes or good quantisers?
-Since lattices that produce good codes and quantisers often also describe dense~\emph{sphere packings}~\cite{SPLAG}, a related question is: do there exist lattices of Voronoi's first kind that produce dense sphere packings?
-These questions do not appear to have trivial answers.
-The questions have heightened importance due to the algorithm described in this paper.
-
-It is straightforward to construct an `arbitrary' lattice of Voronoi's first kind.
-One approach is to construct the $n+1$ by $n+1$ symmetric matrix $Q = B^\prime B$ with elements $Q_{ij} = q_{ij} = \dotprod{b_{i}}{b_j}$ given by the Selling parameters.
-Choose the off diagonal entries of $Q$ to be nonpositive with $q_{ij}=q_{ji}$ and set the diagonal elements $q_{ii} = -\sum_{j=1}^{n+1} q_{ij}$.
-%We then have $Q\onebf = \zerobf$ and so $\onebf$ is in the null space of $Q$.
-The matrix $Q$ is diagonally dominant, that is, $\abs{q_{ii}} \geq \sum_{j=1}^{n+1} \abs{q_{ij}}$, and so $Q$ is positive semidefinite.
-A rank deficient Cholesky decomposition~\cite{Higham90analysisof} can now be used to recover a matrix $B$ such that $B^\prime B  = Q$.
-The columns of $B$ are vectors of the obtuse superbasis.
-
-A number applications such as phase unwrapping~\cite{Teunissen_GPS_1995,Hassibi_GPS_1998}, single frequency estimation~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}, and related signal processing problems~\cite{McKilliam2007,Clarkson2007,McKilliam2009IndentifiabliltyAliasingPolyphase,Quinn_sparse_noisy_SSP_2012} also require computing a closest lattice point.
-In these applications the particular lattice arises from the signal processing problem under consideration.
-If that lattice happens to be of Voronoi's first kind then our algorithm can be used.
-An example where this occurs is the problem of computing the \emph{sample intrinsic mean} in circular statistics~\cite{McKilliam_mean_dir_est_sq_arc_length2010}.
-In this particular problem the lattice $A_n^*$ is involved.
-A fast closest point algorithm requiring only $O(n)$ operations exists for $A_n^*$~\cite{McKilliam2009CoxeterLattices,McKilliam2008b} and so the algorithm described in this paper is not needed in this particular case.
-However, there many exist other signal processing problems where lattices of Voronoi's first kind arise.
-
-A final remark is that our algorithm assumes that the obtuse superbasis is known in advance.
-It is known that all lattices of dimension less than 4 are of Voronoi's first kind and an algorithm exists to recover the obtuse superbasis in this case~\cite{SPLAG}.
-Lattices of dimension larger than 4 need not be of Voronoi's first kind.
-An interesting question is: given a lattice, is it possible to efficiently decide whether it is of Voronoi's first kind?
-A related question is: is it possible to efficiently find an obtuse superbasis if it exists?
-
-%In some applications, for example, the detection of communications signals involving multiple anntanea~\cite{Ryan2008,Wubben_2011} the lattice basis is not known in advance.
-It is then of interest whether the
-
-\section{Conclusion}
-
-The paper describes an algorithm to compute a closest lattice point in a lattice of Voronoi's first kind when the obtuse superbasis is known~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.
-The algorithm requires $O(n^4)$ operations where $n$ is the dimension of the lattice.
-The algorithm iteratively computes a series of relevant vectors that converges to a closest lattice point after at most $n$ terms.
-Each relevant vector in the series can be efficiently computed in $O(n^3)$ operations by computing a minimum cut in an undirected flow network.
-The algorithm has potential application in communications engineering problems such as coding and quantisation.
-An interesting problem for future research is to find lattices of Voronoi's first kind that produce good codes, good quantisers, or dense sphere packings~\cite{SPLAG,Conway1982VoronoiRegions}.
-
-%\small
-\bibliography{bib}
+\bibliography{../bib}
 
 \end{document}
-
-%%% Local Variables: 
-%%% mode: latex
-%%% TeX-master: t
-%%% End:

### 2014-06-02 findsuperbase/findsuperbase.tex
minor

@@
 
+Notes:
+\begin{itemize}
+\item Need definition of `strict' and `lax' relevant vectors. 
+\item How many lax relevant vectors can there be (presumably this comes the integer lattice, so $3^{n}-1$?)
+\item Are vectors from a superbase always strict?  Can probably show this using the arguments in Conway and Sloane's paper.
+\end{itemize}
+
+Our algorithm begins by enumerating a superset $R$ of the set of relevant vectors $\relevant(\Lambda)$.
+This can be achieved using the algorithm of Micciancio and Voulgaris~\cite{Micciancio09adeterministic} or the algorithm described by Agrell~et.~al.~\cite[Sec.~VI.C]{Agrell2002}.
+In many cases these algorithms will return precisely the set of relevant vectors and so $R = \relevant(\Lambda)$, but for what follows it is sufficient if $R \supseteq \relevant(\Lambda)$.
+
+If $\Lambda$ is of first kind then, by Theorem~\ref{thm:revvecssuperbase}, the vectors $b_1,\dots,b_{n+1}$ from the obtuse superbase are contained in $R$.
+A straightforward procedure to determine whether the lattice is of first kind consists of enumerating all $\binom{\abs{R}}{n+1}$ combinations of $n+1$ vectors from $R$ and checking whether any combination forms a superbase.
+That is, let $v_1,\dots,v_{n+1}$ be a set of $n+1$ distinct vectors from $R$.
+If both $v_1+\dots + v_{n+1} = 0$ and $\dotprod{v_{i}}{v_{j}} \leq 0$ for all $i \neq j$ then the lattice is of first kind with superbase is $v_1,\dots,v_{n+1}$.
+If no such combination of vectors exists in $R$, then the lattice is not of first kind.
+
+In general there are as many as $2^{n+1}-2$ relevant vectors (BLERG strict, lax etc?)
+
 \bibliography{../bib}

### 2014-06-03 findsuperbase/code/montecarlo.scala
added code

@@ -1,0 +1,24 @@
+/**
+* Run montecarlo.
+*/
+import pubsim.distributions.RealRandomVariable
+import pubsim.VectorFunctions.randomMatrix
+
+val iters = 10
+val Ns = List(1,2,3,4,5)
+
+val starttime = (new java.util.Date).getTime
+
+Ns.map{ n =>
+
+}
+
+
+
+val runtime = (new java.util.Date).getTime - starttime
+println("Finished in " + (runtime/1000.0) + " seconds.\n") 
+
+
+
+
+

### 2014-06-03 findsuperbase/code/montecarlo.scala
added plot.  Need to fill in code and generate data

@@ -9,16 +9,18 @@ val Ns = List(1,2,3,4,5)
 
 val starttime = (new java.util.Date).getTime
 
-Ns.map{ n =>
+val datalist = Ns.map{ n =>
+val p = 1.0
+(n,p,0.1)
+}.toList
 
+val file = new java.io.FileWriter("data/testname")
+file.write("n" + "\t" + "proportion" + "\t" + "stddev" + "\n")
+datalist.foreach{ v => 
+  val (n,p,d) = v
+  file.write(n + "\t" + p + "\t" + d + "\n")
 }
-
-
+file.close
 
 val runtime = (new java.util.Date).getTime - starttime
 println("Finished in " + (runtime/1000.0) + " seconds.\n") 
-
-
-
-
-

### 2014-06-04 findsuperbase/code/montecarlo.scala
minor

@@ -2,6 +2,9 @@
 * Run montecarlo.
 */
 import pubsim.distributions.RealRandomVariable
+import pubsim.distributions.UniformNoise
+import pubsim.distributions.GaussianNoise
+import pubsim.lattices.firstkind.FirstKindCheck
 import pubsim.VectorFunctions.randomMatrix
 
 val iters = 10

### 2014-06-04 findsuperbase/code/montecarlo.scala
added java libs

@@ -1,6 +1,6 @@
 /**
-* Run montecarlo.
-*/
+  * Run montecarlo.
+  */
 import pubsim.distributions.RealRandomVariable
 import pubsim.distributions.UniformNoise
 import pubsim.distributions.GaussianNoise
@@ -13,17 +13,23 @@ val Ns = List(1,2,3,4,5)
 val starttime = (new java.util.Date).getTime
 
 val datalist = Ns.map{ n =>
-val p = 1.0
-(n,p,0.1)
+
+  //run all the MonteCarlo iterations
+  iters.map { i =>
+    val B = randomMatrix(n,n,UniformNoise.constructFromMinMax(0,1))
+  }.toList
+
+  val p = 1.0
+  (n,p,0.1)
 }.toList
 
 val file = new java.io.FileWriter("data/testname")
 file.write("n" + "\t" + "proportion" + "\t" + "stddev" + "\n")
-datalist.foreach{ v => 
+datalist.foreach{ v =>
   val (n,p,d) = v
   file.write(n + "\t" + p + "\t" + d + "\n")
 }
 file.close
 
 val runtime = (new java.util.Date).getTime - starttime
-println("Finished in " + (runtime/1000.0) + " seconds.\n") 
+println("Finished in " + (runtime/1000.0) + " seconds.\n")

### 2014-06-04 findsuperbase/findsuperbase.tex
added code
added plot.  Need to fill in code and generate data
added lib and data

--- 2014-06-02 findsuperbase/findsuperbase.tex
@@
 
+\usepackage{tikz}
+\usepackage{pgfplots}
+\usetikzlibrary{calc}
+
 %opening
-\title{Finding a obtuse superbase}
+\title{The number of lattices of Voronoi's first kind}
 \author{Robby McKilliam}
@@
 
-This document describes an algorithm to determine whether a lattice, specified by a basis matrix, is of Voronoi's first kind.
+% This document describes an algorithm to determine whether a lattice, specified by a basis matrix, is of Voronoi's first kind.
 The algorithm returns an obtuse superbase for the lattice if the lattice is of Vornoi's first kind.
 
-An $n$-dimensional lattice $\Lambda$ is said to be of \emph{Voronoi's first kind} if it has what is called an \emph{obtuse superbasis}~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.
+% An $n$-dimensional lattice $\Lambda$ is said to be of \emph{Voronoi's first kind} if it has what is called an \emph{obtuse superbasis}~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.
 That is, there exists a set of $n+1$ vectors $b_1,\dots,b_{n+1}$ such that $b_1,\dots,b_n$ are a basis for $\Lambda$,
-\begin{equation}\label{eq:superbasecond}
-b_1 + b_2 \dots + b_{n+1} = 0
-\end{equation}
-(the \emph{superbasis} condition), and the inner products satisfy
-\begin{equation}\label{eq:obtusecond}
-q_{ij} = b_i \cdot b_j \leq 0, \qquad \text{for} \qquad i,j = 1,\dots,n+1, i \neq j
-\end{equation}
-(the \emph{obtuse} condition).
+% \begin{equation}\label{eq:superbasecond}
+% b_1 + b_2 \dots + b_{n+1} = 0
+% \end{equation}
+% (the \emph{superbasis} condition), and the inner products satisfy
+% \begin{equation}\label{eq:obtusecond}
+% q_{ij} = b_i \cdot b_j \leq 0, \qquad \text{for} \qquad i,j = 1,\dots,n+1, i \neq j
+% \end{equation}
+% (the \emph{obtuse} condition).
 The $q_{ij}$ are called the \emph{Selling parameters}~\cite{Selling1874}.
@@
 
-\begin{theorem} \label{thm:revvecssuperbase} (Conway and Sloane~\cite[Theorem~3]{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases})
-The relevant vectors of $\Lambda$ are of the form,
+% \begin{theorem} \label{thm:revvecssuperbase} (Conway and Sloane~\cite[Theorem~3]{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases})
+% The relevant vectors of $\Lambda$ are of the form,
+% \[
+% \sum_{i \in I} b_i
+% \]
+% where $I$ is a strict subset of $\{1, 2, \dots, n+1\}$ that is not empty, i.e., $I \subset \{1, 2, \dots, n+1\}$ and $I \neq \emptyset$.
+% \end{theorem}
+
+% Notes:
+% \begin{itemize}
+% \item Need definition of `strict' and `lax' relevant vectors. 
+% \item How many lax relevant vectors can there be (presumably this comes the integer lattice, so $3^{n}-1$?)
+% \item Are vectors from a superbase always strict?  Can probably show this using the arguments in Conway and Sloane's paper.
+% \end{itemize}
+
+% Our algorithm begins by enumerating a superset $R$ of the set of relevant vectors $\relevant(\Lambda)$.
+This can be achieved using the algorithm of Micciancio and Voulgaris~\cite{Micciancio09adeterministic} or the algorithm described by Agrell~et.~al.~\cite[Sec.~VI.C]{Agrell2002}.
+In many cases these algorithms will return precisely the set of relevant vectors and so $R = \relevant(\Lambda)$, but for what follows it is sufficient if $R \supseteq \relevant(\Lambda)$.
+
+% If $\Lambda$ is of first kind then, by Theorem~\ref{thm:revvecssuperbase}, the vectors $b_1,\dots,b_{n+1}$ from the obtuse superbase are contained in $R$.
+A straightforward procedure to determine whether the lattice is of first kind consists of enumerating all $\binom{\abs{R}}{n+1}$ combinations of $n+1$ vectors from $R$ and checking whether any combination forms a superbase.
+That is, let $v_1,\dots,v_{n+1}$ be a set of $n+1$ distinct vectors from $R$.
+If both $v_1+\dots + v_{n+1} = 0$ and $\dotprod{v_{i}}{v_{j}} \leq 0$ for all $i \neq j$ then the lattice is of first kind with superbase is $v_1,\dots,v_{n+1}$.
+If no such combination of vectors exists in $R$, then the lattice is not of first kind.
+
+% In general there are as many as $2^{n+1}-2$ relevant vectors (BLERG strict, lax etc?)
+
+The plot shows the estimated (after 500 trials) probability of obtaining a lattice of Voronoi's first kind for $n=1,2,3,4$.
+Three different methods for generating the $n \times n$ basis matrix $B$ are considered:
+\begin{enumerate}
+\item entries drawn independently and uniformly in the interval $[0,1]$, 
+\item entries drawn independently and Gaussian with zero mean and variance one, and 
+\item the $i,j$th entry satisfies
 \[
-\sum_{i \in I} b_i
+B_{i,j} = \begin{cases}
+R_{i,j} & 1 \leq i,j \leq h \\
+-C_{i-h,j} & h < i \leq n, 1 \leq j \leq h \\
+C_{i,j-h} & h < j \leq n, 1 \leq i \leq h \\
+R_{i-h,j-h} & h < i,j \leq n,
+\end{cases}
 \]
-where $I$ is a strict subset of $\{1, 2, \dots, n+1\}$ that is not empty, i.e., $I \subset \{1, 2, \dots, n+1\}$ and $I \neq \emptyset$.
-\end{theorem}
+where $h = \ceil{n/2}$ is the smallest integer greater than or equal to $n/2$ and $R_{i,j}$ and $C_{i,j}$ for $i,j = 1,\dots,h$ are independently Gaussian distributed with zero mean and variance one. 
+\end{enumerate}
+When $n$ is even, the third case corresponds with the model commonly used for MIMO channels.
+In this case $R$ and $C$ are $\tfrac{n}{2} \times \tfrac{n}{2}$ matrices containing the real and imaginary parts of the channel coefficients.
 
-\bibliography{../bib}
+As expected, all lattices generated in dimensions 1, 2, and 3 were found to be of Voronoi's first kind.
+When $n=4$ roughtly 25\% of the lattices generated were of Voronoi's first kind.
+This was the case for all 3 basis generation methods.
+These simulations suggest that the lattices of Voronoi's first kind do not lie within a space of measure zero (in some sense!) in dimension 4.
+
+I was too oppimistic when I thought I would be able to run this with $n=6$.
+This would take too long with the existing code.
+There are a number of optimisations that can be made that might make running simulations with $n=6$ possible.
+
+\begin{figure}[tp]
+\centering 
+  \begin{tikzpicture}
+    \begin{axis}[font=\footnotesize,clip mode=individual,height=7cm,width=9cm,legend style={draw=none,fill=none,legend pos=south west,font=\footnotesize}]    
+      \addplot[mark=*,color=red,only marks,/pgfplots/error bars/.cd,
+        x dir=none,
+        y dir=both,
+        y explicit] table[x=n,y=proportion,y error=stddev] {code/data/uniform};
+      \addplot[mark=*,color=blue,only marks,/pgfplots/error bars/.cd,
+        x dir=none,
+        y dir=both,
+        y explicit] table[x=n,y=proportion,y error=stddev] {code/data/gaussian};
+      \addplot[mark=x,color=green,only marks,/pgfplots/error bars/.cd,
+        x dir=none,
+        y dir=both,
+        y explicit] table[x=n,y=proportion,y error=stddev] {code/data/mimo};
+\legend{Uniform, Gaussian}
+    \end{axis}
+  \end{tikzpicture}
+\caption{Estimated probability of generating a lattices of Voronoi's first kind for different basis distributions and dimension $n=1,2,3,4$.}
+\end{figure}
+
+%\bibliography{../bib}
 

### 2014-06-04 findsuperbase/code/montecarlo.scala

@@ -6,30 +6,49 @@ import pubsim.distributions.UniformNoise
 import pubsim.distributions.GaussianNoise
 import pubsim.lattices.firstkind.FirstKindCheck
 import pubsim.VectorFunctions.randomMatrix
+import pubsim.VectorFunctions.randomMIMObasis
+import Jama.Matrix
+import scala.math.sqrt
 
-val iters = 10
-val Ns = List(1,2,3,4,5)
+val z = 1.96 //corresponds with 90% confidence intervals
+val L = 500 //the number of Monte Carlo iterations
+val Ns = List(1,2,3,4)
 
-val starttime = (new java.util.Date).getTime
+runsim( n=>randomMatrix(n,n,UniformNoise.constructFromMinMax(0,1)), "uniform")
+runsim( n=>randomMatrix(n,n,new GaussianNoise(0,1)), "gaussian")
+runsim( n=>randomMIMObasis(n), "mimo")
 
-val datalist = Ns.map{ n =>
+def runsim( basisgen : Int => Matrix, name : String ) {
 
-  //run all the MonteCarlo iterations
-  iters.map { i =>
-    val B = randomMatrix(n,n,UniformNoise.constructFromMinMax(0,1))
+  val starttime = (new java.util.Date).getTime
+
+  print(name + " ")
+
+  val datalist = Ns.map{ n =>
+
+    //run all the MonteCarlo iterations
+    val bdata = (1 to L).map { i =>
+      val B = basisgen(n)
+      val fk = new FirstKindCheck(B)
+      if(fk.isFirstKind) 1 else 0
+    }.toList
+
+    val phat = bdata.foldLeft(0.0) { (sum, v) => sum + v } / L
+    val ci = z * sqrt(phat*(1-phat)/ L )
+
+    print(".")
+    (n,phat,ci)
   }.toList
 
-  val p = 1.0
-  (n,p,0.1)
-}.toList
+  val file = new java.io.FileWriter("data/" + name)
+  file.write("n" + "\t" + "proportion" + "\t" + "stddev" + "\n")
+  datalist.foreach{ v =>
+    val (n,p,d) = v
+    file.write(n + "\t" + p + "\t" + d + "\n")
+  }
+  file.close
 
-val file = new java.io.FileWriter("data/testname")
-file.write("n" + "\t" + "proportion" + "\t" + "stddev" + "\n")
-datalist.foreach{ v =>
-  val (n,p,d) = v
[…]

### 2014-06-04 findsuperbase/findsuperbase.tex
interestly the mimo matrices don't ever appear to be of first kind!

@@
 \end{enumerate}
-When $n$ is even, the third case corresponds with the model commonly used for MIMO channels.
-In this case $R$ and $C$ are $\tfrac{n}{2} \times \tfrac{n}{2}$ matrices containing the real and imaginary parts of the channel coefficients.
+When $n$ is even, the third case corresponds with the model commonly used for MIMO channels, that is, the basis matrix is of the form
+\[
+B = \left(\begin{array}{cc}
+R & -C \\
+C & R
+\end{array}\right)
+\]
+where $R$ and $C$ are $\tfrac{n}{2} \times \tfrac{n}{2}$ matrices containing the real and imaginary parts of the channel coefficients.
 
 As expected, all lattices generated in dimensions 1, 2, and 3 were found to be of Voronoi's first kind.
-When $n=4$ roughtly 25\% of the lattices generated were of Voronoi's first kind.
-This was the case for all 3 basis generation methods.
-These simulations suggest that the lattices of Voronoi's first kind do not lie within a space of measure zero (in some sense!) in dimension 4.
+When $n=4$ roughtly 25\% of the lattices generated by methods 1 and 2 were of Voronoi's first kind, but none of the 500 bases generated by method 3 were of first kind.
+I find this very interesting!
+The simulations suggest that the lattices of Voronoi's first kind do not lie within a space of measure zero with respect to the measures induced by the first two generation methods, but that they do lie on a space of measure zero with respect to the lattices that occur when converting complex valued MIMO channel matrices into real valued lattices bases.
+This might suggest that lattices of first kind will (atleast not directly) be useful for MIMO.
 
@@
         y explicit] table[x=n,y=proportion,y error=stddev] {code/data/uniform};
-      \addplot[mark=*,color=blue,only marks,/pgfplots/error bars/.cd,
+      \addplot[mark=square,color=blue,only marks,/pgfplots/error bars/.cd,
         x dir=none,
@@
         y explicit] table[x=n,y=proportion,y error=stddev] {code/data/mimo};
-\legend{Uniform, Gaussian}
+\legend{Uniform, Gaussian, MIMO}
     \end{axis}

### 2014-06-04 findsuperbase/findsuperbase.tex
added minor

@@
 As expected, all lattices generated in dimensions 1, 2, and 3 were found to be of Voronoi's first kind.
-When $n=4$ roughtly 25\% of the lattices generated by methods 1 and 2 were of Voronoi's first kind, but none of the 500 bases generated by method 3 were of first kind.
+When $n=4$ roughly 25\% of the lattices generated by methods 1 and 2 were of Voronoi's first kind, but none of the 500 bases generated by method 3 were of first kind.
 I find this very interesting!
-The simulations suggest that the lattices of Voronoi's first kind do not lie within a space of measure zero with respect to the measures induced by the first two generation methods, but that they do lie on a space of measure zero with respect to the lattices that occur when converting complex valued MIMO channel matrices into real valued lattices bases.
-This might suggest that lattices of first kind will (atleast not directly) be useful for MIMO.
+The simulations suggest that the lattices of Voronoi's first kind do not lie within a space of measure zero with respect to the measures induced by the first two generation methods, but that they do lie on a space of measure zero with respect to the lattices that occur when converting complex valued MIMO channel matrices into real valued lattice bases.
+This might suggest that lattices of first kind will not be useful for MIMO (at least not directly).
 
-I was too oppimistic when I thought I would be able to run this with $n=6$.
+It would be interesting to see if these conclusion hold in higher dimension.
+I was too optimistic when I thought I would be able to run this with $n=6$.
 This would take too long with the existing code.

### 2014-06-06 findsuperbase/findsuperbase.tex
added neew build and ran new simulations.  This code probably doesn't work for special lattices with a combination of strict and lax relevant vectors

--- 2014-06-04 findsuperbase/findsuperbase.tex
@@
 
-The plot shows the estimated (after 500 trials) probability of obtaining a lattice of Voronoi's first kind for $n=1,2,3,4$.
+The plot shows the estimated (after 5000 trials) probability of obtaining a lattice of Voronoi's first kind for $n=1,2,3,4,5,6$.
 Three different methods for generating the $n \times n$ basis matrix $B$ are considered:
@@
 As expected, all lattices generated in dimensions 1, 2, and 3 were found to be of Voronoi's first kind.
-When $n=4$ roughtly 25\% of the lattices generated by methods 1 and 2 were of Voronoi's first kind, but none of the 500 bases generated by method 3 were of first kind.
-I find this very interesting!
-The simulations suggest that the lattices of Voronoi's first kind do not lie within a space of measure zero with respect to the measures induced by the first two generation methods, but that they do lie on a space of measure zero with respect to the lattices that occur when converting complex valued MIMO channel matrices into real valued lattices bases.
-This might suggest that lattices of first kind will (atleast not directly) be useful for MIMO.
+When $n=4$ roughly 25\% of the lattices generated by methods 1 and 2 were of Voronoi's first kind, but none of the 5000 bases generated by method 3 were of first kind for $n=4$.
+When $n=5$ approximately 3\% of the lattices generated were of Voronoi's first kind for all 3 generation methods.
+When $n=6$ less than 1\% of the lattices generated by methods 1 and 2 where of first kind and approximately 2\% of the lattices generated by methods 3 where of first kind.
 
-I was too oppimistic when I thought I would be able to run this with $n=6$.
+It is interesting that none of the lattices corresponding with MIMO channel matrices are of first kind in dimension 4, but there are some in dimension 6.
+
+%I find this very interesting!
+The simulations suggest that the lattices of Voronoi's first kind do not lie within a space of measure zero with respect to the measures induced by the first two generation methods, but that they do lie on a space of measure zero with respect to the lattices that occur when converting complex valued MIMO channel matrices into real valued lattice bases.
+This might suggest that lattices of first kind will not be useful for MIMO (at least not directly).
+
+%It would be interesting to see if these conclusion hold in higher dimension.
+I was too optimistic when I thought I would be able to run this with $n=6$.
 This would take too long with the existing code.
@@
   \begin{tikzpicture}
-    \begin{axis}[font=\footnotesize,clip mode=individual,height=7cm,width=9cm,legend style={draw=none,fill=none,legend pos=south west,font=\footnotesize}]    
+    \begin{axis}[ymode=log,font=\footnotesize,clip mode=individual,height=7cm,width=9cm,legend style={draw=none,fill=none,legend pos=south west,font=\footnotesize}]    
       \addplot[mark=*,color=red,only marks,/pgfplots/error bars/.cd,

### 2014-06-06 findsuperbase/code/montecarlo.scala

@@ -10,9 +10,9 @@ import pubsim.VectorFunctions.randomMIMObasis
 import Jama.Matrix
 import scala.math.sqrt
 
-val z = 1.96 //corresponds with 90% confidence intervals
-val L = 500 //the number of Monte Carlo iterations
-val Ns = List(1,2,3,4)
+val z = 1.96 //corresponds with 95% confidence intervals
+val L = 5000 //the number of Monte Carlo iterations
+val Ns = List(1,2,3,4,5,6)
 
 runsim( n=>randomMatrix(n,n,UniformNoise.constructFromMinMax(0,1)), "uniform")
 runsim( n=>randomMatrix(n,n,new GaussianNoise(0,1)), "gaussian")
@@ -27,7 +27,7 @@ def runsim( basisgen : Int => Matrix, name : String ) {
   val datalist = Ns.map{ n =>
 
     //run all the MonteCarlo iterations
-    val bdata = (1 to L).map { i =>
+    val bdata = (1 to L).par.map { i =>
       val B = basisgen(n)
       val fk = new FirstKindCheck(B)
       if(fk.isFirstKind) 1 else 0

### 2014-06-06 findsuperbase/findsuperbase.tex
added start of java  to matlab code

@@
 When $n=4$ roughly 25\% of the lattices generated by methods 1 and 2 were of Voronoi's first kind, but none of the 5000 bases generated by method 3 were of first kind for $n=4$.
-When $n=5$ approximately 3\% of the lattices generated were of Voronoi's first kind for all 3 generation methods.
-When $n=6$ less than 1\% of the lattices generated by methods 1 and 2 where of first kind and approximately 2\% of the lattices generated by methods 3 where of first kind.
+When $n=5$ approximately 2\% of the lattices generated were of first kind for all 3 generation methods.
+When $n=6$ less than 1\% of the lattices generated by methods 1 and 2 where of first kind and approximately 1\% of the lattices generated by method 3 where of first kind.
 
@@
   \end{tikzpicture}
-\caption{Estimated probability of generating a lattices of Voronoi's first kind for different basis distributions and dimension $n=1,2,3,4$.}
+\caption{Estimated probability of generating a lattices of Voronoi's first kind for different basis distributions and dimension $n=1,2,3,4,5,6$.}
 \end{figure}

### 2014-06-06 findsuperbase/code/matlab/to_java_matrix.m

@@ -1,0 +1,9 @@
+function out = to_java_matrix(B)
+
+m = size(B,1)
+n = size(B,2)
+
+strObj = javaObjectEDT('Jama.Matrix',m,n);
+
+
+end

### 2014-06-06 findsuperbase/code/montecarlo.m

@@ -1,0 +1,16 @@
+%approximates the proportion of lattices of Voronoi's first kind
+
+if( usejava('jvm') == 0 )
+    disp(['It appears you do not have a java virtual machine installed. ' ...
+          ' This is not going to work.']);
+end
+javaclasspath('./lib/Jama-1.0.3.jar'); %add java libraries in lib folder to matlab
+addpath('./matlab') %add the matlab functions we will use
+
+mat = [1,2,3;4,5,6];
+jmat = to_java_matrix(mat)
+
+z = 1.96; %corresponds with 95% confidence interval
+L = 100; %number of Monte-Carlo trials
+Ns = [1,2,3,4]; %the dimensions we will run simulations for
+

### 2014-06-06 findsuperbase/code/montecarlo.scala

@@ -49,6 +49,6 @@ def runsim( basisgen : Int => Matrix, name : String ) {
   file.close
 
   val runtime = (new java.util.Date).getTime - starttime
-  println(" finished in " + (runtime/1000.0) + " seconds.\n")
+  println(" finished in " + (runtime/1000.0) + " seconds.")
 
 }

### 2014-06-06 findsuperbase/code/matlab/runmontecarlo.m
added matlab stuff

@@ -1,0 +1,27 @@
+function out = runmontecarlo( matrixgenerator, Ns, L, z, name ) 
+%runmontecarlo run L Monte-Carlo simulations to estimate the proportion
+%of lattices of Voronoi's first kind in a given dimension. For each
+%simulation the lattice basis a generated by the function 
+%argument matrixgenerator.
+
+    tic
+    
+    disp(name);
+    
+    ps = zeros(size(Ns));
+    for n = Ns
+        for k = 1:L 
+            B = to_java_matrix(matrixgenerator(n));
+            fck = javaObjectEDT('pubsim.lattices.firstkind.FirstKindCheck',B);
+            if(fck.isFirstKind)
+                ps(n) = ps(n) + 1;
+            end
+        end
+        disp('.');
+    end
+
+    out = ps/L;
+
+    toc
+
+end

### 2014-06-06 findsuperbase/code/matlab/to_java_matrix.m

@@ -1,9 +1,7 @@
 function out = to_java_matrix(B)
-
-m = size(B,1)
-n = size(B,2)
-
-strObj = javaObjectEDT('Jama.Matrix',m,n);
-
+% to_java_matrix converts the matrix B to a Jama.Matrix
+%from the java matrix library
+    
+out = javaObjectEDT('Jama.Matrix',B);
 
 end

### 2014-06-06 findsuperbase/code/montecarlo.m

@@ -1,16 +1,32 @@
 %approximates the proportion of lattices of Voronoi's first kind
 
+clear all;
+
 if( usejava('jvm') == 0 )
     disp(['It appears you do not have a java virtual machine installed. ' ...
           ' This is not going to work.']);
 end
-javaclasspath('./lib/Jama-1.0.3.jar'); %add java libraries in lib folder to matlab
+
+%add java libraries in lib folder to matlab
+javaclasspath('./lib/PubSim.jar');
+javaaddpath('./lib/JGraphT.jar');
+javaaddpath('./lib/Jama-1.0.3.jar');
+javaaddpath('./lib/bignums.jar');
+javaaddpath('./lib/flanagan.jar');
+javaaddpath('./lib/junit-4.11.jar');
+javaaddpath('./lib/RngPack.jar');
+javaaddpath('./lib/colt.jar');
+javaaddpath('./lib/jtransforms-2.4.jar');
+
 addpath('./matlab') %add the matlab functions we will use
 
-mat = [1,2,3;4,5,6];
-jmat = to_java_matrix(mat)
+javaclasspath('-dynamic')
 
 z = 1.96; %corresponds with 95% confidence interval
-L = 100; %number of Monte-Carlo trials
+L = 10; %number of Monte-Carlo trials
 Ns = [1,2,3,4]; %the dimensions we will run simulations for
+
+%start simulations for uniform, gaussian and MIMO matrices
+runmontecarlo( @(n) rand(n), Ns, L, z, 'uniform' ); 
+runmontecarlo( @(n) randn(n), Ns, L, z, 'gaussian' ); 
 

### 2014-06-07 findsuperbase/code/matlab/runmontecarlo.m
added function for generating mimo matrices

@@ -1,12 +1,10 @@
-function out = runmontecarlo( matrixgenerator, Ns, L, z, name ) 
+function out = runmontecarlo( matrixgenerator, Ns, L ) 
 %runmontecarlo run L Monte-Carlo simulations to estimate the proportion
 %of lattices of Voronoi's first kind in a given dimension. For each
 %simulation the lattice basis a generated by the function 
 %argument matrixgenerator.
 
     tic
-    
-    disp(name);
     
     ps = zeros(size(Ns));
     for n = Ns

### 2014-06-07 findsuperbase/code/montecarlo.m

@@ -2,11 +2,6 @@
 
 clear all;
 
-if( usejava('jvm') == 0 )
-    disp(['It appears you do not have a java virtual machine installed. ' ...
-          ' This is not going to work.']);
-end
-
 %add java libraries in lib folder to matlab
 javaclasspath('./lib/PubSim.jar');
 javaaddpath('./lib/JGraphT.jar');
@@ -17,16 +12,23 @@ javaaddpath('./lib/junit-4.11.jar');
 javaaddpath('./lib/RngPack.jar');
 javaaddpath('./lib/colt.jar');
 javaaddpath('./lib/jtransforms-2.4.jar');
-
 addpath('./matlab') %add the matlab functions we will use
 
-javaclasspath('-dynamic')
+%javaclasspath('-dynamic')
 
-z = 1.96; %corresponds with 95% confidence interval
-L = 10; %number of Monte-Carlo trials
-Ns = [1,2,3,4]; %the dimensions we will run simulations for
+L = 500; %number of Monte-Carlo trials
+Ns = [1,2,3,4,5]; %the dimensions we will run simulations for
 
 %start simulations for uniform, gaussian and MIMO matrices
-runmontecarlo( @(n) rand(n), Ns, L, z, 'uniform' ); 
-runmontecarlo( @(n) randn(n), Ns, L, z, 'gaussian' ); 
+disp('uniform');
+puniform = runmontecarlo( @(n) rand(n), Ns, L ); 
+disp('gaussian');
+pgaussian = runmontecarlo( @(n) randn(n), Ns, L ); 
+disp('mimo');
+pmimo = runmontecarlo( @(n) randomMIMOmatrix(n), Ns, L ); 
+
+
+plot(Ns, puniform, '.r'); hold on;
+plot(Ns, pgaussian, 'sb'); hold on;
+plot(Ns, pmimo, 'xg'); hold on;
 

### 2014-07-05 reviews/response.tex
added response

\documentclass[a4paper,10pt]{article}
%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
%\documentclass[journal]{IEEEtran}

\usepackage{../mathbf-abbrevs}
\input{../defs}

\usepackage{zref-xr,zref-user}
\zexternaldocument*{paper}

\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}

\usepackage{amsmath,amsfonts,amssymb, amsthm, bm}

%\usepackage[square,comma,numbers,sort&compress]{natbib}

\usepackage{color}
\newcommand{\comment}[1]{\textcolor{red}{#1}}

\newcommand{\sgn}{\operatorname{sgn}}
\newcommand{\sinc}{\operatorname{sinc}}
\newcommand{\rect}[1]{\operatorname{rect}\left(#1\right)}

%opening
\title{Finding a closest point in a lattice of Voronoi's first kind}
\author{Robby~G.~McKilliam, Alex Grant and I. Vaughan L. Clarkson
%\thanks{}
}

\renewcommand{\theenumii}{(\alph{enumii})}
\renewcommand{\labelenumii}{\theenumii}

\begin{document}

\maketitle

I would like to thank reviwers for there thoughtful comments.  %The authors feel the paper has improved as a result.

\section*{Reviewer 1}\label{sec:reviewer-1}

\begin{enumerate}
\item\textbf{Comment:} \label{comment:shortersec2}
	\\\textbf{Response:}
%\comment{look at shortening sec 2}

\end{enumerate}

\bibstyle{../siam}
\bibliography{bib}

\end{document}

### 2014-07-06 reviews/response.tex
started splitting up reviewer comments

@@
 \begin{enumerate}
-\item\textbf{Comment:} \label{comment:shortersec2}
+
+\item\textbf{Comment:} 
+From the perspective of quadratic forms, the quadratic forms induced by obtuse superbases are in one to one correspondance with graph Laplacians (this is not
+explicitly stated in the paper for some reason).
+Indeed, the quadratic form $\|Bx\|^2$ is exactly $x^T L_G x$, where $L_G$ is the Laplacian the weighted graph $G$ on vertices $\{ 1, .
+.
+.
+, n + 1 \}$ , with edges weights $w_{ij} = -b_i \cdot b_j \geq 0, i = j$ (which they call the Selling parameters).
+More precisely:
+\[
+\|Bx\|_2^2 = x^T L_G x = \sum_{ \{i,j\} \in E[G] } w_{ij} (x_{i} - x_{j} )^2.
+\]
+\\ \textbf{Response:}
+
+\item\textbf{Comment:} 
+BLERG: NEED TO CHECK REFS MATHS ETC To present the result from the perspective of Computational Complexity (also not given in the
+paper), this paper can be viewed as giving a polynomial time algorithm for the Closest Vector
+Problem with Preprocessing (CVPP) over a special class of lattices. In the preprocessing model,
+we are allowed to compute any polynomial amount of advice about the lattice (i.e. a good ba-
+sis, or many short vectors in the lattice, etc.) before answering any CVP queries, where the re-
+sources needed to compute this advice are not counted in the runtime of the algorithm. In its
+gap decisional version, where we need only decide whether a target is at distance $\leq d$ or $> \alpha d$
+(for some approximation factor $α \geq 1$), it is was shown [1, 4] to be NP-Hard for any constant $\alpha = 2^{\log^{1-\epsilon}n}$, and to not be polynomial time solvable for $\alpha = 2 \log n$ , for any fixed $\epsilon > 0$, under the assumption  that $NP \subsetneq \text{RTIME}(2^{\log^{O(1/\epsilon)}n})$ (quasi-polynomial time).
+For its approximate search version (i.e.
+find a lattice vector whose distance to the target is some bounded factor away from optimal), an
+$O( n^{1.5} )$ approximation algorithm was (implicitly) given in [5], which was only recently improved to $O( n \log n )$ in [3].
 \\\textbf{Response:}
-%\comment{look at shortening sec 2}
+
+\item\textbf{Comment:} 
+BLERG CHECK REFS Hence, from the standpoint of CVPP, the algorithm in this paper gives essentially the first
+exact polynomial time algorithm over a reasonably large and natural class of lattices, i.e. those
+of Voronoi’s first kind. Lastly, removing the need for preprocessing for these lattices, at least for
+solving SVP without being given an obtuse superbasis, would require a substantial breakthrough
+in the theory of lattice problems (not stated in paper). Indeed, by a simple reduction this would
+allow us to decide whether a lattice (given any of its bases) is a rotation of $Z_n$ (which is clearly
+of Voronoi’s first kind). This is a long standing open problem in the theory of lattices, which has
+only recently been solved in certain very special cases [6].
+\\\textbf{Response:}
+
+\end{enumerate}
+
+\section*{Reviewer 2}\label{sec:reviewer-2}
+
+\begin{enumerate}
+\item\textbf{Comment:}  
+page 1, last line "The closest lattice point problem is known to be NP-hard under certain condition ...": I found this paragraph very confusing, and somehow inaccurate.
+What "certain conditions"?
+I think a better way to describe the sitation regaring the NP-hardness of CVP in relation to this paper is the following.
+CVP is known to be NP-hard even to approximate within almost polynomial factor (see 35) when the lattice is arbitrary and not known in advance.
+Interestingly (see 35 and 17) there are families of lattices when CVP remain NP-hard even when given the best possible basis (or any hard-to-compute side information about the lattice).
+This version of the problem is called CVP with preprocessing (CVPP), and it is particularly relevant to this work, where it is shown that CVPP can be solved exactly and in polynomial time for the class of lattices of Voronoi first kind, when given as auxiliary information a special basis (namely, an obtuse superbasis.)
+\\\textbf{Response:}
+
+\item\textbf{Comment:}  
+Additional comments about the above paragraph: one author is missing in the reference for 12.
+Of [1,23,40] only 23 achieves $n^{O(n)}$ running time.
+(I think the algorithms in 40 take $2^{O(n^2)}$.) There is a journal version of 37 in SIAM J.
+Computing, and the paper is by two authors (not just Micciancio).
+\\\textbf{Response:}
+
+\item\textbf{Comment:}  
+page 2, second paragraph: Aren't the results of 3,29 and 30 subsumed by 34?
+(The authors sure know better than me.
+A a short clarification may direct the interested reader to the best paper to read.) Also, CVP is always solvable in polynomial time when the dimension is fixed.
+So, it is unclear the significance of CVP algorithms for Leech lattice (in fixed dimension 24) in the context of this discussion.
+When saying "if the obtuse superbasis is known", you may want to cast this requirement in the context of CVPP. (The obtuse superbasis is the advice provided by the preprocessing.) 
+\\\textbf{Response:}
+
+\item\textbf{Comment:}  
+page 2,bottom, displayed equation "$v \cdot x \leq x \cdot x$". Maybe you mean strict inequality 
+\\\textbf{Response:}
+
+\item\textbf{Comment:}  
+page 3, and maybe elsewhere "short vector in a lattice" should probably be "shortest nonzero vector in a lattice". 
+\\\textbf{Response:}
+
+\item\textbf{Comment:}  
+page 3, definition of ceiling(r) as smallest integer strctly larger than 3 seems unconventional.
+The standard definition is not strict.
+Write floor(r)+1 instead, which is consistent with the definition of floor(r) given later in the paper.
+\\\textbf{Response:}
+
+\item\textbf{Comment:}  
+page 3, and maybe elsewhere "short vector in a lattice" should probably be "shortest nonzero vector in a lattice". 
+\\\textbf{Response:}
+
+\item\textbf{Comment:}  
+page 3, proof of Prop.
+2.1.
+Your definition of "half-open voronoi cell" is not a proper definition, and also seems incorrect.
+E.g., take the integer lattice Z, then only one of the corners (rathern than half) of the corners of the voronoi cell should be included if you want a tiling.
+\\\textbf{Response:}
+
+\item\textbf{Comment:}  
+page 3, prop 2.2, again definition of ceiling(r) is non-standard. Use floor(r)+1 instead. 
+\\\textbf{Response:}
+
+\item\textbf{Comment:}  
+page 3, section 3. 
+The process described in (3.1) is almost identical to those in [45,37], so those papers should probably be referenced at the very beginning, and a comparison provided.
+As far as I remember, the difference is that 37 picks the v that minimizes the voronoi norm (i.e., the norm defined by the voronoi cell of the origin) instead of the euclidean norm.
+On the other hand, 45 uses the euclidean norm, but selects an arbitrary relevant vector that reduces the euclidean norm, rather than the one that reduces it the most.
+In this respect, your process is a special case of the one described in 45, and termination (your propositions 3.1 and 3.2) is already proved in 45.
+\\\textbf{Response:}
+
+\item\textbf{Comment:}  
+page 5, as before, you should probably mention also Voulgaris when citing 37. 
+\\\textbf{Response:}
+
+\item\textbf{Comment:}  
+page 5, last line, "dimension less than 4", does it include 4 or not? 
+\\\textbf{Response:}
+
+\item\textbf{Comment:}  
+page 5, as before, you should probably mention also Voulgaris when citing 37. 
+\\\textbf{Response:}
+
+\item\textbf{Comment:}  
+page 6, Theorem 4.1 (this follows from [ 9, Theorem 2+3], not just theorem 3. 
+Also, make the theorem statement self contained by adding "Let L be a lattice of voronoi first kind with obtuse superbasis b1..bn+1". 
+\\\textbf{Response:}
+
+\item\textbf{Comment:}  
+page 8, lemma 5.1. I think you forgot a factor 2 preceding the double summations over S and T (at end of lemma statement, and also in the proof.) 
+\\\textbf{Response:}
+
+\item\textbf{Comment:}  
+page 11, "with be" -> "will be" Displayed equation before "Putting p=...", A | should be || 
+(5.7) again, missing factor 2 before double summation, and second set should be T, not S. 
+\\\textbf{Response:}
+
+\item\textbf{Comment:}  
+page 12, first displayed equation, usual problem: missing factor 2, and summation over S\&S instead of S\&T both on line 2 and 3. 
+\\\textbf{Response:}
+
+\item\textbf{Comment:}  
+page 12, last line, definition of $s_i$, missing (-2) factor. 
+\\\textbf{Response:}
+
+\item\textbf{Comment:}  
+page 13, after (6.1), "be computing" -> "by computing" 
+\\\textbf{Response:}
+
+\item\textbf{Comment:}  
+page 14, "have no effect" -> "has no effect" 
+\\\textbf{Response:}
+
+\item\textbf{Comment:}  
+page 15, definition of $Q$, should the summation defininig $q_{ii}$ be restricted to $j\neq i$? 
+\\\textbf{Response:}
+
+\item\textbf{Comment:}  
+page 15, "there many" -> "there may" 
+\\\textbf{Response:}
+
+\item\textbf{Comment:}  
+References: 
+- 12: missing author (R. Raz) 
+- 18: "A V" -> "A. V." etc, 
+- 37: journal version in SIAM J. computing, 42(3):1364-1391 (2013) 
+\\\textbf{Response:}
+
+\item\textbf{Comment:}  
+Some additional references on CVPP you may find interesting 
+(cite at your own discretion):
+
+- O. Regev (2004) Improved Inapproximability of Lattice and Coding Problems With Preprocessing. IEEE Transactions on Information Theory 50(9): 2031-2037 
+- "Hardness of Approximating the Closest Vector PRoblem with Pre-Processing", M. Alekhnovich, S. Khot, G. Kindler, N. Vishnoi, Computational Complexity 20(4):741-753, 2011.
+
+(Improve NP-hardness result of [35,17] to any constant approximation factor)
+
+- D. Aharonov and Oded Regev (2005). Lattice problems in NP ∩ co-NP. J. ACM 52, 749-765.
+
+(Gives polynomial time solution to approximate the distance estimation version of CVPP within $\gamma=\sqrt(n \log n)$ approximation factor, i.e., for any lattice there is an appropriate advice (more complex than just a good basis) that allows to approximate in polynomial time the distance of any target point to the lattice within gamma, but without actually finding a nearby lattice point.)
+
+- On the Closest Vector Problem with a Distance Guarantee 
+D. Dadush, O. Regev, N. Stephens-Davidowitz, Computational Complexity Conference 2014, to appear (next week!)
+
+I am not sure, as the paper has not appeared yet, but I suspect it is the same as the paper on Dadush webpage with title "On BDD and CVPP". 
+See link 
+http://cs.nyu.edu/~dadush/papers/cvpp.pdf 
+and read abstract for more information.
+\\\textbf{Response:}
 

### 2014-07-06 reviews/response.tex
it's difficult to get all the pieces of this in my head.

@@
 
-\begin{enumerate}
+\begin{itemize}
 
@@
 \[
-\|Bx\|_2^2 = x^T L_G x = \sum_{ \{i,j\} \in E[G] } w_{ij} (x_{i} - x_{j} )^2.
+\|B\xbf\|_2^2 = \xbf^T L_G \xbf = \sum_{ \{i,j\} \in E[G] } w_{ij} (\xbf_{i} - \xbf_{j} )^2.
 \]
@@
 only recently been solved in certain very special cases [6].
+\\\textbf{Response:}
+
+\item\textbf{Comment:} 
+This paper makes a solid contribution to the theory of lattice problems. I think the results here are
+well-motivated (I encourage the authors to include some of the motivation mentioned in the sum-
+mary), and point to the possibility that there might be large and useful classes of lattices where
+solving CVP is “easy”.
+\\\textbf{Response:}
+
+\item\textbf{Comment:} 
+On a technical level, I found the main convergence proof to be interesting, as I’ve rarely seen
+combinatorial arguments successfully applied to lattice problems. On the other hand, while I’ve
+convinced myself the authors have the “right” convergence proof, I found its presentation to be
+rather awkward and unintuitive. I highly recommend that the authors rephrase the proof using
+the language of Laplacians, which would make things more clear (see the additional comments
+below for some technical suggestions).
+\\\textbf{Response:}
+
+\end{itemize}
+
+We now respond to the reviewers numbered comments.
+
+\begin{enumerate}
+
+\item\textbf{Comment:} 
+BLERG REFS.  I believe Proposition 2.1 and 2.2 appear essentially verbatim in [8]. Same for Propositions
+3.1 (which is essentially by definition) and 3.2, they appear in [9]. I don’t think it adds any
+readibility/anything of value to reprove them here.
+\\\textbf{Response:}
+
+\newcommand{\rng}{\operatorname{rng}}
+\newcommand{\subrng}{\operatorname{subr}}
+\newcommand{\decrng}{\operatorname{decrng}}
+
+\item\textbf{Comment:} 
+Almost all the inequalities in the paper can be derived from the following simple obser-
+vation.
+For and $n$ vertex weighted graph $G$ with non-negative edge weights, and vectors, $\xbf,\ybf \in \reals^n$, if there exists a permutation $\pi$ of $[n]$ such that $\xbf_{\pi[1]} \leq \dots \leq \xbf_{\pi[n]}$ and $\ybf_{\pi[1]} \leq \dots \leq \ybf_{\pi[n]}$ then
+\[
+\xbf^T L_G \ybf = \sum_{ \{i,j\} \in E[G] } \wbf_{ij} (\xbf_{i} - \xbf_{j}) (\ybf_{i} - \ybf_{j}) \geq 0.
+\]
+Furthermore, if $\xbf_{\pi[1]} \leq \dots \leq \xbf_{\pi[n]}$ and $\ybf_{\pi[1]} \geq \dots \geq \ybf_{\pi[n]}$ then the above inequality if reversed, i.e $\xbf^T L_G \ybf \leq 0$.
+
+I believe using the above will make many of your proofs more transparent.
+For example, if you redefine $\subrng(\xbf)$ to return the largest set $S$ such that $\min_{i\in S} \xbf_i - \max_{j \notin S} \xbf_j \geq 1$, then noice that if $S \neq \emptyset$ then there is a permutation that puts both $\xbf - 1_S$ and $1_S$ in non-decreasing order.
+In particular this implies that
+\[
+\xbf L_G \xbf = (\xbf - 1_S)^T L_G (\xbf - 1_S) + 2(\xbf - 1_S)^T L_G 1_S + 1_S^T L_G 1_S \geq (\xbf - 1_S)^T L_G (\xbf - 1_S),
+\]
+i.e. $\xbf$ can be made ``shorter'' by squeezing its components closer together.
+
+Furthermore, one can use this to give a two line proof of Theorem 4.1.
+This theorem I actually do think make sense to reprove here, because modifications of this proof are what drives the convergence bound.
+\\\textbf{Response:}
+BLERG.  I think there's an error in the inquality above.
+
+\item\textbf{Comment:} 
+Lemma 5.1: there is a typo here.  The last identity should read
+\[
+\|B\pbf\|^2 - \|B(\pbf + 1_S - 1_T)\|^2 = \Phi(S,\pbf) + \Phi(\bar{T}, \pbf) + 2\sum_{i \in S}\sum_{j\in S} q_{ij}.
+\]
+Notice that factor 2 in front of the last term.
+There are also two related typos on the top of page 12 where $\sum_{i \in S}\sum_{j\in S} q_{ij}$ (this appears twice on difference lines) should be $2\sum_{i \in S}\sum_{j\in T} q_{ij}$.
+\\\textbf{Response:}
+
+\item\textbf{Comment:} 
+Lemma 5.7: $h$ is ill-defined if $\rng(\vbf) = 0$.  You should explicitly state that $\rng(\vbf) \geq 1$.
 \\\textbf{Response:}

### 2014-07-07 reviews/response.tex
added more to responses.  I think this is starting to make sense.  I wonder if I can improved Lemma 5.8?

@@
 \usepackage{zref-xr,zref-user}
-\zexternaldocument*{paper}
+\zexternaldocument*{../paper}
 
@@
 \item\textbf{Comment:}  
-page 2,bottom, displayed equation "$v \cdot x \leq x \cdot x$". Maybe you mean strict inequality 
+page 2,bottom, displayed equation "$v \cdot x \leq x \cdot x$". Maybe you mean strict inequality.
 \\\textbf{Response:}
@@
 \\\textbf{Response:}
+We have instead introduced the term `short vector' in the introduction.
+The following sentence is included in the first paragraph of the introduction: ``This is called the \emph{shortest vector problem} and a solution is called a~\emph{short vector}.''
 
@@
 \\\textbf{Response:}
-
-\item\textbf{Comment:}  
-page 3, and maybe elsewhere "short vector in a lattice" should probably be "shortest nonzero vector in a lattice". 
-\\\textbf{Response:}
+Done.
 
@@
 2.1.
-Your definition of "half-open voronoi cell" is not a proper definition, and also seems incorrect.
-E.g., take the integer lattice Z, then only one of the corners (rathern than half) of the corners of the voronoi cell should be included if you want a tiling.
+Your definition of ``half-open voronoi cell'' is not a proper definition, and also seems incorrect.
+E.g., take the integer lattice $\ints^n$, then only one of the corners (rather than half) of the corners of the Voronoi cell should be included if you want a tiling.
 \\\textbf{Response:}
+The reviewer is correct.
+We only require that some modification of the boundary of the closed Voronoi cell results in a tiling.
+The proof of Proposition~\zref{eq:latticepointsinvorcvell} has been modified accordingly.
 
@@
 \\\textbf{Response:}
+Done.
 
@@
 \\\textbf{Response:}
+It does not include 4.
+We have modified the text here to read ``less than or equal to $3$''.
+We have also modified similar text in the last paragraph of Section~\zref{sec:discussion} to read ``It is known that all lattices of dimension less than or equal to 3 are of Voronoi's first kind and an algorithm exists to recover the obtuse superbasis in this case~\cite{SPLAG}.
+Lattices of dimension larger than 3 need not be of Voronoi's first kind.''
 

### 2014-07-07 paper.tex
added more to responses.  I think this is starting to make sense.  I wonder if I can improved Lemma 5.8?

@@
 \]
-where $\Lambda \backslash  \{\zerobf\}$ denotes the set of lattice points not equal to the origin $\zerobf$.  This is called the \emph{shortest vector problem}.
+where $\Lambda \backslash  \{\zerobf\}$ denotes the set of lattice points not equal to the origin $\zerobf$.
+This is called the \emph{shortest vector problem} and a solution is called a~\emph{short vector}.
 
@@
 These algorithms all require a number of operations that grows as $O(n^{O(n)})$ where $n$ is the dimension of the lattice.
-Recently, Micciancio~\cite{Micciancio09adeterministic} described a solution for the closest lattice point problem that requires a number of operations that grows as $O(2^{2n})$.
+Recently, Micciancio and Voulgaris~\cite{MicciancioVoulgaris_deterministic_jv_2013} described a solution for the closest lattice point problem that requires a number of operations that grows as $O(2^{2n})$.
 This single exponential growth in complexity is the best known.
 
-Although the problem is NP-hard in general, fast algorithms are known for specific highly regular lattices, such as the root lattices $A_n$ and $D_n$, their dual lattices $A_n^*$ and $D_n^*$, the integer lattice $\ints^{n}$, and the Leech lattice~\cite[Chap.
-4]{SPLAG}\cite{Conway1982FastQuantDec,Conway1986SoftDecLeechGolay, Clarkson1999:Anstar, McKilliam2008, McKilliam2008b, McKilliam2009CoxeterLattices,Vardy1993_leech_lattice_MLD}.
+Although the problem is NP-hard in general, fast algorithms are known for specific highly regular lattices, such as the root lattices $A_n$ and $D_n$, their dual lattices $A_n^*$ and $D_n^*$, the integer lattice $\ints^{n}$, and the Leech lattice~\cite[Chap.~4]{SPLAG}\cite{Conway1982FastQuantDec,Conway1986SoftDecLeechGolay, Clarkson1999:Anstar, McKilliam2008, McKilliam2008b, McKilliam2009CoxeterLattices,Vardy1993_leech_lattice_MLD}.
 In this paper we consider a particular class of lattices, those of \emph{Voronoi's first kind}~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases,Valentin2003_coverings_tilings_low_dimension,Voronoi1908_main_paper}.
@@
 Let $\Lambda \subset \reals^{m}$ be an $n$-dimensional lattice.
-For $r\in \reals$ let $\ceil{r}$ denote the smallest integer strictly larger than $r$.
+For $r\in \reals$ let $\floor{r}$ denote the largest integer less than or equal to $r$.
 Let $t \in \reals^m$.
-The number of lattice points inside the scaled and translated Voronoi cell $r \vor(\Lambda) + t$ is at most $\ceil{r}^n$.
+The number of lattice points inside the scaled and translated Voronoi cell $r \vor(\Lambda) + t$ is at most $(\floor{r}+1)^n$.
 \end{proposition}
 \begin{proof}
-It is convenient to modify the boundary of the Voronoi cell so that it tessellates $\reals^m$ under translations by $\Lambda$.
-With this aim we let $V \subset \vor(\Lambda)$ contain all those points from the interior of $\vor(\Lambda)$, but have opposing faces open and closed.
-That is, if $x \in V$ is on the boundary of $V$ then $-x \notin V$.
-With this definition it can be asserted that $V$ tessellates $\reals^m$ under translations by $\Lambda$.
+%It is convenient to modify the boundary of the Voronoi cell so that it tessellates $\reals^m$ under translations by $\Lambda$.  With this aim we 
+Let $V \subset \vor(\Lambda)$ contain all those points from the interior of the closed Voronoi cell $\vor(\Lambda)$, but with boundaries defined so that $V$ tessellates $\reals^m$ under translations by $\Lambda$.
 That is, $\reals^{m} = \cup_{x \in \Lambda}(V + x)$ and the intersection $(V + x)\cap(V+y)$ is empty for distinct lattice points $x$ and $y$.
-Now, for positive integer $k$, the scaled and translated cell $kV + t$ contains precisely one coset representative for each element of the quotient group $\Lambda/k\Lambda$~\cite[Sec.~2.4]{McKilliam2010thesis}.
+For positive integer $k$, the scaled and translated cell $kV + t$ contains precisely one coset representative for each element of the quotient group $\Lambda/k\Lambda$~\cite[Sec.~2.4]{McKilliam2010thesis}.
 There are $k^n$ coset representatives.
-Thus, the number of lattice points inside $r \vor(\Lambda) + t \subset \ceil{r}V + t$ is at most $\ceil{r}^n$.
+Thus, the number of lattice points inside $r \vor(\Lambda) + t \subset (\floor{r}+1)V + t$ is at most $(\floor{r}+1)^n$.
 \end{proof}
@@
 Let $S$ be an $m$-dimensional hypersphere of radius $r$ centered at $t \in \reals^{m}$.
-The number of lattice points from $\Lambda$ in the sphere $S$ is at most $\ceil{r/\rho}^{n}$.
+The number of lattice points from $\Lambda$ in the sphere $S$ is at most $(\floor{r/\rho}+1)^{n}$.
 %where $\ceil{\cdot}$ denotes the smallest integer strictly larger than its argument.
@@
 That is, $S \subset r/\rho \vor(\Lambda) + t$.
-The proof follows because the number of lattice points in $r/\rho \vor(\Lambda) + t$ is at most $\ceil{r/\rho}^n$ by Proposition~\ref{eq:latticepointsinvorcvell}.
+The proof follows because the number of lattice points in $r/\rho \vor(\Lambda) + t$ is at most $(\floor{r/\rho}+1)^n$ by Proposition~\ref{eq:latticepointsinvorcvell}.
 \end{proof}
@@
 and so $x_0,x_1,\dots$ is an infinite sequence of distinct (due to the strict inequality) lattice points  all contained inside an $n$-dimensional hypersphere of radius $r = \|y - x_0\|$ centered at $y$.
-This is a contradiction because, if $\rho$ is the packing radius of the lattice, then less than $\ceil{r/\rho}^n$ lattice points lie inside this sphere by Proposition~\ref{eq:latticepointsinsphere}.
+This is a contradiction because, if $\rho$ is the packing radius of the lattice, then less than $(\floor{r/\rho}+1)^n$ lattice points lie inside this sphere by Proposition~\ref{eq:latticepointsinsphere}.
 \end{proof}
@@
 This simple iterative approach to compute a closest lattice point is related to what is called the \emph{iterative slicer}~\cite{Shalvi_iterativeslicer_2009}.
-Micciancio~\cite{Micciancio09adeterministic} describes a related, but more sophisticated, iterative algorithm that can compute a closest lattice point in a number of operations that grows exponentially as $O(2^{2 n})$.
+Micciancio and Voulgaris~\cite{MicciancioVoulgaris_deterministic_jv_2013} describes a related, but more sophisticated, iterative algorithm that can compute a closest point in a number of operations that grows exponentially as $O(2^{2 n})$.
 This single exponential growth in complexity is the best known.
@@
 In general a lattice can have as many as $2^{n+1}-2$ relevant vectors so computing a minimiser directly can require a number of operations that grows exponentially with $n$.
-To add to this it is often the case that the set of relevant vectors $\relevant(\Lambda)$ must be stored in memory so the algorithm can require an amount of memory that grows exponentially with $n$~\cite[Sec.~6]{Micciancio09adeterministic}\cite{Shalvi_iterativeslicer_2009}.
+To add to this it is often the case that the set of relevant vectors $\relevant(\Lambda)$ must be stored in memory so the algorithm can require an amount of memory that grows exponentially with $n$~\cite{MicciancioVoulgaris_deterministic_jv_2013}\cite{Shalvi_iterativeslicer_2009}.
 We will show that for a lattice of Voronoi's first kind the set of relevant vectors has a compact representation in terms of what is called its \emph{obtuse superbasis}.
@@
 The other factor affecting the complexity is the number of iterations required before the algorithm arrives at a closest lattice point, that is, the size of $K$.
-Proposition~\ref{eq:latticepointsinsphere} suggests that this number might be as large as $\ceil{r/\rho}^n$ where $r = \|y - x_0\|^2$ and $\rho$ is the packing radius of the lattice.
+Proposition~\ref{eq:latticepointsinsphere} suggests that this number might be as large as $(\floor{r/\rho}+1)^n$ where $r = \|y - x_0\|^2$ and $\rho$ is the packing radius of the lattice.
 Thus, the number of iterations required might grow exponentially with $n$.
@@
 The $q_{ij}$ are called the \emph{Selling parameters}~\cite{Selling1874}.
-It is known that all lattices in dimensions less than $4$ are of Voronoi's first kind~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.
+It is known that all lattices in dimensions less than or equal to $3$ are of Voronoi's first kind~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.
 An interesting property of lattices of Voronoi's first kind is that their relevant vectors have a straightforward description.
@@
 \end{equation}
-be the $n+1$ by $n+1$ matrix with columns given by $b_1,\dots,b_{n+1}$ and let $z \in \reals^{n+1}$ be a column vector such that $y = Bz$.
+be the $m$ by $n+1$ matrix with columns given by $b_1,\dots,b_{n+1}$ and let $z \in \reals^{n+1}$ be a column vector such that $y = Bz$.
 We now want to find a column vector $w = (w_1,\dots,w_{n+1})^\prime$ of integers such that
@@
 is minimised.
-Define the column vector $u_0 = \floor{z}$ where $\floor{\cdot}$ denotes the largest integer less than or equal to its argument and operates on vectors elementwise.
+Define the column vector $u_0 = \floor{z}$ where $\floor{\cdot}$ operates on vectors elementwise.
 In view of Theorem~\ref{thm:revvecssuperbase} the iterative procedure~\eqref{eq:relvectsminimslicer} to compute a closest lattice point can be written in the form
@@
 \[
-\|B(z-u-g-w)\|^2 > |B(z-u-v)\|^2.
+\|B(z-u-g-w)\|^2 > \|B(z-u-v)\|^2.
 \]
@@
 A significant hurdle in the practical application of lattices as codes or as quantisers is that computing a closest lattice point is computationally difficult in general~\cite{micciancio_hardness_2001}.
-The best known general purpose algorithms require a number of operations of order $O(2^{2n})$~\cite{Micciancio09adeterministic}.
+The best known general purpose algorithms require a number of operations of order $O(2^{2n})$~\cite{MicciancioVoulgaris_deterministic_jv_2013}.
 In this paper we have focused on the class of lattices of Voronoi's first kind.
@@
 A final remark is that our algorithm assumes that the obtuse superbasis is known in advance.
-It is known that all lattices of dimension less than 4 are of Voronoi's first kind and an algorithm exists to recover the obtuse superbasis in this case~\cite{SPLAG}.
-Lattices of dimension larger than 4 need not be of Voronoi's first kind.
+It is known that all lattices of dimension less than or equal to 3 are of Voronoi's first kind and an algorithm exists to recover the obtuse superbasis in this case~\cite{SPLAG}.
+Lattices of dimension larger than 3 need not be of Voronoi's first kind.
 An interesting question is: given a lattice, is it possible to efficiently decide whether it is of Voronoi's first kind?

### 2014-07-07 reviews/response.tex
added more to responses.

@@
 
-I would like to thank reviwers for there thoughtful comments.  %The authors feel the paper has improved as a result.
+We thank both reviwers for their thoughtful comments.  %The authors feel the paper has improved as a result.
 
@@
 explicitly stated in the paper for some reason).
-Indeed, the quadratic form $\|Bx\|^2$ is exactly $x^T L_G x$, where $L_G$ is the Laplacian the weighted graph $G$ on vertices $\{ 1, .
+Indeed, the quadratic form $\|B\xbf\|^2$ is exactly $\xbf^T L_G \xbf$, where $L_G$ is the Laplacian the weighted graph $G$ on vertices $\{ 1, .
 .
 .
-, n + 1 \}$ , with edges weights $w_{ij} = -b_i \cdot b_j \geq 0, i = j$ (which they call the Selling parameters).
+, n + 1 \}$ , with edges weights $w_{ij} = -\bbf_i \cdot \bbf_j \geq 0, i = j$ (which they call the Selling parameters).
 More precisely:
@@
 Problem with Preprocessing (CVPP) over a special class of lattices. In the preprocessing model,
-we are allowed to compute any polynomial amount of advice about the lattice (i.e. a good ba-
-sis, or many short vectors in the lattice, etc.) before answering any CVP queries, where the re-
-sources needed to compute this advice are not counted in the runtime of the algorithm. In its
+we are allowed to compute any polynomial amount of advice about the lattice (i.e.
+a good basis, or many short vectors in the lattice, etc.) before answering any CVP queries, where the resources needed to compute this advice are not counted in the runtime of the algorithm.
+In its
 gap decisional version, where we need only decide whether a target is at distance $\leq d$ or $> \alpha d$
-(for some approximation factor $α \geq 1$), it is was shown [1, 4] to be NP-Hard for any constant $\alpha = 2^{\log^{1-\epsilon}n}$, and to not be polynomial time solvable for $\alpha = 2 \log n$ , for any fixed $\epsilon > 0$, under the assumption  that $NP \subsetneq \text{RTIME}(2^{\log^{O(1/\epsilon)}n})$ (quasi-polynomial time).
+(for some approximation factor $\alpha \geq 1$), it is was shown [1, 4] to be NP-Hard for any constant $\alpha = 2^{\log^{1-\epsilon}n}$, and to not be polynomial time solvable for $\alpha = 2 \log n$ , for any fixed $\epsilon > 0$, under the assumption  that $NP \subsetneq \text{RTIME}(2^{\log^{O(1/\epsilon)}n})$ (quasi-polynomial time).
 For its approximate search version (i.e.
-find a lattice vector whose distance to the target is some bounded factor away from optimal), an
-$O( n^{1.5} )$ approximation algorithm was (implicitly) given in [5], which was only recently improved to $O( n \log n )$ in [3].
+find a lattice vector whose distance to the target is some bounded factor away from optimal), an $O(n^{1.5})$ approximation algorithm was (implicitly) given in [5], which was only recently improved to $O( \sqrt{ n / \log n} )$ in [3].
 \\\textbf{Response:}
@@
 in the theory of lattice problems (not stated in paper). Indeed, by a simple reduction this would
-allow us to decide whether a lattice (given any of its bases) is a rotation of $Z_n$ (which is clearly
+allow us to decide whether a lattice (given any of its bases) is a rotation of $\ints^n$ (which is clearly
 of Voronoi’s first kind). This is a long standing open problem in the theory of lattices, which has
-only recently been solved in certain very special cases [6].
+only recently been solved in certain very special cases~\cite{Lenstra_Silverberg_revisting_gentra_szydlo_2014}.
 \\\textbf{Response:}
+The final paragraph of Section~\zref{sec:discussion} has been modified to included that the correction with the existing problem of determining a rectangular basis should it exist.
+
+``A final remark is that our algorithm assumes that the obtuse superbasis is known in advance.
+It is known that all lattices of dimension less than or equal to 3 are of Voronoi's first kind and an algorithm exists to recover the obtuse superbasis in this case~\cite{SPLAG}.
+Lattices of dimension larger than 3 need not be of Voronoi's first kind.
+An interesting question is: given a lattice, is it possible to efficiently decide whether it is of Voronoi's first kind?
+A related question is: is it possible to efficiently find a obtuse superbasis if it exists?
+It is suspected that the answer to both of these questions is no.
+An efficient solution to either question would immediately yield a solution to a known problem, that of determining whether a lattice is rectangular (has a basis consisting of pairwise orthogonal vectors) given an arbitrary basis~\cite{Lenstra_Silverberg_revisting_gentra_szydlo_2014}.''
 
@@
 This paper makes a solid contribution to the theory of lattice problems. I think the results here are
-well-motivated (I encourage the authors to include some of the motivation mentioned in the sum-
-mary), and point to the possibility that there might be large and useful classes of lattices where
+well-motivated (I encourage the authors to include some of the motivation mentioned in the summary), and point to the possibility that there might be large and useful classes of lattices where
 solving CVP is “easy”.
 \\\textbf{Response:}
+The following paragraph has been added to Section~\zref{sec:discussion} to suggest that similar fast algorithms for other classes of lattices might exist:
+
+``Another interesting question is: are there subfamilies of Voronoi's first kind that admit even faster algorithms?
+Both $A_n$ and $A_n^*$ are examples of this, but there might exist other subfamilies with algorithms faster than $O(n^4)$.
+A related question is: can the techniques developed in this paper be applied to other families of lattices, i.e., beyond just those of Voronoi's first kind?''
 
@@
 combinatorial arguments successfully applied to lattice problems. On the other hand, while I’ve
-convinced myself the authors have the “right” convergence proof, I found its presentation to be
+convinced myself the authors have the ``right'' convergence proof, I found its presentation to be
 rather awkward and unintuitive. I highly recommend that the authors rephrase the proof using
@@
 \\\textbf{Response:}
+This error has been fixed.
 
@@
 \\\textbf{Response:}
+This error has been fixed.
 
 \item\textbf{Comment:}  
-page 11, "with be" -> "will be" Displayed equation before "Putting p=...", A | should be || 
-(5.7) again, missing factor 2 before double summation, and second set should be T, not S. 
+page 11, "with be" $\to$ "will be" Displayed equation before "Putting p=...", A | should be || (5.7) again, missing factor 2 before double summation, and second set should be T, not S. 
 \\\textbf{Response:}
+These errors have been fixed.
 
@@
 \\\textbf{Response:}
+These errors have been fixed.
 
@@
 \\\textbf{Response:}
+This error has been fixed.
 
 \item\textbf{Comment:}  
-page 13, after (6.1), "be computing" -> "by computing" 
+page 13, after (6.1), "be computing" $\to$ "by computing" 
 \\\textbf{Response:}
+Fixed.
 
 \item\textbf{Comment:}  
-page 14, "have no effect" -> "has no effect" 
+page 14, "have no effect" $\to$ "has no effect" 
 \\\textbf{Response:}
+Fixed.
 
 \item\textbf{Comment:}  
-page 15, definition of $Q$, should the summation defininig $q_{ii}$ be restricted to $j\neq i$? 
+page 15, definition of $Q$, should the summation defining $q_{ii}$ be restricted to $j\neq i$? 
 \\\textbf{Response:}
+Yes. This has been fixed.
 
 \item\textbf{Comment:}  
-page 15, "there many" -> "there may" 
+page 15, "there many" $\to$ "there may" 
 \\\textbf{Response:}
+Fixed.
 
@@
 \bibstyle{../siam}
-\bibliography{bib}
+\bibliography{../bib}
 

### 2014-07-07 paper.tex
added more to responses.

@@
 \item  $\|Bp\|^2 - \|B(p - \onebf_S)\|^2 = \Phi(\bar{S}, p)$, \label{eq:lem:decSellingsdec}
-\item  ${\displaystyle \|Bp\|^2 - \|B(p + \onebf_S - \onebf_T)\|^2 = \Phi(S, p) + \Phi(\bar{T},p) + \sum_{i\in S}\sum_{j\in T} q_{ij} }$.  \label{eq:lem:decSellingsincdec}
+\item  ${\displaystyle \|Bp\|^2 - \|B(p + \onebf_S - \onebf_T)\|^2 = \Phi(S, p) + \Phi(\bar{T},p) + 2\sum_{i\in S}\sum_{j\in T} q_{ij} }$.  \label{eq:lem:decSellingsincdec}
 \end{enumerate}
@@
 \|Bp\|^2 - &\|B(p + \onebf_S - \onebf_T)\|^2 \\
-&= \|Bp\|^2-\|B(p + \onebf_S)\|^2 +  \|Bp\|^2-\|B(p - \onebf_T)\|^2 + \sum_{i\in S}\sum_{j\in T} q_{ij}.
+&= \|Bp\|^2-\|B(p + \onebf_S)\|^2 +  \|Bp\|^2-\|B(p - \onebf_T)\|^2 + 2\sum_{i\in S}\sum_{j\in T} q_{ij}.
 \end{align*}
@@
 \begin{lemma}\label{eq:integergreaterless}
-Let $h \in \{0,1\}^{n+1}$ and $v \in \ints^{n+1}$.  Suppose that
+Let $h \in \{0,1\}^{n+1}$ and $v \in \ints^{n+1}$ with $\rng(v) \geq 1$.  Suppose that
 $h_i = 0$ whenever $v_i = \min(v)$ and that $h_i = 1$ whenever $v_i = \max(v)$.  Then
@@
 \begin{equation}\label{eq:BpBpdecincineq}
-\|Bp\|^2 - \|B(p+\onebf_S-\onebf_T)\|^2 = \Phi(S,p) + \Phi(\bar{T}, p) + \sum_{i\in S}\sum_{j \in S} q_{ij} < 0.
+\|Bp\|^2 - \|B(p+\onebf_S-\onebf_T)\|^2 = \Phi(S,p) + \Phi(\bar{T}, p) + 2\sum_{i\in S}\sum_{j \in T} q_{ij} < 0.
 \end{equation}
@@
 \|B(z-u-h)\|^2 - \|B(z-u-g)\|^2 &= \|Br\|^2 - \|B(r+\onebf_S - \onebf_T)\|^2 \\
-&= \Phi(S,r) + \Phi(\bar{T},r) + \sum_{i\in S}\sum_{j \in S} q_{ij} \\
-&\leq \Phi(S,p) + \Phi(\bar{T}, p) + \sum_{i\in S}\sum_{j \in S} q_{ij} < 0 
+&= \Phi(S,r) + \Phi(\bar{T},r) + 2\sum_{i\in S}\sum_{j \in T} q_{ij} \\
+&\leq \Phi(S,p) + \Phi(\bar{T}, p) + 2\sum_{i\in S}\sum_{j \in T} q_{ij} < 0 
 \end{align*}
@@
 \]
-The first sum above is independent of $t$ and can be ignored for the purpose of minimisation.  Letting $s_i = \sum_{j=1}^{n+1} q_{ij}p_j$, we can equivalently minimise the binary quadratic form
+The first sum above is independent of $t$ and can be ignored for the purpose of minimisation.  Letting $s_i = -2\sum_{j=1}^{n+1} q_{ij}p_j$, we can equivalently minimise the binary quadratic form
 \begin{equation}\label{eq:quadformnp}
@@
 \end{equation}
-We will show that a minimiser of $Q(t)$ can be found efficiently be computing a minimum cut in an undirected flow network.
+We will show that a minimiser of $Q(t)$ can be found efficiently by computing a minimum cut in an undirected flow network.
 %\section{Quadratic $\{0,1\}$ programs and minimium cuts in graphs}
@@
 We require some properties of the weights $w_{ij}$ in relation to $W$.
-If the graph is allowed to contain loops, that is, edges from a vertex to itself, then the weight of these edges $w_{ii}$ have no effect on the weight of any cut.
+If the graph is allowed to contain loops, that is, edges from a vertex to itself, then the weight of these edges $w_{ii}$ has no effect on the weight of any cut.
 We may choose any values for the $w_{ii}$ without affecting $W$.
@@
 One approach is to construct the $n+1$ by $n+1$ symmetric matrix $Q = B^\prime B$ with elements $Q_{ij} = q_{ij} = \dotprod{b_{i}}{b_j}$ given by the Selling parameters.
-Choose the off diagonal entries of $Q$ to be nonpositive with $q_{ij}=q_{ji}$ and set the diagonal elements $q_{ii} = -\sum_{j=1}^{n+1} q_{ij}$.
+Choose the off diagonal entries of $Q$ to be nonpositive with $q_{ij}=q_{ji}$ and set the diagonal elements $q_{ii} = -\sum_{j \neq i} q_{ij}$.
 %We then have $Q\onebf = \zerobf$ and so $\onebf$ is in the null space of $Q$.
-The matrix $Q$ is diagonally dominant, that is, $\abs{q_{ii}} \geq \sum_{j=1}^{n+1} \abs{q_{ij}}$, and so $Q$ is positive semidefinite.
+The matrix $Q$ is diagonally dominant, that is, $\abs{q_{ii}} \geq \sum_{j \neq i} \abs{q_{ij}}$, and so $Q$ is positive semidefinite.
 A rank deficient Cholesky decomposition~\cite{Higham90analysisof} can now be used to recover a matrix $B$ such that $B^\prime B  = Q$.
@@
 A fast closest point algorithm requiring only $O(n)$ operations exists for $A_n^*$~\cite{McKilliam2009CoxeterLattices,McKilliam2008b} and so the algorithm described in this paper is not needed in this particular case.
-However, there many exist other signal processing problems where lattices of Voronoi's first kind arise.
+However, there may exist other signal processing problems where lattices of Voronoi's first kind arise.
+
+Another interesting question is: are there subfamilies of Voronoi's first kind that admit even faster algorithms?
+Both $A_n$ and $A_n^*$ are examples of this, but there might exist other subfamilies with algorithms faster than $O(n^4)$.
+A related question is: can the techniques developed in this paper be applied to other families of lattices, i.e., beyond just those of Voronoi's first kind?
+%More generally, do there exist large and useful familes of lattices where computing a closest point has polynomial complexity?
 
@@
 An interesting question is: given a lattice, is it possible to efficiently decide whether it is of Voronoi's first kind?
-A related question is: is it possible to efficiently find an obtuse superbasis if it exists?
+A related question is: is it possible to efficiently find a obtuse superbasis if it exists?
+It is suspected that the answer to both of these questions is no.
+An efficient solution to either question would immediately yield a solution to a known problem, that of determining whether a lattice is rectangular (has a basis consisting of pairwise orthogonal vectors) given an arbitrary basis~\cite{Lenstra_Silverberg_revisting_gentra_szydlo_2014}.
 

### 2014-07-08 reviews/response.tex
minor

@@
 So, it is unclear the significance of CVP algorithms for Leech lattice (in fixed dimension 24) in the context of this discussion.
+\\\textbf{Response:}
+This sentence has been reworded and the reference list has been reduced.  It now reads:
+
+``Although the problem is NP-hard in general, fast algorithms are known for specific highly regular lattices, such the integer lattice $\ints^{n}$, the root lattices $A_n$ and $D_n$, their dual lattices $A_n^*$ and $D_n^*$, and the related Coxeter lattices~\cite[Chap.~4]{SPLAG}\cite{Conway1982FastQuantDec, McKilliam2008, McKilliam2009CoxeterLattices}.''
+
+The interested reader could read these references in the order given.
+%, i.e., first Chapter 4 of~\cite{SPLAG} and~\cite{Conway1982FastQuantDec} for $D_n$ and $D_n^*$ and older algorithms for $A_n$ and $A_n^*$.
+Then~\cite{McKilliam2008} for a fast algorithm for $A_n^*$, then~\cite{McKilliam2009CoxeterLattices} for fast algorithms for all of the Coxeter lattices that includes $A_n$ and $A_n^*$, $E_8$, etc.
+Technically, the algorithm from~\cite{McKilliam2009CoxeterLattices} does include that from~\cite{McKilliam2008}.
+However~\cite{McKilliam2008} is simpler and shorter to read first.
+
+\item\textbf{Comment:}  
 When saying "if the obtuse superbasis is known", you may want to cast this requirement in the context of CVPP. (The obtuse superbasis is the advice provided by the preprocessing.) 

### 2014-07-08 paper.tex
minor

@@
 
-The closest lattice point problem is known to be NP-hard under certain conditions when the lattice itself, or rather a basis thereof, is considered as an additional input parameter~\cite{micciancio_hardness_2001, Dinur2003_approximating_CVP_NP_hard, Jalden2005_sphere_decoding_complexity}.
+The closest lattice point problem is known to be NP-hard in general~\cite{micciancio_hardness_2001, Dinur2003_approximating_CVP_NP_hard, Jalden2005_sphere_decoding_complexity}.
 Nevertheless, algorithms exist that can compute a closest lattice point in reasonable time if the dimension is small~\cite{Pohst_sphere_decoder_1981,Kannan1987_fast_general_np,Agrell2002}.
-These algorithms all require a number of operations that grows as $O(n^{O(n)})$ where $n$ is the dimension of the lattice.
+These algorithms require a number of operations that grows as $O(n^{O(n)})$ or $O(n^{O(n^2)})$ where $n$ is the dimension of the lattice.
 Recently, Micciancio and Voulgaris~\cite{MicciancioVoulgaris_deterministic_jv_2013} described a solution for the closest lattice point problem that requires a number of operations that grows as $O(2^{2n})$.
@@
 
-Although the problem is NP-hard in general, fast algorithms are known for specific highly regular lattices, such as the root lattices $A_n$ and $D_n$, their dual lattices $A_n^*$ and $D_n^*$, the integer lattice $\ints^{n}$, and the Leech lattice~\cite[Chap.~4]{SPLAG}\cite{Conway1982FastQuantDec,Conway1986SoftDecLeechGolay, Clarkson1999:Anstar, McKilliam2008, McKilliam2008b, McKilliam2009CoxeterLattices,Vardy1993_leech_lattice_MLD}.
+Although the problem is NP-hard in general, fast algorithms are known for specific highly regular lattices, such the integer lattice $\ints^{n}$, the root lattices $A_n$ and $D_n$, their dual lattices $A_n^*$ and $D_n^*$, and the related Coxeter lattices~\cite[Chap.~4]{SPLAG}\cite{Conway1982FastQuantDec, McKilliam2008, McKilliam2009CoxeterLattices}.
 In this paper we consider a particular class of lattices, those of \emph{Voronoi's first kind}~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases,Valentin2003_coverings_tilings_low_dimension,Voronoi1908_main_paper}.
@@
 This result extends upon a recent result by some of the authors showing that a short vector in a lattice of Voronoi's first kind can be found by computing a minimum cut in a weighted graph~\cite{McKilliam_short_vectors_first_type_isit_2012}.
+
+%BLERG incase I do want to introduce graph Laplacian in the introduction.
+%We show that if the obtuse superbasis is known, then a closest lattice point can be computed in $O(n^4)$ operations.
+This is achieved by enumerating a series of \emph{relevant vectors} of the lattice.
+We show that the series converges to a closest lattice point after at most $n$ terms and, using a connection between the obtuse superbasis and the Laplacian of a graph with positive weights, we show how each relevant vector in the series can be computed in $O(n^3)$ operations by an algorithm that computes a minimum cut in an undirected flow network~\cite{Picard_min_cuts_1974,Sankaran_solving_CDMA_mincut_1998,Ulukus_cdma_mincut_1998,Cormen2001}.
 

### 2014-07-08 reviews/response.tex
added more to responses.  More to bib

@@
 \\\textbf{Response:}
+Fixed.
 
@@
 \item\textbf{Comment:}  
-When saying "if the obtuse superbasis is known", you may want to cast this requirement in the context of CVPP. (The obtuse superbasis is the advice provided by the preprocessing.) 
+When saying ``if the obtuse superbasis is known'', you may want to cast this requirement in the context of CVPP. (The obtuse superbasis is the advice provided by the preprocessing.) 
 \\\textbf{Response:}
@@
 \\\textbf{Response:}
+Our definition includes what Conway and Sloane call the `lax' relevant vectors.
+To make this clear the following sentences have been added below the definition of the relevant vectors in Section~\zref{sec:voron-cells-relev}:
+
+``This definition includes what Conway and Sloane call the `lax' relevant vectors~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.
+If the definition above is replaced by one with strict inequality then only the `strict' relevant vectors would be considered.
+It will be of benefit to include all of the strict and lax relevant vectors here.''
 
@@
 \\\textbf{Response:}
+We have included the following text in the first paragraph of Section~\zref{sec:iterative-slicer} when introducing the interative algorithm.
+
+``Motivated by Shalvi~et.~al.~\cite{Shalvi_iterativeslicer_2009} and  Micciancio and Voulgaris~\cite{MicciancioVoulgaris_deterministic_jv_2013}, we consider the following iteration,''
+
+Among the iterative proceedures in this paper, in~\cite{Shalvi_iterativeslicer_2009}, and in~\cite{MicciancioVoulgaris_deterministic_jv_2013}, ours is the simplest.
+The choice of our iteration is not motivated by what might be fastest in general lattices.
+It is motivated by what we can prove to be polynomial time for those lattices of Vornoi's first kind.
+
+The~\emph{iterative slicer}~\cite{Shalvi_iterativeslicer_2009} is similar but uses the (presumably accelerated?) iteration
+\begin{align}
+x_{k+1} &= x_k + \alpha_k v_k \nonumber \\
+(v_k, \alpha_k) &= \arg\min_{ v \in \relevant(\Lambda) \cup \{\zerobf\}, \alpha \in \ints } \|y - x_k - \alpha v \|.
+\end{align} 
+So it allows adding integer multiples of a relevant vector.  We do not believe using this will improve the complexity of our algorithm beyond $O(n^4)$.  It would greatly complicate our proofs.
+
+The iterative algorithm of Micciancio and Voulgaris~\cite{MicciancioVoulgaris_deterministic_jv_2013} is quite different altogther.
+As we understand it, is not to add relevant vectors, but to instead scale the lattice
+
+We have not included a description of the iterative approaches using in
 

### 2014-07-08 paper.tex
added more to responses.  More to bib

@@
 Section~\ref{sec:iterative-slicer} describes a procedure to find a closest lattice point by enumerating a series of relevant vectors.
-The series is guaranteed to converge to a closest lattice point after a finite number of terms.
+The series is guaranteed to converge to a closest point after a finite number of terms.
 In general the procedure might be computationally expensive because the number of terms required might be large and because computation of each relevant vector in the series might be expensive.
@@
 
-Equivalently the Voronoi cell can be defined as the intersection of the half spaces 
+Equivalently, the Voronoi cell can be defined as the intersection of the half spaces 
 \begin{align*}
@@
 \]
-We denote by $\relevant(\Lambda)$ the set of relevant vectors of the lattice $\Lambda$.  The Voronoi cell is the intersection of the halfspaces corresponding with the relevant vectors, that is, 
+This definition includes what Conway and Sloane call the `lax' relevant vectors~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.
+If the definition above is replaced by one with strict inequality then only the `strict' relevant vectors would be considered.
+It will be of benefit to include all of the strict and lax relevant vectors here.
+Denote by $\relevant(\Lambda)$ the set of relevant vectors of the lattice $\Lambda$.
+The Voronoi cell is the intersection of the halfspaces corresponding with the relevant vectors, that is,
 \[
@@
 Let $x_0$ be some lattice point from $\Lambda$, for example the origin.
-Consider the following iteration,
+Motivated by Shalvi~et.~al.~\cite{Shalvi_iterativeslicer_2009} and  Micciancio and Voulgaris~\cite{MicciancioVoulgaris_deterministic_jv_2013}, we consider the following iteration,
 \begin{align}
@@
 
-Proposition~\ref{obs:2} above asserts that after some finite number $K$ of iterations the procedure arrives at $x_K$, a closest lattice point to $y$.
+Proposition~\ref{obs:2} asserts that after some finite number $K$ of iterations the procedure arrives at $x_K$, a closest lattice point to $y$.
 Using Proposition~\ref{obs:1} we can detect that $x_K$ is a closest lattice point by checking whether $\|y - x_K\| \leq \| y - x_{K+1} \|$.
 This simple iterative approach to compute a closest lattice point is related to what is called the \emph{iterative slicer}~\cite{Shalvi_iterativeslicer_2009}.
-Micciancio and Voulgaris~\cite{MicciancioVoulgaris_deterministic_jv_2013} describes a related, but more sophisticated, iterative algorithm that can compute a closest point in a number of operations that grows exponentially as $O(2^{2 n})$.
+Micciancio and Voulgaris~\cite{MicciancioVoulgaris_deterministic_jv_2013} describe a related, but more sophisticated, iterative algorithm that can compute a closest point in a number of operations that grows exponentially as $O(2^{2 n})$.
 This single exponential growth in complexity is the best known.
@@
 The number of iterations required depends on the lattice point that starts the iteration $x_0$.
-It is helpful for $x_0$ to be, in some sense, a close approximation of the closest lattice point $x_K$.
-Unfortunately, computing close approximations of a closest lattice point is known to be computationally difficult~\cite{feige_inapproximability_2004}.
+It is helpful for $x_0$ to be, in some sense, a close approximation of the closest point $x_K$.
+Unfortunately, computing close approximations of a closest lattice point is known to be computationally difficult~\cite{feige_inapproximability_2004,Regev_2004_inappox_lattice_with_preprocessing}.
 We will show that for a lattice of Voronoi's first kind a simple and easy to compute choice for $x_0$ ensures that a closest lattice point is reached in at most $n$ iterations and so $K \leq n$.

### 2014-07-08 reviews/response.tex
more comments.

@@
 \\\textbf{Response:}
-BLERG.  I think there's an error in the inquality above.
+BLERG. This is indeed the property used to prove Theorem 4.1, but not what's needed for the convergence
 
@@
 \\\textbf{Response:}
+We have fixed these errors.  This paragraph now reads:
+
+``The closest lattice point problem is known to be NP-hard in general~\cite{micciancio_hardness_2001, Dinur2003_approximating_CVP_NP_hard,Regev_2004_inappox_lattice_with_preprocessing,feige_inapproximability_2004,Jalden2005_sphere_decoding_complexity}.
+Nevertheless, algorithms exist that can compute a closest lattice point in reasonable time if the dimension is small~\cite{Pohst_sphere_decoder_1981,Kannan1987_fast_general_np,Agrell2002}.
+These algorithms require a number of operations that grows as $O(n^{O(n)})$ or $O(n^{O(n^2)})$ where $n$ is the dimension of the lattice.
+Recently, Micciancio and Voulgaris~\cite{MicciancioVoulgaris_deterministic_jv_2013} described a solution for the closest lattice point problem that requires a number of operations that grows as $O(2^{2n})$.
+This single exponential growth in complexity is the best known.''
 
@@
 
-Among the iterative proceedures in this paper, in~\cite{Shalvi_iterativeslicer_2009}, and in~\cite{MicciancioVoulgaris_deterministic_jv_2013}, ours is the simplest.
+Among the iterative proceedures in our paper, in~\cite{Shalvi_iterativeslicer_2009}, and in~\cite{MicciancioVoulgaris_deterministic_jv_2013}, ours is the simplest.
 The choice of our iteration is not motivated by what might be fastest in general lattices.
@@
 The~\emph{iterative slicer}~\cite{Shalvi_iterativeslicer_2009} is similar but uses the (presumably accelerated?) iteration
-\begin{align}
-x_{k+1} &= x_k + \alpha_k v_k \nonumber \\
-(v_k, \alpha_k) &= \arg\min_{ v \in \relevant(\Lambda) \cup \{\zerobf\}, \alpha \in \ints } \|y - x_k - \alpha v \|.
-\end{align} 
+\begin{align*}
+x_{k+1} &= x_k + \alpha_k v_k  \\
+(v_k, \alpha_k) &= \arg\min_{ (v, \alpha) \in ( \relevant(\Lambda) \cup \{\zerobf\}, \ints) } \|y - x_k - \alpha v \|.
+\end{align*} 
 So it allows adding integer multiples of a relevant vector.  We do not believe using this will improve the complexity of our algorithm beyond $O(n^4)$.  It would greatly complicate our proofs.
 
-The iterative algorithm of Micciancio and Voulgaris~\cite{MicciancioVoulgaris_deterministic_jv_2013} is quite different altogther.
-As we understand it, is not to add relevant vectors, but to instead scale the lattice
+The iterative algorithm of Micciancio and Voulgaris~\cite{MicciancioVoulgaris_deterministic_jv_2013} is different again.
+To our (admittedly rough) knowledge, the proceedure computes an $\alpha_k$ (different, but related to that above) and then proceeds by the same iteration as ours to compute a closest point in the scaled lattice $\alpha_k \Lambda$.
+The next $\alpha_{k+1}$ is computed and the proceedure repeats.
 
-We have not included a description of the iterative approaches using in
+We beleive that detailing the differences between these iterative proceedures would only detract from the core purpose of our paper to decscribe an $O(n^4)$ time algorithm to compute a closest point in lattice of Vornoi's first kind.
 
@@
 - 12: missing author (R. Raz) 
-- 18: "A V" -> "A. V." etc, 
+- 18: "A V" $\to$ "A. V." etc, 
 - 37: journal version in SIAM J. computing, 42(3):1364-1391 (2013) 
@@
 \\\textbf{Response:}
+We thank the reviewer for these interesting references highlighting the difficultly of computing close approximations of a closest lattice point.
+These have been referenced in Section~\zref{sec:iterative-slicer} with the following text:
+
+``Unfortunately, computing close approximations of a closest lattice point is known to be computationally difficult~\cite{feige_inapproximability_2004,Regev_2004_inappox_lattice_with_preprocessing,Aharonov_Regev_2005,Aleknovish_hardness_with_preprocessing_2011,Dadush_cvp_with_distance_guarantee_2014}.''
 

### 2014-07-08 paper.tex
more comments.

@@
 
-\author{Robby~G.~McKilliam, Alex Grant and I. Vaughan L. Clarkson}
+\author{Robby~G.~McKilliam, Alex~Grant and I.~Vaughan~L.~Clarkson}
 
@@
 
-The closest lattice point problem is known to be NP-hard in general~\cite{micciancio_hardness_2001, Dinur2003_approximating_CVP_NP_hard, Jalden2005_sphere_decoding_complexity}.
+The closest lattice point problem is known to be NP-hard in general~\cite{micciancio_hardness_2001, Dinur2003_approximating_CVP_NP_hard,Regev_2004_inappox_lattice_with_preprocessing,feige_inapproximability_2004,Jalden2005_sphere_decoding_complexity}.
 Nevertheless, algorithms exist that can compute a closest lattice point in reasonable time if the dimension is small~\cite{Pohst_sphere_decoder_1981,Kannan1987_fast_general_np,Agrell2002}.
@@
 It is helpful for $x_0$ to be, in some sense, a close approximation of the closest point $x_K$.
-Unfortunately, computing close approximations of a closest lattice point is known to be computationally difficult~\cite{feige_inapproximability_2004,Regev_2004_inappox_lattice_with_preprocessing}.
+Unfortunately, computing close approximations of a closest lattice point is known to be computationally difficult~\cite{feige_inapproximability_2004,Regev_2004_inappox_lattice_with_preprocessing,Aharonov_Regev_2005,Aleknovish_hardness_with_preprocessing_2011,Dadush_cvp_with_distance_guarantee_2014}.
 We will show that for a lattice of Voronoi's first kind a simple and easy to compute choice for $x_0$ ensures that a closest lattice point is reached in at most $n$ iterations and so $K \leq n$.
@@
 It is known that all lattices in dimensions less than or equal to $3$ are of Voronoi's first kind~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.
+%BLERG.  THIS MIGHT BETTER GO IN THE GRAPH SECTION
+% There is a one-to-one correspondence between the obtuse superbasis of a lattice of Vornoi's first kind and what is called the \emph{Laplacian} of a weighted graph.  To se this, let
+% \[
+% B = (b_1\,\,b_2\,\,\dots\,\,b_{n+1})
+% \]
+% be the matrix with columns given by $b_1,\dots,b_{n+1}$ and define the $n+1$ by $n+1$ \emph{Gram matrix} $Q = B^\prime B$.
+The elements of $Q$ are given by the Selling parameters, that is, $Q_{ij} = q_{ij}$.
+Let $G$ be a graph with $n+1$ vertices $v_1,\dots,v_{n+1}$ and edges $e_{ij}$ connecting $v_i$ to $v_j$.
+To each edge assign the \emph{weight} $w_{ij} = -q_{ij} \leq 0$.
+Let $d_{ii} = \sum_{j \neq i} w_{ij}$ be the sum of weights of edges connected to vertex $v_i$.
+The Laplacian of the graph $G$ is the $n+1$ by $n+1$ matrix with elements
+% \[
+% L_{ij} = \begin{cases}
+% d_{ii} & i = j \\
+% -w_{ij} & i \neq j. 
+% \end{cases}
+% \]
 An interesting property of lattices of Voronoi's first kind is that their relevant vectors have a straightforward description.
@@
 \end{theorem}
+
+BLERG: It will be instructive for us to give a proof of this theorem.  In doing so we will introduce some of the notation that will be used in Section~\ref{}.
 
@@
 Let
-\begin{equation}\label{eq:matrxBobtusebasis}
+\[
 B = (b_1\,\,b_2\,\,\dots\,\,b_{n+1})
-\end{equation}
+\]
 be the $m$ by $n+1$ matrix with columns given by $b_1,\dots,b_{n+1}$ and let $z \in \reals^{n+1}$ be a column vector such that $y = Bz$.
@@
 \begin{lemma}\label{lem:decrngpreservesclosestpoints}
-If $v \in \ints^{n+1}$ such that $B(\floor{z} + v)$ is a closest lattice point to $y = Bz$, then $B\big(\floor{z} + \decrng(v)\big)$ is also a closest lattice point to $y$.
+If $v \in \ints^{n+1}$ is such that $B(\floor{z} + v)$ is a closest lattice point to $y = Bz$, then $B\big(\floor{z} + \decrng(v)\big)$ is also a closest lattice point to $y$.
 \end{lemma}

### 2014-07-09 paper.tex
minor

@@
 
-\begin{figure}[tp] 
+\begin{figure}[p] 
 	\centering      
@@
 		\label{lattices:fig:vorregion}   
+\end{figure}
+
+\begin{figure}[p] 
+	\centering      
+		\includegraphics{figs/iterativeseriesexample-1.mps} 
+		\caption{Example of the iterative procedure described in~\eqref{eq:relvectsminimslicer} to compute a closest lattice point to $y = (4,3.5)$ (marked with a cross) in the $2$-dimensional lattice generated by basis vectors $(2,0.4)$ and $(0.4,2)$.
+The initial lattice point for the iteration is $x_0 = (-4.4,-2.8)$.
+The shaded region is the Voronoi cell surrounding the closest lattice point $x_6 = (4.4,2.8)$.}
+		\label{lattices:fig:iterativeexample} 
 \end{figure}
@@
 
-\begin{figure}[tp] 
-	\centering      
-		\includegraphics{figs/iterativeseriesexample-1.mps} 
-		\caption{Example of the iterative procedure described in~\eqref{eq:relvectsminimslicer} to compute a closest lattice point to $y = (4,3.5)$ (marked with a cross) in the $2$-dimensional lattice generated by basis vectors $(2,0.4)$ and $(0.4,2)$.
-The initial lattice point for the iteration is $x_0 = (-4.4,-2.8)$.
-The shaded region is the Voronoi cell surrounding the closest lattice point $x_6 = (4.4,2.8)$.}
-		\label{lattices:fig:iterativeexample} 
-\end{figure}
-
 \section{Lattices of Voronoi's first kind} \label{sec:latt-voron-first}
@@
 It is known that all lattices in dimensions less than or equal to $3$ are of Voronoi's first kind~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.
-%BLERG.  THIS MIGHT BETTER GO IN THE GRAPH SECTION
-% There is a one-to-one correspondence between the obtuse superbasis of a lattice of Vornoi's first kind and what is called the \emph{Laplacian} of a weighted graph.  To se this, let
+%%%%%%%%%%%%%%%%%%%%
+% There is a one-to-one correspondence between the obtuse superbasis of a lattice of Vornoi's first kind and what is called the \emph{Laplacian} of a weighted graph.  To see this, let
 % \[
@@
 % \]
-% be the matrix with columns given by $b_1,\dots,b_{n+1}$ and define the $n+1$ by $n+1$ \emph{Gram matrix} $Q = B^\prime B$.
+% be the matrix with columns given by $b_1,\dots,b_{n+1}$ and define the $n+1$ by $n+1$ symmetric \emph{Gram matrix} $Q = B^\prime B$.
 The elements of $Q$ are given by the Selling parameters, that is, $Q_{ij} = q_{ij}$.
+Because of the superbasis condition the row and column sums of $Q$ are zero.
+In other words, $Q\onebf = \zerobf$ and $\onebf^\prime Q = \zerobf^\prime$ where $\onebf$ and $\zerobf$ denote column vectors of ones and zeros respectively.
 Let $G$ be a graph with $n+1$ vertices $v_1,\dots,v_{n+1}$ and edges $e_{ij}$ connecting $v_i$ to $v_j$.
@@
 % \]
+%%%%%%%%%%%%%%%%%%%
+% Lattices of Vornoi's first kind have some other interesting properties.  Let
+% \[
+% B = (b_1\,\,b_2\,\,\dots\,\,b_{n+1})
+% \]
+% be the matrix with columns given by $b_1,\dots,b_{n+1}$ and define the $n+1$ by $n+1$ symmetric \emph{Gram matrix} $Q = B^\prime B$ where superscript $^\prime$ indicates the vector or matrix transpose.
+The elements of $Q$ are given by the Selling parameters, that is, $Q_{ij} = q_{ij}$.
+Because of the superbasis condition the row and column sums of $Q$ are zero, i.e.,
+% \[
+% \sum_{i = 1}^n q_{ij} = \sum_{j=1}^n q_{ij} = 0.
+% \]  
+% Equivalently, $Q\onebf = \zerobf$ and $\onebf^\prime Q = \zerobf^\prime$ where $\onebf$ and $\zerobf$ denote column vectors of ones and zeros respectively.
+The following formula will be useful.
+For $x\in\reals^{n+1}$ and $y \in \reals^{n+1}$, the quadratic form
+% \begin{align}
+% x^\prime Q y &= \frac{1}{2} ( x^\prime Qy + y^\prime Qx ) \nonumber \\ 
+% &= \frac{1}{2} \sum_{i=1}^{n+1} \sum_{j=1}^{n+1} q_{ij}(x_iy_j + x_jy_i) \nonumber \\
+% &= \frac{1}{2} \sum_{i=1}^{n+1} \sum_{j=1}^{n+1} q_{ij}\big( x_iy_i + x_jy_j - (x_i - x_j)(y_i - y_j) \big) \nonumber \\
+% &= -\sum_{i=1}^{n+1} \sum_{j=i+1}^{n+1} q_{ij}(x_i - x_j)(y_i - y_j). \label{eq:extendedsellings}
+% \end{align}
+% The first line in the above equation follows from the symmetry of $Q$.  The last line follows because
+% \[
+% \sum_{i=1}^{n+1} \sum_{j=1}^{n+1} q_{ij}x_jy_j = \sum_{i=1}^{n+1} \sum_{j=1}^{n+1} q_{ij}x_iy_i = \sum_{i=1}^{n+1} x_iy_i \sum_{j=1}^{n+1} q_{ij} = 0
+% \]
+% because the row and column sums of $Q$ are zero.
+Putting $y = x$ yields what is called~\emph{Selling's formula}~\cite{Selling1874,ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}\cite[Proposition~2.3.1]{Valentin2003_coverings_tilings_low_dimension},
+% \begin{equation}\label{eq:sellings}
+% \|Bx\|^2 = x^\prime Q x = -\sum_{i=1}^{n+1} \sum_{j=i+1}^{n+1} q_{ij}(x_i - x_j)^2.
+% \end{equation}
+%%%%%%%%%%%%%%%%%%%%%
 An interesting property of lattices of Voronoi's first kind is that their relevant vectors have a straightforward description.
@@
 
-BLERG: It will be instructive for us to give a proof of this theorem.  In doing so we will introduce some of the notation that will be used in Section~\ref{}.
+% \begin{align*}
+% x^\prime Q y &= \sum_{i=1}^n \sum_{j=1}^n q_{ij} x_i y_j \\
+% &= \sum_{i=1}^n \sum_{j=1}^n q_{ij} x_i y_j  - 2
+% \end{align*}
+% \[
+% (x-y)^\prime Q (x-y) = \sum_{i=1}^n \sum_{j=1}^n q_{ij} (x_i-y_i)^2
+% \]
+
+%BLERG: It will be instructive for us to give a proof of this theorem.  In doing so we will introduce some of the notation that will be used in Section~\ref{}.
 
@@
 which is precisely $\Phi(S,p)$.
-% Observe that
-% \[
-% \sum_{i=1}^{n+1}\sum_{j=1}^{n+1} q_{ij} p_i^2 =  \sum_{i=1}^{n+1}p_i^2\sum_{j=1}^{n+1} q_{ij} = 0
-% \]
-% and
-% \[
-% \sum_{i=1}^{n+1}\sum_{j=1}^{n+1} q_{ij} p_j^2 =  \sum_{j=1}^{n+1}p_j^2\sum_{i=1}^{n+1} q_{ij} = 0
-% \]
-% since $\sum_{i=1}^{n+1} q_{ij} = \sum_{j=1}^{n+1} q_{ij} = 0$ as a result of the superbasis condition~\eqref{eq:superbasecond}.  Now
-% \begin{align*}
-% \|Bp\|^2 &= \sum_{i=1}^{n+1} \sum_{j=1}^{n+1} q_{ij} p_i p_j \\
-% &= \sum_{i=1}^{n+1} \sum_{j=1}^{n+1} q_{ij} (\tfrac{1}{2}p_j^2 + p_i p_j -  \tfrac{1}{2}p_i^2) \\
-% &= -\frac{1}{2} \sum_{i=1}^{n+1} \sum_{j=1}^{n+1} q_{ij} (p_i - p_j)^2.
-% \end{align*}
-% This is closely related to what is known as Sellings formula~\cite[Proposition 2.3.1]{Valentin2003_coverings_tilings_low_dimension}~\cite{Selling1874}.  Put $g = p - \onebf_S$.  It follows that
-% \[
-% \|Bp\|^2 - \|Bg\|^2 = \frac{1}{2} \sum_{i=1}^{n+1}\sum_{j=1}^{n+1}q_{ij}d_{ij},
-% \]
-% where $d_{ij} = (g_i - g_j)^2 - (p_i - p_j)^2$.  If both $i \notin S$ and $j \notin S$ then $p_i = g_i$, $p_j = g_j$ and so $d_{ij} = 0$.  Similarly if both $i \in S$ and $j \in S$ then 
-% \[
-% d_{ij} = (p_i-1 - p_j+1)^2 - (p_i - p_j)^2 = 0.
-% \]
-% If $i \in S$ and $j \notin S$ then
-% \[
-% d_{ij} = (p_i-1 - p_j)^2 - (p_i - p_j)^2= 1 + 2p_j - 2p_i,
-% \]
-% while if $i \notin S$ and $j \in S$ then
-% \[
-% d_{ij} = (p_i - p_j+1)^2 - (p_i - p_j)^2 = 1 + 2p_i - 2p_j.
-% \]
-% We thus have
-% \begin{align*}
-% \|Bp\|^2 - \|Bg\|^2 &= \frac{1}{2} \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_j - 2p_i) + \frac{1}{2} \sum_{i \notin S}\sum_{j \in S}q_{ij}(1 + 2p_i - 2p_j) \\
-% &= \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_j - 2p_i),
-% \end{align*}
-% where the last line holds because $q_{ij} = q_{ji}$.
+% % Observe that
+% % \[
+% % \sum_{i=1}^{n+1}\sum_{j=1}^{n+1} q_{ij} p_i^2 =  \sum_{i=1}^{n+1}p_i^2\sum_{j=1}^{n+1} q_{ij} = 0
+% % \]
+% % and
+% % \[
+% % \sum_{i=1}^{n+1}\sum_{j=1}^{n+1} q_{ij} p_j^2 =  \sum_{j=1}^{n+1}p_j^2\sum_{i=1}^{n+1} q_{ij} = 0
+% % \]
+% % since $\sum_{i=1}^{n+1} q_{ij} = \sum_{j=1}^{n+1} q_{ij} = 0$ as a result of the superbasis condition~\eqref{eq:superbasecond}.  Now
+% % \begin{align*}
+% % \|Bp\|^2 &= \sum_{i=1}^{n+1} \sum_{j=1}^{n+1} q_{ij} p_i p_j \\
+% % &= \sum_{i=1}^{n+1} \sum_{j=1}^{n+1} q_{ij} (\tfrac{1}{2}p_j^2 + p_i p_j -  \tfrac{1}{2}p_i^2) \\
+% % &= -\frac{1}{2} \sum_{i=1}^{n+1} \sum_{j=1}^{n+1} q_{ij} (p_i - p_j)^2.
+% % \end{align*}
+% % This is closely related to what is known as Sellings formula~\cite[Proposition 2.3.1]{Valentin2003_coverings_tilings_low_dimension}~\cite{Selling1874}.  Put $g = p - \onebf_S$.  It follows that
+% % \[
+% % \|Bp\|^2 - \|Bg\|^2 = \frac{1}{2} \sum_{i=1}^{n+1}\sum_{j=1}^{n+1}q_{ij}d_{ij},
+% % \]
+% % where $d_{ij} = (g_i - g_j)^2 - (p_i - p_j)^2$.  If both $i \notin S$ and $j \notin S$ then $p_i = g_i$, $p_j = g_j$ and so $d_{ij} = 0$.  Similarly if both $i \in S$ and $j \in S$ then 
+% % \[
+% % d_{ij} = (p_i-1 - p_j+1)^2 - (p_i - p_j)^2 = 0.
+% % \]
+% % If $i \in S$ and $j \notin S$ then
+% % \[
+% % d_{ij} = (p_i-1 - p_j)^2 - (p_i - p_j)^2= 1 + 2p_j - 2p_i,
+% % \]
+% % while if $i \notin S$ and $j \in S$ then
+% % \[
+% % d_{ij} = (p_i - p_j+1)^2 - (p_i - p_j)^2 = 1 + 2p_i - 2p_j.
+% % \]
+% % We thus have
+% % \begin{align*}
+% % \|Bp\|^2 - \|Bg\|^2 &= \frac{1}{2} \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_j - 2p_i) + \frac{1}{2} \sum_{i \notin S}\sum_{j \in S}q_{ij}(1 + 2p_i - 2p_j) \\
+% % &= \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_j - 2p_i),
+% % \end{align*}
+% % where the last line holds because $q_{ij} = q_{ji}$.
 \end{proof}

### 2014-07-09 reviews/response.tex
minor

@@
 
-We thank both reviwers for their thoughtful comments.  %The authors feel the paper has improved as a result.
+We thank both reviewers for their thoughtful comments.
+Below, citations within the reviewer comments are left as given by the reviewer.
+Citations in our responses refer to the references at the end of this document.
+
+%The authors feel the paper has improved as a result.
 
@@
 explicitly stated in the paper for some reason).
-Indeed, the quadratic form $\|B\xbf\|^2$ is exactly $\xbf^T L_G \xbf$, where $L_G$ is the Laplacian the weighted graph $G$ on vertices $\{ 1, .
+Indeed, the quadratic form $\|B\xbf\|^2$ is exactly $\xbf^T L_G \xbf$, where $L_G$ is the Laplacian of the weighted graph $G$ on vertices $\{ 1, .
 .
@@
 \]
-\\ \textbf{Response:}
+\\ \textbf{Response:} We thank the reviewer for pointing out the connection with the Laplacian matrix of a graph.
+The following text has been included in Section~\zref{sec:comp-clos-relev} to highlight this link.
+
+``At the core of this technique is the fact that a one-to-one correspondence exists between the obtuse superbasis of a lattice of Voronoi's first kind and the~\emph{Laplacian matrix}~\cite{Chung_spectral_graph_theory_1997,Cvetković_spectra_graphs_1998} of a simple weighted graph with $n+1$ vertices and positive edge weights equal to the negated Selling parameters $-q_{ij}$.''
+
+The formula for $\|B\xbf\|_2^2$ stated by the reviewer is known as~\emph{Selling's formula} in literature on quadratic forms and the geometry of numbers~\cite[Proposition 2.3.1]{Valentin2003_coverings_tilings_low_dimension}~\cite{Selling1874}.
+We do indirectly use this formula in the form of the function $\Phi(S,p)$ and our Lemma~\zref{lem:decSellings}.
+A very early draft of our paper attempted to use Selling's formula more directly, but we ultimately found that it was $\Phi(S,p)$ that was used repeatedly and not Selling's formula.
 
 \item\textbf{Comment:} 
-BLERG: NEED TO CHECK REFS MATHS ETC To present the result from the perspective of Computational Complexity (also not given in the
+To present the result from the perspective of Computational Complexity (also not given in the
 paper), this paper can be viewed as giving a polynomial time algorithm for the Closest Vector
@@
 \item\textbf{Comment:} 
-BLERG CHECK REFS Hence, from the standpoint of CVPP, the algorithm in this paper gives essentially the first
+Hence, from the standpoint of CVPP, the algorithm in this paper gives essentially the first
 exact polynomial time algorithm over a reasonably large and natural class of lattices, i.e. those
@@
 of Voronoi’s first kind). This is a long standing open problem in the theory of lattices, which has
-only recently been solved in certain very special cases~\cite{Lenstra_Silverberg_revisting_gentra_szydlo_2014}.
+only recently been solved in certain very special cases~[6].
 \\\textbf{Response:}
@@
 \item\textbf{Comment:} 
-BLERG REFS.  I believe Proposition 2.1 and 2.2 appear essentially verbatim in [8]. Same for Propositions
+I believe Proposition 2.1 and 2.2 appear essentially verbatim in [8]. Same for Propositions
 3.1 (which is essentially by definition) and 3.2, they appear in [9]. I don’t think it adds any
@@
 \item\textbf{Comment:} 
-Almost all the inequalities in the paper can be derived from the following simple obser-
-vation.
+Almost all the inequalities in the paper can be derived from the following simple observation.
 For and $n$ vertex weighted graph $G$ with non-negative edge weights, and vectors, $\xbf,\ybf \in \reals^n$, if there exists a permutation $\pi$ of $[n]$ such that $\xbf_{\pi[1]} \leq \dots \leq \xbf_{\pi[n]}$ and $\ybf_{\pi[1]} \leq \dots \leq \ybf_{\pi[n]}$ then
@@
 i.e. $\xbf$ can be made ``shorter'' by squeezing its components closer together.
+\\\textbf{Response:}
 
+The property decribed by the reviewer is indeed needed to give a proof of Theorem 4.1 (see our response to the next comment number~\ref{com:provethm4.1}).
+However, this is not the property that is needed to prove that our algorithm converges in atmost $n$ steps.
+Let $\subrng_k(p)$ denote the largest set $S$ such that $\min_{i\in S} \xbf_i - \max_{j \notin S} \xbf_j \geq k$.
+Thus $\subrng_1$ is that suggested by the reviewer and $\subrng_2$ is that which is used in the paper.
+Our proof works by considering a $v \in \ints^n$ that can be added to $\floor{z}$ so that $B(\floor{z} + v)$ is a closest lattice point to $y = Bz$.
+Our Lemma~\zref{lem:roundzclose} shows that there exists such a $v$ with $\subrng_2(v) = \emptyset$ and so $\rng(v) \leq n$.
+It is not true that there always exists such a $v$ with  $\subrng_1(v) = \emptyset$.
+
+To consider this in more detail it is worth looking at the proof of Lemma~\zref{lem:decrngpreservesclosestpoints}.
+The lemma shows that $B(\floor{z} + v)$ is a closest point to $y = Bz$ only if $B\big(\floor{z} + \decrng(v)\big)$ is.
+A key part of the proof is the line
+\[
+1 + 2(\zeta_i-\zeta_j) - 2(v_i - v_j) < 1 + 2 - 4 = -1 < 0 \qquad \text{for $i \in S$ and $j \notin S$}
+\]
+where $\zeta_i = z_i - \floor{z_i} \in [0,1)$.
+Since $\zeta_i-\zeta_j \in (-1,1)$ we require that the gap between $v_i - v_j$ to be greater than $\tfrac{3}{2}$.
+Since $v \in \ints^{n+1}$ this means $v_i - v_j \geq 2$.
+We cannot replace the current $\subrng_2$ with the suggested $\subrng_1$ and have this proof hold.
+
+Very early on (more than 2 years ago now!) when first looking at lattices of Voronoi's first kind, the first author believed that a single iteration of the proceedure would surfice.  That is, compute
+\[
+v = \arg\min_{t \in \{0,1\}^{n+1}}\| B(z - \floor{z} - t) \|^2
+\]
+and then $B(\floor{z} + v)$ would be a closest lattice point.  Numerical counterexamples dispelled this possibility.
+
+\item\textbf{Comment:} \label{com:provethm4.1}
 Furthermore, one can use this to give a two line proof of Theorem 4.1.
@@
 \\\textbf{Response:}
-BLERG. This is indeed the property used to prove Theorem 4.1, but not what's needed for the convergence
+
+The property decribed by the reviewer is indeed needed to give a proof of Theorem 4.1.
+To our knowledge, one also needs the property that the relevant vectors of a lattice are given by the shortest coset representatives of $\Lambda / 2\Lambda$.
+This result is attributed to Voronoi (see for example \cite[Theorem~2]{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases} and \cite[Theorem~3.2]{MicciancioVoulgaris_deterministic_jv_2013}).
+This is used by Conway and Sloane~\cite[Theorem~3]{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases} to prove our Theorem~4.1.
+
+We believe that the proof, whilst not difficult, is ulitimately a distraction from our main results.
+This would be different if properties needed for the proof, such as the function $\subrng_1$ and Selling's formula were also needed for the proofs for our algorithmm.
+However, as we have stated, Selling's formula is not directly needed, and $\subrng_2$, not $\subrng_1$ is the function required.
 
@@
 \bibstyle{../siam}
-\bibliography{../bib}
+%\bibliography{../bib}
+\input{../paper.bbl}
 

### 2014-07-09 paper.tex
minor

@@
 This technique has appeared previously~\cite{Picard_min_cuts_1974,Sankaran_solving_CDMA_mincut_1998,Ulukus_cdma_mincut_1998,Cormen2001} but we include the derivation here so that this paper is self contained.
+At the core of this technique is the fact that a one-to-one correspondence exists between the obtuse superbasis of a lattice of Voronoi's first kind and the~\emph{Laplacian matrix}~\cite{Chung_spectral_graph_theory_1997,Cvetković_spectra_graphs_1998} of a simple weighted graph with $n+1$ vertices and positive edge weights equal to the negated Selling parameters $-q_{ij}$.
 

### 2014-07-09 reviews/response.tex
minor

@@
 allow us to decide whether a lattice (given any of its bases) is a rotation of $\ints^n$ (which is clearly
-of Voronoi’s first kind). This is a long standing open problem in the theory of lattices, which has
+of Voronoi's first kind). This is a long standing open problem in the theory of lattices, which has
 only recently been solved in certain very special cases~[6].
 \\\textbf{Response:}
-The final paragraph of Section~\zref{sec:discussion} has been modified to included that the correction with the existing problem of determining a rectangular basis should it exist.
+The final paragraph of Section~\zref{sec:discussion} has been modified to make the connection with the existing problem of determining whether a rectangular basis exists.
 
@@
 It is suspected that the answer to both of these questions is no.
-An efficient solution to either question would immediately yield a solution to a known problem, that of determining whether a lattice is rectangular (has a basis consisting of pairwise orthogonal vectors) given an arbitrary basis~\cite{Lenstra_Silverberg_revisting_gentra_szydlo_2014}.''
+An efficient solution to either question would immediately yield a solution to a known problem, that of determining whether a lattice is rectangular (has a basis consisting of pairwise orthogonal vectors) given an arbitrary basis~\cite{Lenstra_Silverberg_revisting_gentra_szydlo_2014}.
+It is suspected that this simpler problem is difficult.''
 
@@
 \item\textbf{Comment:} 
-On a technical level, I found the main convergence proof to be interesting, as I’ve rarely seen
+On a technical level, I found the main convergence proof to be interesting, as I've rarely seen
 combinatorial arguments successfully applied to lattice problems. On the other hand, while I’ve
@@
 \\\textbf{Response:}
+We thank the reviewer for their comments.  We respond to the authors concerns below.
 
@@
 \\\textbf{Response:}
+
+We agree that these propositions are well known/straightforward, but would prefer to keep them so that the paper is self contained.
+We do not know of references to specific proofs that we can give for these (unlike Theorem~4.1 for which a proof is directly given in \cite[Theorem~3]{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}).
+Proposition 2.2 does appear informally in~\cite[Sec~5]{Shalvi_iterativeslicer_2009}.
+As far as we are aware the remarkable algorithm in~\cite{MicciancioVoulgaris_deterministic_jv_2013} does not rely upon bounding the number of lattice points inside a sphere (or any other volume) as in Proposition~2.2.
+Instead, the proceedure operates on scaled versions of the lattice.
+The scaling is chosen to ensure that the target point always lies inside twice the Voronoi cell.
+
+A motivation for Propositions 3.1 and 3.2 is as follows.
+The stopping criterion in our proceedure is slightly different from those in~\cite{Shalvi_iterativeslicer_2009} and~\cite{MicciancioVoulgaris_deterministic_jv_2013}.
+The stopping criterions in~\cite{Shalvi_iterativeslicer_2009} and~\cite{MicciancioVoulgaris_deterministic_jv_2013} essentially involve a direct check that a given point, say $x$, is contained inside the Voronoi cell.
+That is, one directly asserts that $2 x \cdot v \leq \|v\|^2$ for all relevant vectors $v$.
+We cannot use this approach since it would amount to checking this inequality over as many as $2^{n+1}-2$ relevant vectors.
+Instead we simply check that $\|y - x_K\| \leq \|y - x_{K+1}\|$.
+That is, we check that the proceedure converges in the sense that
+\[
+\|y - x_K\| = \|y - x_{K+1}\| = \|y - x_{K+2}\| = \dots.
+\]
+This is a straightforward, but important, difference and it motivates Propostions 3.1 and 3.2.
 

### 2014-07-10 reviews/response.tex
added minor

@@
 We thank both reviewers for their thoughtful comments.
-Below, citations within the reviewer comments are left as given by the reviewer.
+Below, the citations within the reviewer comments are left as given by the reviewer.
 Citations in our responses refer to the references at the end of this document.
@@
 
-The formula for $\|B\xbf\|_2^2$ stated by the reviewer is known as~\emph{Selling's formula} in literature on quadratic forms and the geometry of numbers~\cite[Proposition 2.3.1]{Valentin2003_coverings_tilings_low_dimension}~\cite{Selling1874}.
+The formula for $\|B\xbf\|_2^2$ stated by the reviewer is known as~\emph{Selling's formula} in the literature on quadratic forms and the geometry of numbers~\cite[Proposition 2.3.1]{Valentin2003_coverings_tilings_low_dimension}~\cite{Selling1874}.
 We do indirectly use this formula in the form of the function $\Phi(S,p)$ and our Lemma~\zref{lem:decSellings}.
@@
 exact polynomial time algorithm over a reasonably large and natural class of lattices, i.e. those
-of Voronoi’s first kind. Lastly, removing the need for preprocessing for these lattices, at least for
+of Voronoi's first kind. Lastly, removing the need for preprocessing for these lattices, at least for
 solving SVP without being given an obtuse superbasis, would require a substantial breakthrough
@@
 \\\textbf{Response:}
-The final paragraph of Section~\zref{sec:discussion} has been modified to make the connection with the existing problem of determining whether a rectangular basis exists.
+The final paragraph of Section~\zref{sec:discussion} has been modified to make the connection with the existing problem of determining whether a rectangular basis exists in a given lattice.
 
@@
 An interesting question is: given a lattice, is it possible to efficiently decide whether it is of Voronoi's first kind?
-A related question is: is it possible to efficiently find a obtuse superbasis if it exists?
-It is suspected that the answer to both of these questions is no.
-An efficient solution to either question would immediately yield a solution to a known problem, that of determining whether a lattice is rectangular (has a basis consisting of pairwise orthogonal vectors) given an arbitrary basis~\cite{Lenstra_Silverberg_revisting_gentra_szydlo_2014}.
+A related question is: is it possible to efficiently find an obtuse superbasis if it exists?
+It is suspected that the answer to this second question is no.
+An efficient solution would yield a solution to a known problem, that of determining whether a lattice is rectangular (has a basis consisting of pairwise orthogonal vectors) given an arbitrary basis~\cite{Lenstra_Silverberg_revisting_gentra_szydlo_2014}.
 It is suspected that this simpler problem is difficult.''
@@
 On a technical level, I found the main convergence proof to be interesting, as I've rarely seen
-combinatorial arguments successfully applied to lattice problems. On the other hand, while I’ve
+combinatorial arguments successfully applied to lattice problems. On the other hand, while I've
 convinced myself the authors have the ``right'' convergence proof, I found its presentation to be
@@
 We do not know of references to specific proofs that we can give for these (unlike Theorem~4.1 for which a proof is directly given in \cite[Theorem~3]{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}).
-Proposition 2.2 does appear informally in~\cite[Sec~5]{Shalvi_iterativeslicer_2009}.
-As far as we are aware the remarkable algorithm in~\cite{MicciancioVoulgaris_deterministic_jv_2013} does not rely upon bounding the number of lattice points inside a sphere (or any other volume) as in Proposition~2.2.
-Instead, the proceedure operates on scaled versions of the lattice.
+The proof of Proposition 2.1 has been shortened in response to comment~\ref{com:rev2halfopen} of reviewer 2.
+A result like Proposition 2.2 does appear informally in~\cite[Sec~5]{Shalvi_iterativeslicer_2009}.
+As far as we are aware, the remarkable algorithm in~\cite{MicciancioVoulgaris_deterministic_jv_2013} does not rely upon bounding the number of lattice points inside a sphere (or any other volume).
+Instead, the procedure operates recursively on scaled versions of the lattice.
 The scaling is chosen to ensure that the target point always lies inside twice the Voronoi cell.
 
-A motivation for Propositions 3.1 and 3.2 is as follows.
-The stopping criterion in our proceedure is slightly different from those in~\cite{Shalvi_iterativeslicer_2009} and~\cite{MicciancioVoulgaris_deterministic_jv_2013}.
+Motivating Propositions 3.1 and 3.2 is that the stopping criterion in our proceedure is slightly different from those in~\cite{Shalvi_iterativeslicer_2009} and~\cite{MicciancioVoulgaris_deterministic_jv_2013}.
 The stopping criterions in~\cite{Shalvi_iterativeslicer_2009} and~\cite{MicciancioVoulgaris_deterministic_jv_2013} essentially involve a direct check that a given point, say $x$, is contained inside the Voronoi cell.
@@
 
-\item\textbf{Comment:} 
+\item\textbf{Comment:} \label{com:uselaplaicesubr1}
 Almost all the inequalities in the paper can be derived from the following simple observation.
-For and $n$ vertex weighted graph $G$ with non-negative edge weights, and vectors, $\xbf,\ybf \in \reals^n$, if there exists a permutation $\pi$ of $[n]$ such that $\xbf_{\pi[1]} \leq \dots \leq \xbf_{\pi[n]}$ and $\ybf_{\pi[1]} \leq \dots \leq \ybf_{\pi[n]}$ then
+For an $n$ vertex weighted graph $G$ with non-negative edge weights, and vectors, $\xbf,\ybf \in \reals^n$, if there exists a permutation $\pi$ of $[n]$ such that $\xbf_{\pi[1]} \leq \dots \leq \xbf_{\pi[n]}$ and $\ybf_{\pi[1]} \leq \dots \leq \ybf_{\pi[n]}$ then
 \[
@@
 However, this is not the property that is needed to prove that our algorithm converges in atmost $n$ steps.
-Let $\subrng_k(p)$ denote the largest set $S$ such that $\min_{i\in S} \xbf_i - \max_{j \notin S} \xbf_j \geq k$.
+To see this, let $\subrng_k(p)$ denote the largest set $S$ such that $\min_{i\in S} \xbf_i - \max_{j \notin S} \xbf_j \geq k$.
 Thus $\subrng_1$ is that suggested by the reviewer and $\subrng_2$ is that which is used in the paper.
@@
 Since $v \in \ints^{n+1}$ this means $v_i - v_j \geq 2$.
-We cannot replace the current $\subrng_2$ with the suggested $\subrng_1$ and have this proof hold.
+We cannot replace the current $\subrng_2$ with the suggested $\subrng_1$ and have this proof hold because then we would only have $v_i - v_j \geq 1$.
 
-Very early on (more than 2 years ago now!) when first looking at lattices of Voronoi's first kind, the first author believed that a single iteration of the proceedure would surfice.  That is, compute
-\[
-v = \arg\min_{t \in \{0,1\}^{n+1}}\| B(z - \floor{z} - t) \|^2
-\]
-and then $B(\floor{z} + v)$ would be a closest lattice point.  Numerical counterexamples dispelled this possibility.
+% Very early on (more than 2 years ago now!) when first looking at lattices of Voronoi's first kind, the first author believed that a single iteration of the proceedure would surfice.
+That is, compute
+% \[
+% v = \arg\min_{t \in \{0,1\}^{n+1}}\| B(z - \floor{z} - t) \|^2
+% \]
+% and then $B(\floor{z} + v)$ would be a closest lattice point.  Numerical counterexamples dispelled this possibility.
 
@@
 This result is attributed to Voronoi (see for example \cite[Theorem~2]{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases} and \cite[Theorem~3.2]{MicciancioVoulgaris_deterministic_jv_2013}).
-This is used by Conway and Sloane~\cite[Theorem~3]{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases} to prove our Theorem~4.1.
+This is used by Conway and Sloane in~\cite[Theorem~3]{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.
 
 We believe that the proof, whilst not difficult, is ulitimately a distraction from our main results.
-This would be different if properties needed for the proof, such as the function $\subrng_1$ and Selling's formula were also needed for the proofs for our algorithmm.
-However, as we have stated, Selling's formula is not directly needed, and $\subrng_2$, not $\subrng_1$ is the function required.
+This would be different if properties needed for the proof, such as the function $\subrng_1$ and Selling's formula were also needed for proving the convergence of our algorithmm.
+However, as we have stated in comments~\ref{com:uselaplaicesubr1}, Selling's formula is not directly needed, and $\subrng_2$, not $\subrng_1$ is the function required.
 
@@
 Notice that factor 2 in front of the last term.
-There are also two related typos on the top of page 12 where $\sum_{i \in S}\sum_{j\in S} q_{ij}$ (this appears twice on difference lines) should be $2\sum_{i \in S}\sum_{j\in T} q_{ij}$.
+There are also two related typos on the top of page 12 where $\sum_{i \in S}\sum_{j\in S} q_{ij}$ (this appears twice on different lines) should be $2\sum_{i \in S}\sum_{j\in T} q_{ij}$.
 \\\textbf{Response:}
@@
 \item\textbf{Comment:}  
-page 1, last line "The closest lattice point problem is known to be NP-hard under certain condition ...": I found this paragraph very confusing, and somehow inaccurate.
-What "certain conditions"?
+page 1, last line ``The closest lattice point problem is known to be NP-hard under certain condition ...'': I found this paragraph very confusing, and somehow inaccurate.
+What ``certain conditions''?
 I think a better way to describe the sitation regaring the NP-hardness of CVP in relation to this paper is the following.
@@
 ``This definition includes what Conway and Sloane call the `lax' relevant vectors~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.
-If the definition above is replaced by one with strict inequality then only the `strict' relevant vectors would be considered.
-It will be of benefit to include all of the strict and lax relevant vectors here.''
+If the definition above is replaced by one with strict inequality then only the `strict' relevant vectors would be considered.'' %It will be of benefit to include all of the strict and lax relevant vectors here.''
 
 \item\textbf{Comment:}  
-page 3, and maybe elsewhere "short vector in a lattice" should probably be "shortest nonzero vector in a lattice". 
+page 3, and maybe elsewhere ``short vector in a lattice'' should probably be ``shortest nonzero vector in a lattice''. 
 \\\textbf{Response:}
@@
 
-\item\textbf{Comment:}  
+\item\textbf{Comment:}  \label{com:rev2halfopen}
 page 3, proof of Prop.

### 2014-07-10 paper.tex
minor
added minor

@@
 If the definition above is replaced by one with strict inequality then only the `strict' relevant vectors would be considered.
-It will be of benefit to include all of the strict and lax relevant vectors here.
-Denote by $\relevant(\Lambda)$ the set of relevant vectors of the lattice $\Lambda$.
-The Voronoi cell is the intersection of the halfspaces corresponding with the relevant vectors, that is,
+%It will be of benefit to include all of the strict and lax relevant vectors here.
+Denote by $\relevant(\Lambda)$ the set of relevant vectors of the lattice $\Lambda$.  The Voronoi cell is the intersection of the halfspaces corresponding with the relevant vectors, that is, 
 \[
@@
 An interesting question is: given a lattice, is it possible to efficiently decide whether it is of Voronoi's first kind?
-A related question is: is it possible to efficiently find a obtuse superbasis if it exists?
-It is suspected that the answer to both of these questions is no.
-An efficient solution to either question would immediately yield a solution to a known problem, that of determining whether a lattice is rectangular (has a basis consisting of pairwise orthogonal vectors) given an arbitrary basis~\cite{Lenstra_Silverberg_revisting_gentra_szydlo_2014}.
+A related question is: is it possible to efficiently find an obtuse superbasis if it exists?
+It is suspected that the answer to this second question is no.
+An efficient solution would yield a solution to a known problem, that of determining whether a lattice is rectangular (has a basis consisting of pairwise orthogonal vectors) given an arbitrary basis~\cite{Lenstra_Silverberg_revisting_gentra_szydlo_2014}.
+It is suspected that this simpler problem is difficult.
 

### 2014-07-11 reviews/response.tex
minor

@@
 The iterative algorithm of Micciancio and Voulgaris~\cite{MicciancioVoulgaris_deterministic_jv_2013} is different again.
-To our (admittedly rough) knowledge, the proceedure computes an $\alpha_k$ (different, but related to that above) and then proceeds by the same iteration as ours to compute a closest point in the scaled lattice $\alpha_k \Lambda$.
-The next $\alpha_{k+1}$ is computed and the proceedure repeats.
+To our (admittedly rough) knowledge, the proceedure computes an $\alpha_k$ (different, but related to that above) and then proceeds to compute a closest point in the scaled lattice $\alpha_k \Lambda$.
+The scaling $\alpha_k$ is chosen so that the target point lies withing $2\alpha_k \vor(\Lambda)$.
+Once a closest lattice point in $\alpha_k\Lambda$ is found, the next $\alpha_{k+1}$ is computed and the proceedure repeats.
 
-We beleive that detailing the differences between these iterative proceedures would only detract from the core purpose of our paper to decscribe an $O(n^4)$ time algorithm to compute a closest point in lattice of Vornoi's first kind.
+We beleive that detailing the differences between these iterative proceedures would only detract from the core purpose of our paper to decscribe an $O(n^4)$ time algorithm to compute a closest point in lattice of Voronoi's first kind.
 
@@
 \item\textbf{Comment:}  
-page 11, "with be" $\to$ "will be" Displayed equation before "Putting p=...", A | should be || (5.7) again, missing factor 2 before double summation, and second set should be T, not S. 
+page 11, ``with be'' $\to$ ``will be'' Displayed equation before ``Putting p=...'', A | should be || (5.7) again, missing factor 2 before double summation, and second set should be T, not S. 
 \\\textbf{Response:}
@@
 \item\textbf{Comment:}  
-page 13, after (6.1), "be computing" $\to$ "by computing" 
+page 13, after (6.1), ``be computing'' $\to$ ``by computing'' 
 \\\textbf{Response:}
@@
 \item\textbf{Comment:}  
-page 14, "have no effect" $\to$ "has no effect" 
+page 14, ``have no effect'' $\to$ ``has no effect'' 
 \\\textbf{Response:}
@@
 \item\textbf{Comment:}  
-page 15, "there many" $\to$ "there may" 
+page 15, ``there many'' $\to$ ``there may'' 
 \\\textbf{Response:}
@@
 - 12: missing author (R. Raz) 
-- 18: "A V" $\to$ "A. V." etc, 
+- 18: ``A V'' $\to$ ``A. V.'' etc, 
 - 37: journal version in SIAM J. computing, 42(3):1364-1391 (2013) 

### 2014-07-11 reviews/response.tex
minor

@@
 - O. Regev (2004) Improved Inapproximability of Lattice and Coding Problems With Preprocessing. IEEE Transactions on Information Theory 50(9): 2031-2037 
-- "Hardness of Approximating the Closest Vector PRoblem with Pre-Processing", M. Alekhnovich, S. Khot, G. Kindler, N. Vishnoi, Computational Complexity 20(4):741-753, 2011.
+- ``Hardness of Approximating the Closest Vector PRoblem with Pre-Processing'', M. Alekhnovich, S. Khot, G. Kindler, N. Vishnoi, Computational Complexity 20(4):741-753, 2011.
 
@@
 
-I am not sure, as the paper has not appeared yet, but I suspect it is the same as the paper on Dadush webpage with title "On BDD and CVPP". 
+I am not sure, as the paper has not appeared yet, but I suspect it is the same as the paper on Dadush webpage with title ``On BDD and CVPP''. 
 See link 

### 2014-07-11 reviews/response.tex
minor

@@
 \\\textbf{Response:}
+
+The following paragraph has been included in the introduction to put our results into context with the closest vector problem with preprocessing:
+
+``Our results can be placed in the context of a modification of the closest lattice point problem called the \emph{closest vector problem with preprocessing}~\cite{micciancio_hardness_2001,feige_inapproximability_2004,Regev_2004_inappox_lattice_with_preprocessing,Aleknovish_hardness_with_preprocessing_2011,knot_hardness_cvvp_2014,Dadush_cvp_with_distance_guarantee_2014}.
+In this problem some ``advice'' of polynomial size about the lattice is assumed to be given.
+The advice might come in the form of a particular basis for the lattice or might take other forms.
+The advice may be used to compute a closest lattice point, hopefully with reduced complexity.
+%Even this modification of the closest lattice point problem is known to be NP-hard~\cite{micciancio_hardness_2001,feige_inapproximability_2004}.
+Our algorithm can be viewed as an efficient solution for the closest vector problem with preprocessing for the lattices of Voronoi's first kind.  The advice given is the obtuse superbasis.''
 
@@
 
+The following paragraph has been included in the introduction to put our results into context with the closest vector problem with preprocessing:
+
+``Our results can be placed in the context of a modification of the closest lattice point problem called the \emph{closest vector problem with preprocessing}~\cite{micciancio_hardness_2001,feige_inapproximability_2004,Regev_2004_inappox_lattice_with_preprocessing,Aleknovish_hardness_with_preprocessing_2011,knot_hardness_cvvp_2014,Dadush_cvp_with_distance_guarantee_2014}.
+In this problem some ``advice'' of polynomial size about the lattice is assumed to be given.
+The advice might come in the form of a particular basis for the lattice or might take other forms.
+The advice may be used to compute a closest lattice point, hopefully with reduced complexity.
+%Even this modification of the closest lattice point problem is known to be NP-hard~\cite{micciancio_hardness_2001,feige_inapproximability_2004}.
+Our algorithm can be viewed as an efficient solution for the closest vector problem with preprocessing for the lattices of Voronoi's first kind.  The advice given is the obtuse superbasis.''
+
 \item\textbf{Comment:}  
@@
 
-``The closest lattice point problem is known to be NP-hard in general~\cite{micciancio_hardness_2001, Dinur2003_approximating_CVP_NP_hard,Regev_2004_inappox_lattice_with_preprocessing,feige_inapproximability_2004,Jalden2005_sphere_decoding_complexity}.
+``The closest lattice point problem is known to be NP-hard~\cite{micciancio_hardness_2001, Dinur2003_approximating_CVP_NP_hard,Regev_2004_inappox_lattice_with_preprocessing,feige_inapproximability_2004,Jalden2005_sphere_decoding_complexity}.
 Nevertheless, algorithms exist that can compute a closest lattice point in reasonable time if the dimension is small~\cite{Pohst_sphere_decoder_1981,Kannan1987_fast_general_np,Agrell2002}.
@@
 \\\textbf{Response:}
-Our definition includes what Conway and Sloane call the `lax' relevant vectors.
+%In our original definition we intended to include what Conway and Sloane call the `lax' relevant vectors~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.  This we not necessary
+Yes, this has been fixed.  The text now reads reads:
+
+The \emph{relevant vectors} are those lattice points $v \in \Lambda \backslash  \{\zerobf\}$ for which  
+\[
+\dotprod{v}{x} < \dotprod{x}{x} \qquad \text{for all $x \in \Lambda \backslash  \{\zerobf\}$}.
+\]
+
+% Our definition includes what Conway and Sloane call the `lax' relevant vectors.
 To make this clear the following sentences have been added below the definition of the relevant vectors in Section~\zref{sec:voron-cells-relev}:
 
-``This definition includes what Conway and Sloane call the `lax' relevant vectors~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.
+% ``This definition includes what Conway and Sloane call the `lax' relevant vectors~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.
 If the definition above is replaced by one with strict inequality then only the `strict' relevant vectors would be considered.'' %It will be of benefit to include all of the strict and lax relevant vectors here.''
@@
 page 6, Theorem 4.1 (this follows from [ 9, Theorem 2+3], not just theorem 3. 
-Also, make the theorem statement self contained by adding "Let L be a lattice of voronoi first kind with obtuse superbasis b1..bn+1". 
+Also, make the theorem statement self contained by adding "Let $L$ be a lattice of voronoi first kind with obtuse superbasis $b_1,\dots,b_{n+1}$". 
 \\\textbf{Response:}
+We beleive that it is only Theorem 3 from~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases} that needs to cited.
+Theorem~2 in \cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases} asserts that the relevant vectors in any lattice $\Lambda$ are the shortest coset representatives of the quotient $\Lambda/2\Lambda$.
+Theorem~3 uses this, Selling's formula, and the fact that the Selling parameters are nonpositive to prove that the relevant vectors in a lattice of Voronoi's first kind take the form
+\[
+\sum_{i \in I} b_i
+\]
+where $I$ is a strict subset of $\{1, 2, \dots, n+1\}$ that is not empty.
+
+We have included the text ``Let $\Lambda$ be a lattice of Voronoi's first kind with obtuse superbasis $b_1,\dots,b_{n+1}$.'' into the statement of Theorem 4.1.
+
+%BLERG: Check that Theorems 2 and 3 in the published version of\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases} are what I think they are!
 

### 2014-07-11 paper.tex
minor

@@
 If the lattice is used as a code, then the closest lattice point corresponds to what is called \emph{lattice decoding}\index{lattice decoding} and has been shown to yield arbitrarily good codes~\cite{Erex2004_lattice_decoding,Erez2005}.
-The closest lattice point problem also occurs in communications systems involving multiple antennas~\cite{Ryan2008,Wubben_2011}.
-The unwrapping of phase data for location estimation can also be posed as a closest lattice point problem and this has been applied to the global positioning system~\cite{Teunissen_GPS_1995,Hassibi_GPS_1998}.
+The closest lattice point problem occurs in communications systems involving multiple antennas~\cite{Ryan2008,Wubben_2011}.
+The unwrapping of phase data for location estimation can be posed as a closest lattice point problem and this has been applied to the global positioning system~\cite{Teunissen_GPS_1995,Hassibi_GPS_1998}.
 The problem has also found applications to circular statistics~\cite{McKilliam_mean_dir_est_sq_arc_length2010}, single frequency estimation~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}, and related signal processing problems~\cite{McKilliam2007,Clarkson2007,McKilliam2009IndentifiabliltyAliasingPolyphase,Quinn_sparse_noisy_SSP_2012}.
 
-The closest lattice point problem is known to be NP-hard in general~\cite{micciancio_hardness_2001, Dinur2003_approximating_CVP_NP_hard,Regev_2004_inappox_lattice_with_preprocessing,feige_inapproximability_2004,Jalden2005_sphere_decoding_complexity}.
+The closest lattice point problem is known to be NP-hard~\cite{micciancio_hardness_2001, Dinur2003_approximating_CVP_NP_hard,Regev_2004_inappox_lattice_with_preprocessing,feige_inapproximability_2004,Jalden2005_sphere_decoding_complexity}.
 Nevertheless, algorithms exist that can compute a closest lattice point in reasonable time if the dimension is small~\cite{Pohst_sphere_decoder_1981,Kannan1987_fast_general_np,Agrell2002}.
@@
 We show that the series converges to a closest lattice point after at most $n$ terms and, using a connection between the obtuse superbasis and the Laplacian of a graph with positive weights, we show how each relevant vector in the series can be computed in $O(n^3)$ operations by an algorithm that computes a minimum cut in an undirected flow network~\cite{Picard_min_cuts_1974,Sankaran_solving_CDMA_mincut_1998,Ulukus_cdma_mincut_1998,Cormen2001}.
+
+Our results can be placed in the context of a modification of the closest lattice point problem called the \emph{closest vector problem with preprocessing}~\cite{micciancio_hardness_2001,feige_inapproximability_2004,Regev_2004_inappox_lattice_with_preprocessing,Aleknovish_hardness_with_preprocessing_2011,knot_hardness_cvvp_2014,Dadush_cvp_with_distance_guarantee_2014}.
+In this problem some ``advice'' of polynomial size about the lattice is assumed to be given.
+The advice might come in the form of a particular basis for the lattice or might take other forms.
+The advice may be used to compute a closest lattice point, hopefully with reduced complexity.
+%Even this modification of the closest lattice point problem is known to be NP-hard~\cite{micciancio_hardness_2001,feige_inapproximability_2004}.
+Our algorithm can be viewed as an efficient solution for the closest vector problem with preprocessing for the lattices of Voronoi's first kind.  The advice given is the obtuse superbasis.
 
@@
 \[
-\dotprod{v}{x} \leq \dotprod{x}{x} \qquad \text{for all $x \in \Lambda$}.
+\dotprod{v}{x} < \dotprod{x}{x} \qquad \text{for all $x \in \Lambda \backslash  \{\zerobf\}$}.
 \]
-This definition includes what Conway and Sloane call the `lax' relevant vectors~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.
+%This definition includes what Conway and Sloane call the `lax' relevant vectors~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.
 If the definition above is replaced by one with strict inequality then only the `strict' relevant vectors would be considered.
@@
 
-\begin{theorem} \label{thm:revvecssuperbase} (Conway and Sloane~\cite[Theorem~3]{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases})
+\begin{theorem} \label{thm:revvecssuperbase} (Conway and Sloane~\cite[Theorem~3]{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}) Let $\Lambda$ be a lattice of Voronoi's first kind with obtuse superbasis $b_1,\dots,b_{n+1}$.
 The relevant vectors of $\Lambda$ are of the form,

### 2014-07-11 reviews/response.tex
minor

@@
 \\\textbf{Response:}
-We thank the reviewer for their comments.  We respond to the authors concerns below.
+We thank the reviewer for their comments.
+The order of some paragraphs and lemmas has been changed in Section~\zref{sec:seri-relev-vect} with the intent of improving the readability of the proofs and presenting the intuition behind the proofs before formalities begin.
+We response to the authors other concerns below.
 

### 2014-07-11 paper.tex
minor

@@
 Are there lattices of Voronoi's first kind that are useful in applications such as coding, quantisation, or signal processing?
-We discuss these questions in Section~\ref{sec:discussion}.
-We now focus on the problem of computing a closest lattice point in a lattice of Voronoi's first kind.
+These questions are discussed in Section~\ref{sec:discussion}.
+We now focus on the problem of computing a closest point in a lattice of Voronoi's first kind.
 
@@
 
-We are interested in solving the closest lattice point problem for lattices of Voronoi's first kind.
+%We are interested in solving the closest lattice point problem for lattices of Voronoi's first kind.  
 Let $\Lambda \subset \reals^m$ be an $n$ dimensional lattice of Voronoi's first kind with obtuse superbasis $b_1,\dots,b_{n+1}$ and let $y \in \reals^m$.
@@
 In what follows it is assumed that $y$ lies in the space spanned by the basis vectors $b_1,\dots,b_{n}$.
-This assumption is without loss of generality because $x$ is a closest lattice point to $y$ if and only if $x$ is a the closest lattice point to the orthogonal projection of $y$ into the space spanned by $b_1,\dots,b_{n}$.
+This assumption is without loss of generality because $x$ is a closest lattice point to $y$ if and only if $x$ is a closest lattice point to the orthogonal projection of $y$ into the space spanned by $b_1,\dots,b_{n}$.
 Let
@@
 \newcommand{\rng}{\operatorname{rng}}
+
+We first provide an intuitive description of our proof technique.
+Denote by $\min(p)$ and $\max(p)$ the minimum and maximum values obtained by the elements of the vector $p \in \reals^{n+1}$ and define the function
+\[
+\rng(p) = \max(p) - \min(p).
+\] 
+%to return the difference between the maximum and minimum of $p$.  %For example, if $p = (2,-1,4) \in \ints^3$ then $\rng(p) = 4 - (-1)=5$.  We might also 
+%For example $\rng(2,-1,4)= 4 - (-1) = 5$.  
+Observe that $\rng(p)$ cannot be negative and that if $\rng(p) = 0$ then all of the elements of $p$ are equal.
+Let $\ell$ be a nonegative integer.
+Say that a lattice point $x$ is $\ell$-\emph{close} to $y$ if there exists a $v \in \ints^{n+1}$ with $\rng(v) = \ell$ such that $x + Bv$ is a closest point to $y$.
+We will show (Lemma~\ref{lem:roundzclose}) that the lattice point $x_0 = B\floor{z}$ that initialises the iterative procedure~\eqref{eq:xseqfirsttype} is $K$-close to $y$ where $K \leq n$.
+We then prove (Lemma~\ref{lem:rngdecreases}) that if the lattice point $x_k$ obtained on the $k$th iteration of the procedure is $\ell$-close, then the lattice point $x_{k+1}$ obtained on the next iteration is $(\ell-1)$-close.
+Since $x_0$ is $K$-close it follows that after $K \leq n$ iterations the lattice point $x_K$ is $0$-close.
+At this stage it is guaranteed that $x_{K}$ itself is a closest lattice point to $y$.
+This is shown in the following lemma.
+
+\begin{lemma}\label{lem:rngzeroclosestpoint}
+If $x$ is a lattice point that is $0$-close to $y$ then $x$ is a closest lattice point to $y$.
+\end{lemma}
+\begin{proof}
+Because $x$ is $0$-close there exists a $v \in \ints^{n+1}$ with $\rng(v) = 0$ such that $x + Bv$ is a closest point to $y$.
+Because $\rng(v) = 0$ all elements from $v$ are identical, that is, $v_1=v_2=\dots=v_{n+1}$.
+In this case $Bv = \sum_{i=1}^{n+1} v_n b_n = v_1\sum_{i=1}^{n+1}b_n = 0$
+as a result of the superbasis condition~\eqref{eq:superbasecond}.  Thus $x = x + Bv$ is a closest point to $y$. 
+\end{proof}
+
 \newcommand{\subrng}{\operatorname{subr}}
@@
 
-It is necessary to introduce some notation.
+We now proceed with a formal proof culminating in Theorem~\ref{thm:mainitrconvthm} stated at the end of this section.
+It is necessary to first introduce some notation.
 For $S$ a subset of indices $\{1,\dots,n+1\}$ let $\onebf_S$ denote the column vector of length $n+1$ with $i$th element equal to one if $i \in S$ and zero otherwise.
 %Thus, for example, if $p \in \reals^{n+1}$, then $p+\onebf_S$ denotes the vector $p$ with those elements having index from $S$ incremented by one.
-For $S \subseteq \{1,\dots,n+1\}$ and $p \in \reals^{n+1}$ we define the function
+For $S \subseteq \{1,\dots,n+1\}$ and $p \in \reals^{n+1}$ define the function
 \[
@@
 where $q_{ij} = \dotprod{b_i}{b_j}$ are the Selling parameters from~\eqref{eq:obtusecond}.
-We denote by $\bar{S}$ the complement of the set of indices $S$, that is $\bar{S} = \{ i \in \{1,\dots,n+1\} \mid i \notin S\}$.
+Denote by $\bar{S}$ the complement of the set of indices $S$, that is $\bar{S} = \{ i \in \{1,\dots,n+1\} \mid i \notin S\}$.
 
@@
 
-Denote by $\min(p)$ and $\max(p)$ the minimum and maximum values obtained by the elements of the vector $p$ and define the function
-\[
-\rng(p) = \max(p) - \min(p).
-\] 
-%to return the difference between the maximum and minimum of $p$.  %For example, if $p = (2,-1,4) \in \ints^3$ then $\rng(p) = 4 - (-1)=5$.  We might also 
-%For example $\rng(2,-1,4)= 4 - (-1) = 5$.  
-Observe that $\rng(p)$ cannot be negative and that if $\rng(p) = 0$ then all of the elements of $p$ are equal.
-We define the function $\subrng(p)$ to return the largest subset, say $S$, of the indices of $p$ such that $\min\{p_i, i \in S\} - \max\{p_i, i \notin S\} \geq 2.$  If no such subset exists then $\subrng(p)$ is the empty set $\emptyset$.
+Define the function $\subrng(p)$ to return the largest subset, say $S$, of the indices of $p$ such that $\min\{p_i, i \in S\} - \max\{p_i, i \notin S\} \geq 2.$  If no such subset exists then $\subrng(p)$ is the empty set $\emptyset$.
 For example,
@@
 \[
-p_{\sigma(n+1)} \leq p_{\sigma(n)} + 1 \leq p_{\sigma(n-1)} + 2 \leq \dots \leq p_{\sigma(1)} + n.
+p_{\sigma(n+1)} \leq p_{\sigma(n)} + 1 \leq p_{\sigma(n-1)} + 2 \leq \dots \leq p_{\sigma(1)} + n
 \]
@@
 
-Finally we define the function
+Finally, define the function
 \[
@@
 \begin{proof}
-Assume that no such $T$ exists.  Then $\decrng(d_k) \neq d_k$ for all positive integers $k$ and so 
+Assume that no such $T$ exists.  In this case $\decrng(d_k) \neq d_k$ for all positive integers $k$ and so 
 \[
@@
 
-We are now ready to study properties of a closest lattice point in a lattice of Voronoi's first kind.
+We are now ready to study properties of the lattice point $x_0 = B\floor{z}$ that initialises the iterative procedure~\eqref{eq:xseqfirsttype}.
 
@@
 
-Let $\ell$ be a nonegative integer.
-We say that a lattice point $x$ is $\ell$-\emph{close} to $y$ if there exists a $v \in \ints^{n+1}$ with $\rng(v) = \ell$ such that $x + Bv$ is a closest lattice point to $y$.
-Lemma~\ref{lem:roundzclose} asserts that the lattice point $x_0 = B\floor{z}$ that initialises the iterative procedure~\eqref{eq:xseqfirsttype} is $K$-close to $y$ where $K \leq n$.
-From Lemma~\ref{lem:rngdecreases} stated below it will follow that if the lattice point $x_k$ obtained on the $k$th iteration of the procedure is $\ell$-close, then the lattice point $x_{k+1}$ obtained on the next iteration is $(\ell-1)$-close.
-Since $x_0$ is $K$-close it will then follow that after $K \leq n$ iterations the lattice point $x_K$ is $0$-close.
-At this stage it is guaranteed that $x_{K}$ itself is a closest lattice point to $y$.
-This is shown in the following lemma.
-
-\begin{lemma}\label{lem:rngzeroclosestpoint}
-If $x$ is a lattice point that is $0$-close to $y$, then $x$ is a closest lattice point to $y$.
-\end{lemma}
-\begin{proof}
-Because $x$ is $0$-close there exists a $v \in \ints^{n+1}$ with $\rng(v) = 0$ such that $x + Bv$ is a closest lattice point to $y$.
-Because $\rng(v) = 0$ all elements from $v$ are identical, that is, $v_1=v_2=\dots=v_{n+1}$.
-In this case $Bv = \sum_{i=1}^{n+1} v_n b_n = v_1\sum_{i=1}^{n+1}b_n = 0$
-as a result of the superbasis condition~\eqref{eq:superbasecond}.  Thus $x = x + Bv$ is a closest point to $y$. 
-\end{proof}
-
+%BLERG: will want a connecting sentence here.
+
+Recall that a lattice point $x$ is said to by $\ell$-close to $y$ if there exists a $v \in \ints^{n+1}$ with $\rng(v) = \ell$ such that $x + Bv$ is a closest point to $y$.
+Lemma~\ref{lem:roundzclose} above asserts that the lattice point $x_0 = B\floor{z}$ that initialises the iterative procedure is $K$-close to $y$ where $K \leq n$.
+We now prove Lemma~\ref{lem:rngdecreases} from which it will follow that if the lattice point $x_k$ obtained on the $k$th iteration of the procedure is $\ell$-close, then the lattice point $x_{k+1}$ obtained on the next iteration is $(\ell-1)$-close.
 Before giving the proof of Lemma~\ref{lem:rngdecreases} we require the following simple result.
@@
 
-The next theorem asserts that the iterative procedure~\eqref{eq:xseqfirsttype} converges to a closest lattice point in $K \leq n$ iterations.  This is the primary result of this section.
-
-\begin{theorem}
-Let $x_0,x_1,\dots$ be the sequence of lattice points given by the iterative procedure~\eqref{eq:xseqfirsttype}.  There exists $K \leq n$ such that $x_K$ is a closest lattice point to $y = Bz$.
+We are now ready to prove our main theorem asserting that the iterative procedure~\eqref{eq:xseqfirsttype} converges to a closest lattice point in $K \leq n$ iterations.
+This is the primary result of this section.
+
+\begin{theorem}\label{thm:mainitrconvthm}
+Let $\Lambda \subset \reals^m$ be a lattice of Voronoi's first kind with obtuse superbasis $b_1,\dots,b_{n+1}$.
+Let $y \in \reals^m$ and let $x_0,x_1,\dots$ be a sequence of lattice points given by the iterative procedure~\eqref{eq:xseqfirsttype}.
+There exists $K \leq n$ such that $x_K$ is a closest lattice point to $y$.
 \end{theorem}
 \begin{proof}
+Recall that we, without loss of generality, suppose that $y$ lies in the space spanned by the basis vectors $b_1,\dots,b_{n}$ and so $y = Bz$ where $z \in \ints^{n+1}$.
 Let $x_k = B u_k$ be the lattice point obtained on the $k$th iteration of the procedure.
-Suppose that $x_k$ is $\ell$-close to $y=Bz$ with $\ell > 0$.
+Suppose that $x_k$ is $\ell$-close to $y$ with $\ell > 0$.
 The procedure computes $t_{k} \in \{0,1\}^{n+1}$ satisfying
@@
 By calling the vertex $v_0$ the \emph{source} and the vertex $v_{n+2}$ the \emph{sink} the graph $G$ is what is called a \emph{flow network}.
-The flow network is \emph{undirected} since the weights assigned to each edge are undirected.
+The flow network is \emph{undirected} because the weights assigned to each edge are undirected.
 A \emph{cut} in the flow network $G$ is a subset $C \subset V$ of vertices with its complement $\bar{C} \subset V$ such that the source vertex $v_0 \in C$ and the sink vertex $v_{n+2} \in \bar{C}$.

### 2014-07-12 reviews/response.tex
added changes

@@
 \item\textbf{Comment:} 
-From the perspective of quadratic forms, the quadratic forms induced by obtuse superbases are in one to one correspondance with graph Laplacians (this is not
+From the perspective of quadratic forms, the quadratic forms induced by obtuse superbases are in one to one correspondence with graph Laplacians (this is not
 explicitly stated in the paper for some reason).
@@
 \\\textbf{Response:}
-We thank the reviewer for their comments.
-The order of some paragraphs and lemmas has been changed in Section~\zref{sec:seri-relev-vect} with the intent of improving the readability of the proofs and presenting the intuition behind the proofs before formalities begin.
-We response to the authors other concerns below.
+We thank the reviewer for their comments.  We response to the authors other concerns below.
 
@@
 3.1 (which is essentially by definition) and 3.2, they appear in [9]. I don’t think it adds any
-readibility/anything of value to reprove them here.
+readability/anything of value to reprove them here.
 \\\textbf{Response:}
@@
 
-Motivating Propositions 3.1 and 3.2 is that the stopping criterion in our proceedure is slightly different from those in~\cite{Shalvi_iterativeslicer_2009} and~\cite{MicciancioVoulgaris_deterministic_jv_2013}.
-The stopping criterions in~\cite{Shalvi_iterativeslicer_2009} and~\cite{MicciancioVoulgaris_deterministic_jv_2013} essentially involve a direct check that a given point, say $x$, is contained inside the Voronoi cell.
+Motivating Propositions 3.1 and 3.2 is that the stopping criterion in our procedure is slightly different from those in~\cite{Shalvi_iterativeslicer_2009} and~\cite{MicciancioVoulgaris_deterministic_jv_2013}.
+The stopping criteria in~\cite{Shalvi_iterativeslicer_2009} and~\cite{MicciancioVoulgaris_deterministic_jv_2013} essentially involve a direct check that a given point, say $x$, is contained inside the Voronoi cell.
 That is, one directly asserts that $2 x \cdot v \leq \|v\|^2$ for all relevant vectors $v$.
@@
 Instead we simply check that $\|y - x_K\| \leq \|y - x_{K+1}\|$.
-That is, we check that the proceedure converges in the sense that
+That is, we check that the procedure converges in the sense that
 \[
@@
 \]
-This is a straightforward, but important, difference and it motivates Propostions 3.1 and 3.2.
+This is a straightforward, but important, difference and it motivates Propositions 3.1 and 3.2.
 
@@
 I believe using the above will make many of your proofs more transparent.
-For example, if you redefine $\subrng(\xbf)$ to return the largest set $S$ such that $\min_{i\in S} \xbf_i - \max_{j \notin S} \xbf_j \geq 1$, then noice that if $S \neq \emptyset$ then there is a permutation that puts both $\xbf - 1_S$ and $1_S$ in non-decreasing order.
+For example, if you redefine $\subrng(\xbf)$ to return the largest set $S$ such that $\min_{i\in S} \xbf_i - \max_{j \notin S} \xbf_j \geq 1$, then notice that if $S \neq \emptyset$ then there is a permutation that puts both $\xbf - 1_S$ and $1_S$ in non-decreasing order.
 In particular this implies that
@@
 
-The property decribed by the reviewer is indeed needed to give a proof of Theorem 4.1 (see our response to the next comment number~\ref{com:provethm4.1}).
-However, this is not the property that is needed to prove that our algorithm converges in atmost $n$ steps.
+The property described by the reviewer is indeed needed to give a proof of Theorem 4.1 (see our response to the next comment number~\ref{com:provethm4.1}).
+However, this is not the property that is needed to prove that our algorithm converges in at most $n$ steps.
 To see this, let $\subrng_k(p)$ denote the largest set $S$ such that $\min_{i\in S} \xbf_i - \max_{j \notin S} \xbf_j \geq k$.
@@
 where $\zeta_i = z_i - \floor{z_i} \in [0,1)$.
-Since $\zeta_i-\zeta_j \in (-1,1)$ we require that the gap between $v_i - v_j$ to be greater than $\tfrac{3}{2}$.
-Since $v \in \ints^{n+1}$ this means $v_i - v_j \geq 2$.
+Since $\zeta_i-\zeta_j \in (-1,1)$ we require that the gap between $v_i - v_j$ to be greater than $\tfrac{3}{2}$ for this inequality to hold.
+Because $v \in \ints^{n+1}$ this means $v_i - v_j \geq 2$.
 We cannot replace the current $\subrng_2$ with the suggested $\subrng_1$ and have this proof hold because then we would only have $v_i - v_j \geq 1$.
@@
 
-The property decribed by the reviewer is indeed needed to give a proof of Theorem 4.1.
+The property described by the reviewer is indeed needed to give a proof of Theorem 4.1.
 To our knowledge, one also needs the property that the relevant vectors of a lattice are given by the shortest coset representatives of $\Lambda / 2\Lambda$.
@@
 
-We believe that the proof, whilst not difficult, is ulitimately a distraction from our main results.
-This would be different if properties needed for the proof, such as the function $\subrng_1$ and Selling's formula were also needed for proving the convergence of our algorithmm.
-However, as we have stated in comments~\ref{com:uselaplaicesubr1}, Selling's formula is not directly needed, and $\subrng_2$, not $\subrng_1$ is the function required.
+We believe that the proof, whilst not difficult, is ultimately a distraction from our main results.
+This would be different if properties needed for the proof, such as the function $\subrng_1$ and Selling's formula were also needed for proving the convergence of our algorithm.
+However, as we have stated in comment~\ref{com:uselaplaicesubr1} above, Selling's formula is not directly needed, and $\subrng_2$, not $\subrng_1$ is the function required.
+
+The order of some paragraphs and lemmas has been changed in Section~\zref{sec:seri-relev-vect} with the intent of improving the readability of the proofs and presenting the intuition behind the proofs before formalities begin.
 
@@
 \begin{enumerate}
-\item\textbf{Comment:}  
+\item\textbf{Comment:}  \label{com:cvvp}
 page 1, last line ``The closest lattice point problem is known to be NP-hard under certain condition ...'': I found this paragraph very confusing, and somehow inaccurate.
 What ``certain conditions''?
-I think a better way to describe the sitation regaring the NP-hardness of CVP in relation to this paper is the following.
+I think a better way to describe the situation regarding the NP-hardness of CVP in relation to this paper is the following.
 CVP is known to be NP-hard even to approximate within almost polynomial factor (see 35) when the lattice is arbitrary and not known in advance.
@@
 \\\textbf{Response:}
+Please see our response to comment~\ref{com:cvvp}.
 
 \item\textbf{Comment:}  
-page 2,bottom, displayed equation "$v \cdot x \leq x \cdot x$". Maybe you mean strict inequality.
+page 2,bottom, displayed equation ``$v \cdot x \leq x \cdot x$''. Maybe you mean strict inequality.
 \\\textbf{Response:}
@@
 \item\textbf{Comment:}  
-page 3, definition of ceiling(r) as smallest integer strctly larger than 3 seems unconventional.
+page 3, definition of ceiling(r) as smallest integer strictly larger than 3 seems unconventional.
 The standard definition is not strict.
@@
 2.1.
-Your definition of ``half-open voronoi cell'' is not a proper definition, and also seems incorrect.
+Your definition of ``half-open Voronoi cell'' is not a proper definition, and also seems incorrect.
 E.g., take the integer lattice $\ints^n$, then only one of the corners (rather than half) of the corners of the Voronoi cell should be included if you want a tiling.
@@
 The process described in (3.1) is almost identical to those in [45,37], so those papers should probably be referenced at the very beginning, and a comparison provided.
-As far as I remember, the difference is that 37 picks the v that minimizes the voronoi norm (i.e., the norm defined by the voronoi cell of the origin) instead of the euclidean norm.
-On the other hand, 45 uses the euclidean norm, but selects an arbitrary relevant vector that reduces the euclidean norm, rather than the one that reduces it the most.
+As far as I remember, the difference is that 37 picks the v that minimizes the Voronoi norm (i.e., the norm defined by the Voronoi cell of the origin) instead of the Euclidean norm.
+On the other hand, 45 uses the Euclidean norm, but selects an arbitrary relevant vector that reduces the Euclidean norm, rather than the one that reduces it the most.
 In this respect, your process is a special case of the one described in 45, and termination (your propositions 3.1 and 3.2) is already proved in 45.
 \\\textbf{Response:}
-We have included the following text in the first paragraph of Section~\zref{sec:iterative-slicer} when introducing the interative algorithm.
+We have included the following text in the first paragraph of Section~\zref{sec:iterative-slicer} when introducing the iterative algorithm.
 
@@
 
-Among the iterative proceedures in our paper, in~\cite{Shalvi_iterativeslicer_2009}, and in~\cite{MicciancioVoulgaris_deterministic_jv_2013}, ours is the simplest.
+Among the iterative procedures in our paper, in~\cite{Shalvi_iterativeslicer_2009}, and in~\cite{MicciancioVoulgaris_deterministic_jv_2013}, ours is the simplest.
 The choice of our iteration is not motivated by what might be fastest in general lattices.
-It is motivated by what we can prove to be polynomial time for those lattices of Vornoi's first kind.
+It is motivated by what we can prove to be polynomial time for those lattices of Voronoi's first kind.
 
@@
 The iterative algorithm of Micciancio and Voulgaris~\cite{MicciancioVoulgaris_deterministic_jv_2013} is different again.
-To our (admittedly rough) knowledge, the proceedure computes an $\alpha_k$ (different, but related to that above) and then proceeds to compute a closest point in the scaled lattice $\alpha_k \Lambda$.
-The scaling $\alpha_k$ is chosen so that the target point lies withing $2\alpha_k \vor(\Lambda)$.
-Once a closest lattice point in $\alpha_k\Lambda$ is found, the next $\alpha_{k+1}$ is computed and the proceedure repeats.
+To our (admittedly rough) knowledge, the procedure computes an $\alpha_k$ (different, but related to that above) and then proceeds to compute a closest point in the scaled lattice $\alpha_k \Lambda$.
+The scaling $\alpha_k$ is chosen so that the target point lies within $2\alpha_k \vor(\Lambda)$.
+Once a closest lattice point in $\alpha_k\Lambda$ is found, the next $\alpha_{k+1}$ is computed and the procedure repeats.
 
-We beleive that detailing the differences between these iterative proceedures would only detract from the core purpose of our paper to decscribe an $O(n^4)$ time algorithm to compute a closest point in lattice of Voronoi's first kind.
+We believe that detailing the differences between these iterative procedures would only detract from the core purpose of our paper to describe an $O(n^4)$ time algorithm to compute a closest point in lattice of Voronoi's first kind.
 
@@
 \item\textbf{Comment:}  
-page 5, last line, "dimension less than 4", does it include 4 or not? 
+page 5, last line, ``dimension less than 4'', does it include 4 or not? 
 \\\textbf{Response:}
@@
 page 6, Theorem 4.1 (this follows from [ 9, Theorem 2+3], not just theorem 3. 
-Also, make the theorem statement self contained by adding "Let $L$ be a lattice of voronoi first kind with obtuse superbasis $b_1,\dots,b_{n+1}$". 
+Also, make the theorem statement self contained by adding ``Let $L$ be a lattice of Voronoi first kind with obtuse superbasis $b_1,\dots,b_{n+1}$''. 
 \\\textbf{Response:}
-We beleive that it is only Theorem 3 from~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases} that needs to cited.
+We have included the text ``Let $\Lambda$ be a lattice of Voronoi's first kind with obtuse superbasis $b_1,\dots,b_{n+1}$.'' into the statement of Theorem 4.1.
+
+We believe that it is only Theorem 3 from~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases} that needs to cited.
 Theorem~2 in \cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases} asserts that the relevant vectors in any lattice $\Lambda$ are the shortest coset representatives of the quotient $\Lambda/2\Lambda$.
@@
 where $I$ is a strict subset of $\{1, 2, \dots, n+1\}$ that is not empty.
-
-We have included the text ``Let $\Lambda$ be a lattice of Voronoi's first kind with obtuse superbasis $b_1,\dots,b_{n+1}$.'' into the statement of Theorem 4.1.
 
@@
 - O. Regev (2004) Improved Inapproximability of Lattice and Coding Problems With Preprocessing. IEEE Transactions on Information Theory 50(9): 2031-2037 
-- ``Hardness of Approximating the Closest Vector PRoblem with Pre-Processing'', M. Alekhnovich, S. Khot, G. Kindler, N. Vishnoi, Computational Complexity 20(4):741-753, 2011.
+- ``Hardness of Approximating the Closest Vector Problem with Pre-Processing'', M. Alekhnovich, S. Khot, G. Kindler, N. Vishnoi, Computational Complexity 20(4):741-753, 2011.
 
@@
 \\\textbf{Response:}
-We thank the reviewer for these interesting references highlighting the difficultly of computing close approximations of a closest lattice point.
+We thank the reviewer for these interesting references highlighting the difficulty of computing close approximations of a closest lattice point.
 These have been referenced in Section~\zref{sec:iterative-slicer} with the following text:

### 2014-07-12 paper.tex
added changes

@@
 u_{k+1} &= u_k + t_k \nonumber \\
-t_k &= \arg\min_{t \in \{0,1\}^{n+1}}\| B(z - u_k - t) \|^2, \label{eq:pvecmin}
+t_k &= \arg\min_{t \in \{0,1\}^{n+1}}\| B(z - u_k - t) \|^2 \label{eq:pvecmin}
 \end{align}
@@
 We now proceed with a formal proof culminating in Theorem~\ref{thm:mainitrconvthm} stated at the end of this section.
-It is necessary to first introduce some notation.
+We first introduce some notation.
 For $S$ a subset of indices $\{1,\dots,n+1\}$ let $\onebf_S$ denote the column vector of length $n+1$ with $i$th element equal to one if $i \in S$ and zero otherwise.
@@
 Let $\Lambda \subset \reals^m$ be a lattice of Voronoi's first kind with obtuse superbasis $b_1,\dots,b_{n+1}$.
-Let $y \in \reals^m$ and let $x_0,x_1,\dots$ be a sequence of lattice points given by the iterative procedure~\eqref{eq:xseqfirsttype}.
+Let $z \in \reals^{n+1}$, let $y = Bz$, and let $x_0,x_1,\dots$ be a sequence of lattice points given by the iterative procedure~\eqref{eq:xseqfirsttype}.
 There exists $K \leq n$ such that $x_K$ is a closest lattice point to $y$.
@@
 \begin{proof}
-Recall that we, without loss of generality, suppose that $y$ lies in the space spanned by the basis vectors $b_1,\dots,b_{n}$ and so $y = Bz$ where $z \in \ints^{n+1}$.
+%Recall that we suppose, without loss of generality, that $y$ lies in the space spanned by the basis vectors $b_1,\dots,b_{n}$ and so $y = Bz$ where $z \in \ints^{n+1}$.  
 Let $x_k = B u_k$ be the lattice point obtained on the $k$th iteration of the procedure.
@@
 By calling the vertex $v_0$ the \emph{source} and the vertex $v_{n+2}$ the \emph{sink} the graph $G$ is what is called a \emph{flow network}.
-The flow network is \emph{undirected} because the weights assigned to each edge are undirected.
+The flow network is undirected because the weights assigned to each edge are undirected.
 A \emph{cut} in the flow network $G$ is a subset $C \subset V$ of vertices with its complement $\bar{C} \subset V$ such that the source vertex $v_0 \in C$ and the sink vertex $v_{n+2} \in \bar{C}$.
@@
 That is, $W(C,\bar{C})$ is the sum of the weights on the edges crossing from the vertices in $C$ to the vertices in $\bar{C}$.
-In what follows we will often drop the argument and write $W$ rather than $W(C,\bar{C})$.
+In what follows we often drop the argument and write $W$ rather than $W(C,\bar{C})$.
 A \emph{minimum cut} is a $C$ and $\bar{C}$ that minimise the weight $W$.
@@
 We may choose any values for the $w_{ii}$ without affecting $W$.
-We will find it convenient to set $w_{0,0} = w_{n+2,n+2} = 0$.
-The remaining $w_{ii}$ we shall specify shortly.
+It will be convenient to set $w_{0,0} = w_{n+2,n+2} = 0$.
+The remaining $w_{ii}$ shall be specified shortly.
 The edge $e_{0,n+2}$ is in every cut.
@@
 In particular, the subsets $C$ and $\bar{C}$ corresponding to a minimum cut are not changed.
-We will find it convenient to choose $w_{0,n+2} = w_{n+2,0} = 0$.
+It will be convenient to choose $w_{0,n+2} = w_{n+2,0} = 0$.
 
@@
 In this paper we have focused on the class of lattices of Voronoi's first kind.
-We have shown that computing a closest point in a lattice of Voronoi's first kind can be achieved in a comparatively modest number of operations of order $O(n^4)$.
+We have shown that computing a closest point in a lattice of Voronoi's first kind can be achieved in a comparatively modest number of operations of order $O(n^4)$ under the assumption that the obtuse superbasis is known.
 Besides being of theoretical interest, the algorithm has potential for practical application.

### 2014-07-13 reviews/response.tex
minor edits in response to Vaughan

@@
 Lattices of dimension larger than 3 need not be of Voronoi's first kind.
-An interesting question is: given a lattice, is it possible to efficiently decide whether it is of Voronoi's first kind?
-A related question is: is it possible to efficiently find an obtuse superbasis if it exists?
-It is suspected that the answer to this second question is no.
-An efficient solution would yield a solution to a known problem, that of determining whether a lattice is rectangular (has a basis consisting of pairwise orthogonal vectors) given an arbitrary basis~\cite{Lenstra_Silverberg_revisting_gentra_szydlo_2014}.
-It is suspected that this simpler problem is difficult.''
+Given a lattice, is it possible to efficiently decide whether it is of Voronoi's first kind?
+Is it possible to efficiently find an obtuse superbasis if it exists?
+It is suspected that the answer to this second question is no because an efficient solution would yield a solution to a known problem, that of determining whether a lattice is rectangular (has a basis consisting of pairwise orthogonal vectors) given an arbitrary basis~\cite{Lenstra_Silverberg_revisting_gentra_szydlo_2014}.''  %It is suspected that this simpler problem is difficult.''
 

### 2014-07-13 paper.tex
minor edits in response to Vaughan

@@
 Lattices of dimension larger than 3 need not be of Voronoi's first kind.
-An interesting question is: given a lattice, is it possible to efficiently decide whether it is of Voronoi's first kind?
-A related question is: is it possible to efficiently find an obtuse superbasis if it exists?
-It is suspected that the answer to this second question is no.
-An efficient solution would yield a solution to a known problem, that of determining whether a lattice is rectangular (has a basis consisting of pairwise orthogonal vectors) given an arbitrary basis~\cite{Lenstra_Silverberg_revisting_gentra_szydlo_2014}.
-It is suspected that this simpler problem is difficult.
+Given a lattice, is it possible to efficiently decide whether it is of Voronoi's first kind?
+Is it possible to efficiently find an obtuse superbasis if it exists?
+It is suspected that the answer to this second question is no because an efficient solution would yield a solution to a known problem, that of determining whether a lattice is rectangular (has a basis consisting of pairwise orthogonal vectors) given an arbitrary basis~\cite{Lenstra_Silverberg_revisting_gentra_szydlo_2014}.
+%It is suspected that this simpler problem is difficult.
 

### 2014-07-22 paper.tex
minor edits to paper

@@
 The unwrapping of phase data for location estimation can be posed as a closest lattice point problem and this has been applied to the global positioning system~\cite{Teunissen_GPS_1995,Hassibi_GPS_1998}.
-The problem has also found applications to circular statistics~\cite{McKilliam_mean_dir_est_sq_arc_length2010}, single frequency estimation~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}, and related signal processing problems~\cite{McKilliam2007,Clarkson2007,McKilliam2009IndentifiabliltyAliasingPolyphase,Quinn_sparse_noisy_SSP_2012}.
+The problem has also found application to circular statistics~\cite{McKilliam_mean_dir_est_sq_arc_length2010}, single frequency estimation~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}, and related signal processing problems~\cite{McKilliam2007,Clarkson2007,McKilliam2009IndentifiabliltyAliasingPolyphase,Quinn_sparse_noisy_SSP_2012,McKilliam_pps_unwrapping_tsp_2014}.
 
@@
 Section~\ref{sec:latt-voron-first} describes lattices of Voronoi's first kind and their obtuse superbasis.
-In Section~\ref{sec:seri-relev-vect} it is shown that for these lattices, the series of relevant vectors results in a closest lattice point after at most $n$ terms.
+In Section~\ref{sec:seri-relev-vect} it is shown that for these lattices the series of relevant vectors results in a closest lattice point after at most $n$ terms.
 Section~\ref{sec:comp-clos-relev} shows that each relevant vector in the series can be computed in $O(n^3)$ operations by computing a minimum cut in an undirected flow network.
@@
 \begin{equation}\label{eq:relvectnearpointieq}
-\dotprod{(y - x)}{v} \leq \tfrac{1}{2} \dotprod{v}{v} \qquad \text{for all $v \in \relevant(\Lambda)$}.  
+%\dotprod{(y - x)}{v} \leq \tfrac{1}{2} \dotprod{v}{v} \qquad \text{for all $v \in \relevant(\Lambda)$}.  
+\|y - x\| \leq  \|y - x - v\|  \qquad \text{for all $v \in \relevant(\Lambda)$}.
 \end{equation}
@@
 On the other hand if $x_k$ is not a closest lattice point to $y$ we have $y - x_k \notin \vor(\Lambda)$ and from~\eqref{eq:relvectnearpointieq} there exists a relevant vector $v$ such that
-\[
-0 > \dotprod{v}{v} - 2\dotprod{(y - x_k)}{v}.
-\]
-Adding $\|y - x_k\|^2$ to both sides of this inequality gives
 \begin{align*}
-\|y - x_k\|^2 &> \dotprod{v}{v} - 2\dotprod{(y - x_k)}{v} + \|y - x_k\|^2 \\
-&= \|y - x_k - v\|^2 \\
-&\geq \arg\min_{ v \in \relevant(\Lambda) \cup \{\zerobf\}}\|y - x_k - v \|^2 \\
-&= \|y - x_k - v_k \|^2 \\
-&= \|y - x_{k+1}\|^2. 
+\|y - x_k\| &>  \|y - x_k - v\| \\
+&\geq \arg\min_{ v \in \relevant(\Lambda) \cup \{\zerobf\}}\|y - x_k - v \| \\
+&= \|y - x_k - v_k \| \\
+&= \|y - x_{k+1}\|. 
 \end{align*}
@@
 \begin{equation}\label{eq:superbasecond}
-b_1 + b_2 \dots + b_{n+1} = 0
+b_1 + b_2 \dots + b_{n+1} = \zerobf
 \end{equation}
@@
 The $q_{ij}$ are called the \emph{Selling parameters}~\cite{Selling1874}.
-It is known that all lattices in dimensions less than or equal to $3$ are of Voronoi's first kind~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.
+It is known that all lattices of dimension less than or equal to $3$ are of Voronoi's first kind~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.
 %%%%%%%%%%%%%%%%%%%%
@@
 u_{k+1} &= u_k + t_k \nonumber \\
-t_k &= \arg\min_{t \in \{0,1\}^{n+1}}\| B(z - u_k - t) \|^2 \label{eq:pvecmin}
+t_k &= \arg\min_{t \in \{0,1\}^{n+1}}\| B(z - u_k - t) \| \label{eq:pvecmin}
 \end{align}
@@
 Because $\rng(v) = 0$ all elements from $v$ are identical, that is, $v_1=v_2=\dots=v_{n+1}$.
-In this case $Bv = \sum_{i=1}^{n+1} v_n b_n = v_1\sum_{i=1}^{n+1}b_n = 0$
+In this case $Bv = \sum_{i=1}^{n+1} v_n b_n = v_1\sum_{i=1}^{n+1}b_n = \zerobf$
 as a result of the superbasis condition~\eqref{eq:superbasecond}.  Thus $x = x + Bv$ is a closest point to $y$. 
@@
 If $v_i = \max(v)$ then $h_i = 1$ and we need only show that $1-h_j \leq \max(v) - v_j$ for all $j$.
-If $v_j = \max(v)$ then $h_j = 1$ and the results holds since $1 - h_j = 0 = \min(v) - v_j$.
+If $v_j = \max(v)$ then $h_j = 1$ and the results holds since $1 - h_j = 0 = \max(v) - v_j$.
 Otherwise if $v_j < \max(v)$ then $1-h_j \leq 1 \leq \max(v) - v_j$ because $\max(v)$ and $v_j$ are integers.
@@
 \[
-\|B(z-u-g)\|^2 > \|B(z-u-h)\|^2 \geq \min_{t \in \{0,1\}^{n+1}}\|B(z - u - t)\|^2
+\|B(z-u-g)\| > \|B(z-u-h)\| \geq \min_{t \in \{0,1\}^{n+1}}\|B(z - u - t)\|
 \]
@@
 \[
-\|B(z - u_k - t_{k})\|^2 = \min_{t \in \{0,1\}^{n+1}}\|B(z - u_k - t)\|^2
+\|B(z - u_k - t_{k})\| = \min_{t \in \{0,1\}^{n+1}}\|B(z - u_k - t)\|
 \]
@@
 In this particular problem the lattice $A_n^*$ is involved.
-A fast closest point algorithm requiring only $O(n)$ operations exists for $A_n^*$~\cite{McKilliam2009CoxeterLattices,McKilliam2008b} and so the algorithm described in this paper is not needed in this particular case.
+A fast closest point algorithm requiring only $O(n)$ operations exists for $A_n^*$~\cite{McKilliam2009CoxeterLattices,McKilliam2008} and so the algorithm described in this paper is not needed in this particular case.
 However, there may exist other signal processing problems where lattices of Voronoi's first kind arise.

### 2014-07-22 reviews/response.tex
minor edits to paper
minor

@@
 \\\textbf{Response:}
-We thank the reviewer for their comments.  We response to the authors other concerns below.
+We thank the reviewer for their comments.
+The order of some paragraphs and lemmas has been changed in Section~\zref{sec:seri-relev-vect} with the intent of improving the readability of the proofs and presenting the intuition behind the proofs before formalities begin.
+We agree that a simpler proof would always be preferable, but, after considerable effort, have not found a simpler rigorous proof.
+We respond to the reviewers numbered comments below.
 
 \end{itemize}
-
-We now respond to the reviewers numbered comments.
 
@@
 We do not know of references to specific proofs that we can give for these (unlike Theorem~4.1 for which a proof is directly given in \cite[Theorem~3]{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}).
-The proof of Proposition 2.1 has been shortened in response to comment~\ref{com:rev2halfopen} of reviewer 2.
+The proof of Proposition 2.1 has been shortened in response to comment~\ref{com:rev2halfopen} of reviewer 2 and the proof of Proposition 2.2 has also been shortened.
 A result like Proposition 2.2 does appear informally in~\cite[Sec~5]{Shalvi_iterativeslicer_2009}.
@@
 
-The order of some paragraphs and lemmas has been changed in Section~\zref{sec:seri-relev-vect} with the intent of improving the readability of the proofs and presenting the intuition behind the proofs before formalities begin.
+%The order of some paragraphs and lemmas has been changed in Section~\zref{sec:seri-relev-vect} with the intent of improving the readability of the proofs and presenting the intuition behind the proofs before formalities begin.
 
@@
 \item\textbf{Comment:}  
-page 11, ``with be'' $\to$ ``will be'' Displayed equation before ``Putting p=...'', A | should be || (5.7) again, missing factor 2 before double summation, and second set should be T, not S. 
+page 11, ``with be'' $\to$ ``will be'' 
 \\\textbf{Response:}
-These errors have been fixed.
+Fixed.
+
+\item\textbf{Comment:}  
+Displayed equation before ``Putting p=...'', A $|$ should be $\|$
+\\\textbf{Response:}
+Fixed.
+
+\item\textbf{Comment:}  
+(5.7) again, missing factor 2 before double summation, and second set should be T, not S. 
+\\\textbf{Response:}
+Fixed.
 

### 2014-07-22 paper.tex
minor

@@
 \]
-The lemma will follow if we show that $B(u+g+w)$ is a closest lattice point to $y$ since then the lattice point $B(u+g)$ with be $(\ell-1)$-close to $y$.
+The lemma will follow if we show that $B(u+g+w)$ is a closest lattice point to $y$ since then the lattice point $B(u+g)$ will be $(\ell-1)$-close to $y$.
 The proof is by contradiction.
@@
 Because the $q_{ij}$ are nonpositive for $i \neq j$ the weights $w_{ij}$ are nonnegative for all $i \neq j$ with $i,j = 1,\dots,n+1$.
-As discussed the value of the weights $w_{ii}$ have no effect on the weight of any cut $W$ so setting $q_{ii} = - w_{ii}$ for  $i = 1,\dots,n+1$ is of no consequence.
+As discussed the value of the weights $w_{ii}$ has no effect on the weight of any cut $W$ so setting $q_{ii} = - w_{ii}$ for  $i = 1,\dots,n+1$ is of no consequence.
 Finally the weights $w_{i,n+2}$ and $w_{0,i}$ can be chosen so that both are nonnegative and

### 2014-07-27 reviews/response.tex
minor

@@
 
-We agree that these propositions are well known/straightforward, but would prefer to keep them so that the paper is self contained.
-We do not know of references to specific proofs that we can give for these (unlike Theorem~4.1 for which a proof is directly given in \cite[Theorem~3]{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}).
+We agree that these propositions are well known/straightforward, but would prefer to keep them.
+%We do not know of references to specific proofs that we can give for these (unlike Theorem~4.1 for which a proof is directly given in \cite[Theorem~3]{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}).
 The proof of Proposition 2.1 has been shortened in response to comment~\ref{com:rev2halfopen} of reviewer 2 and the proof of Proposition 2.2 has also been shortened.
-A result like Proposition 2.2 does appear informally in~\cite[Sec~5]{Shalvi_iterativeslicer_2009}.
-As far as we are aware, the remarkable algorithm in~\cite{MicciancioVoulgaris_deterministic_jv_2013} does not rely upon bounding the number of lattice points inside a sphere (or any other volume).
+%A result like Proposition 2.2 does appear informally in~\cite[Sec~5]{Shalvi_iterativeslicer_2009}.
+It would also be possible to instead used Micciancio and Voulgaris~\cite[Lemma~3.7]{MicciancioVoulgaris_deterministic_jv_2013}%As far as we are aware, the remarkable algorithm in~\cite{MicciancioVoulgaris_deterministic_jv_2013} does not rely upon bounding the number of lattice points inside a sphere (or any other volume).
 Instead, the procedure operates recursively on scaled versions of the lattice.
@@
 \item\textbf{Comment:}  
-Displayed equation before ``Putting p=...'', A $|$ should be $\|$
+Displayed equation before ``Putting p=...'', a $|$ should be $\|$
 \\\textbf{Response:}

### 2014-07-27 paper.tex
minor

@@
 
-\begin{theorem} \label{thm:revvecssuperbase} (Conway and Sloane~\cite[Theorem~3]{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}) Let $\Lambda$ be a lattice of Voronoi's first kind with obtuse superbasis $b_1,\dots,b_{n+1}$.
+\begin{theorem} \label{thm:revvecssuperbase} \emph{(Conway and Sloane~\cite[Theorem~3]{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases})} Let $\Lambda$ be a lattice of Voronoi's first kind with obtuse superbasis $b_1,\dots,b_{n+1}$.
 The relevant vectors of $\Lambda$ are of the form,

### 2014-08-26 reviews/proofedits.tex
added proofedits

\documentclass[a4paper,10pt]{article}
%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
%\documentclass[journal]{IEEEtran}

\usepackage{../mathbf-abbrevs}
\input{../defs}

\usepackage{zref-xr,zref-user}
\zexternaldocument*{../paper}

\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}

\usepackage{amsmath,amsfonts,amssymb, amsthm, bm}

%\usepackage[square,comma,numbers,sort&compress]{natbib}

\usepackage{color}
\newcommand{\comment}[1]{\textcolor{red}{#1}}

\newcommand{\sgn}{\operatorname{sgn}}
\newcommand{\sinc}{\operatorname{sinc}}
\newcommand{\rect}[1]{\operatorname{rect}\left(#1\right)}

%opening
\title{Finding a closest point in a lattice of Voronoi's first kind}
\author{Robby~G.~McKilliam, Alex Grant and I. Vaughan L. Clarkson
%\thanks{}
}

\renewcommand{\theenumii}{(\alph{enumii})}
\renewcommand{\labelenumii}{\theenumii}

\begin{document}

\begin{enumerate}

\item Page 1, AMS subject classifications: 11Y16

\item Page 2, paragraph 4, ``In this problem some ``advice'' about the lattice polynomial size is assumed to be given.''  This should be modified to ``In this problem some ``advice'' about the lattice is assumed to be given.''  That is, remove ``polynomial size''.

\item Page 3, the equation
\[
\rho = \frac{\|s\|}{2} = \frac{1}{2} \min_{x \in \Lambda / \{\zerobf\} } \|x\|
\]
The forward slash in $x \in \Lambda / \{\zerobf\}$ should be backslash like $x \in \Lambda \backslash \{\zerobf\}$.  The equation should appear as
\[
\rho = \frac{\|s\|}{2} = \frac{1}{2} \min_{x \in \Lambda \backslash \{\zerobf\} } \|x\|
\]

\item Page 14.  It is written:

\vspace{0.5cm}
\hspace{1cm}\parbox{10cm}{
The weight can now be written as
\begin{align*}
W(C,\bar{C}) = \sum_{i \in C} \sum_{j \in \bar{C}} w_{ij} = \sum_{i =0}^{n+2} \sum_{j =0}^{n+2} w_{ij} t_i (1 - t_j) = F(t),
\end{align*}
for example. Finding a minimum \dots
}
\vspace{0.5cm}

Please remove ``for example'' and end the equation with a full stop.  The text should appear as

\vspace{0.5cm}
\hspace{1cm}\parbox{10cm}{
The weight can now be written as
\begin{align*}
W(C,\bar{C}) = \sum_{i \in C} \sum_{j \in \bar{C}} w_{ij} = \sum_{i =0}^{n+2} \sum_{j =0}^{n+2} w_{ij} t_i (1 - t_j) = F(t).
\end{align*}
Finding a minimum \dots
}
\vspace{0.5cm}

\end{enumerate}

\end{document}

### 2014-10-21 talk/presentation.tex
modified the talk a little

@@
 \item Computing a short vector is called the \term{shortest vector problem}.
-\item Applications in cryptography and number theory.
+\item NP-hard for arbitrary lattices.
+\end{itemize}
+}
+
+\frame{
+\frametitle{The shortest vector problem}
+Longstanding applications in computer science:
+\begin{itemize}
+\item integer programming {\tiny (Lenstra~et.~al. 1983, Kannan 1987)}
+\item the factoring of polynomials over rationals {\tiny (Lenstra~et.~al. 1982)}
+\item check solvability by radicals {\tiny (Landau and Miller 1985)}
+\item solving low density subset sum problems {\tiny (Coster~et.~al. 1992)}
+\end{itemize}
+}
+
+\frame{
+\frametitle{The shortest vector problem}
+Central problem in modern cryptography enabling:
+\begin{itemize}
+\item efficient public key encryption {\tiny (Gentry~et.~al. 2008, Micciancio 2012)}
+\item post quantum cryptography {\tiny (Perlner~et.~al. 2009, Guneysu 2014, Peikert 2014)}
+\item efficient hash functions {\tiny (Peikert and Rosen 2006, Micciancio 2007)}
+\item fully homomorphic encryption {\tiny (Gentry~et.~al. 2009, 2010)}
+\end{itemize}
+}
+
+\frame{
+\frametitle{The shortest vector problem}
+Communications applications:
+\begin{itemize}
+\item minimum distance of lattice codes {\tiny (Conway and Sloane 1998, Erez~et.~al. 2004)}
+\item capacity analysis of network codes {\tiny (Nazer and Gastpar 2011)}
+\end{itemize}
+}
+
+\frame{
+\frametitle{The shortest vector problem}
+\begin{itemize}
 \item NP-hard for arbitrary lattices.

### 2014-10-22 talk/presentation.tex
added to talk

@@
 \begin{itemize}
-\item Computing a short vector is called the \term{shortest vector problem}.
+\item The computational task of finding a short vector is called the \term{shortest vector problem}.
 \item NP-hard for arbitrary lattices.
@@
 \frametitle{The shortest vector problem}
-Longstanding applications in computer science:
+
+Longstanding problem in number theory and computer science:
 \begin{itemize}
@@
 \item the factoring of polynomials over rationals {\tiny (Lenstra~et.~al. 1982)}
-\item check solvability by radicals {\tiny (Landau and Miller 1985)}
+\item checking solvability by radicals {\tiny (Landau and Miller 1985)}
 \item solving low density subset sum problems {\tiny (Coster~et.~al. 1992)}
+\end{itemize}
+}
+
+\frame{
+\frametitle{The shortest vector problem}
+
+Communications applications:
+\begin{itemize}
+\item minimum distance of lattice codes {\tiny (Conway and Sloane 1999, Erez~et.~al. 2004)}
+\item capacity analysis of network codes {\tiny (Nazer and Gastpar 2011)}
 \end{itemize}
@@
 \item post quantum cryptography {\tiny (Perlner~et.~al. 2009, Guneysu 2014, Peikert 2014)}
-\item efficient hash functions {\tiny (Peikert and Rosen 2006, Micciancio 2007)}
+%\item efficient hash functions {\tiny (Peikert and Rosen 2006, Micciancio 2007)}
 \item fully homomorphic encryption {\tiny (Gentry~et.~al. 2009, 2010)}
 \end{itemize}
-}
 
-\frame{
-\frametitle{The shortest vector problem}
-Communications applications:
-\begin{itemize}
-\item minimum distance of lattice codes {\tiny (Conway and Sloane 1998, Erez~et.~al. 2004)}
-\item capacity analysis of network codes {\tiny (Nazer and Gastpar 2011)}
-\end{itemize}
+Security depends upon the complexity of the shortest vector problem.
 }
@@
 \frametitle{The closest lattice point problem}
-Applications to:
+Communications applications:
 \begin{itemize}
-\item coding and quantisation,
-\item multi-antenna communication (MIMO),
-\item unwrapping of phase data for electronic distance measurement in GPS and surveying,
-\item single frequency estimation,
-\item polynomial phase estimation,
-\item circular statistics.
+\item spectrally efficient codes {\tiny (Conway and Sloane 1983, Erez~et.~al.~2004)},
+\item vector quantisation and compression {\tiny (Conway~and~Sloane~1982, Erez~et.~al.~2005)},
+\item multi-antenna communication (MIMO) {\tiny (El Gamal~et.~al.~2004, Wubben~et.~al.~2011)},
+\item equalisation of delay spread channels {\tiny (Zamir~et.~al.~2002, Damen~et.~al.~2003)},
+\item network coding {\tiny (Nazer and Gastpar 2011)},
+\item blind channel estimation {\tiny (Ryan~et.~al. 2007, McKilliam~2014)}.
 \end{itemize}
+
+}
+
+\frame{
+\frametitle{The closest lattice point problem}
+Signal processing applications:
+\begin{itemize}
+\item unwrapping of phase data for electronic distance measurement in GPS and surveying {\tiny (Tuenissen~1995, Hassibi~and~Boyd~1998)},
+\item single frequency estimation {\tiny (McKilliam~et.~al.~2010)},
+\item polynomial phase estimation {\tiny (McKilliam~et.~al.~2014)},
+\item circular statistics {\tiny (McKilliam~et.~al.~2014)}.
+\end{itemize}
+
 }
@@
 \begin{itemize}
+\item Applications require fast solutions to the closest lattice point problem.
 \item NP-hard for arbitrary lattices.
@@
 \begin{itemize}
+\item Are any existing cryptographic schemes weakened?
 \item Can good codes or quantisers be constructed from lattices of Voronoi's first kind?
-\item Do applications such as global positioning, phase unwrapping, etc., involve lattices of Voronoi's first kind?
+\item Can MIMO communications systems benefit?
+\item Do applications such as global positioning, frequency estimation, etc., involve lattices of Voronoi's first kind?
 \item Are there subfamilies of Voronoi's first kind that admit even faster algorithms?

### 2014-10-22 talk/presentation.tex
minor

@@
 \begin{itemize}
+\item sphere packing {\tiny (Voronoi 1908, Conway and Sloane 1999)}
 \item integer programming {\tiny (Lenstra~et.~al. 1983, Kannan 1987)}
@@
 \begin{itemize}
-\item minimum distance of lattice codes {\tiny (Conway and Sloane 1999, Erez~et.~al. 2004)}
+\item minimum distance of lattice codes {\tiny (Conway and Sloane 1982, Erez~et.~al. 2004)}
 \item capacity analysis of network codes {\tiny (Nazer and Gastpar 2011)}
@@
 \item Short vectors are easy to find in the \term{root lattices} $\ints^n$, $A_n$, and $D_n$.
-\item We will show that the problem is relatively easy to solve for lattices of \term{Voronoi's first kind}.
+\item We will show that the problem is relatively easy to solve for lattices of \term{Voronoi's first kind} {\tiny (Voronoi~1908, Conway~and~Sloane~1992, McKilliam~and~Grant~2012)}.
 \end{itemize}
@@
 \item spectrally efficient codes {\tiny (Conway and Sloane 1983, Erez~et.~al.~2004)},
-\item vector quantisation and compression {\tiny (Conway~and~Sloane~1982, Erez~et.~al.~2005)},
+\item quantisation and compression {\tiny (Conway~and~Sloane~1982, Erez~et.~al.~2005)},
 \item multi-antenna communication (MIMO) {\tiny (El Gamal~et.~al.~2004, Wubben~et.~al.~2011)},
@@
 \begin{itemize}
-\item Applications require fast solutions to the closest lattice point problem.
+\item Require fast solutions to the closest lattice point problem.
 \item NP-hard for arbitrary lattices.
@@
 \frametitle{Lattices of Voronoi's first kind}
-\begin{theorem}[Conway and Sloane (1992)]
+\begin{theorem}[Conway and Sloane 1992]
 Let $\Lambda$ be a $n$-dimensional lattice of Voronoi's first kind with obtuse superbase $b_1, \dots, b_{n+1}$.  The relevant vectors in $\Lambda$ are of the form 
@@
 \item Can good codes or quantisers be constructed from lattices of Voronoi's first kind?
-\item Can MIMO communications systems benefit?
+\item Can MIMO communication systems benefit?
 \item Do applications such as global positioning, frequency estimation, etc., involve lattices of Voronoi's first kind?

### 2015-07-02 talk/presentation.tex
mino

--- 2014-10-21 talk/presentation.tex
@@
 {\small
-Institute for Telecommunications Research \\ University of South Australia 
+Institute for Telecommunications Research \\ School of Information Technology and Mathematical Sciences \\ University of South Australia 
 \\ \vspace{0.2cm}
@@
 \item Computing a short vector is called the \term{shortest vector problem}.
-\item NP-hard for arbitrary lattices.
-\end{itemize}
-}
-
-\frame{
-\frametitle{The shortest vector problem}
-Longstanding applications in computer science:
-\begin{itemize}
-\item integer programming {\tiny (Lenstra~et.~al. 1983, Kannan 1987)}
-\item the factoring of polynomials over rationals {\tiny (Lenstra~et.~al. 1982)}
-\item check solvability by radicals {\tiny (Landau and Miller 1985)}
-\item solving low density subset sum problems {\tiny (Coster~et.~al. 1992)}
-\end{itemize}
-}
-
-\frame{
-\frametitle{The shortest vector problem}
-Central problem in modern cryptography enabling:
-\begin{itemize}
-\item efficient public key encryption {\tiny (Gentry~et.~al. 2008, Micciancio 2012)}
-\item post quantum cryptography {\tiny (Perlner~et.~al. 2009, Guneysu 2014, Peikert 2014)}
-\item efficient hash functions {\tiny (Peikert and Rosen 2006, Micciancio 2007)}
-\item fully homomorphic encryption {\tiny (Gentry~et.~al. 2009, 2010)}
-\end{itemize}
-}
-
-\frame{
-\frametitle{The shortest vector problem}
-Communications applications:
-\begin{itemize}
-\item minimum distance of lattice codes {\tiny (Conway and Sloane 1998, Erez~et.~al. 2004)}
-\item capacity analysis of network codes {\tiny (Nazer and Gastpar 2011)}
-\end{itemize}
-}
-
-\frame{
-\frametitle{The shortest vector problem}
-\begin{itemize}
+\item Applications in cryptography and number theory.
 \item NP-hard for arbitrary lattices.

### 2015-07-02 talk/presentation.tex
minor mods to talk

--- 2014-10-22 talk/presentation.tex
@@
 \begin{itemize}
-\item sphere packing {\tiny (Voronoi 1908, Conway and Sloane 1999)}
-\item integer programming {\tiny (Lenstra~et.~al. 1983, Kannan 1987)}
-\item the factoring of polynomials over rationals {\tiny (Lenstra~et.~al. 1982)}
-\item checking solvability by radicals {\tiny (Landau and Miller 1985)}
-\item solving low density subset sum problems {\tiny (Coster~et.~al. 1992)}
+\item sphere packing %{\tiny (Voronoi 1908, Conway and Sloane 1999)}
+\item integer programming %{\tiny (Lenstra~et.~al. 1983, Kannan 1987)}
+\item the factoring of polynomials over rationals %{\tiny (Lenstra~et.~al. 1982)}
+\item checking solvability by radicals %{\tiny (Landau and Miller 1985)}
+\item solving low density subset sum problems %{\tiny (Coster~et.~al. 1992)}
 \end{itemize}
@@
 \begin{itemize}
-\item minimum distance of lattice codes {\tiny (Conway and Sloane 1982, Erez~et.~al. 2004)}
-\item capacity analysis of network codes {\tiny (Nazer and Gastpar 2011)}
+\item minimum distance of lattice codes %{\tiny (Conway and Sloane 1982, Erez~et.~al. 2004)}
+\item capacity analysis of network codes %{\tiny (Nazer and Gastpar 2011)}
 \end{itemize}
@@
 \begin{itemize}
-\item efficient public key encryption {\tiny (Gentry~et.~al. 2008, Micciancio 2012)}
-\item post quantum cryptography {\tiny (Perlner~et.~al. 2009, Guneysu 2014, Peikert 2014)}
+\item efficient public key encryption %{\tiny (Gentry~et.~al. 2008, Micciancio 2012)}
+\item post quantum cryptography %{\tiny (Perlner~et.~al. 2009, Guneysu 2014, Peikert 2014)}
 %\item efficient hash functions {\tiny (Peikert and Rosen 2006, Micciancio 2007)}
-\item fully homomorphic encryption {\tiny (Gentry~et.~al. 2009, 2010)}
+\item fully homomorphic encryption %{\tiny (Gentry~et.~al. 2009, 2010)}
 \end{itemize}
@@
 \item Short vectors are easy to find in the \term{root lattices} $\ints^n$, $A_n$, and $D_n$.
-\item We will show that the problem is relatively easy to solve for lattices of \term{Voronoi's first kind} {\tiny (Voronoi~1908, Conway~and~Sloane~1992, McKilliam~and~Grant~2012)}.
+\item We will show that the problem is relatively easy to solve for lattices of \term{Voronoi's first kind} %{\tiny (Voronoi~1908, Conway~and~Sloane~1992, McKilliam~and~Grant~2012)}.
 \end{itemize}
@@
 \begin{itemize}
-\item spectrally efficient codes {\tiny (Conway and Sloane 1983, Erez~et.~al.~2004)},
-\item quantisation and compression {\tiny (Conway~and~Sloane~1982, Erez~et.~al.~2005)},
-\item multi-antenna communication (MIMO) {\tiny (El Gamal~et.~al.~2004, Wubben~et.~al.~2011)},
-\item equalisation of delay spread channels {\tiny (Zamir~et.~al.~2002, Damen~et.~al.~2003)},
-\item network coding {\tiny (Nazer and Gastpar 2011)},
-\item blind channel estimation {\tiny (Ryan~et.~al. 2007, McKilliam~2014)}.
+\item spectrally efficient codes %{\tiny (Conway and Sloane 1983, Erez~et.~al.~2004)},
+\item quantisation and compression %{\tiny (Conway~and~Sloane~1982, Erez~et.~al.~2005)},
+\item multi-antenna communication (MIMO) %{\tiny (El Gamal~et.~al.~2004, Wubben~et.~al.~2011)},
+\item equalisation of delay spread channels %{\tiny (Zamir~et.~al.~2002, Damen~et.~al.~2003)},
+\item network coding %{\tiny (Nazer and Gastpar 2011)},
+\item blind channel estimation % {\tiny (Ryan~et.~al. 2007, McKilliam~2014)}.
 \end{itemize}
@@
 \begin{itemize}
-\item unwrapping of phase data for electronic distance measurement in GPS and surveying {\tiny (Tuenissen~1995, Hassibi~and~Boyd~1998)},
-\item single frequency estimation {\tiny (McKilliam~et.~al.~2010)},
-\item polynomial phase estimation {\tiny (McKilliam~et.~al.~2014)},
-\item circular statistics {\tiny (McKilliam~et.~al.~2014)}.
+\item unwrapping of phase data for electronic distance measurement in GPS and surveying %{\tiny (Tuenissen~1995, Hassibi~and~Boyd~1998)},
+\item single frequency estimation %{\tiny (McKilliam~et.~al.~2010)},
+\item polynomial phase estimation %{\tiny (McKilliam~et.~al.~2014)},
+\item circular statistics %{\tiny (McKilliam~et.~al.~2014)}.
 \end{itemize}
@@
 }
+
+% \frame{
+% \frametitle{The closest lattice point problem}
+% Central problem in modern cryptography enabling:
+% \begin{itemize}
+% \item efficient public key encryption %{\tiny (Gentry~et.~al. 2008, Micciancio 2012)}
+% \item post quantum cryptography %{\tiny (Perlner~et.~al. 2009, Guneysu 2014, Peikert 2014)}
+% %\item efficient hash functions {\tiny (Peikert and Rosen 2006, Micciancio 2007)}
+% \item fully homomorphic encryption %{\tiny (Gentry~et.~al. 2009, 2010)}
+% \end{itemize}
+
+% Security depends upon the complexity of the closest lattice point vector problem.
+% }
 
@@
 \item The number of iterations depends on $x_0$ and might be large.
-\item Minimising over the set of relevant vectors, that is computing
+\item Minimising over the set of relevant vectors, that is, computing
 \[
